### 2022-05-24 head/acknowledgements.tex
Add files via upload

\chapter*{Acknowledgements}
\markboth{Acknowledgements}{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}

I wish to express my deepest gratitude to my supervisor, Damir Filipović, for his guidance and continuous support during these five years. This thesis greatly benefited from his many valuable comments and flow of smart ideas. He guided me and I learned a lot from him about doing research. For this, I am truly grateful.

Special thanks go to professor Michael Rockinger who convinced me to pursue a doctorate degree. He took me under his wing since my Bachelor studies and helped me becoming the researcher I am today. This thesis would not exist without him.

I would like to acknowledge the rest of my thesis committee, the professors Pierre Collin-Dufresne, Andreas Fuster and Simon Scheidegger for their advice and comments.

I am also thankful to the professors Elena Perazzi and Damien Challet. I was their TA for several years and they helped me become a better teacher.

During these years I had the chance to work with extremely nice and intelligent people. I am thankful to my colleagues and friends Antoine Didisheim and Coralie Jaunin for the countless chess games and `bridge tours` around the campus.

I would like to thank my family: my parents Jacques and Sonia, my sister Isabelle and her husband Edward and my fiancée Cathy for their unconditional love and for bearing all my endless monologues about my latest algorithms. I dedicate this thesis to them.

\bigskip

\noindent\textit{Lausanne, \today}
\hfill M.-A.~D.

### 2022-05-24 head/settings_epfl_template.tex
Add files via upload

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%		Thesis Settings
%
%		EDOC Template
%		2011
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[a4paper,11pt]{book}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
% % Uncomment for bibliography
% % Bibliography using Biblatex
%\usepackage{doi}
%\usepackage[autostyle]{csquotes}
% \usepackage[
%     backend=biber,
%     style=authoryear,
%     natbib=true,
%     firstinits=true,
%     sortlocale=en_US,
%     url=false, 
%     doi=true,
%     eprint=false,
%     isbn=false
% ]{biblatex}
%\addbibresource{tail/bibliography.bib}
% % OR Bibliography management for Bibtex 
% Load natbib before babel
\usepackage[round]{natbib}

\usepackage[shorthands=off,french,german,english]{babel}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% EDOC THESIS TEMPLATE: Variant 1.0 -> Latin modern, large text width&height
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{lmodern} % use this to fix blurry typewriter text font
%\usepackage[a4paper,top=22mm,bottom=28mm,inner=35mm,outer=25mm]{geometry}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% EDOC THESIS TEMPLATE: Variant 2.0 -> Utopia, Gabarrit A (lighter pages)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{fourier} % Utopia font-typesetting including mathematical formula compatible with newer TeX-Distributions (>2010)
%\usepackage{utopia} % on older systems -> use this package instead of fourier in combination with mathdesign for better looking results
%\usepackage[adobe-utopia]{mathdesign}
\setlength{\textwidth}{146.8mm} % = 210mm - 37mm - 26.2mm
\setlength{\oddsidemargin}{11.6mm} % 37mm - 1in (from hoffset)
\setlength{\evensidemargin}{0.8mm} % = 26.2mm - 1in (from hoffset)
\setlength{\topmargin}{-2.2mm} % = 0mm -1in + 23.2mm 
\setlength{\textheight}{221.9mm} % = 297mm -29.5mm -31.6mm - 14mm (12 to accomodate footline with pagenumber)
\setlength{\headheight}{14pt}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{setspace} % increase interline spacing slightly
\setstretch{1.1}

\makeatletter
\setlength{\@fptop}{0pt}  % for aligning all floating figures/tables etc... to the top margin
\makeatother

\usepackage{graphicx,xcolor}
\graphicspath{{images/}}

%\usepackage{subfig}
\usepackage{booktabs}
\usepackage{lipsum}
\usepackage{microtype}
\usepackage{url}

\usepackage{fancyhdr}
\renewcommand{\sectionmark}[1]{\markright{\thesection\ #1}}
\pagestyle{fancy}
	\fancyhf{}
	\renewcommand{\headrulewidth}{0.4pt}
	\renewcommand{\footrulewidth}{0pt}
	\fancyhead[OR]{\bfseries \nouppercase{\rightmark}}
	\fancyhead[EL]{\bfseries \nouppercase{\leftmark}}
	\fancyfoot[EL,OR]{\thepage}
\fancypagestyle{plain}{
	\fancyhf{}
	\renewcommand{\headrulewidth}{0pt}
	\renewcommand{\footrulewidth}{0pt}
	\fancyfoot[EL,OR]{\thepage}}
\fancypagestyle{addpagenumbersforpdfimports}{
	\fancyhead{}
	\renewcommand{\headrulewidth}{0pt}
	\fancyfoot{}
	\fancyfoot[RO,LE]{\thepage}
}

\usepackage{listings}
\lstset{language=[LaTeX]Tex,tabsize=4, basicstyle=\scriptsize\ttfamily, showstringspaces=false, numbers=left, numberstyle=\tiny, numbersep=10pt, breaklines=true, breakautoindent=true, breakindent=10pt}

\usepackage{hyperref}
\hypersetup{pdfborder={0 0 0},
	colorlinks=true,
	linkcolor=black,
	citecolor=black,
	urlcolor=black}
\urlstyle{same}
\ifpdf
\usepackage[final]{pdfpages}
\else
\usepackage{calc}
\usepackage{breakurl}
\usepackage[nlwarning=false]{hypdvips}
\usepackage{backref}
\renewcommand*{\backref}[1]{}
\fi
\usepackage{bookmark}

\makeatletter
\renewcommand\@pnumwidth{20pt}
\makeatother

\makeatletter
\def\cleardoublepage{\clearpage\if@twoside \ifodd\c@page\else
    \hbox{}
    \thispagestyle{empty}
    \newpage
    \if@twocolumn\hbox{}\newpage\fi\fi\fi}
\makeatother \clearpage{\pagestyle{plain}\cleardoublepage}

%%%%% CHAPTER HEADER %%%%
\usepackage{color}
\usepackage{tikz}
\usepackage[explicit]{titlesec}
\newcommand*\chapterlabel{}
%\renewcommand{\thechapter}{\Roman{chapter}}
\titleformat{\chapter}[display]  % type (section,chapter,etc...) to vary,  shape (eg display-type)
	{\normalfont\bfseries\Huge} % format of the chapter
	{\gdef\chapterlabel{\thechapter\ }}     % the label 
 	{0pt} % separation between label and chapter-title
 	  {\begin{tikzpicture}[remember picture,overlay]
    \node[yshift=-8cm] at (current page.north west)
      {\begin{tikzpicture}[remember picture, overlay]
        \draw[fill=black] (0,0) rectangle(35.5mm,15mm);
        \node[anchor=north east,yshift=-7.2cm,xshift=34mm,minimum height=30mm,inner sep=0mm] at (current page.north west)
        {\parbox[top][30mm][t]{15mm}{\raggedleft \rule{0cm}{0.6cm}\color{white}\chapterlabel}};  %the empty rule is just to get better base-line alignment
        \node[anchor=north west,yshift=-7.2cm,xshift=37mm,text width=\textwidth,minimum height=30mm,inner sep=0mm] at (current page.north west)
              {\parbox[top][30mm][t]{\textwidth}{\rule{0cm}{0.6cm}\color{black}#1}};
       \end{tikzpicture}
      };
   \end{tikzpicture}
   \gdef\chapterlabel{}
  } % code before the title body
\titlespacing*{name=\chapter,numberless}{-3.7cm}{83.2pt-\parskip}{-3.2pt+\parskip}
\titlespacing*{\chapter}{-3.7cm}{50pt-\parskip-\parskip}{30pt+\parskip+\parskip}
\titlespacing*{\section}{0pt}{13.2pt}{1em-\parskip}  % 13.2pt is line spacing for a text with 11pt font size
\titlespacing*{\subsection}{0pt}{13.2pt}{1em-\parskip}
\titlespacing*{\subsubsection}{0pt}{13.2pt}{1em-\parskip}
\titlespacing*{\paragraph}{0pt}{13.2pt}{1em-\parskip}

\newcounter{myparts}
\newcommand*\partlabel{}
\titleformat{\part}[display]  % type (section,chapter,etc...) to vary,  shape (eg display-type)
	{\normalfont\bfseries\Huge} % format of the part
	{\gdef\partlabel{\thepart\ }}     % the label 
 	{0pt} % separation between label and part-title
 	  {\ifpdf\setlength{\unitlength}{20mm}\else\setlength{\unitlength}{0mm}\fi
	  \addtocounter{myparts}{1}
	  \begin{tikzpicture}[remember picture,overlay]
    \node[anchor=north west,xshift=-65mm,yshift=-6.9cm-\value{myparts}*20mm] at (current page.north east) % for unknown reasons: 3mm missing -> 65 instead of 62
      {\begin{tikzpicture}[remember picture, overlay]
        \draw[fill=black] (0,0) rectangle(62mm,20mm);   % -\value{myparts}\unitlength
        \node[anchor=north west,yshift=-6.1cm-\value{myparts}*\unitlength,xshift=-60.5mm,minimum height=30mm,inner sep=0mm] at (current page.north east)
        {\parbox[top][30mm][t]{55mm}{\raggedright \color{white}Part \partlabel \rule{0cm}{0.6cm}}};  %the empty rule is just to get better base-line alignment
        \node[anchor=north east,yshift=-6.1cm-\value{myparts}*\unitlength,xshift=-63.5mm,text width=\textwidth,minimum height=30mm,inner sep=0mm] at (current page.north east)
              {\parbox[top][30mm][t]{\textwidth}{\raggedleft \rule{0cm}{0.6cm}\color{black}#1}};
       \end{tikzpicture}
      };
   \end{tikzpicture}
   \gdef\partlabel{}
  } % code before the title body
\titlespacing*{\part}{11.06cm}{26.4pt-\parskip-\parskip}{0pt}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{mathtools}
% Fix the problem with delimiter size caused by fourier and amsmath packages.
\makeatletter
\def\resetMathstrut@{%
  \setbox\z@\hbox{%
    \mathchardef\@tempa\mathcode`\(\relax
      \def\@tempb##1"##2##3{\the\textfont"##3\char"}%
      \expandafter\@tempb\meaning\@tempa \relax
  }%
  \ht\Mathstrutbox@1.2\ht\z@ \dp\Mathstrutbox@1.2\dp\z@
}
\makeatother

### 2022-05-24 head/titlepage.tex
Add files via upload

\begin{titlepage}
\begin{otherlanguage}{french}
\begin{center}
%\large
\sffamily

\null\vspace{2cm}
{\huge Machine Learning and Risk Management} \\[24pt] 
\textcolor{gray}{\small{THIS IS A TEMPORARY TITLE PAGE \\ It will be replaced for the final print by a version \\ provided by the registrar's office.}}

\vfill

\begin{tabular} {cc}
\parbox{0.3\textwidth}{\includegraphics[width=4cm]{images/epfl}}
&
\parbox{0.7\textwidth}{%
	Thèse n. XXXX \the\year\\
	présentée le XXXX\\
	Collège du Management de la Technologie\\
	Chaire Swissquote en finance quantitative\\
	programme doctoral en finance\\
%
%	ÉCOLE POLYTECHNIQUE FÉDÉRALE DE LAUSANNE\\
	pour l'obtention du grade de Docteur ès Sciences\\
	par\\ [4pt]
	\null \hspace{3em} Marc-Aurèle Antoine DIVERNOIS\\[9pt]
%
\small
Proposition du jury:\\[4pt]
%
    Prof Name Surname, président du jury\\
    Prof Name Surname, directeur de thèse\\
    Prof Name Surname, rapporteur\\
    Prof Name Surname, rapporteur\\
    Prof Name Surname, rapporteur\\[12pt]
%
Lausanne, EPFL, \the\year}
\end{tabular}
\end{center}
\vspace{2cm}
\end{otherlanguage}
\end{titlepage}

### 2022-05-24 main/paper1.tex
Add files via upload

\chapter{A deep learning approach to estimate forward default intensities}

\section{Introduction}
\label{S:1}
The first default prediction models appeared forty years ago with the first generation model presented by \citet{altman}. This work led to the so-called Altman Z-score formula which uses accounting data to compute the default probability of a firm in the next two years. However, when used for financial firms, Altman's Z-score formula needs to be used with care  because, as I will discuss in this paper, financial firms have to be treated carefully due to their frequent use of off-balance sheet financing. Twenty years later, a second generation of reduced-form models used econometrical tools such as maximum likelihood, probit, and logit regressions. The major drawback of these models is that they do not provide multi-period forecasts. One innovative recent development is the use of doubly stochastic Poisson intensity model combined with multiple logit regression to account for multi-period default probability estimation. This model has been proposed by \citet{DSW} in \textit{Multi-period corporate default prediction with stochastic covariates}. The main contribution over prior work was to exploit the time-series dynamics of the explanatory covariates in order to estimate the likelihood of default over several future periods. Their model employs firm-specific and macroeconomic data to create a Markov state vector $X_t$ in order to compute independent firm default intensities $\lambda(t)$ and other types of exit intensities $\phi(t)$.
\citet{DSW} is the first model capable of multi-period default probability estimation using time dynamics of covariates $X_t$. Applications of \citet{DSW} works are various. We can find them in credit rating by credit rating agencies, banks who want to calculate the minimal amount of capital to be held and other researches analyzing the link between macroeconomic cycles and firm's default probabilities. Covariates used by \citet{DSW} are firm's distance to default, firm's trailing one-year stock return, three-month Treasury bill rate and trailing one-year return on S\&P500. Estimating their model on US-listed Industrial firms between 1980 and 2004, they find that distance to default and the current state of the economy have a significant impact on default hazard rates.

The two papers the closest to my paper are the work from \citet{DSW} and \citet{Duan2012}. The first model uses the doubly stochastic argument to derive multi-period default probabilities. To do so, it requires some strong assumptions (i.e Vectorial Autoregressive process) regarding the behavior of the time-series of covariates to generate future random values for the covariates. Obviously, if the process is misspecified, biases are introduced both in the forecasted covariates and in the future default probabilities. Five years later, \citet{Duan2012} shows that we can relax the VAR assumption with the use of forward intensities. This paper explains how we can reduce biases by projecting current event realizations on past data. For convenience, they specify the intensities as a linear function of state variables. I wish to extend the latter by removing the assumption of linear intensities and use an artificial neural network to estimate the intensities of the Poisson processes governing both default and other exits. 
Many papers on estimating default probabilities with machine learning techniques have already been written. For instance, \citet{barboza2017machine} test several machine learning models (support vector machines, bagging, boosting, and random forest) to predict bankruptcy one year prior to the event. They report a substantial improvement in prediction accuracy using these machine learning techniques especially when, in addition to the original Altman’s Z-score variables, six complementary financial indicators are included. To the best of my knowledge, such types of models do not rely on theoretical foundations and this is where the forward intensity model can contribute to the literature. It is able to provide multi-period predictions and is supported by a solid mathematical and econometrical background.

In \citet{DSW}, one of the main assumption is that the covariates governing both default and other exits intensities follow a high-dimensional vector auto-regressive (VAR) process. Using this type of process forces the model to greatly reduce either the number of firms in the sample or the number of state variables explaining firm attributes; if one does not restrict the number of firms or variables in the estimation, the dimension of the model will simply be too high and it will considerably increase computational time. A major step forward made in \citet{Duan2012} was to get rid off the VAR process in order to reduce computational time by using ``a new reduced-form approach based on a \textit{forward} intensity model''. These forward intensities produce a term structure of bankruptcy probabilities without using any sort of high-dimensional process. Using this method allows the model to incorporate a lot more state variables or individual firms in the sample. Moreover, \citet{Duan2012} state that their model may also improve robustness to misspecification because in a VAR model, estimation of future values are highly sensitive to any biases. On the other side, the forward intensity model approach is a ``direct projection of past data on current realizations'' which does not involve random estimation of future values. \\
Regarding covariates used, both \citet{Duan2012} and \citet{DSW} estimate their own Distance-to-Default (hereafter DtD). An important aspect to highlight is that distance to defaults specified in \citet{DSW} differ from those estimated in \citet{Duan2012} since the first one are estimated using the Variance Restriction method (see \textit{Measuring Distance-to-Default for Financial and Non-Financial Firms} \citep{Duan2012DTD}) and the second one using the transformed-data maximum likelihood estimation method to account for financial firms. The variance restriction method is a popular way to implement the \citet{Merton1974} model but fails to estimate properly the default point for financial firms. Following the KMV assumption (see \citet{KMV}), the default point in this method is specified as short term debt plus one half of long term debt and does not take into account other liabilities. However, it is well-known that financial firms such as banks specify a high portion of their debt as ``other liabilities". Hence, to include financial firms in the sample, the default point has to be adjusted to the sum of short term debt, one half of long term debt and a fraction $\delta$ of other liabilities. The method used by \citet{Duan2012} is using a maximum likelihood estimation in order to estimate the unknown fraction $\delta$. The Appendix provides additional methodological information on the DtD estimation using the variance restriction method and the maximum likelihood estimation. \\ 
To summarize, the forward intensity models require an assumption to link covariates to intensities. \citet{Duan2012} uses a linear assumption and a maximum likelihood to estimate those parameters. This paper contributes to previous literature by using neural networks to relax the linear assumption of forward intensities. Neural networks allow the estimation of highly non-linear functions without specifying the form of the relationships. The organization of the paper is as follows. Section 2 sets up the reduced-form model of default, develops the likelihood function used as loss function later on and describes the neural network approach. Section 3 discusses summary statistics of the dataset used. Section 4 presents the results. Section 5 concludes. Appendices at the end of the paper contains several proofs and more details on the Distance-to-Default estimation.

\section{Methodology}
\label{S:2}

\subsection{Default model}
\label{SS:2-1}

The model adds to the literature of reduced-form models of default for multiperiod corporate prediction using the doubly stochastic formulation as in \citet{DSW} (DSW henceforth) and \citet{Duan2012} (Duan henceforth). The default's time is modeled as the stopping time 
\begin{equation}
\tau_D = \inf\{t : N_t >0 , M_t = 0\},
\end{equation}
where $N_t$ and $M_t$ are the counting processes governing default and other exits respectively. Similarly, the stopping time for combined exits is denoted by
\begin{equation}
\tau_C = \inf\{t : N_t >0 \wedge M_t > 0\}.
\end{equation}

\noindent Naturally, we have the following :
\begin{equation*}
 \left.\begin{aligned}
        \text{if the firm exits due to default, } \tau_{Ci} = \tau_{Di},\\
        \text{if the firm does not exits due to default, } \tau_{Ci} < \tau_{Di}.
       \end{aligned}
 \right\}
 \qquad \text{$\Rightarrow \tau_{Ci} \leq \tau_{Di}$}
\end{equation*}

Let us denote by $Z_{it}$ the set of firm-specific variables at time t for the firm i and $Y_t$ the set of macroeconomic variables at time t. The first time of entry of the firm i is denoted  $t_i^0$. The econometrician's information set $\mathscr{F}_t$ at time t is thus
\begin{equation}
    \mathscr{F}_t = \{ Y_s : s \leq t \} \cup \mathscr{G}_{1t}  \cup \mathscr{G}_{2t} ... \cup \mathscr{G}_{Nt},
\end{equation}
where 
\begin{equation}
    \mathscr{G}_{it}  = \{  (1_{\tau_{C_i} < u} , 1_{\tau_{D_i} < u} , Z_{iu}) : t_i^0 \leq u \leq \min(\tau_D, \tau_C, t)  \}.
\end{equation}

\begin{figure}[H]
\label{lifespan}
\centering
    \begin{tikzpicture}[
dot/.style = {circle, fill=black,inner sep=0pt, minimum size=4pt},
every label/.append style = {inner sep=0pt, rotate around={0:(-0.5,1.5)}},
    thick      
                        ]
\draw ( 0,0.2) -- + (0,-0.4) node[below] {Start Data};
\draw (10,0.2) -- + (0,-0.4) node[below] {End Data};
%
\draw[thick] (0,0) -- node[below=2mm] {} + (10,0);
%
    \foreach \p in {0.1, 0.3, 0.4, 0.5, 0.6,0.7, 0.9}
{
    \node[dot] at (10*\p,0) {};
}
    \node[dot] at (10*0.1,0) {};
    \node[dot,label={$t_i^0$}] at (10*0.2,0) {};
    \node[dot,label={$\tau_{D_i}$}] at (10*0.8,0) {};

\end{tikzpicture}
\caption{Example of the lifespan of a firm. Each dot corresponds to a time period where the econometrician gathers firm-specific (if available) and macroeconomic variables. $t^0_i$ is the entry time of the firm i and $\tau_{D_i}$ denotes the default time of the firm i.}
\label{fig:lifespan}
\end{figure}

In Figure \ref{fig:lifespan}, each dot corresponds to a period. At each period, the econometrician gathers firm-specific variables (DTD, Cash/TA, NI/TA, ...) and macroeconomic variables (S\&P500 return, treasury rate). For a particular point in time t, the econometrician knows the whole time-serie of macroeconomic variables until t irrespective of $t_i^0$, and the time-series of firm-specific variables from $t_i^0$ to $\tau_{D_i}$ if $t_i^0 \leq t$ and $t \leq \tau_{D_i}$.
Following Proposition 1 in \citet{DSW}, the conditional probability of default within $s$ years can be computed as 
\begin{equation} \label{eq:3.6}
\mathbb{P}[ \tau_D <t+s |\mathscr{F}_t] = E_t \left[ \int_t^{t+s}  e^{-\int_t^z (\lambda(u)+\phi(u)) du} \cdot \lambda(z) dz \right].
\end{equation}
The probability of default is a function of intensities $\lambda$ (default) and $\phi$ (other exits). However, these intensities are unknown and unobservable. In DSW model, the state variables governing Poisson intensities are assumed to follow a specific vector autoregressive (VAR) process. This assumption is relaxed in Duan's paper by using forward intensity rates. Instead of modeling $\lambda_{it}$ and $\phi_{it}$ as some function of state variables available at time t, \citet{Duan2012} propose to deal with $f_{it}(\tau)$ and $g_{it}(\tau)$ directly as functions of state variables available at time t and the forward starting time of interest $\tau$. The analogy to interest rates would be that $\lambda_t$ is the short rate and $f_t(u)$ is the forward rate for horizon $u$. \citet{Duan2012} propose a model to predict corporate defaults at multiple horizons by estimating these forward intensities via maximum likelihood. To do so, they use a linear assumption in the relationship between the variables and the forward intensities (i.e $f_{it}(\tau) = \exp(\alpha_0(\tau) + \alpha_1(\tau) x_{it,1} + \alpha_2(\tau) x_{it,2} + ... + \alpha_k(\tau) x_{it,k}$). However, it is highly likely that default intensities depend on those covariates in a non-linear way. I propose to use an artificial neural network (ANN) to find the set of weights governing the process $f_{it}$ and $g_{it}$. I show that I am able to capture potential non-linear relationships between the state variables and the forward intensities which significantly improves forecasts.

\subsubsection{Forward intensities}
\label{SSS:2-1-1}

Since we do not have the exact knowledge of $\lambda$ and $\phi$, \citet{Duan2012} propose to use forward intensity rates. However, we need to translate the default probability given in equation \ref{eq:3.6} (which depends on spot intensities) into a formula that depends on forward intensities. To do so, we first compute the probability of surviving as a function of forward combined intensity, which will be later used to compute the probability of default as a function of both combined and default forward intensities. Let us denote by $F_{it}(\tau)$ the conditional distribution function of the combined (default and other exits) exit time evaluated at $t+\tau$. Hence, 1-$F_{it}(\tau)$ is the probability of surviving in the interval $[t, t+\tau ]$. Therefore, we have :
\begin{equation}
1-F_{it}(\tau) = \mathbb{E}[e^{-\int_t^{t+\tau} (\lambda(s)+\phi(s)) ds}].
\end{equation}
Then, let us introduce the quantity $\psi_{it}(\tau)$ to be :
\begin{equation} \label{eq:3.8}
\psi_{it}(\tau) \equiv -\frac{\ln(1-F_{it}(\tau))}{\tau} \equiv -\frac{\ln(\mathbb{E}[e^{-\int_t^{t+\tau} (\lambda(s)+\phi(s)) ds}])}{\tau}.
\end{equation}
Reverting equation \ref{eq:3.8} gives :
\begin{equation}\label{eq:psi}
e^{-\psi_{it}(\tau)\cdot\tau} = 1-F_{it}(\tau).
\end{equation}
Where $e^{-\psi_{it}(\tau)\cdot\tau}$ is again the survival probability. We now need to compute $\psi_{it}(\tau)\cdot\tau$. At this point, \citet{Duan2012} makes the assumption that $\psi_{it}$ is differentiable and defines the forward combined exit intensity as 
\begin{equation} \label{eq:3.10}
g_{it}(\tau) \equiv \frac{F_{it}'}{1-F_{it}}.
\end{equation}
Equation \ref{eq:3.10} comes from the definition of a hazard rate function. Referring  P.87 of \citet{book}, the definition of a hazard rate function is the following:
%\newtheorem*{mydef}{Definition}
\begin{definition}[Hazard rate]
Let denote by $\tau$ a stopping time and $F(T) \equiv \mathbb{P}[\tau \leq T]$ its distribution function. Assume that $F(T) < 1 \forall T$, and that F(T) has a density f(T). The hazard rate function h of $\tau$ is :
\begin{equation*}
h(T) \equiv \frac{f(T)}{1-F(T)}.
\end{equation*}
\end{definition}
A hazard rate is the local arrival probability of a stopping time per time interval. Please note that under suitable regularity conditions, intensities and hazard rates are closely similar. In particular, in our doubly-stochastic framework, hazard rates and intensities are equivalent. This is why in \citet{Duan2012}, we have $\lambda(t) = h(t)$ and the distinction between hazard rates and intensity is often not made. \\

\noindent The relation between $g_{it}(\tau)$ and $\psi_{it}(\tau)$  is given by :
\begin{align} \label{gphi}
    g_{it}(\tau) &= \frac{F_{it}'(\tau)}{1-F_{it}(\tau)} \nonumber \\
    &= \psi_{it}(\tau) + \psi_{it}'(\tau)\tau.
\end{align}

\noindent Then, we can compute the quantity $\psi_{it}(\tau) \tau$ that we were looking for as :
\begin{equation}
    \psi_{it}(\tau) \tau = \int_0^{\tau}g_{it}(s) ds.
\end{equation}

\noindent For an interested reader, proofs of the above formulations can be found in the Appendix. Hence, the probability of surviving over [t,t+$\tau$] is given by
\begin{equation} \label{survprob}
    \mathbb{P}[\tau_c > t+\tau |\mathbb{F}_t] = \exp(-\int_0^\tau g_{it}(s) ds).
\end{equation}

The forward default intensity for horizon $\tau$ is defined as the limit for a small time step of the probability of defaulting in this small time step given that the firm has survived. The probability is Bayesian and the forward default intensity denoted $f_{it}(\tau)$ is the following :
\begin{equation}
f_{it}(\tau) \equiv \dfrac{\lim_{\Delta t\to 0} \dfrac{\mathbb{P}[t + \tau < \tau_{Di}=\tau_{Ci} \leq t+ \tau + \Delta t]}{\Delta t}}{e^{-\psi_{it}(\tau)\cdot\tau}}.
\end{equation}
Hence, the probability of defaulting between t and t+$\tau$ is given by :
\begin{equation}\label{probdef}
\int_0^{\tau} e^{-\psi_{it}(s)s} f_{it}(s)ds = \int_0^{\tau} e^{-\int_0^s g_{it}(u) du} f_{it}(s)ds.
\end{equation}

\subsubsection{Likelihood function}
\label{SSS:2-1-2}

In this setup, the likelihood function depends on three types of probabilities (default, other exit and surviving) which themselves depend on the types of intensities (default and other exits). The negative log-likelihood function has to be adjusted to the neural network framework and can be used as objective function to be minimized as long as we feed mini-batches and not single data points (``online learning''). Mini-batch feeding is a common practice in machine learning papers and consists of splitting the available data in batches of fixed size. Then, each backward pass takes one batch to perform one gradient decent step.  \\

To allow further comparison with \citet{Duan2012},  I employ the same discretization of time: $t = 0, 1, 2,...$ and $\tau = 0, 1, 2,...$ are time sequences of one month increment. Similarly, $f_{it}(\tau)$ and $g_{it}(\tau)$ are forward intensities computed at time t for the period [t+$\tau$, t+$\tau$+1]. The use of the $\tau$ index is to account for multiperiod prediction. When $\tau = 0$, the forward intensity model is computing spot intensities. When we set $\tau = 1$, the forward intensity model produces estimates one step ahead, and so on so forth. I denote $X_{it} = (x_{it,1}, x_{it,2} ,...)$ the set of firm-specific and macroeconomic variables explaining both default and combined exit intensities. As specified in \citet{Duan2012} $f_{it}(\tau)$ and $g_{it}(\tau)$ are functions of $X_{it}$ and can be specified as any form of function as long as they satisfy the following constraints. Since combined exit intensity has to be greater or equal than default intensity, we need to make sure that the forms specified for $f_{it}(\tau)$ and $g_{it}(\tau)$ satisfy the following conditions : $f_{it}(\tau) \leq g_{it}(\tau)$, $f_{it}(\tau) > 0$, $g_{it}(\tau) > 0$. \\

I have designed two neural networks, one will be trained to compute $f_{it}$ and one will be trained to output $h_{it}$ where $g_{it} = f_{it} + h_{it}$. I impose non-negativity on outputs of both models such that the combined exit intensity will never be smaller than the default intensity for all horizons. Let us denote by $\lambda$ and $\mu$ the set of parameters (weights) tuned in the neural network for $f_{it}$ and $h_{it}$ respectively. $N^{(\lambda)}$ and $N^{(\mu)}$ stand for the output of the neural network for $f_{it}$ and $h_{it}$ respectively. The log-likelihood for prediction horizon $\tau$ is expressed as
\begin{equation} \label{decomp1}
\mathcal{L}(\lambda(s)) = \sum_{i=1}^N \sum_{t=0}^{T-s-1} \mathcal{L}_{i,t}(\lambda(s)), \qquad s = 0,1,...,\tau-1
\end{equation}

\begin{equation} \label{decomp2}
\mathcal{L}(\mu(s)) = \sum_{i=1}^N \sum_{t=0}^{T-s-1} \mathcal{L}_{i,t}(\mu(s)), \qquad s = 0,1,...,\tau-1
\end{equation}
where 
\begin{align} \label{small_lik_f}
\mathcal{L}_{i,t}(\lambda(s)) &= \underbrace{\textbf{1}_{t_{0i} \leq t,\tau_{Ci} > t + s +1}}_\textrm{(1)} \cdot (-N_{it}^{(\lambda)}(s)) \Delta t)\\ \nonumber
&+ \underbrace{\textbf{1}_{t_{0i} \leq t, \tau_{Di}=\tau_{Ci} \leq t+s+1}}_\textrm{(2)} \cdot \ln(1-\exp[-N_{it}^{(\lambda)}(s)\Delta t]) \\
&+ \underbrace{\textbf{1}_{t_{0i} \leq t, \tau_{Di} \neq \tau_{Ci}, \tau_{Ci} \leq t+s+1}}_\textrm{(3)} \cdot (-N_{it}^{(\lambda)}(s)\Delta t) \nonumber,
\end{align}

\begin{align} \label{small_lik_h}
\mathcal{L}_{i,t}(\mu(s)) &= \underbrace{\textbf{1}_{t_{0i} \leq t,\tau_{Ci} > t + s +1}}_\textrm{(1)} \cdot (-N_{it}^{(\mu)}(s) \Delta t)\\ \nonumber
&+ \underbrace{\textbf{1}_{t_{0i} \leq t, \tau_{Di} \neq \tau_{Ci}, \tau_{Ci} \leq t+s+1}}_\textrm{(3)} \cdot \ln(1-\exp(-N_{it}^{(\mu)}(s))\Delta t)).
\end{align}

The likelihoods above are a sum of indicator functions multiplied by their respective probability. The indicator function $\textbf{1}_{A<B}$ is equal to one if $A<B$ or zero if $A \geq B$. Hence, the likelihoods above are specifying three mutually exclusive indicator functions which define three independent cases during time interval $[t,t+\tau+1]$:
\begin{enumerate}
\item The firm does not exit the sample between t and $t+\tau$ and is considered as surviving. This case is specified as (1) in the likelihood because the combined exit time $\tau_{Ci}$ is not in the interval $[t,t+\tau+1]$.
\item The firm exits the sample due to default during the interval and is considered as defaulted. This case is specified as (2) because $\tau_{Ci}=\tau_{Di}$ when the firm exits due to default conjointly with $\tau_{Di}$ being in the interval $[t,t+\tau+1]$.
\item The firm exits due to other reasons specified as (3) during the interval since the stopping time $\tau_{Di} \neq \tau_{Ci}$ conjointly with $\tau_{Ci} \leq t+\tau+1$.
\end{enumerate}

Proof of the above formulation can be found in the Appendix. As in previous studies, the likelihood functions still exhibit the decomposable property which allows to estimate the model for each horizon of prediction independently.

Since the intensities are directly driven by the covariates, \citet{Duan2012} requires an assumption on the mapping from the covariates to these intensities. In \citet{Duan2012} the mapping is made with a linear assumption, whereas in the framework of this paper, the mapping depends on the whole architecture of the neural network. When the neural network has only one hidden layer of one neuron coupled with an exponential activation function, the model boils down to \citet{Duan2012} as the intensities will be a linear combination of the covariates. As the width and depth of the network increases, we depart more and more from the linear assumption and we allow more non-linearities to be incorporated in these intensities. \\

\subsection{Neural Networks}
\label{SS:2-2}

Neural networks can be seen as a very general function to map a given input (in this case a set of macroeconomic and firm-specific variables) into a desired output (forward intensities). It learns how to compute the output by tuning weights in order to minimize a given loss function. A neural network is constructed by juxtaposing several ``hidden'' layers of neurons. The input of each layer is a data transformation of the output of the previous layer. Initially, the weights of the network are assigned random values. Then, the training process starts and consists of many iterations of a forward pass and a backward pass. The forward pass takes as input a batch of data and computes the loss value, the backward pass then computes the gradient and adjust the weights of the network based on a learning rate hyperparameter. As an illustration, figure \ref{fig:net} shows how a neural network looks like with 2 hidden layers : 5 neurons in the first layer and 3 neuron in the second layer. In this example, 3 features (inputs) are fed to the network. Each feature is connected to the first hidden layer by a set of weights. The outputs of the first layer are also weighted to produce the inputs of the second hidden layer. Non-linearity is introduced in each node with a non-linear activation function (i.e sigmoid). Finally, the output of the second hidden layer is aggregated to produce the final output of the model. The neural networks in this paper are implemented in Python using the library TensorFlow. Approximately one hour of computing time is needed to fit all networks on a 32GB RAM quad core 2.7 GhZ computer, and GPU computing is not necessary as the networks are rather small.

\begin{figure}[H]
    \centering
    \begin{tikzpicture}[
   shorten >=1pt,->,
   draw=black!50,
    node distance=\layersep,
    every pin edge/.style={<-,shorten <=1pt},
    neuron/.style={circle,fill=black!25,minimum size=17pt,inner sep=0pt},
    input neuron/.style={neuron, fill=green!50},
    output neuron/.style={neuron, fill=red!50},
    hidden neuron/.style={neuron, fill=blue!50},
    annot/.style={text width=4em, text centered}
]
    % Draw the input layer nodes
    \foreach \name / \y in {1,...,3}
    % This is the same as writing \foreach \name / \y in {1/1,2/2,3/3,4/4}
    \node[input neuron, pin=left:Input \#\y] (I-\name) at (0,-\y-2.5) {};  %%% <-- MODIFIED
    % set number of hidden layers
    \newcommand\Nhidden{2}

    % Draw the hidden layer nodes

       \foreach \y in {1,...,5} { %%% MODIFIED (1,...,12 -> 1,...,5, and the next five lines)
           \node[hidden neuron] (H1-\y) at (1*\layersep,-\y*\nodeinlayersep) {$\frac{1}{1+e^{-x}}$};
       }
    \node[annot,above of=H1-1, node distance=1.5cm] (hl1) {Hidden Layer 1};

        \foreach \y in {1,...,3} { %%% MODIFIED (1,...,12 -> 1,...,5, and the next five lines)
           \node[hidden neuron] (H2-\y) at (2*\layersep,-40-\y*\nodeinlayersep) {$\frac{1}{1+e^{-x}}$};
       }
    \node[annot,above of=H2-1, node distance=2.9cm] (hl2) {Hidden Layer 2};

    % Draw the output layer node
    \node[output neuron,pin={[pin edge={->}]right:Output}, right of=H\Nhidden-2] (O) {}; %%% <-- MODIFIED (from H\Nhidden-6 to H\Nhidden-3) 
    % Connect every node in the input layer with every node in the
    % hidden layer.
    \foreach \source in {1,...,3}
        \foreach \dest in {1,...,5} %%% <-- MODIFIED (1,...,12 -> 1...,3,5)
            \path (I-\source) edge (H1-\dest);
    % connect all hidden stuff
    \foreach [remember=\N as \lastN (initially 1)] \N in {2,...,\Nhidden}
       \foreach \source in {1,...,5} %%% <-- MODIFIED (1,...,12 -> 1...,3,5)
           \foreach \dest in {1,...,3} %%% <-- MODIFIED (1,...,12 -> 1...,3,5)
               \path (H\lastN-\source) edge (H\N-\dest);
    % Connect every node in the hidden layer with the output layer
    \foreach \source in {1,...,3} %%% <-- MODIFIED (1,...,12 -> 1...,3,5)
        \path (H\Nhidden-\source) edge (O);

    % Annotate the layers
    \node[annot,left of=hl1] {Input Layer};
    \node[annot,right of=hl\Nhidden] {Output Layer};
\end{tikzpicture}
    \caption{Illustration of a neural network [5, 3]. In this network, there are three input features, five neurons in the first hidden layer and 3 neurons in the second hidden layer. Each neuron is activated with a sigmoid function.}
    \label{fig:net}
\end{figure}

The use of neural networks can be motivated twofold. First, neural networks are well suited to approximate a function (in our case forward intensities) with the advantage of having different degrees of modularity. By definition, the architecture of the network generates the form of the function approximated. A deeper network allows for more non-linearities in the approximation of the function, at the expense of having more parameters to estimate. For instance, suppose that every observation comes with 12 input features (i.e : $x$ is a vector of shape 1x12), a [5, 3] (i.e 2 layers neural network with 5 neurons in the first hidden layer and 3 neurons in the second hidden layer) can be viewed as a function computing the output $N_{it}^{(\lambda)}$ in the following way :

\begin{equation*}
N_{it}^{(\lambda)} =  \begin{bmatrix}
\phi_1 (\begin{bmatrix}
\phi_2 (\begin{bmatrix}
x
\end{bmatrix}_{1\times 12} 
\begin{bmatrix}
w_1
\end{bmatrix}_{12\times 5} + 
\begin{bmatrix}
b_1
\end{bmatrix}_{1\times 5})
\end{bmatrix}_{1 \times 5 } 
\begin{bmatrix}
w_2
\end{bmatrix}_{5 \times 3} + 
\begin{bmatrix}
b_2
\end{bmatrix}_{1 \times 3})
\end{bmatrix}_{1 \times 3} 
\begin{bmatrix}
w
\end{bmatrix}_{3 \times 1},
\end{equation*}

\noindent  with the activation functions being for instance the sigmoid function $\phi_1(x) = \phi_2(x) = \frac{1}{1+e^{-x}}$, $x$ being the data input, $w_1$, $w_2$ and $w$ weight matrices and $b_1$ and $b_2$ biases matrices. In this setup, the number of parameters to estimate is equal to 12x5 + 1x5 + 5x3 + 1x3 + 3x1 = 86. \\
Neural networks are also well-suited for this paper because they allow an easy replication of the benchmark model \citet{Duan2012}. More specifically, if the activation function $\phi(x)$ is chosen as being an exponential $\exp(x)$, and the network architecture is [1] (i.e a single hidden layer with a single neuron), the output $N_{it}^{(\lambda)}$ becomes

\begin{align*}
N_{it}^{(\lambda)} &=  
\phi 
(
\begin{bmatrix}
x
\end{bmatrix}_{1 \times 12} \cdot
\begin{bmatrix}
w
\end{bmatrix}_{12 \times 1} +
\begin{bmatrix}
b1
\end{bmatrix}_{1 \times 1}
)\\
&= \exp(b1 + w_1 \cdot x_1 + w_2 \cdot x_2 + ... + w_{12} \cdot x_{12}),
\end{align*}

\noindent Results of this architecture are described in section 4.

%HERE
\section{Empirical section}
\label{S:3}

\subsection{Data}
\label{sec3-1}

The accounting data is taken from Wharton Research Data Services (WRDS) using the CRSP/Compustat merged database. The macroeconomic data is taken from CRSP, the Federal Reserve Bank Reports and Datastream. The bankrupcty data is taken from the Compustat database, using the DLRSN item for the reason of deletion and the DLDTE item for the date of deletion. DLRSN contains the code that indicates the reason a company become inactive on the database. I consider firms with a DLRSN code 2 (bankruptcy) or 3 (liquidation) to be defaulted, any other DLRSN code as ``other exits" and no DLRSN code as surviving. For additional information on DLRSN and DLDTE codes, please refer to the Wharton WRDS documentation. I focus on the period from 1991 to 2018 to match the accounting data with the bankruptcy data. Using WRDS database, I have downloaded accounting information for every company that has been listed someday on either NYSE, AMEX or NASDAQ between 1991 and 2018. This means I have 27 years of data where firms entered and/or exited anywhere in this sample. Using this kind of sample brings a problem of cylindric data : firm's entering/exiting time obviously are not the same for each company. We can see the whole dataset as a 3D matrix with x-axis being features, y-axis being time, and z-axis being firm. We only need to fill the matrix with NaN for elements where the firm $i$ doesn't exist or already left at time $t$. 
Since neural networks need plenty of data points to be well trained, I chose not to remove firms even if it has a short lifespan. When a variable is completely missing for a firm, I drop the whole firm because the likelihood is not specified if a variable is fully missing. However, when the variable is not fully missing but only some data points are not available, I use the last available information before the missing entry. I winsorize all variables at the 2.5 and 97.5 percentile. Before dropping firms, the dataset has in total 12527 surviving firms, 1453 defaulted firms and 17950 firms that exited the sample for other reasons. Finally, I standardize all variables by substracting the mean of the variable and dividing the result by its standard deviation. The test set is also standardized using the mean and standard deviation from the training set. Table \ref{tab:observ} shows the number of firms in the three categories for each horizon of prediction $\tau$ which corresponds to the firm-month observations used in the likelihood for horizon $\tau$ (see equations \ref{small_lik_f} and \ref{small_lik_h}). Figure \ref{fig:defsurv} shows the total number of firms that defaulted, survived or exited for other reasons plotted on a year on year basis.

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{def_surv_oe_plot.jpg}
    \caption{Number of firms defaulted, exited for other reasons, surviving each year}
    \label{fig:defsurv}
\end{figure}

\begin{table}[]
    \centering
    \begin{tabular}{l l l l}
    \hline\hline
Horizon & Surviving & Defaults & Other exits \\
    \hline
0 &  2'025'094 & 499 & 10'392\\
%1 &  2'007'311 & 497 & 10'384\\
%2 &  1'989'548 & 483 & 10'173\\
3 &  1'972'063 & 514 & 10'746\\
%4 &  1'953'935 & 511 & 10'743\\
%5 &  1'935'839 & 481 & 10'503\\
6 &  1'918'061 & 499 & 10'713\\
%7 &  1'900'041 & 495 & 10'712\\
%8 &  1'882'058 & 460 & 10'433\\
%9 &  1'864'422 & 468 & 10'611\\
%10 &  1'846'587 & 463 & 10'610\\
%11 &  1'828'772 & 445 & 10'324\\
12 &  1'811'323 & 454 & 10'393\\
%13 &  1'793'797 & 453 & 10'387\\
%14 &  1'776'301 & 433 & 10'037\\
%15 &  1'759'250 & 434 & 10'073\\
%16 &  1'742'157 & 432 & 10'070\\
%17 &  1'725'084 & 416 & 9'770\\
%18 &  1'708'387 & 417 & 9'796\\
%19 &  1'691'651 & 416 & 9'793\\
%20 &  1'674'942 & 391 & 9'434\\
%21 &  1'658'690 & 392 & 9'451\\
%22 &  1'642'418 & 390 & 9'446\\
%23 &  1'626'160 & 381 & 9'183\\
24 &  1'610'242 & 382 & 9'193\\
%25 &  1'594'311 & 381 & 9'189\\
%26 &  1'578'397 & 369 & 8'863\\
%27 &  1'562'910 & 370 & 8'871\\
%28 &  1'547'409 & 370 & 8'867\\
%29 &  1'531'923 & 358 & 8'605\\
%30 &  1'516'819 & 360 & 8'606\\
%31 &  1'501'703 & 359 & 8'601\\
%32 &  1'486'610 & 342 & 8'283\\
%33 &  1'471'935 & 343 & 8'296\\
%34 &  1'457'240 & 343 & 8'295\\
35 &  1'457'240 & 343 & 8'295 \\
\hline
    \end{tabular}
    \caption{Number of firms in each category for each horizon of prediction $\tau$ which corresponds to the firm-month observations used in the likelihood for horizon $\tau$}
    \label{tab:observ}
\end{table}

\citet{leippold} used a theoretical model to show that the most powerful default predictor must incorporate both macroeconomic and accounting data. For more transparency and to allow better comparison with previous literature, I chose to work with a similar set of features similar than my benchmark model \citet{Duan2012}. Firm-specific values are common to each firm, macroeconomics variables are function of market data. The exhaustive list of variables is the following :
\begin{enumerate}
\item SP500 : trailing 1-year return on the S\&500 index. 
\item Treasury : 3-month annualized US Treasury bill rate. I use this variable as the risk-free rate $r$ in the model. 
\item DtD : the Distance-to-default is a volatility adjusted leverage measure for gauging how far is the firm from default. It was first introduced by \citet{Merton1974} model which treat firm's equity value as a call option on the underlying asset and the strike being the debt level. The DtD can be viewed as the number of standard deviations of annual asset growth by which the asset level exceeds the firm's liabilities. In this paper, the DtD is estimated using a maximum likelihood taking into account other liabilities of each firm to handle financial firm's bias. It is well-known in the literature that DtD is a significant measure to estimate default probabilities but has to be used conjointly with other variables. See the Appendix for additional information regarding the estimation procedure of DtD.
\item CASH/TA : ratio between cash and short-term investment to total assets. Both quantities are taken from the balance sheet of the company.
\item NI/TA : ratio between net income to the total assets. A loss is registered as a negative net income.
\item SIZE : logarithm of the ratio between a firm market equity value to the average market equity value of the whole S\&P500. The market equity value is computed as the stock price multiplied by the number of outstanding shares at the end of each quarter. SIZE is negative if the firm has a smaller market capitalization than the average market capitalization and positive otherwise.
\item M/B : market-to-book ratio computed using the ratio between the asset market value from the DtD maximum likelihood estimation to the total book asset value.
\end{enumerate}

To capture momentum of variables, I also compute one step rolling window differences for each firm-specific variables. These variables are called ``$\Delta$" followed by the name of variable. They are telling whether the firm has been improving or deteriorating with respect to this particular variable comparing to the last period performance. Given this model specification, the $\Delta$ is particularly interesting because if a firm shows many consecutive negative delta values, it really means the company is in danger. However, it is also important to look at the level value to compare a defaulted firm with a non-defaulted firm. The intuition tells us that prior default time, a defaulted company should have shown lower level values (for instance, DtD) than a non-defaulted firm.

\subsection{Summary statistics}
\label{sec3-2}

This section depicts the summary statistics of the dataset. Table \ref{tab:meanvar} shows a summary of the means of variables for each three categories of firm. Please note that we have to be very careful when comparing two means in this table. Comparing means needs to be done with confidence intervals and tests of significance which involves standard deviations. This table is presented to give a rough idea without performing any statistical inference. The table shows that the average defaulted firm is smaller, has a lower DtD, a smaller Market-to-book ratio, loses more money and has less cash. The prefixes $\Delta$ in front of the variables stand for the one-lag differences. Finally, Figure \ref{fig:corrmat} shows the correlation matrix for the twelve covariates.

\begin{table}[]
    \centering
    \begin{tabular}{lrrr}
    \hline \hline
                 & Surviving &   Default  & Other exits \\
    \hline
Cash/TA          &  0.1952 &  0.1896  &    0.1851 \\
NI/TA            & -0.0756 & -0.2057  &   -0.129 \\
Size             & -2.7845 & -4.6743  &   -3.621 \\
DtD              & 10.641 &  6.441  &    8.401 \\
MBratio          &  2.3774 &  1.6995  &    2.202 \\
$\Delta$ CASH/TA  & -0.0026 &  0.0023  &   -0.0074 \\
$\Delta$   NI/TA     & -0.0019 & -0.0433  &   -0.0148 \\
$\Delta$  Size      & -0.0305 & -0.5095  &   -0.0939 \\
$\Delta$   DtD       & -0.1366 & -0.7346  &    0.0295 \\
$\Delta$  MBratio   & -0.0285 &  0.0520  &    0.0801 \\
\hline
    \end{tabular}
    \caption{Mean of variables for surviving firms, defaulted firms and other exits. $\Delta$ are one period lagged differences.}
    \label{tab:meanvar}
\end{table}

\begin{figure}[]
    \centering
    \includegraphics[scale=0.6]{corrmatrix.jpeg}
    \caption{Correlation matrix for firm-specific and macroeconomic covariates}
    \label{fig:corrmat}
\end{figure}

\section{Results}
\label{sec4}

As every neural network, we need to chose the optimal hyperparameters to achieve the highest performance. In my setup, this consists mainly of choosing the architecture of the model (i.e the number of neurons and layers). To do so, there is currently no other better method than trial and error. However, I need to be careful to not chose hyperparameters that overfit the test set. To avoid any overfitting, I perform a 5-fold cross validation for each horizon of prediction. I cut 15\% of the dataset as test set, all results that I will talk about in this chapter are out-of-sample and performed on the observations of the test set that the model has never seen before. The remaining set of observations is partitioned into smaller subsets so that in every fold of the validation a different subset is used as validation set and the rest is used as training set. Finally, to measure the discriminatory power of the models, I use the Lorenz curve (\citet{lorenz}) and I use the Gini coefficient as a scalar performance measure to aggregate across folds to get the  measure that I will use to discriminate models. \\

%Cumulative accuracy profile (CAP), also known as power curve, is often employed todetermine the performance of a rating/default prediction model. CAP is only concernedwith rankings and totally ignores the degrees of riskiness. To produce a CAP, we first lineup the obligors ranging from most risky to least risky.  Then, set a percentage and takea group of most risky obligors corresponding to this chosen percentage.  Finally, identifydefaulted obligors in this group and compute the percentage represented by these defaultedones in the population of all defaulted obligors. CAP is a plot that relates the percentageamong defaulted obligors to the percentage among the ranked obligors. In a large sample,the CAP corresponding to a perfect risk ranking model should quickly rise to one and levelat one. In contrast, a completely uninformative risk ranking model will have the CAP plotas a line with the slope equal to one.  Figure 4 presents the CAPs corresponding to thestandard and hierarchical intensity models

\begin{definition}[Lorenz curve]
The Lorenz curve of a predictor P is the two-dimensional graph
\begin{equation*}
    (\mathbb{P}\{P \leq p\},\mathbb{P}\{P\leq p | Y = 1\}),
\end{equation*}
$\forall p \in (-\infty,+\infty)$.
\end{definition}

The Lorenz curve plots on the x-axis the cumulative percentage of observations against the fraction of defaults on the y-axis.
These curves are often used in the literature for default prediction (for instance \citet{leippold}, \citet{Duan2012}) and I have seen very similar plots under two other names : power curves and cumulative accuracy profiles. They are different from ROC curves and precision-recall curves because they don't rely on thresholds to discriminate of true positives against false positives. I believe that they are particularly well-suited as performance measure in this model because of the multiperiod framework involved. The idea is that if the model is outputting a false positive for an horizon x but the true positive is horizon x+1, the model should not be ``too much" penalized. The ROC and precision-recall curves would treat this as a false positive even though the prediction was not that far from the target. The idea behind the Lorenz curve is to order default probabilities and look how they are distributed across defaulted and non-defaulted firms. We can easily see whether the model is outputting high probabilities for defaulted firms and small probabilities for surviving firms. \\
Finally, the Gini coefficient is used as my scalar summary statistic to compare models. It measures the degree of inequality of the Lorenz curve. A perfect model has a Gini coefficient close to 1 (depending on the fraction of defaults in the dataset) and a poor model has a Gini coefficient of 0 (perfect equality).

\subsection{Model choice}
\label{sec4-1}

The choice of the optimal architecture of the neural network has been made using k-fold cross-validation. The average Gini coefficient across all folds for every horizon is used as comparison tool to determine the best model architecture. If the architecture is too deep, the implicit function computed in the network to output forward intensities incorporates too many parameters and the risk of overfitting is larger, resulting in a lower accuracy measure. Similarly, if the network is not deep enough, the forward intensities are computed using a too simplistic representation and result in a low accuracy measure as well. This is usually known in the machine learning literature as the bias-variance tradeoff. Figure \ref{fig:ginicomp} shows the average Gini coefficient compute on the validation test across all folds of the cross validation for each horizon. I compare the scores obtained with different network architectures with those using the linear assumption. In this setup, one can interpret ``architecture'' as ``non-linearity degree'' as higher architecture involves more weights in the output function to be estimated. By disantangling the spaghettis, we can clearly see the bias-variance tradeoff. Increasing the depth of the networks from the simplest neural network [1] to a two layers network [2, 1] increases the Gini coefficients, in particular at mid horizons. This is probably due to the non-linearities introduced via the second layer. Next, it looks like the [3, 2] performs similarly as the [2, 1]; but we clearly see that [5, 3] dominates any other previous models. Finally, increasing the depth again to [10, 5] decreases Gini coefficient at all horizons, suggesting a severe overfitting of the forward intensities function. For next sections, I will now only show out-of-sample results of the [5, 3] architecture using the test set preliminary set on the side. \\

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{duan_vs_NN.jpeg}
    \caption{Gini coefficient : Linear vs Neural Networks}
    \label{fig:ginicomp}
\end{figure}

\subsection{Performance}
\label{sec4-2}

Figure \ref{fig:perf} shows Lorenz curves for horizon 0, 3, 6, 12, 24 and 35. Please note that all results generalize well to all others horizons. The curves are completely out-of-sample since they are computed on the test set that the model has never seen before. Table \ref{fig:ginicomp} shows Gini coefficients for both the linear assumption \citet{Duan2012} and for the [5, 3] for the same horizons. Overall, as expected the neural network clearly outperforms the linear assumption, suggesting that the linear assumption from \citet{Duan2012} can be greatly improved by adding non-linearities in the specification of the intensities. Unfortunately, it is difficult to tell which kind of non-linearities should be taken into account since neural networks are often seen as a black box. However, I will still try to answer this question by looking how the average intensity outputted by the model changes when we change a feature ceteris paribus (see section dedicated to ``sensitivities"). Another attempt at answering this question is described in the section ``computational graph'', where I deep dive into the weights of the network to understand how the output is computed.

For comparison purposes, figure \ref{fig:lor0duan} shows the Lorenz curves for both linear assumption, its replication in the neural network framework, and the [5, 3] model. The linear assumption and its replication have similar performance, suggesting again that the neural network ``NN+exp+[1]" is able to easily replicate the linear framework depicted in \citet{Duan2012}.

\begin{figure*}
        \centering
        \begin{subfigure}[b]{0.48\textwidth}
            \centering
            \includegraphics[width=\textwidth]{0.png}
            \caption*{Horizon 0}
        \end{subfigure}
        \quad
        \begin{subfigure}[b]{0.48\textwidth}  
            \centering 
            \includegraphics[width=\textwidth]{3.png}
            \caption*{Horizon 3}  
        \end{subfigure}
        \vskip\baselineskip
        \begin{subfigure}[b]{0.48\textwidth}   
            \centering 
            \includegraphics[width=\textwidth]{6.png}
            \caption*{Horizon 6}   
        \end{subfigure}
        \quad
        \begin{subfigure}[b]{0.48\textwidth}   
            \centering 
            \includegraphics[width=\textwidth]{12.png}
            \caption*{Horizon 12}   
        \end{subfigure}
                \vskip\baselineskip
        \begin{subfigure}[b]{0.48\textwidth}   
            \centering 
            \includegraphics[width=\textwidth]{24.png}
            \caption*{Horizon 24}
        \end{subfigure}
        \quad
        \begin{subfigure}[b]{0.48\textwidth}   
            \centering 
            \includegraphics[width=\textwidth]{35.png}
            \caption*{Horizon 35}  
        \end{subfigure}
        \caption{Out-of-sample Lorenz Curves for each horizon. The orange line shows the performance of Duan's model and the red line shows the performance of the [5,3] neural network.} 
        \label{fig:perf}
    \end{figure*}

\begin{table}[]
    \centering
    \begin{tabular}{lll}
\hline
Horizon	&	[5, 3]	&	Linear	\\
\hline \hline
0	&	0.85	&	0.70	\\
%1	&	0.83	&	0.66	\\
%2	&	0.80	&	0.71	\\
3	&	0.79	&	0.71	\\
%4	&	0.79	&	0.58	\\
%5	&	0.77	&	0.53	\\
6	&	0.76	&	0.64	\\
%7	&	0.78	&	0.61	\\
%8	&	0.77	&	0.68	\\
%9	&	0.70	&	0.58	\\
%10	&	0.76	&	0.57	\\
%11	&	0.76	&	0.70	\\
12	&	0.70	&	0.57	\\
%13	&	0.74	&	0.65	\\
%14	&	0.71	&	0.46	\\
%15	&	0.66	&	0.54	\\
%16	&	0.66	&	0.45	\\
%17	&	0.67	&	0.47	\\
%18	&	0.68	&	0.50	\\
%19	&	0.65	&	0.46	\\
%20	&	0.61	&	0.56	\\
%21	&	0.66	&	0.54	\\
%22	&	0.66	&	0.44	\\
%23	&	0.66	&	0.54	\\
24	&	0.66	&	0.44	\\
%25	&	0.58	&	0.47	\\
%26	&	0.63	&	0.52	\\
%27	&	0.63	&	0.50	\\
%28	&	0.63	&	0.48	\\
%29	&	0.68	&	0.48	\\
%30	&	0.62	&	0.49	\\
%31	&	0.56	&	0.51	\\
%32	&	0.62	&	0.46	\\
%33	&	0.62	&	0.38	\\
%34	&	0.62	&	0.32	\\
35	&	0.57	&	0.39	\\
\hline
    \end{tabular}
    \caption{Gini coefficients}
    \label{tab:ginicoeff}
\end{table}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{lor0duan.png}
    \caption{Comparison with the benchmark model \citet{Duan2012}. The green line shows the performance of the [1] neural network (i.e a single hidden layer with a single neuron) with an exponential activation function. The orange line shows the performance of Duan's model and the red line shows the performance of the [5,3] neural network.}
    \label{fig:lor0duan}
\end{figure}

\subsection{Computational graph}
\label{sec4-3}

Figure \ref{fig:compgraph} is a representation of a fully trained neural network for the forward default intensity $f$ at horizon 0. Negative weights are drawn as blue lines connecting neurons where orange lines show positive weights. The thicker the line, the higher the weight is in absolute value. Recall that each neuron is activated with a sigmoid function and the biases are not shown on the graph.
In the input layer, all variables seem to be used in the computation of the first layer. In the first layer however, the second neuron presents a higher weight in the network than the others. In return, the inputs connected to the second neuron of the first hidden layer all present low relative weights. Even though the computational graph gives an overview of the neural network and is useful to understand the way the output is computed, it is not trivial to see which variables have more impact on the output. In the following section, I plot sensitivities of each variable in each horizon to better understand the causality of each input.

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{compgraphr.png}
    \caption{Trained [5, 3] Neural Network for forward default intensity at horizon 0. Negative weights are drawn as blue lines connecting neurons where orange lines show positive weights. The thicker the line, the higher the weight is in absolute value.}
    \label{fig:compgraph}
\end{figure}

\subsection{Sensitivities}
\label{sec4-4}

Neural networks are often seen as black boxes because their outputs are coming from a general function involving many parameters. It is incorporating non-linearities via the layers and the activation functions. Figure \ref{fig:sensi} is an attempt at gauging how the model reacts to a change of an input variable. It plots the average default forward intensity against a shift in the specified variable and is computed the following way:

\begin{enumerate}
    \item Compute the forward default intensity for all the observations in the test set and average the result. Then, keeping everything else equal, change the value of one feature by an absolute value and feed the ``updated" observations to the network as new input and compute the new average forward default intensity. 
    \item Repeat step 1 for all absolute values in some interval
    \item Plot the average intensity against the absolute change
    \item Repeat step 1-3 for all 11 other features.
\end{enumerate}

Keep in mind that there is a non-negativity constraint on forward intensities. A decreasing relationship means that an increase of the associated variable decreases the probability of default of the firm. A flat relationship means that the associated variable has a limited impact on the probability of default. We should expect CASH/TA, size, DtD, NI/TA, Market-to-book ratio and all their lagged differences ($\Delta$) to be decreasing. \\

First of all, most graphs show non-linear relationships, which is not surprising given the nature of the neural network specification. Moreover, all relationships are intuitive and expected. Forward intensities in both Size and $\Delta$Size are decreasing, suggesting that small firms tend to have higher likelihood of defaulting which is consistent with the ``too big to fail" paradigm. Similarly, firms with decreasing cash ($\Delta$CASH/TA $<$ 0) or low levels of CASH/TA appear to have higher probability of default. The model also predicts that firms with low NI/TA or decreasing NI/TA should have higher probability of default. Finally, forward intensities in DtD should be decreasing for all horizons to reflect that a higher distance to default makes the firm less likely to default. At horizon 0, a negative change in distance-to-default has a substantially greater effect on forward intensities than any other variables. This result is consistent with \cite{DSW}. Overall,
it appears that the most important predictors of default in this model are in short term Market-to-book ratio, DtD and CASH/TA, and in the long run NI/TA, CASH/TA, $\Delta$CASH/TA and DtD.

\begin{figure*}
\centering
        \begin{subfigure}[b]{\textwidth}   
            \centering 
            \includegraphics[width=\textwidth]{legendsensi.JPG}
            {{\small }}    
        \end{subfigure}
        \quad
        \begin{subfigure}[b]{0.475\textwidth}
            \centering
            \includegraphics[width=\textwidth]{0_corr.png}
            \caption*{Horizon 0}
        \end{subfigure}
        \quad
        \begin{subfigure}[b]{0.475\textwidth}  
            \centering 
            \includegraphics[width=\textwidth]{3_corr.png}
            \caption*{Horizon 3}    
        \end{subfigure}
        \vskip\baselineskip
        \begin{subfigure}[b]{0.475\textwidth}   
            \centering 
            \includegraphics[width=\textwidth]{6_corr.png}
            \caption*{Horizon 6}  
            \captionsetup{position=top}
        \end{subfigure}
        \quad
        \begin{subfigure}[b]{0.475\textwidth}   
            \centering 
            \includegraphics[width=\textwidth]{12_corr.png}
            \caption*{Horizon 12}  
            \captionsetup{position=top}
        \end{subfigure}
                \vskip\baselineskip
        \begin{subfigure}[b]{0.475\textwidth}   
            \centering 
            \includegraphics[width=\textwidth]{24_corr.png}
            \caption*{Horizon 24}  
        \end{subfigure}
        \quad
        \begin{subfigure}[b]{0.475\textwidth}   
            \centering 
            \includegraphics[width=\textwidth]{35_corr.png}
            \caption*{\small Horizon 35}
        \end{subfigure}
        \caption{Sensitivities for each horizon.} 
        \label{fig:sensi}
    \end{figure*}

\section{Conclusion}
\label{S.5}

I propose an approach to estimate default forward intensities which relies on using machine learning techniques. The key improvement over previous estimation methods is the introduction of possible highly non-linear relationships between covariates and forward intensities. Neural networks are nothing else than a very general mapping of input data to an output which is obtained by tuning weights while minimizing a given loss function. Non-linearities are introduced via the juxtaposition of layers and the activation functions. The econometric model governing the forward intensity written by \citet{Duan2012} has been adapted to this new framework to allow the use of neural networks. Neural networks are also well-suited for this paper because they allow an easy replication of the benchmark model \citet{Duan2012}. More specifically, if the network architecture is [1] (i.e : one layer and one neuron) and the activation function is chosen as being an exponential $\exp(x)$, the neural network boils down to a logit regression. The dataset used in this paper is the same as previous literature to allow easier comparison. It consists of 5 firm-specific variables computed from accounting data and 2 macroeconomic variables to control for the health of the economy. I also account for momentum of these variables by feeding the model the one lagged differences of each variable. Looking at summary statistics only, the average defaulted firm is small, has low cash, low market-to-book ratio, low Distance-to-Default and has large and negative profits. To measure the discriminatory power of the models, I follow previous literature and use Lorenz curves (also known as ``Cumulative accuracy profile" or ``Power curves"). The idea behind Lorenz curves is to order default probabilities and look how they are distributed across defaulted and non-defaulted firms. The average Gini coefficient across all folds of the cross-validation is used as comparison tool to gauge the accuracy of the model. Results show that the architecture [5, 3] (i.e 2 layers with 5 neurons in the first hidden layer and 3 neurons in the second hidden layer) seems to outperform others architecture. In this setup, one can interpret ``architecture'' as ``non-linearity degree'' as higher architecture involves more weights in the output function to be estimated. Out-of-sample Lorenz Curves and Gini coefficients show that the neural network approach outperforms the linear assumption for every horizon, suggesting the presence of non-linearities in forward default intensities. Finally, even if neural networks are known as being black boxes, one can try to see how the model reacts to a change of input variables. Most sensitivities plots show non-linear relationships, which is not surprising given the nature of the neural network specification. It appears that the most important predictors of default in this model are in the short term Market-to-book ratio, Distance-to-default and NI/TA, and in the long run NI/TA, CASH/TA, $\Delta$CASH/TA and Distance-to-default. Further works could involve more variables in the estimation. In particular, a challenging (due to lack of data) but nonetheless exciting study would be to gauge the effect of market sentiment on default intensities.

### 2022-05-24 main/paper2.tex
Add files via upload

\chapter{Causal Networks with Neural Networks}

\section{Introduction}\label{S:1}

The causal network of a dynamical system provides important information that may help to better understand its behavior and ultimately design better policies to predict and control it. 
Large number of banks, insurances, hedge funds, and other financial institutions around the globe are interacting daily and thus their causal network is of great importance in econometrics.

There have been many attempts during the past decades to capture and visualize the network of interconnections among a set of financial institutions. 
The most widely used concept of causality in time series econometrics is due to \citet{granger}. 
This is based on statistical analysis of the financial series such as their stock prices over a finite time period. 
Granger’s definition of causality states that a time series $X$ is a cause of another time series $Y$, if the mean squared error of the 1-step ahead forecast for $Y$ is smaller when the history of $X$ is included in the forecasting information set. 
Otherwise, when the forecast does not improve by including the information of $X$, then it is declared that $X$ does not cause $Y$.
This idea is reflected in the information-theoretic measure that we use in this work to infer the causal interactions among a network of time series.

In the great majority of practical applications, Granger-causality has been studied in the context of Vector Autoregressive (VAR) models. 
For instance, \citet{billio2012econometric} proposes several measures  based on Granger-causality to capture the connections between the monthly returns of different financial institutions. 
It uses principle component analysis and ``pairwise'' Granger-causality tests to identify the causal networks. 
Other related works are \citet{diebold2014network} and \citet{barigozzi2016network} in which the authors propose connectedness measures based on generalized variance decomposition.
However, the measures introduced in these works are again limited to linear systems and they are based on pairwise comparison which as we show in Section  \ref{sec:dig} fails to infer the true causal relationships.
%and they are based on pairwise comparison which as we show in Section \ref{sec:dig}, it fail to infer the true causal relationships.
 %focuses on one particular network structure: the long-run variance decomposition network. Such networks are characterized by the infinite vector moving average and similar to \citet{diebold2014network}, they are limited to linear systems.

Contributions of this paper are both in network identification literature as well in finance.
%This paper adds contribution in the network identification literature as well as in finance literature. 
Our contribution to network identification are as follows:
\begin{itemize}
    \item We use an information-theoretic measure known as directed information (DI) to infer the Granger-causalities among a set of time series. 
This measure is non-parametric, i.e., it does not depend on the underlying model of the dynamics and it is capable of capturing causal relationships in both linear and nonlinear systems. The output of this approach is a directed graph known as Directed Information Graph (DIG) that visualizes the interconnections among a set of time series such as stock returns.

\item Computing DI has both high computational and sample complexity which makes it not suitable for inferring the causal structure of large networks. 
To overcome this problem, we develop a novel approach based on Recurrent Neural Networks (RNNs) that reduces the complexity of evaluating DIs in high-dimensional settings. 
\end{itemize}

%show that Directed Information is biased when the information set does not include all true parents, resulting in an imprecise network graph. To overcome this issue, we develop an algorithm to consistently estimate Directed Information while reducing complexity.

Applications of DIG are various in finance. 
In particular, we recommend to use it as a preliminary feature selection of another predictive model. Feature selection is a process often used in machine learning and statistics which consists of keeping only a subset of relevant features, usually to avoid overfitting or to reduce dimensionality. 
For example, \citet{fs1} and \citet{fs2} show that extraneous features are prone to reduce model's performance measures. 
Finance applications of feature selection models are various and include credit scoring, stock market behavior analysis  or even fraud detection (\citet{fs_fraud}).
\citet{tsai2009} states that feature selection preprocessing is not addressed carefully enough in the bankruptcy prediction literature. They compare five feature selection methods used in bankruptcy prediction: t-test, correlation matrix, step-wise regression, Principle Component Analysis (PCA) and factor analysis and show that any of these methods improves performance. 
\citet{fs_stock} uses a Sequential Forward Selection algorithm to select relevant features predicting the Turkish market index. The use of such feature selection model reduces model prediction error compared to the case where all features are used. This is due to information embedded in several economic factors already included in the market index. They show that only the lagged value of the market index is enough to predict the forthcoming value of the index.

%In this work, we use an information-theoretic measure known as directed information (DI) to infer the Granger-causalities among a set of time series. 
%This measure does not depend on the underlying model of the dynamic and it is capable of capturing causal relationships in both linear and nonlinear systems.
%However, computing DI has both high computational and sample complexity which makes it not suitable for inferring the causal structure of large networks. 
%To overcome this problem, we develop a novel approach based on Recurrent Neural Networks (RNNs) that reduces the complexity of evaluating DIs in high-dimensional settings.

\subsection{Related Work}
In recent years, several approaches have been developed to generalize the applicability of Granger-causality to non-linear and large dynamics. 
To mention a few, \citet{psaradakis2005markov} that introduces different terminologies for causality based on Granger’s ideas and provide a set of parametric non-causality constraints in the context of Markov switching VAR models.
In a similar context, \citet{bianchi2019modeling} investigates time-varying systemic risk based on a range of multi-factor asset pricing models and develops a Markov Chain Monte Carlo (MCMC) scheme to infer their model parameters and consequently obtain their corresponding networks.
Another attempt is \citet{bonaccolto2019estimation} in which the authors explore quantile-based methods of Granger causality.
%This work focuses on the linear factor model augmented with network dependence introduced by \citet{billio2018networks} and develops a method to identify causality among quantiles of the time series. 
This method is consistent with \citet{hong2009granger} and \citet{corsi2018measuring} that focus on causality among tail events. 
These methods are suitable for capturing causal relationships that are not in the center of their distributions, or in the mean but they are in the tails of their distributions.
It is important to emphasize that our proposed approach using DI is also capable of capturing such causal relations.

Most of the above aforementioned approaches are developed and tested for small size networks. 
Often, the problem of network identification in high dimensional settings requires more considerations and even its own techniques. 
For instance, \citet{billio2019bayesian} proposes a Bayesian non-parametric Lasso prior (BNP-Lasso) for high-dimensional VAR models that can improve efficiency and accuracy. 
In order to overcome overparametrization and overfitting issues in large VAR models, BNP-Lasso clusters the VAR coefficients into groups and shrinks the coefficients of each group toward a common location.
%Although, the proposed method in this work is suitable for inference in high dimensional VAR models but 
However, this method is limited to linear models with Gaussian innovations.
To overcome this limitation, \citet{kalli2018bayesian} proposes a Bayesian non-parametric approach that allows for nonlinearity in the conditional mean, heteroskedasticity in the conditional variance, and non-Gaussian innovations. However, unlike the BNP-Lasso, it does not allow sparsity in the model.
\citet{petrova2019quasi} proposes yet another non-parametric, quasi-Bayesian likelihood estimation methodology for high dimensional setting with time varying parameters. 
%This method is also non-parametric. 
The work in \citet{iacopini2019bayesian} tackles the curse of dimensionality by a two-stage approach. In the first stage, a spike-and-slab prior distribution is used for each entry of the coefficient matrix which also identifies the interconnection network.
In the second stage, it imposes prior dependence on the coefficients by specifying a Markov process for their random distribution. 
%A side product of this procedure is the interconnection network.
A closely related work is \citet{bernardi2019high} that proposes a shrinkage and selection methodology designed for network inference in high-dimensional settings. It uses a regularized linear regression model with spike-and-slab prior on the parameters.
However, both methods are limited to VAR models.

The rest of the paper is organized as follows. 
Section \ref{Sec1} reviews the notion of Granger causality and formally introduce directed information graphs which is suitable for linear and nonlinear systems. 
In Section \ref{sec:method}, we introduce a novel approach for inferring the Granger-causal network of high dimensional nonlinear systems. 
Finally, in Section \ref{sec:exp}, we apply our method to learn the causal network of both synthetic and real-world dataset. For the real-world experiment, we used the daily stock prices of major US firms.

%Afterwards, we evaluate the performance of our method in both synthetic linear and nonlinear settings and also apply it to infer the interconnections among the major US financial institutions.

\section{Causal Network}\label{Sec1}

In this section, we present a statistical approach to learn the causal interconnections in a dynamical systems based on Granger causality \citet{granger}. 
We begin by introducing some notations.
Plain capital letters denote random variables or processes, while lowercase letters denote their realizations. 
Bold letters are used for column vectors, matrices, and tensors and calligraphy letters are used for sets. 
We use $X_{j,t}$ to denote the value of a time series $X_j$ at time $t$ 
%and $X_{j,t_1}^{t_2}$, when $t_1\leq t_2$, to denote $X_j$ from time $t_1$ to $t_2$, namely, $(X_{j,t_1},...,X_{j,t_2})$. 
%Subsequently, 
and $X^{t}_j$ to denote the time series $X_j$ up to time $t$. 
For a set $\mathcal{A}=\{a_1,...,a_n\}$ and an index set $\mathcal{I}\subseteq\{1,...,n\}$, we define $\mathcal{A}_{-\mathcal{I}}:=\mathcal{A}\setminus\{a_i: i\in\mathcal{I}\}$.

\subsection{Granger Causality}\label{sec:granger}

Researchers from different fields have developed various frameworks and graphical models to capture and represent interconnections among variables or processes.
One of the most popular and widely used frameworks in economics is the notion of Granger causality.
The basic idea in this framework was originally introduced by Wiener \citet{wiener1956theory}, and later formalized by Granger \citet{granger}. 
The idea is as follows: ``we say that $X$ is causing $Y$ if we are better able to predict the future of $Y$ using all available information than if the information apart from the past of $X$ had been used."

Despite broad philosophical viewpoint of \citet{granger1963economic}, his formulation for practical implementation was done using multivariate autoregressive (MVAR) models and linear regression which has been widely adopted in econometric and other disciplines.
More precisely, in order to identify the influence of $X_t$ on $Y_t$ in a MVAR comprises of three time series $\{X,Y,Z\}$, Granger's idea is to compare the performance of two linear regressions: the first one predicts $Y_t$ given $\{X^{t-1},Y^{t-1},Z^{t-1}\}$ and the second one predicts $Y_t$ given $\{Y^{t-1},Z^{t-1}\}$. Clearly, the performance of the second predictor is bounded by the first predictor. If they have the same performance, then we say $X$ does not Granger cause $Y$. It is important to emphasize that this formulation is only applicable in linear systems.

To go beyond linear systems, works such as \citet{directed} and \citet{massey} use information-theoretical measures and generalize Granger causality. 
In this work, we introduce and apply directed information (DI) \citet{directed}, an information-theoretical tools to measure interconnections among firms.
DI has been used in many applications to infer causal relationships. For example, \citet{spike} and \citet{di12} used it for analyzing neuroscience data and \citet{acc2014}  and \citet{etesami2018econometric} applied to market data.

In order to visualize the inferred interconnections among time series using DI, directed information graphs (DIGs) have been developed \citet{directed}.  
DIGs are a type of graphical models in which nodes represent time series and arrows indicate the direction of causation. We use DIG to represent the causal network among the covered firms.

\subsection{Directed Information Graphs (DIGs)}\label{sec:dig}

In the rest of this section, we describe how the DI can capture the interconnections in causal\footnote{In causal systems, given the full past of the system, the present of the processes become statistically independent. In other words, there are no simulations relationships between the time series.} dynamical systems (linear or non-linear) and formally define DIGs.

Consider a dynamical system comprised of three time series $\{X,Y,Z\}$ that we assume they have a joint probability density function $p(X,Y,Z)$. 
To answer whether $X$ has influence on $Y$ or not over time horizon $[1,T]$,  following the idea of Granger, we compare the average performance of two particular predictors over this time horizon. 
The first predictor uses the history of all three time series while the second one uses the history of all processes excluding process $X$. 
On average, the performance of the predictor with less information (the second one) is upper bounded by the performance of the predictor with more information  (the first one). 
However,  if the prediction of both predictors are close over time horizon $[1,T]$,  it is an indication that $X$ does not cause $Y$ in this time horizon.
To rigorously formalize this idea, we need the predictors and a measure to compare their performances.

In the definition of DI,  the predictors belong to the space of probability measures.  
More precisely,  
%let $X^{t-1}$ represents the time series $X$ up to time $t-1$.
the prediction of the first predictor at time $t$ is $p(Y_t|Y^{t-1},Z^{t-1},X^{t-1})$ that is the conditional density function of $Y_t$ given the history of all time series. Similarly, the prediction of the second predictor is $p(Y_t|Y^{t-1},Z^{t-1})$ that is the conditional density function of $Y_t$ given the history of all time series except time series $X$.

Given the predictions of the first and the second predictors at time $t$ for an outcome $y_t\in \mathcal{Y}$, the goodness of these predictions are measured by the log-loss that are defined respectively by 
\begin{align*}
    &-\log p(Y_t=y_t|Y^{t-1},Z^{t-1},X^{t-1}),\\
    &-\log p(Y_t=y_t|Y^{t-1},Z^{t-1}).
\end{align*}
According to the above measures of goodness, the better the predictor is, the smaller its log-loss will be.
This loss function has meaningful information-theoretical interpretations.
Namely, the log-loss is the Shannon's code length\footnote{It is also called the description length of $y_t$. For more information see \citet{cover2012elements}.}, i.e., the number of bits required to efficiently represent $y_t$.

At time $t$ for an outcome $y_t\in \mathcal{Y}$, the difference between the log-losses of the two predictors compares their performances.  This difference is also called \textit{regret},
\begin{align}
r_t&:=-\log{p(Y_t=y_t|Y^{t-1},Z^{t-1})}-\big(-\log {p(Y_t=y_t|Y^{t-1},Z^{t-1},X^{t-1})}\big)\\
&=\log \frac{p(Y_t=y_t|Y^{t-1},Z^{t-1},X^{t-1})}{p(Y_t=y_t|Y^{t-1},Z^{t-1})}.
\end{align}
Note that the regrets are non-negative for all $t$ and all outcomes $y_t$.
The average regret over the time horizon $[1,T]$ is given by 
\begin{align}\label{eq:ave}
   \frac{1}{T}\sum_{t=1}^T \mathbb{E}[r_t],
\end{align}
where the expectation is taken over the joint density function\footnote{For the sake of notational simplicity, we use $p(y^{t},z^{t-1},x^{t-1})$ to denote $p(Y^{t}=y^{t},Z^{t-1}=z^{t-1},X^{t-1}=x^{t-1})$.} of $X$, $Y$, and $Z$, i.e., 
\begin{align}
    \mathbb{E}[r_t]=\int p(y^{t},z^{t-1},x^{t-1})\log \frac{p(y_t|y^{t-1},z^{t-1},x^{t-1})}{p(y_t|y^{t-1},z^{t-1})}dy^tdx^{t-1}dz^{t-1}.
\end{align}
The average regret in \eqref{eq:ave} is called \textit{directed information} (DI) and will be our measure of causation in this work.
This measure is always positive and if it is zero, it indicates that the history of time series $X$ contains no significant information that would help in predicting the future of time series $Y$ given the history of $Y$ and $Z$. 
This definition can be generalized to more than three time series as follows,

 \begin{definition}\label{cdi}
Consider a network of $m$ time series $\mathcal{R}=\{R_1,...,R_m\}$ with the joint probability density function $p$. The directed information from $R_i$ to $R_j$ over time horizon $[1,T]$ is given by
\begin{align}\label{cdif}
I(R_i\rightarrow R_j||\mathcal{R}_{-\{i,j\}}):=\frac{1}{T}\sum_{t=1}^T\mathbb{E}\left[\log\frac{p(R_{j,t}|\mathcal{R}^{t-1}_{-\{i,j\}},R_{j}^{t-1},R_{i}^{t-1})}{p(R_{j,t}|\mathcal{R}^{t-1}_{-\{i,j\}},R_{j}^{t-1})}\right],
\end{align}
where $\mathcal{R}_{-\{i,j\}}^{t-1}\!:=\!\{R_{1}^{t-1},...,R_{m}^{t-1}\}\!\setminus\!\{R_{i}^{t-1},R_{j}^{t-1}\}$.
We declare $R_i$ influences $R_j$ over time horizon $[1,T]$, if and only if
\begin{align}\label{cdif1}
I(R_i\rightarrow R_j||\mathcal{R}_{-\{i,j\}})>0.
\end{align}
 \end{definition}

 An interpretation of $R_i$ influencing $R_j$ is that varying $R_i$ will change the value of $R_j$ even if all the other variables within the network remains unchanged. 
 In another words, if $R_i$ does not influence $R_j$, then varying $R_i$ would not change $R_j$ when the values of the remaining times series are fixed. 
 This can be seen from the fact that DI compares two conditional distributions of $R_j$ over a time horizon of length $T$; one is given the history of all the time series while the other one is given all the history except the history of $R_i$. Thus, if DI in \eqref{cdif} is zero, then these two conditional distributions are equal over this time horizon. This implies that the history of $R_i$ does not contain any useful information for $R_j$.

It is important to emphasize that the definition of DI does not rely on any model assumption, thus DI is capable of inferring the causal relationships in general (linear or non-linear) dynamical systems. Next, we define the graphical model that we use in this work to visualize the causal network among firms.

\begin{definition}\label{DIG1}
Directed information graph (DIG) of a set of $m$ time series $\mathcal{R}=\{R_1,...,R_m\}$ is a directed graph $G=(\mathcal{V},\mathcal{E})$, where nodes represent time series ($\mathcal{V}=\mathcal{R}$) and arrow $(R_i,R_j)\in \mathcal{E}$ denotes that $R_i$ influences $R_j$. 
%with weight $w_{i,j}:=I(R_i\rightarrow R_j||\mathcal{R}_{-\{i,j\}})$. 
%Consequently, $(R_i,R_j)\notin \mathcal{E}$ if and only if its corresponding weight is zero.
\end{definition}

A simple way to represent the DIG $G$ of a dynamical system is via the adjacency matrix $\textbf{DIG}=[d_{i,j}]_{m\times m}$ that is defined by
\begin{align}
    d_{j,i} = \left\{
\begin{array}{ll}
      1 & \text{if}\  I(R_i\rightarrow R_j||\mathcal{R}_{-\{i,j\}})>0,\\
      0 & \text{otherwise}. \\
\end{array} 
\right. 
\end{align}
Given a DIG $G=(\mathcal{V},\mathcal{E})$, we define the parent set of node $R_j$ denoted by $\mathcal{PA}_j\subset \mathcal{V}$ to be the set of all times series that have direct influences on $R_j$, i.e., $\mathcal{PA}_j:=\{R_k: d_{j,k}=1\}$. Similarly, the children set of node $R_j$ is given by $\mathcal{CH}_j:=\{R_k: d_{k,j}=1\}$. 
Next example demonstrates the DIG of a simple linear system.
\begin{example}\label{r1}
Consider a network of three times series $\{X,Y,Z\}$ with the following linear dynamic,
\begin{small}
\begin{equation}\label{model1}
\begin{pmatrix}
X_t \\
Y_t\\
Z_t
\end{pmatrix}=\begin{pmatrix}
0.5 & 0& 0\\
0.4 & 0.5 & 0\\
0 & -0.2 & 0
\end{pmatrix}\begin{pmatrix}
X_{t-1} \\
Y_{t-1}\\
Z_{t-1}
\end{pmatrix}+\begin{pmatrix}
N_{X_t} \\
N_{Y_t}\\
N_{Z_t}
\end{pmatrix},
\end{equation}
\end{small}
where $N_X$, $N_Y$, and $N_Z$ are three independent stationary Gaussian processes with zero mean and a diagonal covariance matrix (1, 0.9, 1). 
Since the dynamic is linear and the exogenous noises are Gaussian, we can compute the DIs using the following expression\footnote{Equation \eqref{simueq} does not hold in general setting.}\citet{acc2014}.
\begin{equation}\label{simueq}
I(Z\rightarrow Y||X)=\frac{1}{2T}\sum_{t=1}^{T}\log\frac{|\Sigma_{Y_{t-1},Y_{t},X_{t-1}}||\Sigma_{Z_{t-1},Y_{t-1},X_{t-1}}|}{|\Sigma_{Y_{t-1},X_{t-1}}||\Sigma_{Z_{t-1},Y_{t-1},Y_{t},X_{t-1}}|},
\end{equation}
where $|\Sigma_{Y_{t-1},Y_{t},X_{t-1}}|$ denotes the determinant of the covariance matrix of $\{Y_{t-1},Y_{t},X_{t-1}\}$. Using \eqref{simueq}, we computed the DIs of this system,
\begin{figure}
\hspace{4cm}
\xygraph{ !{<0cm,0cm>;<1.9cm,0cm>:<0cm,1.7cm>::} 
!{(-.7,1) }*+{X}="x"
!{(.4,1.6) }*+{Y}="y"
!{(1.5,1)}*+{Z}="z"
%"x":@(ld,lu)"x"
%"y":@(lu,ru)"y"
"x":"y"
"y":"z"
}
   \caption{DIG of the system in \eqref{model1}.}\label{fig:example_dig}
\end{figure}
\begin{small}
\begin{align*}
&I(Y\rightarrow X||Z)=0,\ \ I(Z\rightarrow X||Y)=0,\ \ I(Z\rightarrow Y||X)=0,\ \ I(Z\rightarrow Z||X,Y)=0,\\
&I(X\rightarrow Z||Y)=0,\ \ \ \ I(X\rightarrow Y||Z)\approx 0.1, \ \ \ I(Y\rightarrow Z||X)\approx 0.03.
\end{align*}
\end{small}
Figure \ref{fig:example_dig} illustrates the DIG of this system. In this example, $\mathcal{PA}_Z=\{Y\}$ and $\mathcal{CH}_Z=\{\}$.
\end{example}

Inference methods based on pairwise comparison has been developed and applied in the literature to identify the causal structure of time series.  The methods in \citet{billio2012econometric}, \citet{billio2010measuring}, and \citet{allen2010financial} are three such examples.
However,  pairwise comparison is not a correct approach in general and may fail to capture the true underlying network.
% For more details see the work by \citet{directed}.
For instance,  considering the pairwise comparison in Example \ref{r1} between $X$ and $Z$ leads to a conclusion that $X$ directly influences $Z$, which would be inaccurate.
More precisely, without conditioning on $Y$, we obtain
 \begin{align*}
 I(X\rightarrow Z)=\frac{1}{T}\sum_{t=1}^T\mathbb{E}\left[\log\frac{p(Z_t|Z^{t-1},X^{t-1})}{p(Z_t|Z^{t-1})}\right]\approx 0.002>0.
 \end{align*}
Notice that the DI in \eqref{cdif} is not a measure based on pairwise comparison. On the contrary, it measures the influence by conditioning on the remaining time series within the network. 
%his will incorporate the information that the rest of network.

\begin{remark}

A causal model allows a factorization of the joint density function in some specific ways.
It was shown in \citet{directed} that under a mild assumption, the joint density function of a causal discrete-time dynamical system with DIG $G(\mathcal{R},\mathcal{E})$ can be factorized as follows,
\begin{equation}\label{abo}
p(R_1,...,R_m)=\prod_{i=1}^{m}\prod_{t=1}^T p(R_{i,t}|\mathcal{R}_{\mathcal{PA}_i\cup\{i\}}^{t-1}).
\end{equation}
%where $\mathcal{B}_i$ denotes the parent set of node $R_i$ in graph $G$ and $\mathcal{R}_{B_i\cup\{i\}}^{t-1}$ denotes the time series with indices $B_i\cup\{i\}$ up to time $t-1$.  
Such factorization is called a generative model.

\end{remark}

\subsection{Inferring  DIGs}
Inferring the DIG of a dynamical system requires estimating the DIs between all ordered pairs of time series within that system. More precisely, inferring the DIG of a network of $m$ time series requires computing $m(m-1)$ number of DIs. 
On the other hand, estimating DI  requires estimating all the expectation terms in (\ref{cdif}).
In information theory this expectation is known as \textit{conditional mutual information}\footnote{For more details see \citet{cover2012elements}.}, i.e., 
\begin{align}
    I(R_{j,t};R_{i}^{t-1}|\mathcal{R}^{t-1}_{-\{i,j\}},R^{t-1}_{j}):=\mathbb{E}\left[\log\frac{p(R_{j,t}|\mathcal{R}^{t-1}_{-\{i,j\}},R_{j}^{t-1},R_{i}^{t-1})}{p(R_{j,t}|\mathcal{R}^{t-1}_{-\{i,j\}},R_{j}^{t-1})}\right].
\end{align}
Using this notation, \eqref{cdif} can be written as follows 
\begin{align}\label{estdi}
I(R_i\rightarrow R_j||\mathcal{R}_{-\{i,j\}})=\frac{1}{T}\sum_{t=1}^T I(R_{j,t};R_{i}^{t-1}|\mathcal{R}^{t-1}_{-\{i,j\}},R^{t-1}_{j}),
\end{align}
Therefore, parametric and non-parametric estimators for the conditional mutual information can be used to estimate the DIs. 
There are different methods that can be used to estimate the terms in (\ref{estdi}) given i.i.d. samples of the time series such as plug-in empirical estimator and k-nearest neighbor estimator. For an overview of such estimators see the Appendix and the articles in \citet{paninski2003estimation}, \citet{noshad2019scalable}, and \citet{jiao2013universal}.

 In general, estimating the DI in (\ref{estdi}) is a complicated task and has high sample complexity.
 This is due to the fact that it requires estimating high dimensional conditional distributions. 
However, knowing some side information about the underlying dynamic can simplify the learning task of the DIG. 
For instance, in Example \ref{r1}, since the underlying dynamic is linear with Gaussian exogenous noises, the DIs can be computed via the covariance matrices \eqref{simueq}. 
Clearly, the covariance matrix can be estimated with lower complexity compared to conditional mutual information. 
For our experimental results, we used \eqref{simueq} for the linear Gaussian experiment and  the k-nearest method in \citet{sricharan2011k} for the non-linear experiment. 
The main reason for selecting k-nearest method is because it has shown relatively better performance compared to the other non-parametric estimators. 
%More precisely, we estimated each of the above conditional mutual information using the $k$-nearest method of \citet{sricharan2011k}. 
For the sake of completeness, we describe the steps of this method in Appendix.

Side information can also help to infer the DIG of a dynamical system without directly estimating the DIs but instead it provides an alternative approach to identify the DIG. 
For example, if it is given that the underlying dynamic is linear, i.e., $\textbf{X}_t = \textbf{AX}_{t-1} + \textbf{N}_t$, then it has been shown in \citet{acc2014} that the support\footnote{The support of a matrix $\textbf{B}=[b_{i,j}]$ is a binary matrix of the same dimension as $\textbf{B}$ such that its entry $(i,j)$ is one if and only if $b_{i,j}\neq0$.} of the coefficient matrix $\textbf{A}$ is equal to the adjacency matrix of its corresponding DIG.   
This result implies that in linear systems, one can obtain the DIG by estimating the coefficient matrix. The latter problem has lower complexity and it can be done using e.g., linear regression. 
For similar examples in econometric models see \citet{etesami2018econometric}.

\subsection{DIG in High-dimensional Settings}
For large networks with thousands nodes and millions of edges such as social or financial networks, DIGs become too complex to infer and analyze.
The main reason is that without any side information, estimating the DI has high computational and sample complexity. 
Furthermore, the estimating complexity of DI increase with the dimension of the network.
This is due to the fact that the DI in \eqref{cdif} measures the influence from $R_i$ to $R_j$ by conditioning on the information from the remaining network $\mathcal{R}_{-\{i,j\}}$. 
Therefore, the size of the conditioning set grows with the size of the network.
This motivates the prior works to reduce the complexity of estimating  DIs and thus make it more suitable for inferring the DIG of large networks by reducing the size of the conditioning set.

One such approaches is proposed by \citet{quinn2013efficient}, in which they developed an efficient algorithm to identify the best directed tree approximation of a given network.
This means reducing the size of the conditioning set to zero, i.e., no conditioning.  
However, this approach comes with the price of an approximation error and furthermore it fails to identify many interconnections between the processes.

The authors in \citet{quinn2015bounded} presented a more generalized version of the above approximation in which they identify the optimal connected bounded in-degree\footnote{Connected bounded in-degree graphs with bound $k$ are connected directed graphs in which each node has at most $k$ number of parents.} approximations.  
This method reduces the size of the conditioning set in \eqref{cdif} to some constant value (bound of the in-degrees) which is independent of the network size. 
Although, this approach improves upon the approximation error but there is still a trade-off between the sample complexity and the approximation error. In another words, as the in-degree bound increases, the sample complexity increases but the approximation error decreases.

In this work, we propose a new method that reduces the size of the conditioning set in \eqref{cdif} to only one for any given network while introducing less approximation error compared to the prior works. 
In this method, we estimate the directed information from $R_i$ to $R_j$ by conditioning on an auxiliary time series. This auxiliary time series is defined such that it comprises the information that the remaining of the network $\mathcal{R}_{-\{i,j\}}$ has about $R_j$.
Next section explains this idea in more details.

\section{Methodology}\label{sec:method}

In order to present our method, we need the following preliminary result that characterizes an important property of DI in \eqref{cdif}. All the proofs are presented in Appendix A.

\begin{lemma}\label{lemma:small}
Consider a network of $m$ time series $\mathcal{R}=\{R_1,...,R_m\}$ with corresponding DIG $G=(\mathcal{V},\mathcal{E})$. Let $\mathcal{C}$ be a subset of $\mathcal{R}_{-\{i,j\}}$ such that $\mathcal{PA}_j\subseteq \mathcal{C}$. If $R_i\not\in\mathcal{PA}_j$, then we have
\begin{align}\label{eq:eqi}
%I(R_i\rightarrow R_j||\mathcal{R}_{-\{i,j\}})=
I(R_i\rightarrow R_j||\mathcal{C})=0.
\end{align}
\end{lemma}
Note that if $\mathcal{C}=\mathcal{R}_{-\{i,j\}}$ and $R_i$ is not a parent of $R_j$, then by the definition of DIG, Equation  \eqref{eq:eqi} holds.
On the other hand, this result states that to detect whether there is an influence from $R_i$ to $R_j$ in a network of time series, it suffices to find a subset of time series that either contains the parents of $R_j$ or their information.
In the remaining of this section, we first clarify the above statement via a simple linear system and later generalize it to non-linear models using neural networks.

\begin{remark}
It is important to emphasize that the reverse of Lemma \ref{lemma:small} does not hold. In another words, if there exists a subset $\mathcal{C}\subset\mathcal{R}_{-\{i,j\}}$ such that \eqref{eq:eqi} holds, we cannot conclude that $R_i$ has no direct influence on $R_j$.
\end{remark}

%\begin{lemma}
%In a network of $m$ time series $\mathcal{R}=\{R_1,...,R_m\}$ with corresponding DIG $G=(\mathcal{V},\mathcal{E})$, the followings hold
%\begin{itemize}
%\item For any subset $\mathcal{C}\subseteq\mathcal{R}_{-\{i,j\}}$ that contains the parents of $R_j$,  i.e., $\mathcal{PA}_j\subseteq C$, we have
%\begin{align}
%I(R_i\rightarrow R_j||\mathcal{R}_{-\{i,j\}})=I(R_i\rightarrow R_j||\mathcal{C}).
%\end{align}
%\item If $(R_i,R_j)\in \mathcal{E}$, then for any subset %$\mathcal{S}\subseteq\mathcal{R}_{-\{i,j\}}$, we have
%\begin{align}
%I(R_i\rightarrow R_j||\mathcal{S})>0.
%\end{align}

%\item If for some subset $\mathcal{S}\subseteq\mathcal{R}_{-\{i,j\}}$, $I(R_i\rightarrow R_j||\mathcal{S})=0$, then $(R_i,R_j)\not\in \mathcal{E}$.
%\end{itemize}
%\end{lemma}

\subsection{Linear Systems}

Consider a first order vector autoregression model (VAR) with $m$ time series, 
\begin{align}\label{eq:linearsys}
\textbf{X}_t = \textbf{AX}_{t-1} + \textbf{N}_t,
\end{align}
where $\textbf{X}_t, \textbf{N}_t\in\mathbb{R}^m$, $\textbf{A}\in\mathbb{R}^{m\times m}$, and  $\textbf{N}_t$ is a vector of $m$ independent exogenous noises.
As we discussed earlier, the result in \citet{acc2014} implies that the DIG of this VAR model is encoded in the support of its coefficient matrix $\textbf{A}=[a_{i,j}]$, i.e., 
\begin{align}
    I(X_i\rightarrow X_j|| \mathcal{X}_{-\{i,j\}})=0 \Longleftrightarrow a_{j,i}=0.
\end{align}
In another words, the parents of time series $X_j$ are the ones whose corresponding coefficients are non-zero in the $j$-th row of matrix $\textbf{A}$. 
This also can be seen from the $j$-th row of the matrix equation in \eqref{eq:linearsys},
\begin{align}\label{eq:linear_each}
    X_{j,t}=\sum_{k=1}^m a_{j,k}X_{k,t-1}+N_{j,t}.
\end{align}
Another way to interpret the above equation is to say that the information of the network about time series $X_j$ is in the form of a ``portfolio", i.e., a linear combination of the other time series. 
Therefore, it is possible to summarize the network's information about $X_j$ into only one time series, namely a well-designed portfolio. 
Next result shows the form of such portfolio.

\begin{lemma}\label{lemma:lin}
In the linear system of \eqref{eq:linearsys}, $X_i$ has no direct influence on $X_j$ if and only if
\begin{align}\label{eq:lemma_linear}
I(X_{i}\rightarrow X_j || Q) =0,
\end{align}
where $Q$ is a time series which we call the ideal portfolio and it is defined by
%obtained from projecting $X_{j,t}$ on span-$\{X_{1,t-1},...,X_{n,t-1}\}\setminus\{X_{i,t-1}\}$. More precisely, 
$Q_{t-1}:=\textbf{u}_t^T\textbf{X}_{-\{i\},t-1}$, where
\begin{align*}
&\textbf{u}_t:=\arg\min_{\textbf{w}\in\mathbb{R}^{m-1}}\mathbb{E}\left[||X_{j,t}- \textbf{w}^T \textbf{X}_{-\{i\},t-1}||_2^2\right],\\
&\textbf{X}_{-\{i\},t-1} :=[X_{1,t-1},\cdots, X_{i-1,t-1}, X_{i+1,t-1}, \cdots, X_{m,t-1} ]^T.
\end{align*}
\end{lemma}
According to the above Lemma, projecting $X_j$ on $\mathcal{X}_{-\{i\}}$ results in an ideal portfolio $Q$ that contains all the information for deciding whether there is an influence from $X_i$ to $X_j$.
%This result suggests that in order to detect the influence of $X_i$ on $X_j$ in a linear system, it is possible to summarize all the information in the network about $X_j$ into a time series $L$.  
Hence, instead of estimating $I(X_{i}\rightarrow X_j || \mathcal{X}_{-\{i,j\}})$ whose complexity depends on the network size, one can estimate $I(X_{i}\rightarrow X_j || Q)$.
%without introducing any approximation error.
Note that the sample complexity of the latter DI does not grow with the size of the network and thus it is suitable for estimating the DIG of large networks.

%In the remaining of this section, we generalize this result to obtain an auxiliary time series  this result to non-linear systems  .

\subsection{Non-linear Systems with Additive Noise}
Inferring the causal network of non-linear systems is a challenging problem that its complexity increases exponentially with the dimension of the network. 
In this section, we study the causal inference problem in non-linear systems whose dynamic can be captured by
\begin{align}\label{nonlinear}
    X_{j,t} = F_{j}(\mathcal{X}^{t-1}) + \varepsilon_{j,t}, \ \ j=1,...,m,
\end{align}
where $\mathcal{X}^{t-1}=\{X_{1}^{t-1},...,X_{m}^{t-1}\}$, $\{F_j(\cdot)\}$ is a set of non-linear continuous functions, and $\{\varepsilon_{j,t}\}$ is a set of independent exogenous noises.
We call this model non-linear with \textit{additive noise} due to the noise term that is added to the non-linear term\footnote{In contrary to additive noise, there are systems in which the exogenous terms are multiplicative, e.g., $X_{j,t}=X_{i,t-1}\varepsilon_{j,t}+X_{j,t-1}$.}.
This is a general non-linear dynamic that can be used to model the behavior of wide range of physical dynamical systems.
The dynamic is called \textit{Markovian} if $\mathcal{X}^{t-1}$ is replaced by $\mathcal{X}_{t-1}=\{X_{1,t-1},...,X_{m,t-1}\}$.

Below, we generalize the result of Lemma \ref{lemma:lin} to the non-linear system in \eqref{nonlinear} by showing that in such systems, it is possible to reduce the conditioning set in the DI to one time series. 
%Afterward, we propose two approaches for two sub-classes of \eqref{nonlinear} to obtain this times series.
\begin{lemma}\label{lemma:nonlinear}
In \eqref{nonlinear}, $X_i$ has no direct influence on $X_j$ if and only if 
\begin{align}\label{eq:non-q}
I(X_{i}\rightarrow X_j || Q) =0,
\end{align}
where $Q$ is a time series defined by $Q_{t-1}:=F_{j}(\mathcal{X}^{t-1}_{-\{i\}})$. 
\end{lemma}
In the remaining of this section, we propose two methods to obtain the time series $Q$ introduced in the above Lemma.

\paragraph{Koopman-based lifting technique}
Consider a particular sub-class of the non-linear system in \eqref{nonlinear} whose dynamic is defined by
\begin{align}\label{nonlin}
F_{j}(\mathcal{X}^{t-1}) = \sum_{k=1}^K w_{j,k} h_k(\mathcal{X}_{t-1}), \ \ j=1,...,m,
\end{align}
where $\{w_{j,k}\in\mathbb{R}\}$ are the weights and $\{h_k(\cdot)\}$ denotes a set of library functions that are assumed to be known.
This model is Markovian and the library functions can be seen as a set of basis that are used to approximated $F_j(\cdot)$.
Examples of such library functions are monomials and Gaussian radial basis functions. 
%This is a generalization of VAR model, namely, if $\{h_k(X_{1,t-1},...,X_{m,t-1})=\textbf{V}_k^T \textbf{X}_{t-1}\}$ for some vectors $\{\textbf{V}_k\}$, then \eqref{nonlin} becomes a VAR model. 
%Analogous to Lemma \ref{lemma:lin}, we have the following result.

%\begin{lemma}\label{lemma:nonlin}
%In \eqref{nonlinear}, if $X_i$ has no direct influence on $X_j$, then 
%\begin{align}
%I(X_{i}\rightarrow X_j || \mathcal{X}_{-\{i,j\}})=I(X_{i}\rightarrow X_j || Q) =0 ,
%\end{align}
%where $Q=\{Q_t\}$ is defined by $Q_{t-1}:=\sum_{k=1}^K w_{j,k} h_k(X_{1,t-1},...,X_{n,t-1})$.
%\end{lemma}

%This implies that if one can estimate $Q$, then inferring the DIG of \eqref{nonlin} can be done with the lower complexity.

In this setting, the results of Lemma \ref{lemma:nonlinear} implies that the following time series can be substituted in the conditioning of the DI. 
\begin{align}
    Q_t = \sum_{k=1}^K w_{j,k} h_k(\mathcal{X}_{-\{i\},t-1}).
\end{align}
However, in this formulation, the weights $\{w_{,k}\}$ are unknown. 
An approach to obtain the weights is a non-linear filtering technique known as Koopman-based lifting \citet{koopman1931hamiltonian}. 
This technique takes observational data and a set of library functions as inputs and obtains the unknown coefficients $\{w_{,k}\}$.
The main steps of this technique are; transforming the data (lifting the data),  applying a linear identification on the lifted data, and finally applying another transformation to bring down the results into the original vector field.  Figure \ref{fig:koopman} illustrates the main steps. For more details see Appendix and \citet{mauroy2019koopman}. 
\begin{figure}
\centering
\includegraphics[scale=.4]{koopman.pdf}
\caption{Koopman lifting technique compared to classical non-linear identification.}\label{fig:koopman}
\end{figure}

Although, the Koopman-based lifting technique is theoretically sound but it has some shortcomings facing real-world applications.
First, the Koopman's performance depends on the choice of the library functions and second, it often fails to estimate the real time series $Q$.
More precisely, this technique involves the computation of matrix $\textbf{L}:=log(\textbf{P}_x^{\dagger}\textbf{P}_y)/T_s$, where $\textbf{P}_x$ and $\textbf{P}_y$ are estimated from the observational data\footnote{See Appendix for more details.}.
Matrix $\textbf{P}_x^{\dagger}$ denotes the pseudo-inverse, and the function $\log(\cdot)$ denotes the (principal) matrix logarithm.
On the other hand, Koopman lifting technique is applicable for estimating the time series $Q$ only when the resulting matrix $\textbf{L}$ is real\footnote{See \citet{culver1966existence} for conditions under which a real matrix has a real logarithm.}.
However, this is not always the case in real-world applications due to observational noises and lack of sufficient data. 
%In fact, as it is stated in \citet{culver1966existence}, a real matrix has a real logarithm if and only if it is invertible and each Jordan block belonging to a negative eigenvalue occurs an even number of times. 
To overcome such shortcomings, we propose an alternative approach to estimate $Q$ using recurrent neural networks (RNNs).

\paragraph{RNNs method}
 Recurrent neural networks are a specific class of Neural Networks well suited to learn time series. 
 They are distinguished by their memory as they are able to remember information from prior inputs to influence their current outputs.
 The universal approximation theorem states that a neural network with enough hidden layers can approximate any non-linear continuous function such as $F_j(\cdot)$ in \eqref{nonlinear} (see \citet{HORNIK1989359}).

 \begin{algorithm}[t]
 \textbf{Input:} Observational data of $m$ time series up to time $T$, $\mathcal{X}^T$, Threshold $\alpha>0$\;
 \textbf{Output:} Adjacency matrix of $\textbf{DIG}=[d_{i,j}]$\;
 \For{ $i,j=1,...,m$}{
  Train an RNN $R_j(\cdot;\Theta^*)$ that maps $\mathcal{X}_{-\{i\}}^{t-1}$ to $X_{j,t}$\; 
  Define $Q_{t-1}=R_{j}(\mathcal{X}_{-\{i\}}^{t-1};\Theta^*)$\;
  \eIf{$I(X_{i}\rightarrow X_j || Q) > \alpha$}{
   $d_{j,i}=1$
   }{
   $d_{j,i}=0$
  }
  }
 \caption{Infer-DIG}\label{alg}
\end{algorithm}
 Given the aforementioned result, we train an RNN using the observational data to estimate the time series $Q$ defined in Lemma \ref{lemma:nonlinear}.
 More precisely, our RNN maps $\mathcal{X}_{-\{i\}}^{t-1}$ as the inputs to $X_{j,t}$ as the output. 
 Let $R_j(\mathcal{X}_{-\{i\}}^{t-1};\Theta^*)$ denotes the trained RNN with parameters $\Theta^*$. 
 In this case, the time series $Q$ can be written as  $Q_{t-1}=R_{j}(\mathcal{X}_{-\{i\}}^{t-1};\Theta^*)$.
 Finally, we use \eqref{eq:non-q} to detect whether $X_i$ has influence on $X_j$ or not. 
 Algorithm Infer-DIG in \ref{alg} summarizes the steps of our RNN method.

 %\begin{align}
  %   \min_{\theta}||X_{j,t} - F_{j}(\mathcal{X}_{-}^{t-1};\theta)||^2.
 %\end{align}

\section{Experimental Results}\label{sec:exp}

 Since the true empirical DIG of firms is unknown, to evaluate the performance of our approach, we use different simulated environment.
 In this section, we first describe the simulation methodology in a linear Gaussian framework. 
 We then show that our results generalize well to nonlinear setting by conducting an experiment on a nonlinear system. 
 Finally, we apply our approach to a set of empirical data describing the daily stock prices of US firms and obtain their corresponding causal network.

 \subsection{Linear Gaussian Framework} \label{linfr}
 In this experiment, we consider a linear system, a VAR(1) model whose dynamic is given by 
\begin{equation}\label{eq:ex_lin}
    \textbf{X}_t = \textbf{A} \textbf{X}_{t-1} + \textbf{N}_t
\end{equation}
with $m$ being the number of asset returns, $\textbf{X}_t = (X_{1,t}, X_{2,t}, ... , X_{m,t})^\top$ being the vector of returns at time $t$, $\textbf{A}=[a_{i,j}]$ being a $m\times m$ matrix and $\textbf{N}_t$ being a $\mathcal{N}(0,\textbf{I})$ vector of noises. 
As we discussed earlier, in such linear systems, $a_{i,j}$ captures the influence of asset $j$ on asset $i$, i.e., there is an influence from $j$ to $i$ if and only if $a_{i,j}\neq0$.

%A parent is defined as an asset that has an influence on another asset. In this setup, the set of the parents of an asset i is defined as 
%\begin{equation}
%    \textbf{P}_i = \{z : |[\textbf{A}]_{i,z}| > 0 \}
%\end{equation}

%To simulate our VAR, we need that the matrix $\textbf{A}$ is composed of assets that influence others (i.e $a_{i,j} \neq 0$) and some assets that don't ($a_{i,j} = 0$).  $\textbf{A}$ is constructed to allow all cases of potential influences between assets: i $\leftrightharpoons$ j, i $\rightarrow$ j, i $\leftarrow$ j, i independent of j. Let $1_{a>b}$ be the indicator function which is equal to 1 if $a > b$ and 0 otherwise.

To reflect an important property of the market that some firms are more connected than others in our experiment, we divided the $m$ time series into two parts. 
First part ($1\leq i\leq s$) indicates assets with high degrees of connectedness and the second part ($1+s\leq i\leq m$) are the ones with low degrees of connectedness.
Parameter $1<s<m$ denotes the numbers of assets with high degrees of connectedness.
Afterward, for every entry $(i,j)$ of $\textbf{A}$, we independently generated a random number $x\sim U(-0.9, 0.9)$ and decided on value $a_{i,j}$ as follows,
\begin{align}\label{eq:rules}
    a_{i,j}=\left\{
\begin{array}{ll}
      x 1_{|x| > \underline{\epsilon} },  & \text{if}\ \ 1\leq i\leq s, 1\leq j\leq s, \\
      x 1_{|x| > \overline{\epsilon} },  & \text{if}\ \ 1+s\leq i\leq m, 1\leq j\leq m, \\
      0,  & \text{if}\ \ 1\leq i\leq s, 1+s\leq j\leq m, \\
\end{array} 
\right.
\end{align}
where $1_{a>b}$ denotes the indicator function which is equal to 1 when $a>b$ and $0$ otherwise and $\underline{\epsilon}$ and $\bar{\epsilon}$ are thresholds to define non-zero entries in the upper-left and the lower part of $\textbf{A}$, respectively.
Figure \ref{fig:matrix} illustrates the structure of the resulting $\textbf{A}$.
We select these thresholds such that $\underline{\epsilon}<\bar{\epsilon}$. 
This ensures that the upper-left of $\textbf{A}$ is denser than its lower part or equivalently, assets with indices $\{1,...,s\}$ are more connected than the ones with indices $\{1+s,...,m\}$.
In our experiment, we select $(s,m)=(85,100)$ and $(\underline{\epsilon},\bar{\epsilon})=(0.4,0.7)$. 
%further results are robust to changes of parameters.
\begin{figure}
    \centering
    \includegraphics[scale=.2]{fig_matrix.pdf}
    \caption{Structure of matrix $\textbf{A}$ in \eqref{eq:ex_lin} that is build using \eqref{eq:rules}.}
    \label{fig:matrix}
\end{figure}
Finally, to guarantee the stability of the time series, we rescale\footnote{Formally, we use $\textbf{A}/(\rho(\textbf{A}) + \epsilon)$, where $0<\epsilon$.} $\textbf{A}$ such that its spectral radius is strictly less than one, i.e.,  $\rho(\textbf{A})< 1$. 
Once the matrix $\textbf{A}$ is defined, we simulate the time series using \eqref{eq:ex_lin} for a period of $T = 30000$ and use the resulting data for our estimations.

To study the effect of the conditioning set on detecting the influences, in our experiments, we consider four different conditioning sets.
More precisely, to measure whether asset $i$ influences asset $j$, we estimate $I(X_i\rightarrow X_j||\mathcal{C}_j)$ for the following choices of the conditioning set; 
\begin{enumerate}
    \item \textit{True parents:} In this approach, we select $\mathcal{C}_j$ to be the true parents of $X_j$ excluding $X_i$, i.e., $\mathcal{C}_j=\mathcal{PA}_{j}\setminus\{X_i\}$.
    Note that this approach is not practical\footnote{This is because in structural learning problems, we do not know the true parents of each asset. In another words, if we had access to the true parents of each asset, we would have the DIG of the system and there is no need to compute the DIs.} and we use it only as the benchmark to better understand the performances of the other approaches.

    \item \textit{Most correlated:} In this case, we define $\mathcal{C}_j$ to be the set of $k$ most correlated assets with $X_j$ (except $X_i$).

    \item \textit{Ideal portfolio:} In this scenario, $\mathcal{C}_j$ contains the portfolio $Q$, where $Q$ is defined in Lemma \ref{lemma:lin}.
    For further discussion see Appendix. 
    %comprises of assets that influence $j$ except asset $i$, i.e., $\mathcal{C}_j=\{U\}$, where $Q=(Q_t)$ and $Q_t=\sum_{k\neq i}a_{i,k}X_{k,t}$.
    %Note that $Q$ coincides with the time series introduced in Lemma \ref{lemma:lin}.

    \item \textit{RNN:} This method applies Algorithm \ref{alg} to estimate the time series $Q$ and defines $\mathcal{C}_j=\{Q\}$.

    %\item \textit{Maximally correlated:} In this case, $\mathcal{C}_j$ is the projection (ignoring i) maximally correlated with j (see Appendix C) \TODO{Do we really need this, because we have the ideal projection.}
\end{enumerate}
Note that we also applied the Koopman-based lifting techniques but due to its mentioned shortcomings, it was unable to robustly identify the interconnections. 
Hence, we could not compare its performance with the other methods. 
In this experiment, since the dynamic is linear and the noises are Gaussian, we use Equation \eqref{simueq} to estimate the DIs. 
Finally, we obtain the adjacency matrix of the corresponding DIGs by comparing the estimated DIs with a threshold $\alpha>0$, i.e., 
\begin{align}\label{eq:m}
    [\textbf{DIG}]_{j,i}=\left\{
\begin{array}{ll}
      1  & \text{if}\ \ \widehat{I}(X_i\rightarrow X_j||\mathcal{C}_j)>\alpha, \\
      0,  & \text{otherwise}, \\
\end{array} 
\right.
\end{align}
where $\widehat{I}(\cdot\rightarrow\cdot||\cdot)$ denotes the estimated DI from the data. 
In order to compare the performances of the aforementioned four approaches, we use the precision and recall measure between the true DIG (obtained from $\textbf{A}$) and their estimated DIGs. 
Formally, the precision and the recall are defined by
\begin{align*}
    Precision:= \dfrac{TP}{TP+FP}, \ \ Recall:= \dfrac{TP}{TP+FN},
\end{align*}
where 
\begin{align*}
    & TP := \sum_{i,j=1}^m 1_{a_{j,i} \neq 0}1_{ [\textbf{DIG}]_{j,i} \neq 0}, \ \ FP := \sum_{i,j=1}^m 1_{a_{j,i} = 0}1_{ [\textbf{DIG}]_{j,i} \neq 0},\\
    & FN := \sum_{i,j=1}^m 1_{a_{j,i} \neq 0}1_{ [\textbf{DIG}]_{j,i} = 0}.
    %\item $TN := \sum_{i,j=1}^m 1_{a_{j,i} = 0}1_{[\textbf{DIG}]_{j,i} \leq \alpha}$.
\end{align*}

\begin{figure*}
        \centering
        \begin{subfigure}[b]{0.48\textwidth}
            \centering
            \includegraphics[width=\textwidth]{Linear_Performance_True parents.jpg}
        \end{subfigure}
        \quad
        \begin{subfigure}[b]{0.48\textwidth}  
            \centering 
            \includegraphics[width=\textwidth]{Linear_Performance_Most correlated assets.jpg}
        \end{subfigure}
        \vskip\baselineskip
        \begin{subfigure}[b]{0.48\textwidth}   
            \centering 
            \includegraphics[width=\textwidth]{Linear_Performance_Ideal Portfolio.jpg}
        \end{subfigure}
                  \quad
        \begin{subfigure}[b]{0.48\textwidth}   
            \centering 
            \includegraphics[width=\textwidth]{Linear_Performance_RNN.jpg}
        \end{subfigure}
        \caption{Precision and recall curves in the linear framework for the \textit{True parents}, \textit{Most correlated}, \textit{Ideal portfolio}, and the \textit{RNN}, respectively.} 
        \label{perflin}
    \end{figure*}

Figure \ref{perflin} shows the performances of the four aforementioned approaches in the linear framework. 
%In a linear and Gaussian environment, the Gaussian estimator allows to compute the exact value of the DI. 
It is not surprising that the \textit{true parents} approach achieves 100\% accuracy, as it is anticipated by Lemma \ref{lemma:small}.
The \textit{ideal portfolio}'s performance is guaranteed by Lemma \ref{lemma:lin} and it is verified by our experiment.
However, it is important to emphasize that the \textit{ideal portfolio} shows ideal performance because the underlying model is linear. 
As we will see in the next section, its performance declines when the underlying model deviates from being linear. 
For the \textit{most correlated} approach, we used $k=10$ but as it is shown in Figure \ref{perflin}, it has the worst performance among the four conditioning methods. 
This is due to the fact that the set of the ten most correlated assets with a given asset $j$ does not necessarily contains the true parents of asset $j$.
On the other hand, we observe high accuracy from the \textit{RNN} approach which is a striking result.
This result is an evidence that an RNN is capable of estimating the ideal portfolio, i.e., the time series $Q$ in Lemma \ref{lemma:lin} without any side information about the underlying model.

%On the other hand, the \textit{RNN} approach keeps its high performance even in non-linear system.

\subsection{Non-Linear framework} \label{nonlinfr}
To compare the performances of the different approaches from the previous section in non-linear environment, we simulate a set of quadratic processes whose dynamic is given below, 
\begin{equation}\label{eq:ex_non}
        X_{i,t} =b_i  {\textbf{X}^T_{t-1}} \textbf{A}_i \textbf{X}_{t-1} + N_{i,t}, \ i=1,...,m,
\end{equation}
where $\textbf{A}_i\in\mathbb{R}^{m \times m}$, $\textbf{X}_t = (X_{1,t}, X_{2,t},..., X_{m,t})^T$, $N_{i,t}\sim\mathcal{N}(0,\sigma^2)$, and $b_i\sim U(-0.9 , 0.9)$.
%Matrix  is associated to asset $i$. 
Note that the term $|[\textbf{A}_i]_{j,k} + [\textbf{A}_i]_{k,j}|$ captures the effect of $X_{j,t-1}X_{k,t-1}$ on $X_{i,t}$. Thus, it is possible to obtain the true parents of asset $i$ as follows,
\begin{equation}
 \mathcal{PA}_i = \{ X_j :  [\textbf{1}^T \cdot \big(|\textbf{A}^T_i + \textbf{A}_i|\big)]_{j}  > 0\},
\end{equation}
where $\textbf{1}$ denotes all-one vector of length $m$.
Each matrix $\textbf{A}_i$ is simulated independently by following the similar procedure as in Section \ref{linfr}.
In this experiment, since the model is non-linear, we could not apply \eqref{simueq} to estimate the DIs but instead we used the k-nearest method to estimate the mutual information and applied Equation \eqref{estdi}.
%sampling their entries randomly from a uniform distribution (similar to section \ref{linfr}).

Herein, we again compare the performances of the four different conditioning approaches. 
Figure \ref{fig:perfnonlin} shows the precision-recall curves for these approaches in the quadratic model with $m=15$. 
Precision-recall curves are a standard tools to illustrate and compare the performances of different learning methods. 
In this curve the precision is demonstrated in the y-axis vs. the recall on the x-axis for all potential values of the threshold $\alpha$.

Similar to the linear setting, we use the \textit{true parents} as a benchmark since it has the ideal performance. 
%Conditioning on the true parents obviously performs the best as all the information is included in the Directed Information. 
%However, it is also the heaviest in terms of computational time. 
It is however important to emphasize that this conditioning approach has higher complexity compared to the others. 
This is because in the \textit{true parent} approach, the size of the conditioning set is relatively larger than the other approaches.

For the \textit{most correlated} approach, we use $k=5$, i.e., the size of the conditioning set is five.
With this method, we could slightly reduce the estimation complexity of the DIs compared to the  \textit{true parent} approach but this comes with the price of losing the performance.
clearly, the performance of the \textit{most correlated} approach can be improved by increasing $k$ but this will increase the complexity.

The performance of the \textit{ideal portfolio} approach (using the time series in Lemma \ref{lemma:lin} as the conditioning) is worse than all others which is not surprising as the model is no longer linear. 
%\TODO{that is not the ideal portfolio, the ideal portfolio is not linear in the quadratic framework. that is the estimated linear portfolio using (31)}
This means that the information embedded in the linear portfolio is not sufficient to decide the non-linear influences among the time series.

Finally, as it is shown in Figure \ref{fig:perfnonlin}, the \textit{RNN} approach outperforms the \textit{most correlated} and the \textit{ideal portfolio} approaches and it shows close performance to the \textit{true parents} but with the size of the conditioning set equal to one. 
This result once more fortifies our claim that with an RNN we can summarize the information of the network into one time series and use it for detecting the causal relationships. 
This claim is due to Lemma \ref{lemma:nonlinear} and the universal approximation theorem which states that a neural network with enough hidden layers can approximate any non-linear function  \citet{HORNIK1989359}.
The slight difference between the performance of the \textit{RNN} and the \textit{true Parents} is because of the estimation error in the recurrent neural network.

%The latter result suggests that a one-dimension time series conditioning coming from the Neural Network is able to keep as much information as the N-dimensions conditioning coming from the True Parents. This is consistent with the Universal Approximation Theorem  (see ).  Complexity and computational time are also reduced; this reduction is growing as more and more assets are included in the DIG.

\begin{figure}
\centering
\includegraphics[width=\textwidth]{precrecall_15assets.jpg}
\caption{Precision-Recall curves for the quadratic model.}\label{fig:perfnonlin}
\end{figure}

\subsection{Empirical DIG}

This section describes how to apply our approach to empirical data and obtain the DIG of some US firms. 
We extracted the daily stock prices and the daily US Treasury rate as risk free returns from the CRSP database from 1990 to 2020. 
As the market is likely to evolve through these years, 
we chose to divide the dataset into six subsets, each of which has a length of five years and estimate the corresponding DIG of each subset separately. 
%chose to compute a different DIG for every five years. 
Herein, we assume that the causal structure of the market evolved but its rate was slow enough such that during a period of five years, the DIG of the market remained unchanged.

For every subset, we keep the data of the 1000 firms with the highest maximum market capitalization and compute their excess return time series $X_{i,t}$, using the following relationship,
\begin{equation}
X_{i,t} = ln(P_{i,t}) - ln(P_{i,t-1}) - r_t,  
\end{equation}
where $P_{i,t}$ denotes the stock price of the firm $i$ at time $t$ and $r_{t}$ is the risk free rate at time $t$. 
Afterwards, we apply Algorithm \ref{alg} with the excess returns as the input to estimate the corresponding DIG of each subset. 
We use the k-nearest neighbor method to estimate the DIs.
We define the threshold $\alpha$ to be the unconditional mean across the estimated DIs. Note that in this experiment, the true DIGs of the market are not known, hence, we could not compute the precision-recall curves.

For the sake of presentation, instead of the complete DIGs with 1000 nodes, we draw the sub-graphs consisting of the 30 largest firms in Figures\footnote{For a better presentation, interactive plots are available at https://marcaureledivernois.github.io/firm-network/} \ref{empdig1}, \ref{empdig2}, and \ref{empdig3} . 
Each graph consists of 60 nodes illustrating the cause firm on the top hemisphere and the effect firm on the bottom hemisphere.
%Each hemisphere consists of 30 firms such that the north hemisphere denotes the cause and the south  shows all outdegrees of the 30 firms and the south hemisphere shows the indegrees of the 30 firms. 
For instance, if there is an edge between ``from: AAPL'' on the top and ``to: GOOGL'' on the bottom, it means that Apple influences Google. The dynamic evolution of the DIGs through time can often be explained by real events that happened in the market. For instance, in the DIG 2010-2014, Apple was not influencing General Electric (GE). However, on the 17th October 2017, Apple announced a partnership with GE to bring Predix, GE's data and analytics platform, to their iPhones and iPads. We are able to capture this partnership in the DIG 2015-2019 as an edge is now present from Apple to GE. Another example is the announced collaboration between AT\&T and Cisco to manage IoT devices and launch 5G service at the end of the 2010s: 
there was neither an edge from AT\&T to Cisco nor from  Cisco to AT\&T during the first half of the 2010s, but the DIG for the second half of the 2010s shows a mutual influence, reflecting an increased relationship between the two companies.

Table \ref{GrangerCausality} shows the Degree of Granger Causality (DGC) defined as the fraction of relationships in the network among all potential relationships. Formally,
\begin{equation}
    DGC = \dfrac{1}{N^2} \sum_i \sum_j [\textbf{DIG}]_{j,i},
\end{equation}
These results show that the DGC increased both in the DotCom bubble and in the Subprime Crisis, suggesting an increase of the connectedness in turmoil periods. 
This finding is consistent with \citet{correlcrisis} stating that correlation increases in bear markets.

Tables \ref{outdegree} and \ref{indegree} show the outdegree and indegree of every firm in the six subsets. Outdegree is defined as the number of edges going out of a specific node. Indegree is the number of edges going to a specific node.
These tables also reveal interesting facts. For instance, the SPY ticker, an ETF launched in 1993 and aiming at tracking the S\&P500 return, enters in the 30 biggest market capitalizations in 2010 and has the highest number of outdegrees in the periods 2010-2014 and 2015-2019 but relatively low number of indegrees. 
This result suggests that the market return is influencing a high number of firms, but the converse is not necessarily true. One way to intuitively interpret this finding is that bullish and/or bearish markets are likely to influence next period stock returns (as a persistence effect known in business cycles), but individual stock returns struggle at predicting next period market returns. 
%In each of the 5-years subsample, we use the excess returns of the 1000 biggest firms as conditional information to compute the DIs of the 30 biggest firms.
%\TODO{I dont understand the sample and subsample.}
%We used the k-nearest neighbor estimator to estimate the DIs and used algorithm \ref{alg} to  estimate the DIGs.  The threshold $\alpha$ is defined as the unconditional mean across the DIs.

\begin{table}[]
\begin{tabular}{cccccc}
\hline
\multicolumn{1}{|c|}{1990-1994} &
  \multicolumn{1}{c|}{1995-1999} &
  \multicolumn{1}{c|}{2000-2004} &
  \multicolumn{1}{c|}{2005-2009} &
  \multicolumn{1}{c|}{2010-2014} &
  \multicolumn{1}{c|}{2015-2019} \\ \hline
\multicolumn{1}{|c|}{0.23} &
  \multicolumn{1}{c|}{0.27} &
  \multicolumn{1}{c|}{0.18} &
  \multicolumn{1}{c|}{0.28} &
  \multicolumn{1}{c|}{0.32} &
  \multicolumn{1}{c|}{0.25} \\ \hline

\end{tabular}
\caption{Degree of Granger Causality (DGC) for each sub-graph. DGC is defined as the fraction of relationships in the network among all potential relationships.}
\label{GrangerCausality}
\end{table}

\begin{figure*}
        \centering
        \begin{subfigure}[b]{0.6\textwidth}
            \centering
            \includegraphics[width=\textwidth]{net1994.JPG}
        \end{subfigure}
        \vskip\baselineskip
        \begin{subfigure}[b]{0.6\textwidth}   
            \centering 
            \includegraphics[width=\textwidth]{net1999.JPG}
        \end{subfigure}
        \caption{Empirical DIG for the periods 1990-1994 (top) and 1995-1999 (bottom). Interactive graphs can be found at https://marcaureledivernois.github.io/firm-network/} 
        \label{empdig1}
    \end{figure*}

\begin{figure*}[]
        \centering
        \begin{subfigure}[b]{0.6\textwidth}
            \centering
            \includegraphics[width=\textwidth]{net2004.JPG}
        \end{subfigure}
        \vskip\baselineskip
        \begin{subfigure}[b]{0.6\textwidth}   
            \centering 
            \includegraphics[width=\textwidth]{net2009.JPG}
        \end{subfigure}
        \caption{Empirical DIG for the periods 2000-2004 (top) and 2005-2009 (bottom). Interactive graphs can be found at https://marcaureledivernois.github.io/firm-network/} 
        \label{empdig2}
    \end{figure*}

\begin{figure*}
        \centering
        \begin{subfigure}[b]{0.6\textwidth}
            \centering
            \includegraphics[width=\textwidth]{net2014.JPG}
        \end{subfigure}
        \vskip\baselineskip
        \begin{subfigure}[b]{0.6\textwidth}   
            \centering 
            \includegraphics[width=\textwidth]{net2019.JPG}
        \end{subfigure}
        \caption{Empirical DIG for the periods 2010-2014 (top) and 2015-2019 (bottom). Interactive graphs can be found at https://marcaureledivernois.github.io/firm-network/} 
        \label{empdig3}
    \end{figure*}

\begin{table}[]
\resizebox{\textwidth}{!}{
\begin{tabular}{|ll|ll|ll|ll|ll|ll|}
\hline
\multicolumn{2}{|c|}{1990-1994} &
  \multicolumn{2}{c|}{1995-1999} &
  \multicolumn{2}{c|}{2000-2004} &
  \multicolumn{2}{c|}{2005-2009} &
  \multicolumn{2}{c|}{2010-2014} &
  \multicolumn{2}{c|}{2015-2019} \\ \hline
Ticker &
  Out &
  Ticker &
  Out &
  Ticker &
  Out &
  Ticker &
  Out &
  Ticker &
  Out &
  Ticker &
  Out \\ \hline
SBC  & 13 & JNJ  & 12 & XOM  & 12 & PEP  & 15 & SPY   & 17 & SPY   & 14 \\
CHV  & 11 & HWP  & 12 & JNJ  & 10 & WFC  & 13 & T     & 15 & DIS   & 12 \\
XON  & 10 & INTC & 12 & MSFT & 9  & CSCO & 12 & JNJ   & 15 & BRK   & 12 \\
PG   & 10 & BAC  & 12 & DELL & 9  & WMT  & 11 & CVX   & 13 & PFE   & 11 \\
GE   & 10 & PFE  & 11 & WMT  & 9  & IBM  & 11 & AAPL  & 12 & UNH   & 11 \\
AN   & 9  & IBM  & 11 & C    & 9  & PG   & 11 & INTC  & 12 & JPM   & 11 \\
BEL  & 9  & MSFT & 11 & HD   & 8  & CVX  & 11 & CSCO  & 12 & AAPL  & 10 \\
MRK  & 8  & SBC  & 11 & JPM  & 8  & HPQ  & 10 & GE    & 12 & GOOGL & 10 \\
IBM  & 8  & GE   & 10 & PFE  & 8  & VZ   & 10 & IBM   & 11 & WMT   & 9  \\
BMY  & 8  & MRK  & 10 & PG   & 7  & GE   & 9  & GOOGL & 11 & CSCO  & 9  \\
ABT  & 7  & LMG  & 9  & INTC & 7  & UBS  & 9  & XOM   & 11 & VZ    & 9  \\
AHP  & 7  & MO   & 9  & BAC  & 7  & PFE  & 9  & JPM   & 10 & BA    & 9  \\
T    & 7  & NT   & 9  & BMY  & 6  & C    & 9  & KO    & 10 & CVX   & 8  \\
BLS  & 7  & XOM  & 9  & SBC  & 6  & SLB  & 9  & MRK   & 10 & WFC   & 8  \\
AIG  & 7  & KO   & 9  & IBM  & 5  & MSFT & 9  & BRK   & 10 & V     & 8  \\
F    & 7  & HD   & 8  & TWX  & 5  & ORCL & 9  & SLB   & 10 & PG    & 8  \\
HWP  & 7  & BEL  & 8  & AIG  & 5  & MRK  & 8  & WMT   & 10 & T     & 7  \\
MOT  & 7  & BMY  & 8  & CSCO & 4  & KO   & 8  & VZ    & 9  & JNJ   & 7  \\
DD   & 6  & SUNW & 8  & GE   & 4  & GOOG & 7  & DIS   & 9  & MRK   & 7  \\
DIS  & 6  & AIG  & 7  & VZ   & 4  & XOM  & 7  & PFE   & 9  & KO    & 6  \\
MOB  & 6  & WCOM & 7  & MRK  & 4  & JNJ  & 7  & BAC   & 8  & XOM   & 5  \\
PEP  & 6  & CSCO & 6  & MOT  & 4  & MO   & 7  & PM    & 8  & MSFT  & 5  \\
MSFT & 6  & QCOM & 6  & TXN  & 4  & JPM  & 6  & ORCL  & 7  & MA    & 5  \\
INTC & 6  & C    & 6  & HPQ  & 3  & COP  & 6  & PG    & 7  & BABA  & 5  \\
WMT  & 6  & PG   & 6  & ORCL & 3  & T    & 6  & PEP   & 6  & FB    & 4  \\
GTE  & 5  & DELL & 5  & KO   & 2  & INTC & 6  & MSFT  & 6  & BAC   & 4  \\
MO   & 5  & ORCL & 4  & SUNW & 1  & AAPL & 6  & C     & 5  & INTC  & 4  \\
JNJ  & 4  & WMT  & 4  & LU   & 1  & BAC  & 5  & WFC   & 5  & GE    & 3  \\
PFE  & 4  & EMC  & 2  & EMC  & 0  & MT   & 4  & AMZN  & 5  & HD    & 3  \\
KO   & 2  & AOL  & 2  & NT   & 0  & AIG  & 4  & GILD  & 5  & AMZN  & 2  \\ \hline
\end{tabular}}
\caption{Outdegrees (Out) ranked for each sub-graph. Outdegree is defined as the number of edges going out of a specific node.}
\label{outdegree}
\end{table}

\begin{table}[]
\resizebox{\textwidth}{!}{
\begin{tabular}{|ll|ll|ll|ll|ll|ll|}
\hline
\multicolumn{2}{|c|}{1990-1994} &
  \multicolumn{2}{c|}{1995-1999} &
  \multicolumn{2}{c|}{2000-2004} &
  \multicolumn{2}{c|}{2005-2009} &
  \multicolumn{2}{c|}{2010-2014} &
  \multicolumn{2}{c|}{2015-2019} \\ \hline
Tic  & In & Tic  & In & Tic  & In & Tic  & In & Tic   & In & Tic   & In \\ \hline
XON  & 22 & BEL  & 15 & MOT  & 12 & MRK  & 19 & MRK   & 18 & UNH   & 17 \\
CHV  & 17 & JNJ  & 15 & INTC & 11 & ORCL & 19 & PM    & 15 & PFE   & 16 \\
AN   & 13 & AOL  & 14 & C    & 11 & BAC  & 16 & JPM   & 15 & HD    & 14 \\
SBC  & 12 & XOM  & 14 & SUNW & 11 & SLB  & 14 & ORCL  & 14 & AMZN  & 14 \\
HWP  & 11 & NT   & 12 & ORCL & 10 & WFC  & 13 & IBM   & 14 & GOOGL & 12 \\
PEP  & 10 & QCOM & 12 & BMY  & 9  & JPM  & 12 & XOM   & 13 & GE    & 12 \\
AHP  & 10 & SUNW & 11 & DELL & 9  & COP  & 12 & PG    & 13 & PG    & 12 \\
MOB  & 10 & ORCL & 11 & TWX  & 7  & MT   & 11 & JNJ   & 12 & CVX   & 11 \\
GTE  & 9  & CSCO & 11 & HPQ  & 7  & PEP  & 10 & DIS   & 12 & BABA  & 9  \\
IBM  & 9  & C    & 11 & GE   & 7  & C    & 10 & INTC  & 11 & MRK   & 9  \\
BMY  & 8  & DELL & 10 & CSCO & 7  & MO   & 10 & WFC   & 10 & DIS   & 9  \\
ABT  & 8  & SBC  & 10 & LU   & 6  & CVX  & 10 & C     & 10 & WFC   & 8  \\
KO   & 7  & BMY  & 9  & HD   & 6  & AIG  & 9  & VZ    & 10 & BA    & 7  \\
BEL  & 7  & WCOM & 9  & TXN  & 6  & T    & 9  & WMT   & 10 & MA    & 7  \\
MSFT & 7  & GE   & 9  & EMC  & 6  & JNJ  & 8  & GE    & 10 & AAPL  & 7  \\
BLS  & 6  & IBM  & 9  & BAC  & 5  & GOOG & 7  & KO    & 9  & T     & 7  \\
PFE  & 6  & MO   & 9  & PG   & 5  & KO   & 7  & BAC   & 9  & WMT   & 6  \\
DD   & 6  & LMG  & 8  & WMT  & 5  & VZ   & 6  & PFE   & 9  & MSFT  & 6  \\
JNJ  & 5  & HWP  & 6  & VZ   & 5  & MSFT & 6  & AAPL  & 9  & INTC  & 6  \\
WMT  & 4  & MSFT & 6  & JPM  & 4  & INTC & 6  & SPY   & 8  & FB    & 5  \\
F    & 4  & BAC  & 5  & KO   & 3  & AAPL & 6  & CVX   & 7  & KO    & 5  \\
MRK  & 4  & INTC & 5  & NT   & 3  & GE   & 6  & AMZN  & 7  & JPM   & 4  \\
PG   & 4  & WMT  & 5  & XOM  & 2  & WMT  & 6  & GILD  & 7  & XOM   & 4  \\
AIG  & 4  & PFE  & 4  & MSFT & 2  & XOM  & 5  & CSCO  & 7  & JNJ   & 4  \\
T    & 3  & AIG  & 4  & AIG  & 2  & PFE  & 5  & GOOGL & 7  & SPY   & 4  \\
GE   & 2  & KO   & 3  & MRK  & 1  & CSCO & 4  & MSFT  & 6  & BRK   & 4  \\
MO   & 2  & MRK  & 2  & PFE  & 1  & HPQ  & 4  & PEP   & 5  & CSCO  & 3  \\
INTC & 2  & PG   & 2  & SBC  & 1  & UBS  & 2  & SLB   & 5  & BAC   & 2  \\
DIS  & 1  & HD   & 2  & IBM  & 0  & IBM  & 1  & T     & 4  & V     & 1  \\
MOT  & 1  & EMC  & 1  & JNJ  & 0  & PG   & 1  & BRK   & 4  & VZ    & 1  \\ \hline
\end{tabular}}
\caption{Indegrees (In) ranked for each sub-graph. Indegree is defined as the number of edges going to a specific node.}
\label{indegree}
\end{table}

\section{Conclusion}\label{sec:con}

In this work, we introduce an information-theoretic measure known as directed information that is capable of capturing nonlinear Granger-causality in an interactive system. 
We develop a novel algorithm based on recurrent neural network utilized with directed information. 
This algorithm can infer the interconnections within a large network with less complexity than previous works. As a proof of concept, we show that our approach performs well both in a linear and in a non-linear simulated environments. Finally, we apply this algorithm to infer the causal relationships among the major US firms during 1990 to 2020.

### 2022-05-24 main/paper3.tex
Add files via upload

\chapter{StockTwits Classified Sentiment and Stock Returns}

\section{Introduction}
\label{S:1}

Recent developments in artificial intelligence and the growing amount of alternative data have created new areas of research in finance. In particular, often coming from news, social media or annual reports, textual data is increasingly used in the literature. \citet{nikfarjam2010text} documents recent text mining approaches for stock market prediction. The pioneer work from \citet{antweiler2004all} computes a bullishness measure out of 1.5 million messages posted on Yahoo! Finance and Raging Bull and finds that stock messages help predict market volatility. Their results clearly reject the hypothesis that all that talk is just noise. They show that there is financially relevant information present. Growing from that, there are now two main sides of sentiment analysis literature studying the forecasting abilities of natural language processing (NLP hereafter): text mining of annual reports and sentiment analysis using social media. On one side, \citet{shirata2011extracting} and \citet{cecchini2010making} report that extracting phrases from annual reports may be an effective predictor of corporate bankruptcy. Similarly, \citet{loughran2011barron} identify phrases that might be red flags indicating questionable behavior. On the other side, \citet{renault2017intraday} builds an intraday investor sentiment indicator using messages and finds that the change in investor sentiment of the first half-hour of a trading day helps forecast the last half-hour market return of that trading day.

We conjecture that our investor sentiment measure can serve as a proxy of unobservable firm fundamentals. For instance, misaligned managerial and shareholder incentives are not easy to observe quantitatively in firm's annual reports (see \citet{nikolov2014agency}) but analysts may talk about them freely in bearish messages. This flow of data can be used throughout a period to improve forecasts (see \textit{nowcasting} in \citet{challet2013predicting}). \citet{tetlock2008more} uses Wall Street Journal stories to examine whether the usage of language is able to predict individual firms’ accounting earnings and stock returns. They find that some aspects of firms' fundamentals are hard-to-quantify, but investors may use linguistic media content to capture information and incorporate it into stock prices.

Most papers studying social media's predictive power use Twitter as their primary source of data. Twitter has the advantage of being used by a wide range of people across the world and a few influencers can attract attention of many investors. In 2013, Carl Icahn tweeted that following a meeting with Tim Cook (Apple CEO), he bought a large position in Apple and believed that the company is extremely undervalued. This bullish tweet caused the market capitalization of Apple to jump by \$12 billion. In 2019, JPMorgan has created the \textit{Volfefe Index} to track Donald Trump's tweets impact on the stock market. However, it is more difficult to disentangle noise from relevant tweets in Twitter than in other more focused social media. Results from \citet{Ghoshal} show that StockTwits is significantly more informative than Twitter data. This is not surprising as StockTwits is a finance-only social media whereas Twitter also captures irrelevant opinions on a wide range of non-finance related matters.

To the best of our knowledge, our paper is the first that compares predictive power of messages in general and around specific events. \citet{ranco2015effects} is likely the closest paper to our study. However, the finance-tailored data we use allow us to get higher contemporaneous correlations between stock returns and polarity. We also use much more messages (90 million versus 1 million) during a longer period (10 years versus 13 months). We also add contribution by looking at the predictive power of cumulative average abnormal polarity (CAAP hereafter). Our paper also contributes to the Efficient Market Hypothesis (EMH hereafter) literature by gauging how cumulative average abnormal returns (CAAR hereafter) and CAAP behave around sudden peaks of message activity. We collect 90 million messages published on StockTwits from mid-2010 to the end of March 2020. On this microblogging platform, users are invited to identify firms with their ticker when they share opinions. Another useful feature on the platform is the possibility to explicitly label their own messages as \textit{bearish} or \textit{bullish} when users post them. We believe that messages on the platform are reliable for several reasons. First, users have incentives to publish valuable information in order to maintain or increase mentions and/or retweets and thus have a greater share of voice in the forum (\citet{sprenger2014tweets}). In addition, market manipulations happen rarely because malicious users have incentives to post fake news only if they previously traded in the same direction than the news they are creating, which will only benefit them if they already have influence on the platform. Finally, SEC closely monitors large influencers to prevent any market abuse.

The challenge in this context is to create a classifier that understands the vocabulary of the messages posted by the investors. For instance, ``bull" is an animal in everyday language but it is someone optimistic in the financial jargon. Work has already been done in this direction: \citet{loughran2011liability} creates a word list, which helps classify tone in a financial document. However, this might not be sufficient in the context of social media because messages posted present many typos, abbreviations and slang, so one needs to have an additional layer of data preprocessing. For instance, the word ``goooooood" would not be recognized by the model if it is not corrected into ``good" first.

In this paper, we use a logistic regression on Term Frequency-Inverse Document Frequency (TFIDF hereafter) vectorized labeled messages to classify unlabeled messages in either bearish, bullish or neutral class. TFIDF is a weighting scheme gauging the importance of a word in a document (see \citet{erdemlioglu2017market}). As users are prone to post more bullish messages than bearish messages, a good classifier needs to take into account the unbalanced data. Without resampling, the classifier outputs classification scores biased towards the over-represented class. Because we are creating an artificial neutral class, we chose to not resample the data but rather optimally select classification score thresholds by maximizing F1 scores of two distinct classification algorithms: bullish versus non-bullish and bearish versus non-bearish. Then, we aggregate messages daily for every firm to compute sentiment polarity time series for individual firms and for the whole economy. 
%\todo{we should find a concise definition of polarity to be put in the abstract and introduction} \TODO{agree. what about something like : degree of bullishness observed on a given firm on a given day}
%Similarly, \citet{pang2002thumbs} suggests that ``there are certain words people tend to use to express strong sentiments, so that it might suffice to simply produce a list of such words by introspection and rely on them alone to classify the texts."

We then use the daily volume of messages on a given firm to identify sudden peak of activity, indicating a firm event. Using abnormal polarity on each event date, we are able to classify events into three classes: bullish, neutral and bearish. We then compute cumulative average abnormal return and cumulative average abnormal polarity in a 41 days window centered at the identified event. We show that abnormal polarities have significantly higher predictive power than abnormal returns. On average, changes of polarity are associated with changes of contemporaneous return of the same sign; but this result does not hold against next-day return. However, when we focus on specific events, polarity has strong predictive power. It is also interesting to note that polarity tends to be biased towards recent past events. Finally, as robustness check, our event study on CAAR is similar to previous literature and is consistent with Fama’s theory.

The remainder of the paper is structured as follows. Section \ref{S:data} presents the data. Section \ref{S:classification} develops the NLP logistic classifier on TFIDF vectorization. Section \ref{S:polarity} explains our polarity measure. Section \ref{S:event studies} contains the results of the event studies. Section \ref{S:portfolios} shows the portfolio construction and Section \ref{S:conclusion} concludes.

\section{Data}
\label{S:data}

Data is coming from two sources: Compustat/CRSP for the stock prices, StockTwits for the investor sentiment. From the CRSP database, we extract daily stock prices (closing prices), daily volume of transactions and number of shares outstanding from 2010 to 2020 for all US and Canadian firms. Stock prices and number of shares are adjusted to account for any distribution (i.e. dividends, stock splits) so that a comparison can be made on an equivalent basis before and after the distribution.

StockTwits is a large social network similar to Twitter but designed for investors and traders. Users register online and can post messages about any listed firm through the prefix \$ followed by the ticker of the firm. StockTwits was created in 2008 as an app built on the Twitter's API and later detached from Twitter to build a standalone social network. As of April 2019, it has over two million registered users and the number of messages posted is growing exponentially (see Figure \ref{fig:tweet_vol_daily}). StockTwits describes itself as ``the voice of social finance and the best way to find out what is happening right now in the markets and stocks you care about". In practice, it is effectively used by finance professionals to express their opinions on individual firms and the market as a whole. Since mid-2010, users also have the possibility to label their own messages as either \textit{bullish} or \textit{bearish}. This feature is very useful for researchers as even though many messages are unlabeled, it allows for sentiment classification using NLP techniques. The reasons of using StockTwits and not another social media is threefold. First, one of the main challenge in Natural Language Processing is the creation of an appropriate labeled vocabulary. \citet{loughran2011barron}) shows that it is essential to have a specific vocabulary to interpret finance documents (i.e many words have a different meaning in finance than in traditional English (e.g “bear trap”)). In addition to that, social media slang is an additional layer of language complexity. To this extent, the functionality to self-tag \textit{bullish} and \textit{bearish} messages that StockTwits implemented in 2012 is very valuable as it allows the creation of a specific labeled vocabulary out of labeled messages. We are not aware of any other social media platform in finance offering this functionality. Second, StockTwits is less noisy than Twitter because messages focus on finance and economics matters only. Third, extracting data out of StockTwits is easy because of its API. StockTwits' API is designed to query the database to download messages via JSON requests. Using a Python script, we extract all messages since the 10th of July 2009 until the 31st of March 2020 for a preset list of tickers coming from the Compustat database corresponding to all US and Canadian firms. This results in 90 million messages, which we download and store as JSON files. As one message may refer to several tickers, we consider a message with two or more tickers as one message for every ticker identified. We refer to the appendix for more information about this process.

 \begin{figure}
    \centering
    \includegraphics[width=0.75\textwidth]{tweet_vol_daily.JPG}
    \caption{Number of messages posted daily on StockTwits.}
    \label{fig:tweet_vol_daily}
\end{figure}

Every message comes with the following eight features: (1) the ticker discussed (2) the exact timestamp of the message, (3) a unique message identifier, (4) the body of the message, (5) the sentiment label (bearish, bullish, or none) entered by the user, (6) a unique identifier of the user who sent the message, (7) the number of messages published by the user who sent the message, and (8) the number of followers of the user who sent the message. Figure~\ref{screenshot} shows a screenshot of the StockTwits website as of 3rd March 2020, for a query on the AAPL\footnote{AAPL is the ticker for Apple.} ticker. The first message is labeled as bullish by the user ``satkaru", the two next are unlabeled messages that will be classified using a machine learning classifier, and the last message is labeled as bearish by the user ``Etrading''.

\begin{figure*}[h]
\centering
\includegraphics[width=0.75\textwidth]{stocktwits_screeshot_3march2020.JPG}
\caption{Screenshot of StockTwits as of 3rd March 2020. The first message is labeled as bullish by the user ``satkaru", the two next are unlabeled messages that will be classified using a machine learning classifier and the last message is labeled as bearish by the user ``Etrading''.}
\label{screenshot}
\end{figure*}

Figure \ref{categrepartition-preclassif} shows the distribution of user-labeled and unlabeled messages. Overall, around 30 million messages are user-labeled and 60 million messages are unlabeled. Among the user-labeled messages we find five times more bullish than bearish ones. This ratio indicates that investors are on average optimistic about the market, which is consistent with findings in the literature, e.g., \citet{renault2017intraday}. Such class imbalance is a well-known issue in machine learning classification, which we address below. We will classify the unlabeled messages using a machine learning algorithm trained on the set of user-labeled messages. Since not every message contains substantial information, we believe that the sentiment classification should not be a bullish/bearish dichotomy. Hence, we allow for a neutral class to account for messages that do not take a clear stand.

\begin{figure*}[h]
\centering
        \begin{subfigure}[b]{0.75\textwidth}   
            \centering 
            \includegraphics[width=\textwidth]{preclassif.JPG}
        \end{subfigure}
        \caption{Number of user-labeled messages in each category. There are around 60 million unlabeled messages that we classify using the 30 million user-labeled bullish and bearish messages.} 
        \label{categrepartition-preclassif}
\end{figure*}

Figure \ref{fig:data-stocktwits1} shows a log-log histogram of the number of followers per user and a log-log histogram of the number of messages posted by users. As expected, there are few users with many followers (they can be seen as ``influencers"), and many users with a few followers. In addition, most users seem to post on average between 10 and 400 messages and a few post a lot more.

\begin{figure*}[]
\centering
        \begin{subfigure}{0.75\textwidth}   
            \centering 
            \includegraphics[width=\textwidth]{Followers.JPG}
        \end{subfigure}
        \quad
        \begin{subfigure}{0.75\textwidth}
            \centering
            \includegraphics[width=\textwidth]{Tweets.jpg}
        \end{subfigure}

        \caption{User summary statistics. Top graph is a log-log histogram of the number of followers per user and the bottom graph shows the log-log histogram of the number of messages posted by users.} 
        \label{fig:data-stocktwits1}
\end{figure*}

Figure \ref{fig:data-stocktwits2} shows the top 30 most discussed tickers on the platform. Of all tickers extracted, around 75\% are ordinary common share, 15\% are ETFs and the remaining 10\% are other types of securities. Not surprisingly, the S\&P index is the most discussed, followed by Apple and other big tickers. The messages about the 15 (30) biggest tickers represent 20\% (25\%) of the total number of messages, which indicates that users talk about a wide panel of tickers and not only big firms. The bottom graph shows a histogram of the number of messages per ticker. The x-axis is log-scaled because of extreme values, the distribution is highly skewed.

\begin{figure*}[]
\centering
\begin{subfigure}{0.75\textwidth}  
            \centering 
            \includegraphics[width=\textwidth]{top30tickers.jpg}
        \end{subfigure}
        \quad
        \begin{subfigure}{0.75\textwidth}   
            \centering 
            \includegraphics[width=\textwidth]{LogTweetsPerFirm_3.jpg}
        \end{subfigure}
        \caption{Firm summary statistics. Top graph shows the top 30 most discussed tickers on the platform. SPY is the ticker of the S\&P and AAPL is the ticker for Apple. Bottom graph shows the distribution of the number of messages across tickers.} 
        \label{fig:data-stocktwits2}
\end{figure*}

Messages contain qualitative information, which needs to be transformed into quantitative data for the computer to understand. Thereto, we first apply the following preprocessing operations: an apostrophe handler, a contraction form handler (i.e. ``aren't" becomes ``are not"), tickers removal, stop words (i.e. ``a", ``the", ``of") removal\footnote{We follow \citet{renault2020sentiment} and \citet{saif2014stopwords} and use a restrictive list of stopwords to avoid accuracy decrease.}, users removal, lemmatization, URLs removal and a simple spell corrector dealing with more than two repeated characters (i.e. ``soooo goooood" becomes ``soo good"). Table \ref{tab:preprocess} shows five examples of messages before and after preprocessing.

One of the first steps in NLP is the tokenization : the way of slicing a piece of text in smaller units called tokens or terms. In financial lingo, some words only have meaning when associated with other words (i.e ``bad apple" or ``bear flag"). N-gram models allow accounting for words frequently occurring together with other words. The main hyperparameter in N-gram models is the size of the group of words considered : unigram is a term with only one word, bigram is a term with two consecutive words, etc. Bigger N-grams models increase dramatically the size of the vocabulary (i.e : the collection of all terms considered). We are able to control the size of the vocabulary by tuning a hyperparameter keeping only a given number of most frequent terms.  We chose to work with a vocabulary consisting of the one million most frequent unigrams, bigrams or trigrams.

Figure \ref{wordclouds} represents the bullish and bearish word clouds. These correspond to the most frequent terms (up to 3-grams) in user-labeled bullish (bearish) messages relative to their total appearance. The size of the terms represent their importance in the cloud. In the bullish cloud, we see words such as ``bullish divergence", ``room to grow", ``lot potential" which are clearly bullish signals. In the bearish cloud, we find words such as ``recent resistance", ``short setup", ``bad apple" which indeed indicate bearish signals. These findings are reassuring in the sense that the content of the messages on the platform are consistent with their labels. The term ``aldox" in the bullish cloud caught our attention. After some research, it is an abbreviation for Aldoxorubicin, a drug against tumors and is associated with pharmaceutical messages where investors were very enthusiastic about it (example of related message: ``aldox is on the slide. have great faith this is truly world change"). That is why the term is appearing almost exclusively in bullish messages, hence in the bullish cloud.

On another note, most of the times messages are written by humans. It is however possible that some messages are generated by a robot (for instance to spread news or articles). In such cases, we define a neutral class that would absorb these messages if no substantial information is embedded in these messages. The term "long position open" present in the bearish cloud is an anomaly due to bearish user-labeled messages of intraday alerts such as “sell \$labd close labd long position. open labd short position. time: 14:53 ny price: \$13.64 zquant intraday alerts”. However, such anomalies are not an issue, a message “long position open” has a bullish probability of 0.91 and gets classify as bullish as it should. This is the case because when classifying a message, the score of the unigram “long” is much stronger than the score of the trigram “long position open”. From a linguistic point of view our approach is brute force. However, in turn, it works at a massive scale. Overall, the data quality is very good.

\begin{figure}
\begin{center}
    \includegraphics[width=0.75\textwidth]{Positive rate words.jpg}

    \vspace{1cm}

    \includegraphics[width=0.75\textwidth]{Negative rate words.jpg}
   \vspace{0.1cm}
    \caption{Bullish word cloud (top), bearish word cloud (bottom). These correspond to the most frequent terms (up to 3-grams) in user-labeled bullish (bearish) messages relative to their total appearance. The size of the terms represent their importance in the cloud.}
    \label{wordclouds}
  \end{center}
\end{figure}

\newpage
\begin{table}[H]
\begin{tabular*}{\textwidth}{l@{\extracolsep{\fill}}ll}
\hline \multicolumn{2}{l}{Before preprocessing}                         \\   \hline \hline                                                                          
(1) & @CassandraTwit \$uvxy contango 3.5\%...still long. goooooood      \\
(2) & \$FRPT Take profits while you still can. \\
(3) &  \$UVXY \$tvix go time boys and girls. Holding overnight again                                                                                                                                        \\
 (4) & \$dnr Nice upgrade as company goes into its quiet period!      \\
 (5) & \$SPY market won't reverse again towards closing. Get put options. \\ 
 \hline                                                                                          
\end{tabular*}
\end{table}

\begin{table}[H]
\begin{tabular*}{\textwidth}{l@{\extracolsep{\fill}}ll}
\hline \multicolumn{2}{l}{After preprocessing}               \\   \hline \hline                                                                       
(1) & contango still long good            \\
(2) & take profit while you still can \\
(3) & go time boy and girl hold overnight again                                                                  \\
(4) & nice upgrade as company go into its quiet period     \\ 
(5) & market will not reverse again towards closing get put options \\ \hline                                          
\end{tabular*}
\caption{Preprocessing of five sample messages. Preprocessing operations include: punctuation removal, lower casing, apostrophe handling, contraction form handling (i.e. ``won't" becomes ``will not"), tickers removal, users removal, URLs removal, parsing and a simple spell corrector dealing with more than two repeated characters (i.e. ``goooood" becomes ``good")}
\label{tab:preprocess}
\end{table}

\section{Sentiment classification}\label{S:classification}

Since mid-2010, StockTwits users have the choice to label their own messages as either \textit{bearish} or \textit{bullish} or to leave it unlabeled. Unlabeled messages are tricky to deal with because the user either deliberately chose to leave the message neutral by not labeling it or forgot to click on a label. Figure \ref{fig:ts_neutral_vs_uncl} shows the proportions of user-labeled messages in each category across time. In the early years of the platform, most messages are unlabeled, presumably because users were not familiar with the sentiment label yet. Albeit the proportion of unlabeled messages monotonically declines over the years, almost 60\% of the more recent messages are still unlabeled. We believe that by far not all unlabeled messages reflect neutral opinions. Figure \ref{fig:postclassif_acrosstime} shows that a lot of unlabeled messages get classified in either bullish or bearish messages, hence these unlabeled messages had indeed valuable information that we are now able to capture.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.75\textwidth]{preclassif_acrosstime.JPG}
    \caption{Proportions of user-labeled messages in each category: bullish (green), bearish (red), and unlabeled (gray). Proportions are aggregated monthly.}
    \label{fig:ts_neutral_vs_uncl}
\end{figure}

As one of the goal of this paper is to build an accurate time-series sentiment measure for individual firms, it motivates the use of Natural Language Processing to classify all unlabeled messages in either bullish, neutral or bearish class.

To classify unlabeled messages, we use a logistic regression of the labels on TFIDF transformed user-labeled messages, as in \citet{nlp1} and \citet{classifq}. TFIDF stands for Term Frequency-Inverse Document Frequency and is a widely used method to transform a text, in our case a message $m$, into a numerical vector, $TFIDF_m$. The dimension of this vector is equal to the size of the vocabulary (the collection of all terms across all messages). The components of the vector encode the importance of the corresponding terms $t$ in the message $m$, as formally defined by $TFIDF_{m,t} = TF_{m,t} \cdot IDF_t$. The first factor measures how frequently term $t$ appears in the message,
\[ TF_{m,t} =\dfrac{\sum_{i=1}^{N_m} \textbf{1}_{t=t_{m,i}}}{N_m},\]
where $N_m$ denotes the number of terms $t_{m,i}$ in message $m$. The second factor measures how important term $t$ is to the message,
\[ IDF_t = \log \big(\dfrac{V}{\sum_{j=1}^{V} \textbf{1}_{t \in m_j}}\big),\]
where $V$ denotes the total number of messages $m_j$. A term $t$ appearing in many documents (such as ``the", ``is", ``of") is likely to have low information content, hence a low $IDF_t$.

As seen in Figure \ref{categrepartition-preclassif}, user-labeled messages exhibit five times more bullish messages than bearish messages. Such an imbalance is a well-known issue in machine learning classification and needs to be dealt with to avoid biases towards the over-represented class (see \citet{chawla2004special}). To tackle class imbalance, the most common technique is to randomly oversample the minority class, which consists of repeating some samples of the minority class and balance the number of samples between classes in the data. We follow a different approach. In addition to bullish and bearish classes, we define an artificial neutral sentiment class as it is possible that some users are not expressing any opinion and sometimes post finance-irrelevant messages. To do so, we optimally select classification score thresholds and eliminate the class imbalance bias at the same time.

Performance measures widely used in machine learning classification are precision, recall, F1 score and accuracy. The first step is to define a class as the positive class. Instances (messages) are then divided into true positives TP (predicted positive, actual positive), false positives FP (predicted positive, actual negative), true negatives TN (predicted negative, actual negative), and false negatives FN (predicted negative, actual positive). Precision $PRE=\dfrac{TP}{TP+FP}$ is the proportion of true positives among the predicted positives. Recall $REC = \dfrac{TP}{TP+FN}$ is the proportion of true positives among the actual positives. The precision-recall trade-off is captured by the F1 score, $\dfrac{2 \cdot PRE \cdot REC}{PRE + REC}$, the harmonic mean of precision and recall. Accuracy $ACC = \dfrac{TP+TN}{TP+TN+FP+FN}$ is the fraction of correct predictions regardless of the label.

We use 80\% of the user-labeled (bearish and bullish) messages as a training set and keep 20\% as a test set, then we run two binary classifiers. The first (second) classifier sets bullish (bearish) as positive and non-bullish (non-bearish) as negative class. Every message then falls into the following set of labels: \{(non-bullish, bearish), (bullish, bearish), (non-bullish, non-bearish), (bullish, non-bearish)\}. For the two outer cases the two algorithms agree and the final classification is defined to be bearish (non-bullish, bearish) or bullish (bullish, non-bearish), respectively. For the two inner cases, (bullish, bearish) and (non-bullish, non-bearish), the two algorithms disagree, so that the final classification is defined to be neutral. Formally, every message $m$ is then mapped on either

\[ m\mapsto \begin{cases} \text{(non-bullish, bearish)} & =: \text{bearish} \\ \text{(bullish, bearish)}& =: \text{neutral} \\ \text{(non-bullish, non-bearish)}& =: \text{neutral} \\ \text{(bullish, non-bearish)} &=: \text{bullish.} \end{cases}\]

Precision, recall, and F1 scores of the two algorithms differ because they depend on which class is defined as the positive one. To select optimal classification thresholds, we maximize the F1 score of either algorithm. The green (red) line in Figure \ref{fig:optimal_thresholds} is the F1 score for the bullish versus non-bullish (bearish versus non-bearish) classification, respectively. Circles indicate the maximal F1 scores, along with the corresponding classification score thresholds, 0.50 and 0.72, respectively.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.75\textwidth]{optimal thresholds.JPG}
    \caption{Optimal classification thresholds. The green (red) line is the F1 score for the bullish versus non-bullish (bearish versus non-bearish) classification, respectively. Circles indicate the maximal F1 scores, along with the corresponding classification score thresholds.}
    \label{fig:optimal_thresholds}
\end{figure}

If the classification score of a message is bigger (smaller) than 0.72 (0.50), then both classifiers agree on the sentiment and the message is classified as bullish (bearish), respectively. If the classification score is between 0.50 and 0.72, the classifiers disagree, (bullish, bearish), and we consider the message as neutral. Finally, we overwrite the sentiment of a message predicted by the classifier by the user-labeled sentiment whenever the latter is available. Research in sentiment classification shows that human annotators tend to agree about 80 to 85\% of the time when evaluating the sentiment of a document (see e.g. \citet{wilson2005recognizing} and \citet{chen2020large}). This usually represents the accuracy that a sentiment classifier should meet or beat. The accuracy in the test set of our combined classifier is 86\%.

Figure \ref{fig:postclassif_acrosstime} shows the proportions of classified messages in each category. Percentages of bearish (sum of labeled and classified as bearish) and bullish (sum of labeled and classified as bearish) messages are stable over time, suggesting that our classification method is robust. Even if most messages were not user-labeled in the early years of the platform, as seen in Figure \ref{fig:ts_neutral_vs_uncl}, we are now able to classify the sentiment of most messages posted in this period. Consistent with the over-representation of bullish messages observed in the user-labeled messages in Figure \ref{categrepartition-preclassif}, there are many more messages classified as bullish than bearish. Typical messages classified as bullish are messages such as “buy buy” or “hope the pump come soon” whereas typical bearish messages are messages such as “sell everything” or “start short position here”. Neutral messages are either empty, irrelevant to finance (e.g. “political posturing friend”\footnote{This is a reply to the following message : "honestly, how dumb can you be to believe that china was going to buy significant amount of agricultural products after the breakdown in trade talks. even if they buy it will be just a little bit and not significant"}) or ambiguous (e.g. “lol wow”).

\begin{figure}[h]
    \centering
    \includegraphics[width=0.75\textwidth]{postclassif_acrosstime3.JPG}
    \caption{Proportions of classified messages in each category: bullish (light green predicted, green user-labeled), bearish (light red predicted, red user-labeled), and neutral (gray). Proportions are aggregated monthly.}
    \label{fig:postclassif_acrosstime}
\end{figure}

\section{Polarity}\label{S:polarity}

To build sentiment measures for individual firms and the whole market on a given day, we count the sentiments in the messages. To match the timeline on which close-to-close stock returns are computed and to avoid forward-looking bias, we aggregate messages on a close-to-close manner. That is, polarity on day $t$ is computed with messages posted from 4:00 pm on day $t-1$ to 4:00 pm on day $t$. As stock returns are not computed outside of business days, we shift messages posted outside of business days to the next business day available, then remove any non-business day from our sample. We denote by $C_{i,t,j}$ the sentiment of the $j$th message about firm $i$ on day $t$. It is set to 1, 0, or -1 for bullish, neutral, or bearish, respectively. We follow \citet{ranco2015effects} and define the polarity of firm $i$ as
\[    P_{i,t} = \dfrac{\sum_{j=1}^{V_{i,t}} (\textbf{1}_{C_{i,t,j} = 1} - \textbf{1}_{C_{i,t,j} = -1})}{\sum_{j=1}^{V_{i,t}} (\textbf{1}_{C_{i,t,j} = 1} + \textbf{1}_{C_{i,t,j} = -1})},\]
where $V_{i,t}$ denotes the number of messages about firm $i$ on day $t$.\footnote{If $V_{i,t}=0$ then we set $P_{i,t} =0$.}

As an aggregate measure, we define the polarity for the whole market as a weighted average over all firms 
\begin{equation*}
\label{marketpoleq}
    P_t^M = \dfrac{\sum_{i} V_{i,t} \cdot P_{i,t}}{V^M_t} ,
\end{equation*}
where $V^M_t = \sum_{i} V_{i,t}$ denotes the number of messages on day $t$.

Figure \ref{fig:marketpolspy} shows a scatter plot of the market polarity $P_{t}^M$ versus the polarity of the SPY\footnote{SPY is an ETF tracking the S\&P500 return. It is the largest ETF in the world.}. We do not expect the market polarity to be the SPY polarity because the stock universe is not the same (market polarity contains stocks that are not necessarily in the S\&P500 and vice versa). The slope coefficient of the regression line is statistically significantly positive and the contemporaneous Pearson correlation coefficient is 0.53, suggesting that the market polarity is an accurate measure of the aggregated sentiment of the market. Also, consistent with Figure \ref{fig:postclassif_acrosstime}, SPY and market polarities are bullish-biased.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.75\textwidth]{spy_vs_marketpol.JPG}
    \caption{Market polarity on the y-axis versus the polarity of the SPY on the x-axis. The red line shows the linear regression line and coefficients.}
    \label{fig:marketpolspy}
\end{figure}

At this point, for stationary reasons\footnote{Computing the polarity ratio with few daily observations would lead to a spiky time series.}, we compute the median of daily message volume for each ticker and exclude from our sample tickers that have less than a median of 50 daily messages. Our final sample contains 19 tickers. We refer to the appendix for the list of tickers covered as well as more information about the trimming process.

To understand how polarity is related to investor sentiment, we run two linear regressions of contemporaneous daily returns on polarity and 5-days Cumulative Abnormal Polarity :
\begin{align*}
    R_{i,t} &= \alpha + \beta \cdot P_{i,t} + \epsilon_{i,t}, \\
    R_{i,t} &= \alpha + \beta \cdot CAP_{i,5} + \epsilon_{i,t}. 
\end{align*}
Table \ref{regpolret} shows that $\beta$ is positive and significant for both regressions. This indicates that polarity is a good proxy for the sentiment of investors. Further supporting evidence is given by the correlation between polarity and contemporaneous stock returns at the firm level. Figure \ref{fig:ts} shows the time series during 2019 for the top 6 most discussed tickers. In our entire panel of firms, correlations are always positive and range between 0.1 and 0.3.

\begin{table}[]
\centering
\begin{tabular}{l|c|c|c|c|c}
                     & $R_{i,t}$ & $R_{i,t}$   & $R_{i,t+1}$ & $R_{i,t+1}$        \\ \toprule 
Constant             & -0.0047*** & -7.92e-05   & -0.0002    & 7e-06         \\
                     & (0.000)    & (9.81e-05)  & (0.000)    & (9e-05)       \\
$P_{i,t}$           & 0.009***   &               & 0.0003     &               \\
                     & (0.000)    &               & (0.000)    &               \\
$CAP_{i,5}$         &            & 0.0007***     &            & -0.0002       \\
                     &            & (0.000)       &            & (0.000)       \\ \bottomrule
$R^2$                & 0.012      & 0.000         & 0.000      & 0.000              \\
No. Obs.             & 34100      & 34024         & 34100      & 34024             
\end{tabular}
\caption{Results from linear regressions of contemporaneous and next period stock returns on polarity. Stock returns are trimmed at the 5\% percentile on both sides. Standard errors are reported in parentheses. Statistical significance at the 99\%, 95\%, and 90\% level is indicated with ***, **, *, respectively.}
\label{regpolret}
\end{table}

\begin{figure*}[]
        \centering
        \begin{subfigure}[b]{0.48\textwidth}
            \centering
            \includegraphics[width=\textwidth]{ts_pol_ret_aapl.JPG}
        \end{subfigure}
        \quad
        \begin{subfigure}[b]{0.48\textwidth}  
            \centering 
            \includegraphics[width=\textwidth]{ts_pol_ret_amzn.JPG}
        \end{subfigure}
        \vskip\baselineskip
        \begin{subfigure}[b]{0.48\textwidth}   
            \centering 
            \includegraphics[width=\textwidth]{ts_pol_ret_tsla.JPG}
        \end{subfigure}
        \quad
        \begin{subfigure}[b]{0.48\textwidth}   
            \centering 
            \includegraphics[width=\textwidth]{ts_pol_ret_spy.JPG}
        \end{subfigure}
                \vskip\baselineskip
        \begin{subfigure}[b]{0.48\textwidth}   
            \centering 
            \includegraphics[width=\textwidth]{ts_pol_ret_amd.JPG}
        \end{subfigure}
        \quad
        \begin{subfigure}[b]{0.48\textwidth}   
            \centering 
            \includegraphics[width=\textwidth]{ts_pol_ret_fb.JPG}
        \end{subfigure}
        \caption{Time series of daily polarity (red - left axis) and daily stock returns (blue - right axis) since 1st of January 2019 for the top 6 most discussed tickers. Pearson correlation between the two time series is shown in the title.} 
        \label{fig:ts}
    \end{figure*}

We also run two linear regressions of next day returns on polarity and 5-days Cumulative Abnormal Polarity :
\begin{align*}
    R_{i,t+1} &= \alpha + \beta \cdot P_{i,t} + \epsilon_{i,t}, \\
    R_{i,t+1} &= \alpha + \beta \cdot CAP_{i,5} + \epsilon_{i,t}. \\
\end{align*}
Table \ref{regpolret} reveals that polarity has no predictive power for next day stock returns unconditionally. However, the next section depicts how polarity still has embedded information around specific events.

\section{Event studies}\label{S:event studies}

Event studies constitute a statistical method widely used in financial econometrics. In general, they are used to measure the effect of events on the market value of firms. Well known applications of event studies include the testing of various forms of the efficient market hypothesis (EMH) (see \citet{fama1969} and \citet{fama1991efficient}). What's more, as described in \citet{mackinlay1997event}, event studies can also be applied with little modification to other variables than stock returns.

To design an event study, the first step is to define events of interest and the \textit{event window} over which a variable will be examined. Adhering to common practice, we choose the event window expanded from 20 business days before the event to 20 business days after the event. The \textit{estimation window} is used to estimate the parameters of the market model. A common choice is a one-year window ending 20 days before the event. Figure \ref{fig:timeline} shows the corresponding timeline.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.75\textwidth]{timeline.JPG}
    \caption{Timeline of our event studies. $\tau$ is an event date, the event window has 41 business days centered around the event date and the estimation window is a one-year rolling period prior the event.}
    \label{fig:timeline}
\end{figure}

Then, event studies require the distinction between a normal and an abnormal measure. The normal measure is defined as the expected measure conditioning on the event not taking place. In other words, we need an expectation of the variable of interest that would have been measured if the event never existed. The abnormal measure is defined as the measure minus the normal measure. Formally, let $X_{i,t}$ be a measure of interest (i.e. stock return, polarity, ...), $AX_{i,t}$ be the abnormal measure and $E(X_{i,t}|Z_{i,t})$ the normal measure with $Z_{i,t}$ being the conditional information for the normal measure model. We have: 
\[     AX_{i,t} = X_{i,t} - E(X_{i,t}|Z_{i,t}).\]

The normal measure $E(X_{i,t}|Z_{i,t})$ is usually computed using either a constant mean model or a market model. Constant mean models use $E(X_{i,t}|Z_{i,t}) = \mu$ where $\mu$ is the mean of the measure during the estimation window. We chose to work with a market model but all following results hold with the constant mean model as well. As stated in \citet{mackinlay1997event}, the variance of the abnormal measure is not reduced a lot by choosing a more sophisticated model, hence the event study is not sensitive to the choice of the normal model. 
The market model we are using is defined as the following: 
\begin{equation*}\label{eq_mm}
    X_{i,t} = \alpha_i + \beta_i \cdot X_{t}^M + \epsilon_{i,t},
\end{equation*}
with $E(\epsilon_{i,t}) = 0$, $V(\epsilon_{i,t}) = \sigma_{\epsilon}^2$ and $X_{t}^M$ the measure of the whole market. We estimate the parameters of the market model in a one-year rolling estimation window. To avoid overlaps between estimation windows and events, we remove any event day from the estimation windows. This ensures that large event returns do not influence the parameters of the normal measure.

\subsection{Events}
We define an event as an unusual high number of daily messages for a particular firm. We conjecture that a sudden peak in StockTwits message volume indicates that an important firm event is happening on the day of the peak. As a robustness check, we show in Figure \ref{fig:actvol} that increases (decreases) in number of message volume are positively associated with increases (decreases) in contemporaneous weekly volume of stock transactions. These co-movements mean that investors not only post about these stocks but also trade them simultaneously, adjusting their portfolios. This indicates that the message volume peaks is a good proxy to identify event dates as there is no lag between message activity and trades.

\begin{figure}
    \centering
    \includegraphics[width=0.75\textwidth]{activity-volume.jpg}
    \caption{Changes in weekly volume of transactions on the y-axis versus changes in message activity on the x-axis. Activity is measured in weekly messages posted per firm.}
    \label{fig:actvol}
\end{figure}

To measure unusual activity peaks, we use the market model (\ref{eq_mm}) on a one-year rolling estimation window and regress the daily relative change of message volume of individual firms $\dfrac{\Delta V_{i,t}}{V_{i,t-1}}$ on the daily relative change of total message volume $\dfrac{\Delta V_t^M}{V_{t-1}^M}$. Formally,
\[     \dfrac{\Delta V_{i,t}}{V_{i,t-1}} = \alpha_i^V + \beta_i^V \cdot \dfrac{\Delta V_t^M}{V_{t-1}^M} + \epsilon_{i,t}.\]
We can then compute the abnormal volume of messages for firm i as : 
\[     AV_{i,t} = \dfrac{\Delta V_{i,t}}{V_{i,t-1}} - \hat{\alpha}_i^V - \hat{\beta}_i^V \cdot \dfrac{\Delta V_t^M}{V_{t-1}^M}. \]
We define an event for ticker $i$ as a day $t$ where the standardized abnormal volume exceeds 2, 
\[   \dfrac{AV_{i,t} - \mu_{AV_i}}{\sigma_{AV_i}} > 2.\]

%To measure unusual activity peaks, we use the time series of individual firms message volume and identify days where the volume exceeds five times its rolling one-month median. Formally, the event dates for firm $i$ are the days $e$ such that:
%\begin{equation}
%    e = \{d : V_d > 5 \cdot V_m(d) \}
%\end{equation}
%where $V_d$ is the message volume on day d for firm i and $V_m(d)$ is the rolling one-month volume median at day d for firm i.

Then, we define the type of the event as either bullish, neutral or bearish. We use the abnormal polarity $AP_{i,t}$ of the event date to assess how on average investors perceive the event. Figure \ref{fig:abnpol_all} shows the distribution of abnormal polarities. We chose to use the one-third (-0.03) and two-third percentile (0.07) of the distribution of abnormal polarities as thresholds for the type of the event. Conditionally on the existence of an event for firm $i$ at day $t$, we have:

\[     Type_{i,t}= \begin{cases}
      Bullish  & \text{if}\ \ AP_{i,t}> 0.07, \\
      Neutral  & \text{if}\ \ AP_{i,t} \in [-0.03, 0.07], \\
      Bearish  & \text{if}\ \ AP_{i,t} < -0.03. \\
\end{cases} \]

\begin{figure}[h]
    \centering
    \includegraphics[width=0.75\textwidth]{abnpol_distr_all.JPG}
    \caption{Empirical distribution of abnormal polarities. Red dashed lines show the one-third and two-third percentiles.}
    \label{fig:abnpol_all}
\end{figure}

As an illustration, Figure \ref{fig:activityaapl} shows for Apple the time-serie of message volume and the bullish, neutral and bearish events identified as green up-triangles, gray circles and red down-triangles, respectively. Between 2011 and 2020, our algorithm identified 73 events for Apple. Interestingly, our methodology allows us to capture more than earning announcements : about half of these events correspond to earning announcements, but some also correspond to Apple \textit{Keynotes}\footnote{Keynotes are presentations that Apple gives to the press, often presenting new products.} or even CEO letters addressed to investors. Across 19 tickers, we identify 1131 events distributed across the three categories : 454 bullish events, 294 neutral events and 383 bearish events. This coverage is on par with previous studies (i.e: \citet{mackinlay1997event} has 30 firms and 600 events) Figure \ref{fig:events} shows the identified events and their types across the years.
%Table \ref{tab:aaplevent} shows all events for Apple and their associated types.

\begin{figure}[]
    \centering
    \includegraphics[width=0.75\textwidth]{activity_aapl.JPG}
    \caption{Daily message volume for Apple. Events are days with an unusual high number of messages. Green upper-triangles show bullish events, gray circles are neutral events and red down-triangles represent bearish events.}
    \label{fig:activityaapl}
\end{figure}

%\begin{table}[]
%\caption{Identified event dates and associated description for Apple between %2013 and 2020.}
%\centering
%\begin{tabular}{|l|l|}
%\toprule
%Date         & Description\\
%\bottomrule
%2013-10-28  & Earnings announcement\\
%2014-04-23  & Earnings announcement\\
%2014-09-09  & Keynote \\
%2015-10-27  & Earnings announcement\\
%2016-01-26  & Earnings announcement\\
%2016-07-26  & Earnings announcement\\
%2016-09-07  & Keynote \\
%2016-10-25  & Earnings announcement\\
%2017-01-31  & Earnings announcement\\
%2017-05-02  & Earnings announcement \\
%2017-08-01  & Earnings announcement \\
%2017-08-02  & Earnings announcement \\
%2017-09-12  & Keynote \\
%2017-11-02  & Earnings announcement\\
%2018-05-01  & Earnings announcement\\
%2018-07-31  & Earnings announcement\\
%2018-08-01  & First company to reach \$1 trillion\\
%2018-11-01  & Earnings announcement\\
%2019-01-02  & CEO Letter to investors\\
%2019-01-29  & Earnings announcement\\
%2019-03-25  & Keynote \\
%2019-04-30  & Earnings announcement\\
%2019-07-30  & Earnings announcement\\
%2019-09-10  & Keynote\\
%\bottomrule
%\end{tabular}
%\label{tab:aaplevent}
%\end{table}

\begin{figure}[]
    \centering
    \includegraphics[width=0.75\textwidth]{events_across_time.JPG}
    \caption{Number of events in each category across time. Numbers are aggregated monthly.}
    \label{fig:events}
\end{figure}

\subsection{CAAR and CAAP}
We estimate the parameters for the market models using OLS, as it is a consistent and efficient estimator under general conditions.\footnote{An interested reader can refer to \citet{mackinlay1997event} for more details}

\subsubsection{Abnormal Returns}
Given the market models parameters $\hat{\alpha_i}^R$ and $\hat{\beta_i}^R$ estimated in \eqref{eq_mm}, the abnormal return is
\[     AR_{i,t} = R_{i,t} - \hat{\alpha_i}^R - \hat{\beta_i}^R \cdot R_t^M,\]
with $R_t^M$ being the market return. Then, we can compute the cumulative abnormal returns (CAR) around a firm $i$ event $\tau$ as:
\[    CAR_i(\tau,t) = \sum_{s=- 20}^{t} AR_{i,\tau + s},\]
and finally get the cumulative average abnormal returns (CAAR) across the N events as:
\[     CAAR(t) = \dfrac{1}{N} \sum_{j=1}^N CAR_{i_j}(\tau_j,t).\]
Variance of CAR is computed following \citet{mackinlay1997event} as :
\[     var(CAR(\tau,t)) = \dfrac{1}{N^2} \sum_{i=1}^N (CAR_i(\tau,t) - CAAR(t))^2.\]

Figure \ref{fig:ES_CAAR_MM} shows the CAAR around the events identified. This plot is consistent with \citet{mackinlay1997event}. It shows that CAAR related to bearish (bullish) events displays a downward (upward) jump at the event date respectively, and then the jumps are followed by a stable CAAR during the 20 days period after an event. Interestingly, there is a systematic (small) shift in the CAAR already 1 day before an event. The CAAR related to the neutral events exhibits a slight upward shift around the event date but it fades away after a few days. The CAAR related to bearish events shifts already a few days before the event but this shift is not statistically significant. Additionally, as we see in the top-left plot of Figure \ref{fig:boxplotcaar}, box plots are not shifted, indicating that the conditional distributions of the CAR are not statistically different from each other. Mann-Whitney U-test (Table \ref{utest_res}) shows that 5 days before an event, the median of the CAR distribution of the bullish events is neither statistically different from the median of the neutral nor the bearish events.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.75\textwidth]{caar.JPG}
    \caption{Cumulative average abnormal returns around identified events. CAAR related to bearish, neutral and bullish events are displayed with the red, gray and green line, respectively. Areas around lines show confidence intervals at the 95\% level.}
    \label{fig:ES_CAAR_MM}
\end{figure}

\subsubsection{Abnormal Polarity}
Similar to the subsection above, given the market models parameters $\hat{\alpha_i}^P$ and $\hat{\beta_i}^P$ estimated in \eqref{eq_mm}, the abnormal polarity is
\[    AP_{i,t} = P_{i,t} - \hat{\alpha_i}^P - \hat{\beta_i}^P \cdot P_t^M,\]
with $P_t^M$ being the market polarity computed in \eqref{marketpoleq}. Then, we can compute the cumulative abnormal polarity (CAP) and cumulative average abnormal polarity (CAAP) from $\tau - 20$ to $t \le \tau + 20$ as:

\begin{align*}
    CAP_i(\tau,t) &= \sum_{s=- 20}^{t} AP_{i,s+\tau}, \\
    CAAP(t) &= \dfrac{1}{N} \sum_{j=1}^N CAP_{i_j}(\tau_j,t).
\end{align*}

Figure \ref{fig:ES_cum_abn_pol_MM} shows the CAAP around the events identified. The main findings about polarity are twofold. First, conversely to CAAR, CAAP for bullish and bearish events are not constant after the event date, suggesting that users' sentiment about a firm tend to be biased towards recent past events. This can be explained by the fact that users might still post bullish or bearish messages about an event even several days after the event happening, even though the return had already adjusted. Second, and more interestingly, it looks like the CAAP for bullish and bearish events shift several days earlier than the CAAR. This would indicate that the users are on average able to anticipate a future bullish (bearish) event and post positively (negatively) about the firm days before the rise in CAAR. Figure \ref{fig:boxplotcaar} illustrates this striking result with box plots (see \citet{boxplot1} and \citet{boxplot2}) showing the distribution of the CAR and CAP. The line inside a box shows the median while the edges of each box represent the 25\% and 75\% quantile of the distribution. From above the edges of a box, a distance of 1.5 times the interquartile range is measured and a whisker is drawn up to the largest and lowest observed point from the data that falls within this distance.\footnote{Interquartile range is equal to the third quartile minus the first quartile.} The three plots on the left (right) show the boxplots for the CAR (CAP) 5 days before an event, on event date, and 5 days after the event, respectively. To show statistical significance, we use the Mann-Whitney U-test (see \citet{utest} and \citet{sheskin}) on every plot to assess whether the three samples (bullish, neutral and bearish) represent populations with different median values.\footnote{This interpretation only holds under stringent assumptions on the populations, namely that the two population distributions are equal up to a shift.}  Under the null hypothesis, the three samples represent distributions with equal medians.
 Let $\theta_i$ be the median of the distribution i. Formally, we test $H_0 : \theta_{bullish} = \theta_{neutral}$ against $H_1 : \theta_{bullish} > \theta_{neutral}$ and $H_0 : \theta_{neutral} = \theta_{bearish}$ against $H_1 : \theta_{neutral} > \theta_{bearish}$ 5 days before an event, on event date and 5 days after an event. We define U as the Mann-Whitney test statistic, Z as the normal approximation of the Mann-Whitney test statistic for large sample sizes, $n_1$ and $n_2$ as the sample sizes. We refer to \citet{sheskin} for the test statistic computation.
Table \ref{utest_res} shows U-test estimates for pairwise comparisons. The null is rejected in every case except for CAR at $t-5$. 5 days before the event, the boxes of the CAR between bullish, neutral and bearish events are on the same level, suggesting no predictive power of abnormal returns, consistent with the EMH. However, the boxes showing the CAP 5 days before the event are already shifted, meaning that conditionally on an event happening, investors are able on average to anticipate correctly the type of this event. On the event date, the boxes of the CAR shift as the abnormal returns jump for both bullish and bearish events, consistent again with the EMH. Finally, 5 days after the event, the distributions of the CAR are very similar to the distributions on the event dates. Again, this is consistent with the EMH as the abnormal returns adjusted quickly on event date and all information is now embedded in the prices. The distribution of the CAP 5 days after the event continued to shift compared to the event date, as people continue to post about recent past events.
% \TODO{maybe this part of the short story about EMH, which we should provide earlier, at the beginning of the event study part, and then refer to it in the sequel of our text ..}

\begin{figure}[]
    \centering
    \includegraphics[width=0.75\textwidth]{caap.JPG}
    \caption{Cumulative average abnormal polarity around identified events. CAAP related to bearish, neutral and bullish events are displayed with the red, gray and green line, respectively. Areas around lines show confidence intervals at the 95\% level.}
    \label{fig:ES_cum_abn_pol_MM}
\end{figure}

\begin{figure*}[]
        \centering
        \begin{subfigure}[b]{0.42\textwidth}
            \centering
            \includegraphics[width=\textwidth]{box_car_t-5.JPG}
            \caption*{CAR 5 days before the event}
        \end{subfigure}
        \quad
        \begin{subfigure}[b]{0.42\textwidth}  
            \centering 
            \includegraphics[width=\textwidth]{box_cap_t-5.JPG}
            \caption*{CAP 5 days before the event}  
        \end{subfigure}
        \vskip\baselineskip
        \begin{subfigure}[b]{0.42\textwidth}   
            \centering 
            \includegraphics[width=\textwidth]{box_car_t.JPG}
            \caption*{CAR on the event date}   
        \end{subfigure}
        \quad
        \begin{subfigure}[b]{0.42\textwidth}   
            \centering 
            \includegraphics[width=\textwidth]{box_cap_t.JPG}
            \caption*{CAP on the event date}   
        \end{subfigure}
                \vskip\baselineskip
        \begin{subfigure}[b]{0.42\textwidth}   
            \centering 
            \includegraphics[width=\textwidth]{box_car_t+5.JPG}
            \caption*{CAR 5 days after the event}
        \end{subfigure}
        \quad
        \begin{subfigure}[b]{0.42\textwidth}   
            \centering 
            \includegraphics[width=\textwidth]{box_cap_t+5.JPG}
            \caption*{CAP 5 days after the event}  
        \end{subfigure}
        \caption{Distributions of CAP and CAR 5 days before an event, on event date and 5 days after an event. The line inside a box shows the median while the edges of each box represent the 25\% and 75\% quantile of the distribution. From the edges of a box, a distance of 1.5 times the interquartile range is measured and a whisker is drawn up to the largest and lowest observed point from the data that falls within this distance.} 
        \label{fig:boxplotcaar}
    \end{figure*}

% Please add the following required packages to your document preamble:
% \usepackage{multirow}
\begin{table}[]
\centering
\begin{tabular}{|c|c|c|c|c|c|}
\hline
\multicolumn{6}{|c|}{CAR}                                                                                                                                    \\ \hline 
\multicolumn{1}{|l|}{}                     & Alternative Hypothesis                                                         & U     & Z        & $n_1$ & $n_2$ \\ \hline
\multicolumn{1}{|l|}{\multirow{2}{*}{$\tau$-5}} & $H_1$ : $\theta_{bullish} > \theta_{neutral}$ & 61109 & -1.70    & 452  & 292  \\
\multicolumn{1}{|l|}{}                     & $H_1$ : $\theta_{neutral} > \theta_{bearish}$  & 54168 & -0.52    & 292  & 380  \\ \hline
\multirow{2}{*}{$\tau$}                         & $H_1$ : $\theta_{bullish} > \theta_{neutral}$ & 50967 & -5.25*** & 452  & 292  \\
                                           & $H_1$ : $\theta_{neutral} > \theta_{bearish}$  & 43100 & -4.96*** & 292  & 380  \\ \hline
\multirow{2}{*}{$\tau$+5}                       & $H_1$ : $\theta_{bullish} > \theta_{neutral}$ & 52738 & -4.63*** & 452  & 292  \\
                                           & $H_1$ : $\theta_{neutral} > \theta_{bearish}$  & 43239 & -4.91*** & 292  & 380  \\ \hline \hline
\multicolumn{6}{|c|}{CAP}                                                                                                                                    \\ \hline
\multicolumn{1}{|l|}{}                     & Alternative Hypothesis                                                         & U     & Z        & $n_1$ & $n_2$ \\ \hline
\multicolumn{1}{|l|}{\multirow{2}{*}{$\tau$-5}} & $H_1$ : $\theta_{bullish} > \theta_{neutral}$ & 55408 & -3.70*** & 452  & 292  \\
\multicolumn{1}{|l|}{}                     & $H_1$ : $\theta_{neutral} > \theta_{bearish}$  & 47998 & -3.00***   & 292  & 380  \\ \hline
\multirow{2}{*}{$\tau$}                         & $H_1$ : $\theta_{bullish} > \theta_{neutral}$ & 49385 & -8.98*** & 452  & 292  \\
                                           & $H_1$ : $\theta_{neutral} > \theta_{bearish}$  & 42101 & -5.36*** & 292  & 380  \\ \hline
\multirow{2}{*}{$\tau$+5}                       & $H_1$ : $\theta_{bullish} > \theta_{neutral}$ & 44364 & -7.55*** & 452  & 292  \\
                                           & $H_1$ : $\theta_{neutral} > \theta_{bearish}$  & 40515 & -6.00*** & 292  & 380  \\ \hline
\end{tabular}
\caption{Mann-Whitney U-test estimates for pairwise significant differences between distribution medians. Under the null hypothesis, the two samples represent two distributions with equal median values. Statistical significance at the 99\%, 95\%, and 90\% level is indicated with ***, **, *, respectively.}
\label{utest_res}
\end{table}

\section{Portfolios} \label{S:portfolios}

For the construction of a portfolio, in practice, we cannot condition on an event happening over the holding period. But we can repeat an experiment many times by supposing there is an event during the holding period and invest according to this hypothesis (long or short according to CAP). We would hope that repeating this experiment many times, we will be correct a few times and make large gain (or loss).

Our further intuition is that the larger the magnitude of $\mid$CAP$\mid$, the more likely there will be an event in the holding period. We have no statistical evidence for this, but intuitively, it makes sense to assume that the magnitude $\mid$CAP$\mid$ is a  “proxy” for predicting whether an event will happen or not. In that sense, we should better calibrate the thresholds such that the portfolio is zero quite often. Only if we have reason to believe that there will be an event (large magnitude $\mid$CAP$\mid$) then we invest, according to the sign of CAP. Otherwise we stay out of the market.

The data consists of ticker-date instances $(i,t)$, for the universe of tickers $i=1,\dots,19$ and dates $t$ ranging through all business days of the sample period, excluding the first 14 days (for the CAP) and the last day (for the last holding period). Every ticker-date instance $(i,t)$ comes with the feature $CAP_{i,t}=\sum_{s=t-14}^t AP_{i,s}$, which is the running CAP over the last 14 days plus current day $t$ (we do end of day $t$ rebalancing). Note this is different from the definition of ``$CAP_i(\tau,t)$'' above.

   % \item label $L_{it}$, taking 4 categorical values:
   % \[ L_{i,t}=\begin{cases} \text{bullish},&\text{if there is a %bullish event of ticker $i$ during $t+1,\dots,t+5$,}\\ %\text{neutral},&\text{if there is a neutral event of ticker $i$ %during $t+1,\dots,t+5$,}\\ \text{no},&\text{if there is no event %of ticker $i$ during $t+1,\dots,t+5$,}\\ \text{bearish},&\text{if %there is a bearish event of ticker $i$ during %$t+1,\dots,t+5$}\end{cases}\]

\subsection{CAP reset after an event}

As CAP continues to shift after an event, for our portfolio construction, we chose to reset it to 0 after every event. This ensures that we are not exposed to short-term reversals. Let $CAP_{i,t}^{(R)}$ be the feature time-serie of the CAP reset to 0 after every event. Formally, let $\tau_{i,t}\le t$ denote the most recent past event date by $t$ of ticker $i$. Then we define
\[ CAP_{i,t}^{(R)} = \sum_{s=\max\{t-14,\tau_{i,t}+1\}}^t AP_{i,s} = \begin{cases} CAP_{i,t},&\text{if $\tau_{i,t}<t-14$,} \\ \sum_{s= \tau_{i,t}+1}^t AP_{i,s},&\text{if $t-14\le \tau_{i,t}<t$,}\\
0,&\text{if $\tau_{i,t}=t$,}\end{cases}\]
where we used the convention that $\sum_{s=t+1}^t \cdot =0$.

\subsection{Cross-sectional thresholds}

We use time-varying thresholds to build a long and a short portfolio. For every day, we compute the cross-sectional mean and standard deviations of  $CAP_{i,t}^{(R)}$ across the 19 tickers. We will use the mean $\pm$ $x$ standard deviations as and up and down thresholds in the portfolio construction, with $x$ as hyperparameter. We define $U_t(x) = \mu_t + x \cdot \sigma_t$ the upper threshold, and $L_t(x) = \mu_t - x \cdot \sigma_t$ the lower threshold, with $\mu_t$ and $\sigma_t$ being the cross-sectional mean and standard deviations, respectively.  Figure \ref{fig:cross-sec stats} shows the cross-sectional mean and its 99\% confidence interval ($x$ = 2.58). The mean is well centered at zero, which speaks for our method of computing CAP. It also shows that there are essentially two regimes, with a switch in early 2015. In the first regime the standard deviation is much larger (and more volatile) than in the second regime. The Appendix contains the results for 95\% ($x$ = 1.96) and 99.5\% confidence intervals ($x$ = 2.81).

\begin{figure}
    \centering
    \includegraphics[width=0.75\textwidth]{cross_sectional_cap.jpg}
    \caption{Cross-sectional statistics of $CAP^{(R)}$. The blue lines shows the cross-sectional mean of the CAP across the 19 tickers over the year. The green and red lines show the up and down 99\% confidence band. We use them as up and down thresholds for the portfolio construction. }
    \label{fig:cross-sec stats}
\end{figure}

%\subsection{Long/not-long classification}
%
%We want to learn the binary label
%
%\[ L^{long}_{i,t} = \begin{cases} \text{long}, &\text{if %$L_{i,t}=\text{bullish}$,}\\\text{not-long}, &\text{if %$L_{i,t}\in\{\text{neutral},\text{no},\text{bearish}\}$,} \end{cases}
%\]
%Thereto, we choose a threshold $c^{long}$ and predict $\hat %L^{long}_{i,t}=\text{long}$ if $CAP_{i,t}^{(R)}>c^{long}$.
%
%Doing so gives a $2\times 2$ confusion matrix 
%
%\begin{table}[H]
%\centering
%\begin{tabular}{c|c|c|}
%predicted label & \multicolumn{2}{c|}{true label}  \\ \hline
% & long & not-long \\ \hline
%  long &  TP & FP \\\hline
%  not-long & FN & TN
%\end{tabular}
%\end{table}
%
%(this table can be better plotted than here)
%
%Compute precision, recall, F1-score, or the Matthews Correlation Coefficient %MCC (see \url{https://en.wikipedia.org/wiki/Phi_coefficient#Confusion_matrix})
% 
%Then find $c^{long}$ that maximizes F1-score or MCC
%
%
%\subsection{Short/not-short classification}
%
%We want to learn the binary label
%
%\[ L^{short}_{i,t} = \begin{cases} \text{short}, &\text{if %$L_{i,t}=\text{bearish}$,}\\\text{not-short}, &\text{if %$L_{i,t}\in\{\text{bullish},\text{neutral},\text{no}\}$,} \end{cases}
%\]
%Thereto, we choose a threshold $c^{short}$ and predict $\hat %L^{short}_{i,t}=\text{short}$ if $CAP_{i,t}^{(R)}<c^{short}$.
%
%Doing so gives a $2\times 2$ confusion matrix 
%
%\begin{table}[H]
%\centering
%\begin{tabular}{c|c|c|}
%predicted label & \multicolumn{2}{c|}{true label}  \\ \hline
% & short & not-short \\ \hline
%  short &  TP & FP \\\hline
%  not-short & FN & TN
%\end{tabular}
%\end{table}
%
%(this table can be better plotted than here)
%
%Compute precision, recall, F1-score, or the Matthews Correlation Coefficient %MCC (see \url{https://en.wikipedia.org/wiki/Phi_coefficient#Confusion_matrix})
% 
%Then find $c^{short}$ that maximizes F1-score or MCC

\subsection{Portfolio and returns}

We look at two types of portfolios: long only, and short only that we hold for one day.

Denote by $R_{i,t+1}=(S_{i,t+1}-S_{i,t})/S_{i,t} - R_{f,t}$ the excess return of ticker $i$ over $[t,t+1]$.

\paragraph{Long only:}

At the end of any business day $t$, we go long all tickers $i$ with $CAP_{i,t}^{(R)} > U_t(x)$. Denote by $I^{long}_t=\{ i\mid CAP_{i,t}^{(R)} > U_t(x)\}$ the corresponding index set, which could be empty. We then form an equally weighted portfolio and consider the 1-day excess return
\[ R^{long}_{t,t+1} = \begin{cases} \frac{1}{|I^{long}_t|}\sum_{i\in I^{long}_t} R_{i,t,t+1} ,&\text{if $I^{long}_t\neq \emptyset$.} \\
0, &\text{if $I^{long}_t = \emptyset$.} \end{cases}\]

\paragraph{Short only:}

At the end of any business day $t$, we go long all tickers $i$ with $CAP_{i,t}^{(R)} < L_t(x)$. Denote by $I^{short}_t=\{ i\mid CAP_{i,t}^{(R)} < L_t(x)\}$ the corresponding index set, which could be empty. We then form an equally weighted portfolio and consider the 1-day excess return
\[ R^{short}_{t,t+1} = \begin{cases} \frac{1}{|I^{short}_t|}\sum_{i\in I^{short}_t} R_{i,t,t+1} ,&\text{if $I^{short}_t\neq \emptyset$.} \\
0, &\text{if $I^{short}_t = \emptyset$.} \end{cases}\]

Now, we compute these returns for all $t$ and plot a time series.  What we would like to see, of course, is significantly positive excess returns $R^{long}_{t,t+1}$ and negative excess returns $R^{short}_{t,t+1}$.

Figure \ref{fig:cumretpf} shows the cumulative log returns of long and short portfolios as well as the S\&P500. Overall, the strategies are correct in both direction : the long portfolio shows positive returns and the short portfolio shows negative returns. Remarkably, cumulative log returns are most of the time shaped as step functions, suggesting that the strategy is able to take positions before an event, and close the position after the return is earned due to the CAP reset. Figure \ref{fig:pf} shows the number of positions across time of our portfolios. Most of the returns are earned with portfolios consisting of very few tickers. This is a result of our stock picking strategy : we only invest in the top/bottom percentile of CAP, whenever possible.

\begin{figure}
    \centering
    \includegraphics[width=0.75\textwidth]{cumulative_log_return.jpg}
    \caption{Cumulative log returns of both portfolios (x=2.58) and the S\&P500.}
    \label{fig:cumretpf}
\end{figure}

\begin{figure*}[]
        \centering
        %\begin{subfigure}[b]{0.42\textwidth}
        %    \centering
        %    \includegraphics[width=\textwidth]{boxplot.JPG%}
        %    \caption*{Distribution of returns}
        %\end{subfigure}
        %\quad
        %\begin{subfigure}[b]{0.42\textwidth}  
        %    \centering 
        %    \includegraphics[width=\textwidth]{boxplot_zoo%med.JPG}
        %    \caption*{Distribution of returns - zoomed}  
        %\end{subfigure}
        %\vskip\baselineskip
        \begin{subfigure}[b]{0.42\textwidth}   
            \centering 
            \includegraphics[width=\textwidth]{distribution_long_positions.JPG}
            \caption*{Number of positions - Long portfolio}   
        \end{subfigure}
        \quad
        \begin{subfigure}[b]{0.42\textwidth}   
            \centering 
            \includegraphics[width=\textwidth]{distribution_short_positions.JPG}
            \caption*{Number of positions - Short portfolio}   
        \end{subfigure}
                \vskip\baselineskip
        \begin{subfigure}[b]{0.42\textwidth}   
            \centering 
            \includegraphics[width=\textwidth]{positions.JPG}
            \caption*{Number of positions over time}
        \end{subfigure}
        \quad
        \begin{subfigure}[b]{0.42\textwidth}   
            \centering 
            \includegraphics[width=\textwidth]{returns.JPG}
            \caption*{Daily returns}  
        \end{subfigure}
        \caption{Long and Short portfolio for x=2.58.} 
        \label{fig:pf}
    \end{figure*}

%\input{drawdowns}
\newpage
\section{Conclusion} \label{S:conclusion}

We believe that an accurate and timely estimation of investor sentiment on both firms and aggregate market is an excellent proxy of unobservable firm fundamentals. In particular, recent studies on \textit{nowcasting} shows that alternative sources of data can enhance traditional models of stock return and accounting earnings predictions (\citet{challet2013predicting}). In this paper, we scrape 90 million messages out of StockTwits during 2010 to 2020. Messages are either user-labeled as bullish or bearish or left unlabeled. Using the labeled messages as training set, we build a logistic regression on TFIDF vectorized messages to classify the unlabeled messages in either bullish, neutral or bearish class. We observe a 5-for-1 bullish-to-bearish ratio, indicating that investors are on average optimistic. Then, we build daily time-series of polarity for both individual firms and the aggregate market. We show that changes in daily polarity are strongly associated to changes of the same sign in contemporaneous stock returns, but this result loses its significance against next-day returns. However, focused around specific firm events (defined as sudden peak of message volume on a firm), we show that cumulative abnormal polarity has much more predictive power than cumulative abnormal returns. We also note that user's sentiment about a firm tend to be biased towards recent past events. Finally, as robustness check, we show that event studies on CAAR are consistent with previous literature on EMH.

### 2022-05-24 tail/appendix.tex
Add files via upload

\appendix

\chapter{Appendix to Chapter 1}
\section{Relation between $g_{it}(\tau)$ and $\psi_{it}(\tau)$}

The relation between $g_{it}(\tau)$ and $\psi_{it}(\tau)$  is given by :
\begin{align} \label{gphi}
    g_{it}(\tau) &= \frac{F_{it}'(\tau)}{1-F_{it}(\tau)} \nonumber \\
    &= \psi_{it}(\tau) + \psi_{it}'(\tau)\tau.
\end{align}

\newtheorem*{proof1}{Proof}
\begin{proof1}
The last equation above can be computed with the first derivative of $\psi_{it}(\tau)$ with respect to time. Using definition \ref{eq:psi} we have :
\begin{align*}
\psi_{it}'(\tau) &= \dfrac{u' v - u v'}{v^2} = \dfrac{\dfrac{F_{it}'(\tau)}{1-F_{it}(\tau)}\tau + \ln(1-F_{it}(\tau))}{\tau^2} \\
&= \dfrac{\dfrac{F_{it}'(\tau)}{1-F_{it}(\tau)}}{\tau} + \dfrac{\ln(1-F_{it}(\tau))}{\tau^2}\\
\Rightarrow \psi_{it}'(\tau) \tau &= \dfrac{F_{it}'(\tau)}{1-F_{it}(\tau)} + \dfrac{\ln(1-F_{it}(\tau))}{\tau}\\
\Rightarrow \dfrac{F_{it}'(\tau)}{1-F_{it}(\tau)} &= \psi_{it}(\tau) + \psi_{it}'(\tau) \tau.
\end{align*} \qed
\end{proof1}

\section{Computation of $\psi_{it}(\tau) \tau$}

The quantity $\psi_{it}(\tau) \tau$ that we are looking for is :
\begin{equation}
    \psi_{it}(\tau) \tau = \int_0^{\tau}g_{it}(s) ds.
\end{equation}

\newtheorem*{proof2}{Proof}
\begin{proof2}
Integrating by parts $\psi_{it}'(s) s$ between 0 and $\tau$, we have :
\begin{align*}
    \int_0^{\tau}\psi_{it}'(s)\cdot s \cdot ds &= \psi_{it}(s) \cdot s|_{s=0}^{s=\tau} - \int_0^{\tau}\psi_{it}(s) \cdot 1 \cdot ds \\
    &= \psi_{it}(\tau) \tau - \int_0^{\tau}\psi_{it}(s) ds.
\end{align*}
Integrating both sides of equation \ref{gphi} leads to :
\begin{align*}
    \int_0^{\tau} g_{it}(s) ds &= \int_0^{\tau}\psi_{it}(s)ds + \int_0^{\tau}\psi_{it}'(s)\cdot s \cdot ds \\
    &= \int_0^{\tau}\psi_{it}(s)ds + \psi_{it}(\tau) \tau - \int_0^{\tau}\psi_{it}(s) ds \\
    &= \psi_{it}(\tau) \tau.
\end{align*} \qed
\end{proof2}

\section{Likelihood function}
\label{appendix1}

The likelihood function has been developed in \citet{Duan2012}. However, the likelihoods have to be slightly updated to be compatible with the neural network framework. That is, ($\lambda$, $\mu$) is the set of parameters in the neural network for the forward intensity of default $f_{it}$ and one for $h_{it}$ with the combined exit forward intensity being $g_{it} = f_{it} + h_{it}$. I impose non-negativity on both $f_{it}$ and $h_{it}$ to ensure that the combined exit intensity is at least bigger than the forward default intensity. The overall likelihood function for horizon of prediction $\tau$ is by definition given by 
\begin{equation} \label{biglik}
\mathcal{L}_\tau(\lambda,\mu; \tau_C,\tau_D, X) = \prod_{i=1}^N \prod_{t=0}^{T-1} \mathcal{L}_{\tau,i,t}(\lambda,\mu; \tau_{C_i},\tau_{D_i}, X_{it}),
\end{equation}
where
\begin{align}
    \mathcal{L}_{\tau,i,t}(\lambda,\mu; \tau_{C_i},\tau_{D_i}, X_{it}) &= \textbf{1}_{t_{0_i} \leq t, \tau_{C_i} > t+\tau+1} \cdot \mathbb{P}_t(\tau_{C_i} > t + \tau + 1) \\ \nonumber
    &+ \textbf{1}_{t_{0_i} \leq t, \tau_{D_i}=\tau_{C_i} \leq t+\tau} \cdot \mathbb{P}_t(t+\tau < \tau_{D_i} = \tau_{C_i} \leq t + \tau + 1) \\ \nonumber
    &+ \textbf{1}_{t_{0_i} \leq t, \tau_{C_i} \leq t+\tau, \tau_{D_i} \neq \tau_{C_i}} \cdot \mathbb{P}_t(t+\tau < \tau_{C_i} \neq \tau_{C_i} \leq t+\tau + 1) \\ \nonumber
    &+ \textbf{1}_{t_{0_i}>t} + \textbf{1}_{\tau_{C_i} \leq t},
\end{align}
with
\begin{align}
    \mathbb{P}_t(\tau_{C_i} > t + \tau + 1) &= \exp(-\sum_{s=0}^{\tau}g_{it}(s)\Delta t), \\
    \mathbb{P}_t(t+\tau < \tau_{D_i} = \tau_{C_i} \leq t + \tau + 1) &=    
    \begin{cases}
      1-\exp(-f_{it}(0)\Delta t), & \text{if}\ \tau_{C_i} = t+1 \\
      \exp(-\sum\limits_{s=0}^{\tau_{C_i}-t-2}g_{it}(s)\Delta t) \times \\
          \qquad (1-\exp[-f_{it}(\tau_{C_i}-t-1)\Delta t], & \text{otherwise}
    \end{cases} \\
    \mathbb{P}_t(t+\tau < \tau_{C_i} \neq \tau_{C_i} \leq t+\tau + 1)&=
    \begin{cases}
      1-\exp(-g_{it}(0)\Delta t)- \\ 
      \qquad ( 1-\exp(-f_{it}(0)\Delta t)), & \text{if}\ \tau_{C_i} = t+1 \\
      \exp(-\sum\limits_{s=0}^{\tau_{C_i}-t-2}g_{it}(s)\Delta t) \times \\
          \qquad (\exp[-f_{it}(\tau_{C_i}-t-1)\Delta t -  \\ 
          \qquad \exp[-g_{it}(\tau_{C_i}-t-1)\Delta t], & \text{otherwise}
    \end{cases}
\end{align}

Since the indicator functions are all mutually exclusive, taking the log of the likelihood function is very helpful. After the log-linearization, the product terms become summation terms, and the two last indicator functions drop. We are left with the summation of the indicator functions times the logarithm of each probability defined above. Similar to the proposition 2 in \citet{DSW} and subsection 3.2 in \citet{Duan2012}, the pseudo log-likelihood is the product of seperate terms which are function of $f$ and $g$. The first decomposition consists of separating terms involving $f$ and terms involving $g$. The second decomposition consists on separating terms corresponding to different $\tau$. In the end, we get two likelihood functions to estimate for each horizon. For each horizon the parameters of the likelihood function involving $f$ are estimated in a neural network with output $N_{it}^{(\lambda)}$. $g$ is assumed to be of the form $f + h$ as in previous literature. The parameters of the likelihood function involving $h$ are estimated in a neural network with output $N_{it}^{(\mu)}$. Since $\exp(-f)-\exp(-g) = \exp(-f)-\exp(-f-h) = \exp(-f)*(1-\exp(-h))$, the first decomposition is the following :
\begin{align}
    \mathcal{L}_{\tau,i,t}(\lambda; \tau_{C_i},\tau_{D_i}, X_{it}) &= \textbf{1}_{t_{0_i} \leq t, \tau_{C_i} > t+\tau+1} \cdot \exp(-\sum_{s=0}^{\tau} f_{it}(s)\Delta t) \\ \nonumber
    &+ \textbf{1}_{t_{0_i} \leq t, \tau_{D_i}=\tau_{C_i} \leq t+\tau} \cdot \exp(-\sum_{s=0}^{\tau_{C_i}-t-2} f_{it}(s)\Delta t) \cdot (1 - \exp(-f_{it}(\tau_{C_i}-t-1)\Delta t) \\ \nonumber
    &+ \textbf{1}_{t_{0_i} \leq t, \tau_{C_i} \leq t+\tau, \tau_{D_i} \neq \tau_{C_i}} \cdot \exp(-\sum_{s=0}^{\tau_{C_i}-t-2} f_{it}(s) \Delta t) \cdot \exp(-f_{it}(\tau_{C_i}-t-1)\Delta t) \\ \nonumber
    &+ \textbf{1}_{t_{0_i}>t} + \textbf{1}_{\tau_{C_i} \leq t},
\end{align}
\begin{align}
    \mathcal{L}_{\tau,i,t}(\mu; \tau_{C_i},\tau_{D_i}, X_{it}) &= \textbf{1}_{t_{0_i} \leq t, \tau_{C_i} > t+\tau+1} \cdot \exp(-\sum_{s=0}^{\tau} h_{it}(s)\Delta t) \\ \nonumber
    &+ \textbf{1}_{t_{0_i} \leq t, \tau_{D_i}=\tau_{C_i} \leq t+\tau} \cdot \exp(-\sum_{s=0}^{\tau_{C_i}-t-2} h_{it}(s)\Delta t)  \\ \nonumber
    &+ \textbf{1}_{t_{0_i} \leq t, \tau_{C_i} \leq t+\tau, \tau_{D_i} \neq \tau_{C_i}} \cdot \exp(-\sum_{s=0}^{\tau_{C_i}-t-2} h_{it}(s) \Delta t) \cdot (1-\exp(-h_{it}(\tau_{C_i}-t-1)) \\  \nonumber
    &+ \textbf{1}_{t_{0_i}>t} + \textbf{1}_{\tau_{C_i} \leq t}.
\end{align}
Similar to previous literature, we can still decompose the likelihoods into terms involving different horizon of prediction $\tau$. For each $\tau$, we can consider as constant all terms involving previous horizons forward intensities. After log-linearization, the second decomposition is the following :

\begin{align}
    \mathcal{L}_{i,t}(\lambda; \tau_{C_i},\tau_{D_i}, X_{it}) &= \textbf{1}_{t_{0_i} \leq t, \tau_{C_i} > t+\tau+1} \cdot (-f_{it}(s)\Delta t)) \\ \nonumber
    &+ \textbf{1}_{t_{0_i} \leq t, \tau_{D_i}=\tau_{C_i} \leq t+\tau}  \cdot \ln(1 - \exp(-f_{it}(s)\Delta t)) \\ \nonumber
    &+ \textbf{1}_{t_{0_i} \leq t, \tau_{C_i} \leq t+\tau, \tau_{D_i} \neq \tau_{C_i}} \cdot (-f_{it}(s)\Delta t),  \nonumber
\end{align}
\begin{align}
    \mathcal{L}_{i,t}(\mu; \tau_{C_i},\tau_{D_i}, X_{it}) &= \textbf{1}_{t_{0_i} \leq t, \tau_{C_i} > t+\tau+1} \cdot (- h_{it}(s)\Delta t) \\ \nonumber
    &+ \textbf{1}_{t_{0_i} \leq t, \tau_{C_i} \leq t+\tau, \tau_{D_i} \neq \tau_{C_i}} \cdot \ln(1-\exp(-h_{it}(s) \Delta t)). \\  \nonumber
\end{align}
For all horizon of prediction s. We are now left with many small likelihoods that we can maximize separately instead of the huge maximum likelihood \ref{biglik}. Ultimately, we can design two neural networks $NN_{\lambda}$ and $NN_{\mu}$ with outputs $N_{it}^{\lambda}$ and $N_{it}^{\mu}$ respectively to maximize the two above loss functions.

\section{Distance-to-Default estimation}
\label{appendix2}

\subsection{Distance-to-default}

Let $V_T$ be the firm value at time T, $L$ the amount of debt to be repaid and $E_T$ the equity value at time T. In case of bankruptcy debt holders receive money before shareholders in case of bankruptcy. The payoff to shareholders is then given by $E_T = \max(V_T - L,0)$. This payoff is the same as a call option $E_T$ on the underlying $V_T$ with the strike being $L$. \citet{Merton1974} model considers $V_t$ is following a geometric Brownian motion :
\begin{equation*}
dV_t = \mu V_t d_t + \sigma V_t dB_t.
\end{equation*}
We can then use Black-Scholes formula to get the option price and firm's equity value $E_t$ at any time t :
\begin{equation} \label{blackscholes}
E_t = V_t N(d_t) - e^{-r(T-t)}\cdot L \cdot N(d_t - \sigma \sqrt{T-t}),
\end{equation}
\begin{equation*}
d_t = \dfrac{\ln(V_t/L)+(r+\sigma^2/2)(T-t)}{\sigma \sqrt{T-t}},
\end{equation*}
with r being the risk-free rate and N(x) the normal cumulative distribution function.
The distance-to-default is defined as the difference between the expected value of the asset and the default point. After substitution into a normal cumulative distribution function, we get :
\begin{equation}
DtD_t = \dfrac{\ln(\dfrac{V_A}{L})+(\mu-\dfrac{\sigma_A^2}{2})(T-t)}{\sigma_A \sqrt{T-t}}.
\end{equation}
Note that similarly to the variance restriction method, $\mu$ cannot be estimated precisely. Thus, we compute DtD with the following formula :
\begin{equation}
DtD = \dfrac{\ln(\dfrac{V_A}{L})}{\sigma_A \sqrt{T-t}}.
\end{equation}
In the figure \ref{fig:DD} taken from \citet{KMV}, the distance to default is denoted by ``DD". The higher the asset value at horizon H the greater the distance-to-default will get, which lowers the default probability since the firm is more likely to be able to repay the debt owed. Similarly, the more debt the firm has, the smaller the DtD will be for a given level of asset value.

\begin{figure}[h!]
\centering
\includegraphics[width=6in]{DTD_Crosbie_Bohn.JPG} 
\caption{Distance-to-default} 
\label{fig:DD}
\end{figure}

\subsection{Variance restriction method to estimate DtD}
The variance restriction method is used in \citet{DSW} to estimate the distance-to-default (DtD). This method is based on \citet{Merton1974} model which states that firm's equity value can be seen as a call option on the underlying asset and the strike being the amount of debt. This holds because stockholders receive money only once debt holders are fully paid. \\
Applying the Black-Scholes call option formula to equity value, we get the following :
\begin{equation*}
\left\{
  \begin{array}{lll}
    V_E & = & V_A N(d_1) - e^{-r(T-t)} \cdot D \cdot N(d_2)\\
    d_1 & = & \dfrac{\dfrac{V_A}{D}+(r-\dfrac{\sigma_A^2}{2}(T-t)}{\sigma_A \sqrt{T-t}} \\
    d_2 & = & d_1 - \sigma_A \sqrt{T-t}.
  \end{array}
\right.
\end{equation*}
Where $V_E$ is the market equity value, $V_A$ is the market asset value, D is the default point and N the normal cumulative distribution function. Following KMV assumption, the default point D in the variance restriction method is specified as short-term debt plus one half of long-term debt.\\
Using It\^o, we can show that 
\begin{equation*}
\sigma_E = \dfrac{V_A}{V_E}\cdot \frac{\partial V_E}{\partial V_A}\cdot \sigma_A.
\end{equation*}
Hence, the method consists of solving the following system with two equations and two unknowns $V_A$ and $\sigma_A$ :
\begin{equation*}
\left\{
  \begin{array}{lll}
    V_E & = & V_A N(d_1) - e^{-r(T-t)} \cdot D \cdot N(d_2)\\
    \sigma_E & = & \dfrac{V_A}{V_E}\cdot \frac{\partial V_E}{\partial V_A}\cdot \sigma_A,
  \end{array}
\right.
\end{equation*}
Once $V_A$ and $\sigma_A$ are estimated, the DtD is defined as the distance between the expected value of the asset and the default point. After substitution in a normal CDF, we get
\begin{equation}
DtD = \dfrac{\ln(\dfrac{V_A}{L})+(\mu-\dfrac{\sigma_A^2}{2})(T-t)}{\sigma_A \sqrt{T-t}}.
\end{equation}
However, many papers agree to say that $\mu$ is very tedious to estimate. Hence, DtD is often computed as
\begin{equation}
DtD = \dfrac{\ln(\dfrac{V_A}{L})}{\sigma_A \sqrt{T-t}}.
\end{equation}

The major drawback of the variance restriction method used in \citet{DSW} is the definition of the default point. The default point in the variance restriction method is following the so-called KMV assumption. This assumption states that for every firm the default point is exactly equal to short term debt plus one half of long-term debt. However, many financial firms do not account debt as short or long-term debt but as ``other liabilities''. This causes the default point to be abnormally low for financial firms when using KMV assumption. Therefore, to take financial firms into account, we need to adjust the default point by taking into account other liabilities. The method proposed by \citet{Duan2012} employs a maximum likelihood to estimate the optimal fraction $\delta$ of other liabilities to include in the model.

\subsection{Maximum likelihood estimation to estimate DtD}

\citet{Duan2012} and \citet{Duan2012DTD} presented a method to estimate distance-to-defaults without having to exclude financial firms. The method accounts for other liabilities using a maximum likelihood estimation including a parameter $\delta$ to take into account other liabilities. 
The default point in this method becomes :
\begin{equation}
L = \text{short-term debt} + 0.5 \times \text{long-term debt} + \delta \times \text{other liabilities}.
\end{equation}
\citet{Duan2012} and \citet{Duan2012DTD} usually estimate $\delta$ for many firms altogether (i.e. $\delta$ for a whole industry). However, we will improve the methodology by computing $\delta$ for each firm individually. We should obtain higher deltas for financial firms than for non-financial firms. Estimating $\delta$ for each firm is highly time consuming because of greater computation time but it should highly improve the granularity and precision of the model.
The log-likelihood function is given in \citet{Duan2012} and \citet{Duan2012DTD} by :
\begin{align*}
\mathcal{L}_i(\mu,\sigma,\delta) &= -\dfrac{n-1}{2}\ln(2\pi) - \dfrac{1}{2}\sum_{t=2}^n \ln(\sigma^2h_t) - \sum_{t=2}^n \ln(\dfrac{\hat{V}_t(\sigma,\delta)}{A_t}) \\
&-\sum_{t=2}^n \ln(N(\hat{d}_t(\sigma,\delta))) -\sum_{t=2}^n \dfrac{1}{2\sigma^2h_t} \times (\ln(\dfrac{\hat{V}_t(\sigma,\delta)}{\hat{V}_{t-1}(\sigma,\delta)}\cdot \dfrac{A_{t-1}}{A_t}-(\mu - \dfrac{\sigma^2}{2}))^2.
\end{align*}
where n is the number of period observations for each firm $i$. The likelihood above differs from \citet{Duan2012} and \citet{Duan2012DTD} because of the index $i$ since we estimate the likelihood for each of the 2099 firms in our sample to get 2099 estimations of $\delta$. Using this kind of likelihood is complex and very time consuming because we have to solve many inverse Black-Scholes formulas to get the time values of implied asset value $\hat{V}_t(\sigma,\delta)$ for each firm by solving equation \ref{blackscholes}. However, inverse Black-Scholes formula doesn't have any closed form solution. The optimization is even more tedious since the implied asset value $\hat{V}_t(\sigma,\delta)$ depends on the final output of the likelihood $\delta$. $A_t$ is the book asset value and $h_t$ is the ``length of time between two consecutive equity values measured in trading days as a fraction of a year". $h_t$ in the model is set to be 0.25 since we performed a linear interpolation when we had a missing value in the sample (see ``Missing information"). To get rid of the inverse Black-Scholes formula problem, I use a dichotomic algorithm to compute the time series of asset values.\\

\chapter{Appendix to Chapter 2}
\section{Technical proofs}
\subsection{Proof of Lemma \ref{lemma:small}}
From the definition of DIG, we know that for any $R_k\not\in\mathcal{PA}_j$, 
$
 I(R_k\rightarrow R_j||\mathcal{R}_{-\{k,j\}})=0   
$. 
This implies that for all $t$, 
\begin{align}\label{eq:app_l1}
     p(R_{j,t}|\mathcal{R}^{t-1})=p(R_{j,t}|\mathcal{R}_{-\{k\}}^{t-1}).
\end{align}
On the other hand, by the assumption of the Lemma, $R_i\not\in\mathcal{PA}_j$, we have
\begin{align}\label{eq:app_ll2}
 I(R_i\rightarrow R_j||\mathcal{R}_{-\{i,j\}})=0,  
\end{align}
or equivalently, for all $t$,
\begin{align}\label{eq:app_p1}
     p(R_{j,t}|\mathcal{R}^{t-1})=p(R_{j,t}|\mathcal{R}_{-\{i\}}^{t-1}).
\end{align}
 Combining \eqref{eq:app_l1} and \eqref{eq:app_p1} imply that for any pair $\{R_i,R_k\}$ that are not in the parent set of $R_j$, we have
\begin{align}\label{eq:app_l1_2}
   p(R_{j,t}|\mathcal{R}_{-\{k\}}^{t-1})=p(R_{j,t}|\mathcal{R}_{-\{i\}}^{t-1}).
\end{align}
To prove the claim of this lemma, we use \eqref{eq:app_l1_2} to show that all the time series in $\mathcal{R}_{-\{i,j\}}\setminus\mathcal{C}$ can be removed from the conditioning in \eqref{eq:app_ll2}. 
Let $R_k\in \mathcal{R}_{-\{i,j\}}\setminus\mathcal{C}$, by multiplying the above equality with $p(R_{i}^{t-1}|\mathcal{R}_{-\{i,k\}}^{t-1})$ and marginalizing over $R_{i}^{t-1}$, we obtain
\begin{align*}
    &\int p(R_{j,t}|\mathcal{R}_{-\{k\}}^{t-1})p(R_{i}^{t-1}|\mathcal{R}_{-\{i,k\}}^{t-1})dR_i^{t-1}=p(R_{j,t}|\mathcal{R}_{-\{i,k\}}^{t-1})\\
    &=\int p(R_{j,t}|\mathcal{R}_{-\{i\}}^{t-1})p(R_{i}^{t-1}|\mathcal{R}_{-\{i,k\}}^{t-1})dR_i^{t-1}=p(R_{j,t}|\mathcal{R}_{-\{i\}}^{t-1}).
\end{align*}
The above equalities and \eqref{eq:app_p1} imply that for all $t$,
\begin{align*}
p(R_{j,t}|\mathcal{R}^{t-1})=p(R_{j,t}|\mathcal{R}_{-\{i,k\}}^{t-1}),    
\end{align*}
or equivalently,  
\begin{align*}
    I(R_i\rightarrow R_j||\mathcal{R}_{-\{i,j,k\}})=0.
\end{align*}
By repeating the above procedure, we obtain
\begin{align*}
    I(R_i\rightarrow R_j||\mathcal{C})=0.
\end{align*}

\subsection{Proof of Lemma \ref{lemma:lin}}
Consider the VAR model in \eqref{eq:linearsys}. 
First, we assume that $X_i$ has no influence on $X_j$, i.e., $I(X_i\rightarrow X_j||\mathcal{X}_{-\{i,j\}})=0$ or equivalently $a_{j,i}=0$ and show that \eqref{eq:lemma_linear} holds. Given this assumption, we have that for all $t$,
\begin{align*}
    &p(X_{j,t}|\mathcal{X}_{-\{i\}}^{t-1})=p(X_{j,t}|\mathcal{X}^{t-1}).
\end{align*}
Using the equations in \eqref{eq:linearsys} and the assumption that $a_{j,i}=0$, we obtain
\begin{align*}
    &p(X_{j,t}|\mathcal{X}^{t-1})= p(N_{j,t}+\sum_{k}a_{j,k}X_{k,t-1}|\mathcal{X}^{t-1})=p(N_{j,t}+\sum_{k\neq i}a_{j,k}X_{k,t-1}|\mathcal{X}^{t-1})\\ 
    &=p(N_{j,t}+\sum_{k\neq i}a_{j,k}X_{k,t-1}|\sum_{k\neq i}a_{j,k}X_{k,t-1},{X}_{i}^{t-1},{X}_{j}^{t-1})\\
    &=p(N_{j,t}+\sum_{k\neq i}a_{j,k}X_{k,t-1}|\sum_{k\neq i}a_{j,k}X_{k,t-1},{X}_{j}^{t-1}).
\end{align*}
Note that we could replace $\mathcal{X}^{t-1}$ by $\{\sum_{k\neq i}a_{j,k}X_{k,t-1},{X}_{i}^{t-1},{X}_{j}^{t-1}\}$ or\\
$\{\sum_{k\neq i}a_{j,k}X_{k,t-1},{X}_{j}^{t-1}\}$ in the above equations, because given either of them $\sum_{k\neq i}a_{j,k}X_{k,t-1}$ becomes a constant and independent of $N_{j,t}$.
By defining $Q_{t-1}:=\sum_{k\neq i}a_{j,k}X_{k,t-1}$, the above equations can be rewritten as follows 
\begin{align*}
    &p(X_{j,t}|\mathcal{X}^{t-1})=p(X_{j,t}|Q_{t-1},{X}_{i}^{t-1},{X}_{j}^{t-1})=p(X_{j,t}|Q_{t-1},{X}_{j}^{t-1}), \forall t,
\end{align*}
or equivalently, 
\begin{align*}
    &\mathbb{E}\left[\log\frac{p(X_{j,t}|Q_{t-1},{X}_{i}^{t-1},{X}_{j}^{t-1})}{p(X_{j,t}|Q_{t-1},{X}_{j}^{t-1})}\right]=0, \forall t.
\end{align*}
Using the definition of DI, the above equalities can be written in terms of DI as follows
\begin{align*}
   I(X_i\rightarrow X_j||Q)=0.
\end{align*}
On the other hand, we have 
\begin{align*}
&[a_{j,1},...,a_{j,i-1},a_{j,i+1},...,a_{j,m}]=\arg\min_{\textbf{w}\in\mathbb{R}^{m-1}}\mathbb{E}\left[||X_{j,t}- \textbf{w}^T \textbf{X}_{-\{i\},t-1}||_2^2\right]:=\textbf{u}_t.
\end{align*}
where $\textbf{X}_{-\{i\},t-1} :=[X_{1,t-1},..., X_{i-1,t-1}, X_{i+1,t-1},..., X_{m,t-1} ]^T$. This means that $Q_{t-1}=\textbf{u}_t^T\textbf{X}_{-\{i\},t-1}$.

Next, we show the reverse direction, i.e.,  we assume \eqref{eq:lemma_linear} holds, then we show $I(X_i\rightarrow X_j||\mathcal{X}_{-\{i,j\}})=0$. To do so, it suffices to show $a_{j,i}=0$.
Since \eqref{eq:lemma_linear} holds, we have 
\begin{align*}
    {p(X_{j,t}|\textbf{u}_t^T\textbf{X}_{-\{i\},t-1},{X}_{i}^{t-1},{X}_{j}^{t-1})}={p(X_{j,t}|\textbf{u}_t^T\textbf{X}_{-\{i\},t-1},{X}_{j}^{t-1})}, \forall t.
\end{align*}
Using the $j$-th equation of \eqref{eq:linearsys} and the above equalities, for any instances $(\textbf{u}_t^T\textbf{x}_{-\{i\},t-1},{x}_{i}^{t-1},{x}_{j}^{t-1})$ of $(\textbf{u}_t^T\textbf{X}_{-\{i\},t-1},{X}_{i}^{t-1},{X}_{j}^{t-1})$, we obtain $\forall t$,
\begin{align*}
    &\mathbb{E}\left[{X_{j,t}|\textbf{u}_t^T\textbf{x}_{-\{i\},t-1},{x}_{i}^{t-1},{x}_{j}^{t-1}}\right]=\mathbb{E}\left[{X_{j,t}|\textbf{u}_t^T\textbf{x}_{-\{i\},t-1},{x}_{j}^{t-1}}\right],
\end{align*}
which implies
\begin{align*}
    &\mathbb{E}\left[N_{j,t}\right]+\sum_{k\neq i}a_{j,k}x_{k,t-1}+a_{j,i}x_{i,t-1}=\\
    &\mathbb{E}\left[N_{j,t}\right]+\sum_{k\neq i}a_{j,k}x_{k,t-1}
    +a_{j,i}\mathbb{E}\left[X_{i,t-1}|\textbf{u}_t^T\textbf{x}_{-\{i\},t-1},{x}_{j}^{t-1}\right].
\end{align*}
This simplifies to  
\begin{align*}
    a_{j,i}x_{i,t-1}=a_{j,i}\mathbb{E}\left[X_{i,t-1}|\textbf{u}_t^T\textbf{x}_{-\{i\},t-1},{x}_{j}^{t-1}\right], \forall t.
\end{align*}
This equation should hold for any $x_{i,t-1}$. This is only possible if $a_{j,i}=0$.

\subsection{Proof of Lemma \ref{lemma:nonlinear}}
The proof is similar to the linear version and uses the fact that exogenous noises $\{\varepsilon_{j,t}\}$ are independent. More precisely, we have  
\begin{align*}
    &p(X_{j,t}|\mathcal{X}^{t-1})=p(F_{j}(\mathcal{X}^{t-1}) + \varepsilon_{j,t}|F_{j}(\mathcal{X}^{t-1})).
\end{align*}
Since there is no influence from $X_i$ to $X_j$, we can eliminate it from the conditioning and the argument of function $F_j$ and obtain
\begin{align*}
&p\big(F_{j}(\mathcal{X}^{t-1}) + \varepsilon_{j,t}|\mathcal{X}^{t-1}\big)=p\big(F_{j}(\mathcal{X}_{-\{i\}}^{t-1}) + \varepsilon_{j,t}|\mathcal{X}_{-\{i\}}^{t-1}\big).
\end{align*}
On the other hand, because given either $\{F_{j}(\mathcal{X}_{-\{i\}}^{t-1}), X_i^{t-1},X_j^{t-1}\}$ or \\
$\{F_{j}(\mathcal{X}_{-\{i\}}^{t-1}),X_j^{t-1}\}$, the value of $F_{j}(\mathcal{X}_{-\{i\}}^{t-1})$ is no longer a random variable. Using this relationship and the fact that $\varepsilon_{j,t}$ is independent of $\mathcal{X}^{t-1}$, we obtain
\begin{small}
\begin{align*}
    &p\big(F_{j}(\mathcal{X}_{-\{i\}}^{t-1}) + \varepsilon_{j,t}|F_{j}(\mathcal{X}_{-\{i\}}^{t-1}), X_i^{t-1},X_j^{t-1}\big)=
    p\big(F_{j}(\mathcal{X}_{-\{i\}}^{t-1}) + \varepsilon_{j,t}|F_{j}(\mathcal{X}_{-\{i\}}^{t-1}), X_j^{t-1}\big).
\end{align*}
\end{small}
By defining $Q_{t-1}:=F_{j}(\mathcal{X}_{-\{i\}}^{t-1})$, the above equations can be rewritten in terms of DI as follows,
\begin{align*}
    I(X_i\rightarrow X_j||Q)=0.
\end{align*}

To show the reverse, we need to prove that $I(X_i\rightarrow X_j||\mathcal{X}_{-\{i,j\}})=0$ if Equation \eqref{eq:non-q} holds.
Because $I(X_i\rightarrow X_j||Q)=0$ and using Equation \eqref{nonlinear}, for all $t$, we have
\begin{align*}
    &p\big(F_{j}(\mathcal{X}^{t-1}) + \varepsilon_{j,t}|Q_{t-1}, X_i^{t-1},X_j^{t-1}\big)=
    p\big(F_{j}(\mathcal{X}^{t-1}) + \varepsilon_{j,t}|Q_{t-1}, X_j^{t-1}\big),
\end{align*}
where $Q_{t-1}=F_{j}(\mathcal{X}_{-\{i\}}^{t-1})$. Note that the conditioning on the right-hand-side distribution is independent of $X_{i}^{t-1}$. This implies that function $F_j$ does not depend on $X_i$. Therefore, we can remove $X_i^{t-1}$ from the argument of $F_j$, i.e., 
\begin{align*}
    X_{j,t}=F_{j}(\mathcal{X}^{t-1})+\varepsilon_{j,t}=F_{j}(\mathcal{X}_{-\{i\}}^{t-1})+\varepsilon_{j,t},
\end{align*} 
which further implies
\begin{align*}
    &p(F_{j}(\mathcal{X}_{-\{i\}}^{t-1}) + \varepsilon_{j,t}|\mathcal{X}_{-\{i,j\}}^{t-1}, X_i^{t-1},X_j^{t-1})=
    p(F_{j}(\mathcal{X}_{-\{i\}}^{t-1}) + \varepsilon_{j,t}|\mathcal{X}_{-\{i,j\}}^{t-1},X_j^{t-1}). 
\end{align*}
This is equivalent to 
\begin{align*}
    I(X_i\rightarrow X_j||\mathcal{X}_{-\{i,j\}})=0.
\end{align*}

\section{k-Nearest Neighbors estimator of mutual information}
Suppose that $N+M$ i.i.d. realizations $\{\textbf{x}_1, ..., \textbf{x}_{N+M}\}$ are available from $\PP_{X,Y,Z}$, where $\textbf{x}_i$ denotes the $i$-th realization of $(X,Y,Z)$.
The data is randomly divided into two subsets $\mathcal{S}_1$ and $\mathcal{S}_2$ of $N$ and $M$ points, respectively.  
The estimator has two main stages:
In the first stage, a k-nearest density estimator $\widehat{\PP}_{X,Y,Z}$ at the $N$ points of $\mathcal{S}_1$ is estimated using the $M$ realizations of $\mathcal{S}_2$ as follows:

Let $d(\textbf{x},\textbf{y})\in\mathbb{R}$ denote the Euclidean distance between points $\textbf{x}$ and $\textbf{y}$ and $d_k(\textbf{x})\in\mathbb{R}$ denotes the Euclidean distance between a point $\textbf{x}$ and its k-th nearest neighbor among
$\mathcal{S}_2$. The k-nearest region is $\mathcal{S}_k(\textbf{x}) := \{\textbf{y} :d(\textbf{x},\textbf{y})\leq d_k(\textbf{x}) \}$ and the volume of this region is $V_k(\textbf{x}):=\int_{\mathcal{S}_k(\textbf{x})}1dv$. The standard k-nearest density estimator of \cite{sricharan2011k} is defined as 
\begin{small}
\begin{align*}
&\widehat{\PP}_{X,Y,Z}(\textbf{x}):=\frac{k-1}{M\cdot V_k(\textbf{x})}.
\end{align*}
\end{small}
Similarly, we obtain k-nearest density estimators $\widehat{\PP}_{X,Z}, \widehat{\PP}_{Y,Z}$, and $\widehat{\PP}_{Z}$.
Subsequently, the $N$ samples of $\mathcal{S}_1$ is used to approximate the conditional mutual information:
\begin{small}
\begin{align*}
&\widehat{I}(X;Y|Z):=\frac{1}{N}\sum_{i\in \mathcal{S}_1} \log \widehat{\PP}_{X,Y,Z}(\textbf{x}_i) + \log \widehat{\PP}_{Z}(\textbf{x}_i) -\log \widehat{\PP}_{X,Z}(\textbf{x}_i) -\log \widehat{\PP}_{Y,Z}(\textbf{x}_i).
\end{align*}
\end{small}
For more details corresponding this estimator including its bias, variance, and confidence, please see the works by \cite{sricharan2011k} and \cite{loftsgaarden1965nonparametric}.

\section{Koopman-based Lifting Method}
Let $\textbf{X}_t:=\{X_{1,t}, ..., X_{m,t}\}$ denote a network of $m$ time series such that 
\begin{align}\label{dyn}
\dot{\textbf{X}}_t = F(\textbf{X}_t),
\end{align}
where the vector field $F(\textbf{X})=(F_1(\textbf{X}),...,F_m(\textbf{X}))$ is of the form
\begin{align}
F_j(\textbf{X})= \sum_{k=1}^K w_{j,k} h_k(\textbf{X}).
\end{align}
In the above equation, $w_{j,k}\in\mathbb{R}$ are unknown weights and $\{h_k(\textbf{X})\}$ denote a set of known library functions, e.g., monomials. 
Furthermore, let $\varphi^t(\textbf{X}_0)$ denote the solution to \eqref{dyn} associated with the initial condition $\textbf{X}_0$.

Now, suppose that we have $N$ noisy observations $\{(\textbf{x}_1, \textbf{y}_1), ..., (\textbf{x}_N, \textbf{y}_N)\}$ of the system trajectory, where $\textbf{x}_i$ is the initial point and $\textbf{y}_i$ is the final point after $T_s$ steps, i.e., 
\begin{align*}
\textbf{y}_i - \epsilon_i = \varphi^{T_s}(\textbf{x}_i - \varepsilon_i), \ \  i=1,...,N,
\end{align*}
where $\epsilon_i$ and $\varepsilon_i$ are the measurement noises. 
The goal is to estimate the weights $\{w_{j,k}\}$ using these observations and consequently infer the causal network among the time series. To do so, we use the Koopman approach \cite{mauroy2019koopman} that lifts the observation space to another space in which the relationships are linear. 
More precisely, the steps are as follows:
\begin{itemize}
\item Select a set of $M$ basis lifting functions $\{p_1(\textbf{x}), ..., p_M(\textbf{x})\}$, and lift the observations,
\begin{align}
\textbf{P}_x:=\begin{pmatrix}
p_1(\textbf{x}_1) & \cdots & p_M(\textbf{x}_1)\\
p_1(\textbf{x}_2) & \cdots & p_M(\textbf{x}_2)\\
\vdots  &  \ddots & \vdots  \\
p_1(\textbf{x}_N) & \cdots & p_M(\textbf{x}_N)
\end{pmatrix}, 
\textbf{P}_y:=\begin{pmatrix}
p_1(\textbf{y}_1) & \cdots & p_M(\textbf{y}_1)\\
p_1(\textbf{y}_2) & \cdots & p_M(\textbf{y}_2)\\
\vdots  &  \ddots & \vdots  \\
p_1(\textbf{y}_N) & \cdots & p_M(\textbf{y}_N)
\end{pmatrix}\!.
\end{align}

\item Identify the Koopman operator $\textbf{L}:=\frac{1}{T_s}\log(\textbf{P}_x^{\dagger}\textbf{P}_y)$, where $\textbf{P}_x^{\dagger}$ denotes the pseudo-inverse of $\textbf{P}_x$ and the function log denotes the (principal) matrix logarithm.

\item Identify the weights using the following equations: $\widehat{w}_{k,j}:=[\textbf{L}]_{k,l}$, with $l$ such that $p_l(\textbf{x})=x_j$, where $\textbf{x}=(x_1,...,x_m)$.

\end{itemize}

An alternative approach to obtain the weights is the dual lifting method which executes the following steps instead of the above last step.
%Consider the dynamic in \eqref{dyn} in which $F(X)=(F_1(X),...,F_n(X))$, and 
%\begin{align}\label{dy}
%F_j(X) = \sum_{k=1}^K w_{j,k} h_k(X).
%\end{align}
%The goal is to infer the weights $\{w_{j,k}\}$, when the library functions $\{h_k\}$ are assumed to be known.
%In order to present the steps of the Koopman method, we first represent the observational data $X_1,...,X_T$, where $T:=sT_0$ in  the following form. 
%\begin{align}
%D=\begin{pmatrix}
%X_1 & \cdots & X_{T_0}\\
%X_{T_0+1}& \cdots & X_{2T_0}\\
%\vdots  &  \ddots & \vdots  \\
%X_{(s-1)T_0+1} & \cdots & X_{sT_0}
%\end{pmatrix}_{s\times T_0}.
%\end{align}
%The steps of the Koopman method in this case is presented below. 
%\\
%\textbf{Note:} The basis functions $\{g_k\}$ can be monomials or Gaussian ($g_k(x)=\exp(-\gamma||x-x_k||)$). 
%\begin{algorithmic}[H]
%\STATE - Input: $\{(X_1,X_{T_0}), (X_{T_0+1},X_{2T_0}), ..., (X_{(s-1)T_0+1},X_{sT_0})\}$, bases
%\STATE   functions $\{g_1,...,g_M\}$, library functions $\{h_1,...,h_K\}$
%\STATE - Construct the following matrices
%\begin{align*}
%P_x:=\begin{pmatrix}
%g_1(X_1) & \cdots & g_M(X_1)\\
%g_1(X_{T_0+1}) & \cdots & g_M(X_{T_0+1})\\
%\vdots  &  \ddots & \vdots  \\
%g_1(X_{(s-1)T_0+1}) & \cdots & g_M(X_{(s-1)T_0+1})
%\end{pmatrix}_{s\times M},
%\end{align*}
%\begin{align*}
%P_y:=\begin{pmatrix}
%g_1(X_{T_0}) & \cdots & g_M(X_{T_0})\\
%g_1(X_{2T_0}) & \cdots & g_M(X_{2T_0})\\
%\vdots  &  \ddots & \vdots  \\
%g_1(X_{sT_0}) & \cdots & g_M(X_{sT_0})
%\end{pmatrix}_{s\times M}.
%\end{align*}
%\STATE - Compute the Koopman operator 
%$
%\widetilde{L}_{s\times s}:=\frac{1}{T_0}\log\left(P_yP_x^{\dagger}\right)
%$.
At first, it finds matrix $\widehat{\textbf{F}}$ using the following equation, 
\begin{align*}
\widehat{\textbf{F}}_{N\times m}:=
%\begin{pmatrix}
%\widehat{F}(X_1)^T \\
%\widehat{F}(X_{T_0+1})^T\\
%\vdots  \\
%\widehat{F}(X_{(s-1)T_0+1})^T
%\end{pmatrix}_{s\times n} 
\textbf{L}_{N\times N} \begin{pmatrix}
\textbf{x}^T_1 \\
\vdots  \\
\textbf{x}^T_{N} 
\end{pmatrix}_{N\times m}.
\end{align*}
Next, it constructs  
\begin{align*}
\textbf{H}_x:=\begin{pmatrix}
p_1(\textbf{x}_{1}) & \cdots & p_M(\textbf{x}_{1})\\
\vdots  &  \ddots & \vdots  \\
p_1(\textbf{x}_{N}) & \cdots & p_M(\textbf{x}_{N})
\end{pmatrix}_{N\times M},
\end{align*}
 and for each $j$, solve the following regression problem to get the weights
\begin{align*}
\widehat{\textbf{w}}_j :=\arg\min_{\textbf{w}\in\mathbb{R}^M}\left|\left|\textbf{H}_x \textbf{w} - \widehat{\textbf{F}}_{:,j} \right|\right|_2^2 + \rho ||\textbf{w}||_1,
\end{align*}
where $\widehat{\textbf{F}}_{:,j}$ denotes the $j$-th column of matrix $\widehat{\textbf{F}}$  and $\widehat{\textbf{w}}_j =[\hat{w}_{j,1}, ..., \hat{w}_{j,M}]^T$.
%\end{algorithmic}
% The last step of the algorithm is due to Equation \eqref{dy}.
% \begin{align*}
% \widehat{F}_j(X) := \sum_{k=1}^K \hat{w}_j^k h_k(X).
% \end{align*}

\section{Ideal portfolio}
In this section, we show how the ideal portfolio is related to the coefficients of the linear system in \eqref{eq:linearsys}.  
Recall the optimization problem in Lemma \ref{lemma:lin}. 
\begin{align*}
&\textbf{u}_t:=\arg\min_{\textbf{w}\in\mathbb{R}^{m-1}}\mathbb{E}\left[||X_{j,t}- \textbf{w}^T \textbf{X}_{-\{i\},t-1}||_2^2\right],\\
&\textbf{X}_{-\{i\},t-1} :=[X_{1,t-1},\cdots, X_{i-1,t-1}, X_{i+1,t-1}, \cdots, X_{m,t-1} ]^T.
\end{align*}
Consider the $j$-th Equation in \eqref{eq:linearsys}, i.e., 
\begin{align*}
    X_{j,t}=\sum_{k=1}^m a_{j,k}X_{k,t-1}+N_{j,t}.
\end{align*}
If $a_{j,i}=0$, by substituting the above equation into the optimization, we obtain
\begin{align*}
&\min_{\textbf{w}\in\mathbb{R}^{m-1}}\mathbb{E}\left[||\sum_{k\neq i} (a_{j,k}-w_{k})X_{k,t-1}+N_{j,t}||_2^2\right]=\min_{\textbf{w}\in\mathbb{R}^{m-1}}\mathbb{E}\left[||N_{j,t}||_2^2\right]\\
&+\mathbb{E}\left[||\sum_{k\neq i} (a_{j,k}-w_{k})X_{k,t-1}||_2^2\right]+2\mathbb{E}\left[||\left(\sum_{k\neq i} (a_{j,k}-w_{k})X_{k,t-1}\right)N_{j,t}||_2^2\right]\\
&=\min_{\textbf{w}\in\mathbb{R}^{m-1}}\mathbb{E}\left[||\sum_{k\neq i} (a_{j,k}-w_{k})X_{k,t-1}||_2^2\right]+\mathbb{E}\left[||N_{j,t}||_2^2\right]
\end{align*}
The last equality is due to the fact $N_{j,t}$ is independent of $\{X_{k,t-1}\}$ and have zero mean. This implies that the solution is $w_k=a_{j,k}$ for $k\in\{1,...,i-1,i+1,...,m\}$.

\begin{comment}
\section*{Appendix D : Ideal portfolio}
In this section, we present the relation between the ideal portfolio defined in Lemma \ref{lemma:lin} and maximally correlated portfolio. 
We use the following notations and definitions,
\begin{itemize}
\item $r$ denotes the risk-free return (random, but known at beginning of period),
  %\item $X_j$ denotes the excess return of stock $j$ over risk-free $r$,
  \item Candidate excess returns $\textbf{R}=(R_1,\dots,R_{m-1})^T$ over risk-free $r$,
  %(all stocks but $i$ and $j$, e.g., $m=498$)
  \item Covariance matrix of $\begin{pmatrix}
    X \\ \textbf{R}
  \end{pmatrix}$ is denoted by $\mathbf{\Sigma}:=\begin{pmatrix}
    \mathbf{\Sigma}_{XX} & \mathbf{\Sigma}_{X\textbf{R}} \\ \mathbf{\Sigma}_{\textbf{R}X} & \mathbf{\Sigma}_{\textbf{RR}}
  \end{pmatrix}  $. We assume that $\mathbf{\Sigma}_{\textbf{RR}}$ is invertible.

  \item Portfolio with relative weights $\textbf{w} \in \mathbb{R}^{m-1}$ has excess return over $r$ of
\[ Z(\textbf{w}) := \underbrace{\textbf{w}^T  (\textbf{R}+\textbf{1} r)+ (1-\textbf{w}^T \textbf{1}) r}_{\text{portfolio return}}  - r = \textbf{w}^T \textbf{R}.\]
\end{itemize}

Given the above definitions, the correlation of $Z(w)$ with $X$ will be
\begin{align}\label{eqcorr}
  corr(Z(\textbf{w}),X)\!\! &:=\! \dfrac{cov(Z(\textbf{w}),X)}{\sqrt{var(Z(\textbf{w}))var(X)}} = \dfrac{\mathbb{E}[(Z(\textbf{w})-\mathbb{E}[Z(\textbf{w})])^T(X-\mathbb{E}[X])]}{\sqrt{var(Z(\textbf{w}))var(X)}}\\
  &= \dfrac{\textbf{w}^T \mathbf{\Sigma}_{\textbf{R}X}}{\sqrt{\textbf{w}^T\mathbf{\Sigma}_{\textbf{RR}}\textbf{w}}\sqrt{\mathbf{\Sigma}_{XX}}}.
\end{align}
Note that this correlation is invariant under scaling the portfolio weights $w$:
\begin{equation}\label{eqt}
  corr(Z(\textbf{w}),X) = corr(Z(t \textbf{w}),X) \quad \text{for any $t>0$.}
\end{equation}
We can thus assume that $\textbf{w}$ is normalized, $\textbf{w}^T\mathbf{\Sigma}_{\textbf{RR}}\textbf{w}=1$.

Suppose that the goal is to find the portfolio $Z(\textbf{w})$ with normalized weights such that it is maximally correlated with $X$, i.e., 
%\begin{equation}\label{p0}
% \max_{w\in \mathbb{R}^m}  corr(Z(\textbf{w}),X).
% \end{equation}
%In view of \eqref{eqcorr} and \eqref{eqt}, problem \eqref{p0} is equivalent to the constraint optimization problem
\begin{equation}\label{p1}
\begin{aligned}
 \max_{\textbf{w}\in\mathbb{R}^{m-1}}\, &   corr(Z(\textbf{w}),X)   \\
 \text{subject to } &\textbf{w}^T\mathbf{\Sigma}_{\textbf{RR}}\textbf{w}=1,
\end{aligned}
 \end{equation}
 then the solution can be characterize analytically as follows.

\begin{lemma}\label{lemw1}
The problem in \eqref{p1} has a unique solution and it is given by
\begin{align}\label{w1eq}
  {\textbf{w}^*} := \frac{\mathbf{\Sigma}_{\textbf{RR}}^{-1} \mathbf{\Sigma}_{\textbf{R}X}}{\sqrt{\mathbf{\Sigma}_{X\textbf{R}}\mathbf{\Sigma}_{\textbf{RR}}^{-1}\mathbf{\Sigma}_{\textbf{R}X}}}.
\end{align}
\end{lemma}
\begin{proof}
The Lagrangian to \eqref{p1} is
\[ L(\textbf{w},\mu) = \textbf{w}^T \mathbf{\Sigma}_{\textbf{R}X} - \mu( \textbf{w}^T\mathbf{\Sigma}_{\textbf{RR}}\textbf{w}-1)\]
for the Lagrangian multiplier $\mu\in\mathbb{R}$. The first order condition in $\textbf{w}$ reads $2\mu\mathbf{\Sigma}_{\textbf{RR}} \textbf{w}^* = \mathbf{\Sigma}_{\textbf{R}X}$, which yields
\[  \textbf{w}^* = \frac{1}{2\mu} \mathbf{\Sigma}_{\textbf{RR}}^{-1}\mathbf{\Sigma}_{\textbf{R}X}.\]
Choosing $\mu$ such that $\textbf{w}^{*T}\mathbf{\Sigma}_{\textbf{RR}}\textbf{w}^*=1$ gives
\[ 2\mu = \sqrt{\mathbf{\Sigma}_{X\textbf{R}}\mathbf{\Sigma}_{\textbf{RR}}^{-1} \mathbf{\Sigma}_{\textbf{RR}}\mathbf{\Sigma}_{\textbf{RR}}^{-1}\mathbf{\Sigma}_{\textbf{R}X}} =\sqrt{\mathbf{\Sigma}_{X\textbf{R}}\mathbf{\Sigma}_{\textbf{RR}}^{-1}\mathbf{\Sigma}_{\textbf{R}X}},\]
which proves \eqref{w1eq}.
\end{proof}

\paragraph{Relation with Ideal Portfolio:}
Recall the optimization problem in Lemma \ref{lemma:lin}. 
Herein, we rewrite it by substituting $X_{j,t}$ and $\textbf{X}_{-\{i\},t-1}$ by $X$ and $\textbf{R}$, respectively. 
\begin{align*}
&\min_{\textbf{w}\in\mathbb{R}^{m-1}}\mathbb{E}\left[||X- \textbf{w}^T \textbf{R}||_2^2\right].
\end{align*}
Note that this formulation is also related to the projection problem in which the goal is to project $X$ onto the space generated by $\textbf{R}$.
Without loss of generality, we assume that both $X$ and $\textbf{R}$ have zero means. This yields the following solution
\begin{align}\label{w2eq}
    \textbf{w}^{**}:=\mathbf{\Sigma}_{\textbf{RR}}^{-1} \mathbf{\Sigma}_{\textbf{R}X}.
\end{align}
Comparing \eqref{w1eq} and \eqref{w2eq} shows that the only difference is a normalization factor. Therefore, the maximally correlated portfolio contains the same information as the ideal portfolio.

\paragraph{Maximal information in Gaussian case:}
In the case that the joint distribution is Gaussian, the maximally correlated (equivalently, ideal) portfolio $\textbf{w}^{*T} \textbf{R}$ ($\textbf{w}^{**T} \textbf{R}$) captures the same information about $X$ as the entire vector $\textbf{R}$ does, in the following sense.
\begin{lemma}
Assume that $(X,\textbf{R})$ is jointly Gaussian. Then the conditional variances of $X$ given the observation of all excess returns $\textbf{R}$ and of the maximally correlated (ideal) portfolio are the same, i.e., 
\begin{equation}\label{vareqXR}
  var(X\mid \textbf{R} ) = var(X\mid {\textbf{w}^*}^T \textbf{R} )= var(X\mid {\textbf{w}^{**}}^T \textbf{R} ).
\end{equation}
\end{lemma}
\begin{proof}
Gaussian conditioning gives
\[ var(X\mid \textbf{R} ) =\mathbf{\Sigma}_{XX} - \mathbf{\Sigma}_{X\textbf{R}}\mathbf{\Sigma}_{\textbf{RR}}^{-1}\mathbf{\Sigma}_{\textbf{R}X}, \]
and on the other hand
\begin{align*}
    var(X\mid {\textbf{w}^*}^T \textbf{R} ) &= \Sigma_{XX} - \mathbf{\Sigma}_{X\textbf{R}}{\textbf{w}^*}({\textbf{w}^*}^T\mathbf{\Sigma}_{\textbf{RR}}{\textbf{w}^*})^{-1}{\textbf{w}^*}^T\mathbf{\Sigma}_{\textbf{R}X} \\
    &= \Sigma_{XX} - \mathbf{\Sigma}_{X\textbf{R}}{\textbf{w}^*}{\textbf{w}^*}^T\mathbf{\Sigma}_{\textbf{R}X}\\
    &= \Sigma_{XX} - \mathbf{\Sigma}_{X\textbf{R}}\mathbf{\Sigma}_{\textbf{RR}}^{-1}\mathbf{\Sigma}_{\textbf{R}X}.
\end{align*}
Similarly, we have 
\begin{align*}
    &var(X\mid {\textbf{w}^{**}}^T \textbf{R} ) = \Sigma_{XX} - \mathbf{\Sigma}_{X\textbf{R}} {\textbf{w}^{**}}^T({\textbf{w}^*}^T\mathbf{\Sigma}_{\textbf{RR}} {\textbf{w}^{**}}^T)^{-1} {\textbf{w}^{**}}^T\mathbf{\Sigma}_{\textbf{R}X} \\
    &= \Sigma_{XX} - \mathbf{\Sigma}_{X\textbf{R}}\mathbf{\Sigma}_{\textbf{RR}}^{-1}\mathbf{\Sigma}_{\textbf{R}X}\big(\mathbf{\Sigma}_{X\textbf{R}}\mathbf{\Sigma}_{\textbf{RR}}^{-1}\mathbf{\Sigma}_{\textbf{R}X}\big)^{-1}\mathbf{\Sigma}_{X\textbf{R}}\mathbf{\Sigma}_{\textbf{RR}}^{-1}\mathbf{\Sigma}_{\textbf{R}X}\\
    &= \Sigma_{XX} - \mathbf{\Sigma}_{X\textbf{R}}\mathbf{\Sigma}_{\textbf{RR}}^{-1}\mathbf{\Sigma}_{\textbf{R}X}.
\end{align*}
\end{proof}

\begin{remark}
Equality \eqref{vareqXR} is not true in general for non-Gaussian $(X,\textbf{R})$. In particular, using \cite{din_16}, one can check that \eqref{vareqXR} does not hold for multivariate $t$-distributions.
\end{remark}

\end{comment}

\newpage

\chapter{Appendix to Chapter 3}
\section{Number of messages}
As the same information sometimes refer to several companies, users are allowed to identify more than one ticker per message.  Figure \ref{fig:histtic} shows the histogram of the number of tickers tagged per message. As the vast majority of message includes only one ticker, we only show on this plot messages referring to more than one ticker. However, many messages refer to several tickers and this creates duplicates in the database because we consider the same message for all tickers tagged in the message. Figure \ref{fig:uniq} shows the number of messages with and without double counting. In our sample, the number of messages without double counting is 76 million, as opposed to 90 million messages with double counting. Figure \ref{fig:ratio} shows the ratio between the number of messages with double counting divided by the number of messages without double counting. Throughout this paper, we only refer to the number of messages with double counting.

\begin{figure}
    \centering
    \includegraphics[width=0.75\textwidth]{hist_tic.JPG}
    \caption{Histogram of the number of tickers per message, across all messages referring to more than one ticker. The maximum number of tickers per message amounts to 28 and corresponds to 11 messages in the sample.}
    \label{fig:histtic}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.75\textwidth]{uniq.JPG}
    \caption{Total number of messages with double counting (red) compared to the total number of messages without double counting (blue). Numbers are aggregated daily.}
    \label{fig:uniq}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.75\textwidth]{ratio.JPG}
    \caption{Ratio between the number of messages with double counting and the number of messages without double counting. Numbers are aggreaged daily.}
    \label{fig:ratio}
\end{figure}

\newpage

\section{Coverage}

Stocktwits is neither regulated nor moderated, so one needs to filter the information that we use. Even if Stocktwits has valuable information from respected contributors, a blog\footnote{\url{https://www.warriortrading.com/stocktwits-review}} describes the concerns that may rise when using Stocktwits as a financial information provider, namely self-promotion, lack of credibility and other noise. To diversify noise and better extract information, we exclude from our sample tickers that are rarely discussed. Thereto, we compute the median of daily message volume for each ticker and exclude from our sample tickers with a median of less than 50. Decreasing the median threshold increases the coverage at the expense of more noise in the daily polarity. Figure \ref{fig:coverage} shows the coverage as a function of the median threshold. To increase the coverage we need to decrease the threshold a lot (e.g., decreasing the median threshold to 40 from 50 would increase the number of tickers covered to merely 22 from 19). We chose a median threshold of 50 as a balanced trade-off between noise and coverage. Table \ref{tab:listcov} shows the list of tickers covered and associated market capitalization as of 31st of December 2019. To avoid bias in the polarity, we restrict the universe and focus on the most discussed tickers. On top of that, it appears that the most discussed tickers are not always big firms, so it is has the advantage of covering all sizes of firm and we avoid big-firm bias. Also, we do not consider only firms but also ETFs on alternative investments. Finally, we cover several sectors so even if we have a restrictive universe, it is well diversified.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.75\textwidth]{trim.JPG}
    \caption{Coverage as a function of the median threshold. A lower threshold increases the coverage at the expense of a bigger bias in the polarity.}
    \label{fig:coverage}
\end{figure}

\begin{table}[h]
\centering
\begin{tabular}{l|l|c}
Ticker & Name                    & Market capitalization\\ \hline \hline
AAPL   & Apple                   & 1287\\
AMD    & Advanced Micro Devices  & 53 \\
AMRN   & Amarin                  & 7 \\
AMZN   & Amazon                  & 920 \\
BABA   & Alibaba                 & 571 \\
BAC    & Bank of America         & 311 \\
BB     & BlackBerry              & 4 \\
FB     & Facebook                & 585 \\
GLD    & Gold ETF                & 59 \\
IWM    & Small-Cap ETF           & 55 \\
JNUG   & Direxion                & 0.5 \\
MNKD   & MannKind Corporation    & 0.2 \\
NFLX   & Netflix                 & 142 \\
PLUG   & Plug Power              & 1 \\
QQQ    & Nasdaq100 ETF           & 134\\
SPY    & S\&P500 ETF             & 391 \\
TSLA   & Tesla                   & 76 \\
TWTR   & Twitter                 & 25 \\
UVXY   & VIX ETF                 & 0.8 \\
\end{tabular}
\caption{Coverage after the trimming process. List of tickers and corresponding market capitalization as of 31st of December 2019.}
\label{tab:listcov}
\end{table}

%\input{ros.tex}

\newpage
\section{Tutorial for StockTwits messages extraction}
We use CRSP to get stock prices of all US and Canadian listed firms from 1990 to 2020. Out of this dataset, we create the list of unique tickers for which we will extract messages. We will later be able to merge the two datasets using the date and ticker for every observation. We use the StockTwits Application Programming Interface (API) to collect messages from StockTwits. One query on StockTwits API is called a JavaScript Object Notation (JSON) request. Every message on StockTwits has a unique identifier ("msg\_id") posted by a user with a unique identifier ("user\_id"). JSON requests allow to query the database by ticker (called "symbol method") or by user (called "user method"). We use the query by ticker. One query only outputs the latest 30 messages concerning that ticker. However, it is possible to set a parameter ("max") to output the latest 30 messages up to this particular message identifier. This parameter allows us to crawl the message history of a ticker by recursively changing the "max" parameter to the oldest message identifier in the query. To perform a JSON request for Apple (AAPL) up to the message identifier 30'000'000, simply enter the following URL in a browser : https://api.stocktwits.com/api/2/streams/symbol/AAPL.json?\&max=30000000. The page we get looks unreadable but it has always the same structure : several pairs of keys and values. The structure of JSON can easily be interpreted by modern programming languages. We create a Python script to query the API and extract the message history of every ticker in the ticker list. We store the output of every JSON request in .txt files in dedicated ticker folders.

%Good comment to include
%When we think about, what is important in this classification exercise: a risk-averse investor would rather stay out of the market (i.e., realize zero return, no loss) rather than risking a false positive. (leading to a bad investment return, potential loss).
%So the more risk-averse the investor, the larger should be the precision.
%On the other hand, a risky profit seeking investor would regret having missed a good opportunity, i.e., false negative.
%So the risky profit seeking investor prefers a large recall.
\clearpage
\newpage

\section{Portfolio construction - various $x$}
As the thresholds $U_t(x)$ and $L_t(x)$ are functions of the hyperparameter $x$, we provide for robustness check the results of our portfolio construction with $x$ = 1.96  (95\% confidence band) and $x$ = 2.81 (99.5\% confidence band).
\subsection{x = 1.96}

\begin{figure}[h]
    \centering
    \includegraphics[width=0.75\textwidth]{cumulative_log_return1.jpg}
    \caption{Cumulative log returns of both portfolios (x=1.96) and the S\&P500.}
    \label{fig:cumpf1}
\end{figure}

\begin{figure*}[h]
        \centering
       %\begin{subfigure}[b]{0.42\textwidth}
       %    \centering
       %    \includegraphics[width=\textwidth]{boxplot1.jpg}
       %    \caption*{Distribution of returns}
       %\end{subfigure}
       %\quad
       %\begin{subfigure}[b]{0.42\textwidth}  
       %    \centering 
       %    \includegraphics[width=\textwidth]{boxplot_zoomed1.jpg}
       %    \caption*{Distribution of returns - zoomed}  
       %\end{subfigure}
       %\vskip\baselineskip
        \begin{subfigure}[b]{0.42\textwidth}   
            \centering 
            \includegraphics[width=\textwidth]{distribution_long_positions1.jpg}
            \caption*{Number of positions - Long portfolio}   
        \end{subfigure}
        \quad
        \begin{subfigure}[b]{0.42\textwidth}   
            \centering 
            \includegraphics[width=\textwidth]{distribution_short_positions1.jpg}
            \caption*{Number of positions - Short portfolio}   
        \end{subfigure}
                \vskip\baselineskip
        \begin{subfigure}[b]{0.42\textwidth}   
            \centering 
            \includegraphics[width=\textwidth]{positions1.jpg}
            \caption*{Number of positions over time}
        \end{subfigure}
        \quad
        \begin{subfigure}[b]{0.42\textwidth}   
            \centering 
            \includegraphics[width=\textwidth]{returns1.jpg}
            \caption*{Daily returns}  
        \end{subfigure}
        \caption{Long and Short portfolio for x = 1.96.} 
        \label{fig:pf1}
    \end{figure*}

\clearpage
\subsection{x = 2.81}

\begin{figure}[h]
    \centering
    \includegraphics[width=0.75\textwidth]{cumulative_log_return3.jpg}
    \caption{Cumulative log returns of both portfolios (x=2.81) and the S\&P500.}
    \label{fig:cumpf3}
\end{figure}

\begin{figure*}[h]
        \centering
        %\begin{subfigure}[b]{0.42\textwidth}
        %    \centering
        %    \includegraphics[width=\textwidth]{boxplot3.jp%g}
        %    \caption*{Distribution of returns}
        %\end{subfigure}
        %\quad
        %\begin{subfigure}[b]{0.42\textwidth}  
        %    \centering 
        %    \includegraphics[width=\textwidth]{boxplot_zoo%med3.jpg}
        %    \caption*{Distribution of returns - zoomed}  
        %\end{subfigure}
        %\vskip\baselineskip
        \begin{subfigure}[b]{0.42\textwidth}   
            \centering 
            \includegraphics[width=\textwidth]{distribution_long_positions3.jpg}
            \caption*{Number of positions - Long portfolio}   
        \end{subfigure}
        \quad
        \begin{subfigure}[b]{0.42\textwidth}   
            \centering 
            \includegraphics[width=\textwidth]{distribution_short_positions3.jpg}
            \caption*{Number of positions - Short portfolio}   
        \end{subfigure}
                \vskip\baselineskip
        \begin{subfigure}[b]{0.42\textwidth}   
            \centering 
            \includegraphics[width=\textwidth]{positions3.jpg}
            \caption*{Number of positions over time}
        \end{subfigure}
        \quad
        \begin{subfigure}[b]{0.42\textwidth}   
            \centering 
            \includegraphics[width=\textwidth]{returns3.jpg}
            \caption*{Daily returns}  
        \end{subfigure}
        \caption{Long and Short portfolio for x=2.81.} 
        \label{fig:pf3}
    \end{figure*}

### 2022-06-03 head/abstracts.tex
Add files via upload

%\begingroup
%\let\cleardoublepage\clearpage

% English abstract
\cleardoublepage
\chapter*{Abstract}
\markboth{Abstract}{Abstract}
\addcontentsline{toc}{chapter}{Abstract (English/Français)} % adds an entry to the table of contents
% put your text here

This thesis consists of three applications of machine learning techniques to risk management. \\

The first chapter proposes a deep learning approach to estimate physical forward default intensities of companies. Default probabilities are computed using artificial neural networks to estimate the intensities of the inhomogeneous Poisson processes governing default process. The major contribution to previous literature is to allow the estimation of non-linear forward intensities by using neural networks instead of classical maximum likelihood estimation. The model specification allows an easy replication of previous literature using linear assumption and shows the improvement that can be achieved. \\

The second chapter, titled `Causal Networks with Neural Networks` is a co-authored work with Damir Filipović (SFI \& EPFL), Negar Kiyavash (EPFL) and Jalal Etesami (EPFL). We develop a data-driven framework to identify the interconnections between firms using an information-theoretic measure. 
This measure generalizes Granger causality and is capable of detecting nonlinear relationships within a network. 
Moreover, we develop an algorithm using recurrent neural networks and Granger causality to identify the interconnections of high-dimensional nonlinear systems. 
The outcome of this algorithm is the causal graph encoding the interconnections among the firms.
These causal graphs can be used as preliminary feature selection for another predictive model or for systemic risk management.
We evaluate the performance of our algorithm using both synthetic linear and nonlinear experiments and apply it to the daily stock returns of U.S. listed firms and infer their interconnections from 1990 to 2020. \\

The third chapter, titled `StockTwits Classified Sentiment and Stock Returns` is a co-authored work with Damir Filipović (SFI \& EPFL). We scrape 90 million messages from StockTwits over 10 years and classify them into bullish, bearish or neutral classes to create firm-individual sentiment polarity time-series. Polarity is positively associated with contemporaneous stock returns. On average, polarity is not able to predict next-day stock returns but when we focus on specific events (defined as sudden peak of message volume), polarity has predictive power on abnormal returns. \\

\textbf{Keywords:} Risk management, machine learning, neural networks, asset pricing, big data, alternative data.

% French abstract
\begin{otherlanguage}{french}
\cleardoublepage
\chapter*{Résumé}
\markboth{Résumé}{Résumé}
% put your text here

Cette thèse consiste en trois applications de techniques d'apprentissage automatique à la gestion des risques. \\

Le premier chapitre propose une approche d'apprentissage automatique profond pour estimer les probabilités physiques de défaut des entreprises. Les probabilités de défaut sont calculées en utilisant des réseaux de neurones artificiels pour estimer les intensités des processus de Poisson non-homogènes qui gouvernent les processus stochastiques de défaut. La contribution majeure apportée à la littérature existante est de rendre possible l'estimation non-linéaire des intensités en utilisant les réseaux de neurones artificiels au lieu de la classique estimation du maximum de vraisemblance. Les propriétés du modèle autorisent une réplication aisée de la littérature existante (qui utilise l'hypothèse de linéarité de l'intensité) et montre l'amélioration qui peut être obtenue. \\

Le deuxième chapitre, intitulé `Causal Networks wih Neural Networks` est un travail conjoint avec Damir Filipović (SFI \& EPFL), Negar Kiyavash (EPFL) et Jalal Etesami (EPFL). Nous développons un modèle axé sur les données et reposant sur une mesure d'information théorique pour identifier les interconnexions entre les entreprises. Cette mesure utilise la causalité de Granger et est capable de détecter des relations non-linéaires à l'intérieur d'un réseau. De plus, nous développons un algorithme qui utilise les réseaux de neurones récurrents ainsi que la causalité de Granger pour identifier les interconnexions dans les systèmes non-linéaires à haute dimension. Le résultat de cet algorithme est le diagramme causal encodant les interconnexions des entreprises. Ces diagrammes causaux peuvent être utilisés comme modèles préliminaires de sélection de variables d'un autre modèle de prédiction ou pour la gestion de risque systémique. Nous évaluons en premier lieu la performance de notre algorithme en utilisant des expériences synthétiques linéaires et non-linéaires puis nous appliquons notre modèle aux rendements journaliers d'actions américaines cotées pour en déduire leurs interconnexions de 1990 à 2020. \\

Le troisième chapitre, intitulé `StockTwits Classified Sentiment and Stock Returns` est un travail conjoint avec Damir Filipović (SFI \& EPFL). Nous récupérons 90 millions de messages (couvrant 10 ans) provenant de StockTwits et les classifions dans la classe haussière, baissière ou neutre pour créer des séries temporelles de polarité propres à chaque entreprise. La polarité est associée positivement aux rendements d'actions contemporains. En moyenne, la polarité n'est pas capable de prédire les rendements du jour suivant mais lorsque nous nous focalisons sur des évenements spécifiques (définis par une augmentation soudaine du volume de messages), la polarité a de la puissance prédictive sur les rendements anormaux. \\

\textbf{Mots-clés:} Gestion des risques, apprentissage automatique, réseaux de neurones artificiels, évaluation d'actifs, mégadonnées, données alternatives.

\end{otherlanguage}

%\endgroup			
%\vfill

### 2022-06-03 head/acknowledgements.tex
Add files via upload

@@
 
-I wish to express my deepest gratitude to my supervisor, Damir Filipović, for his guidance and continuous support during these five years.
-This thesis greatly benefited from his many valuable comments and flow of smart ideas.
+I wish to express my deepest gratitude to my supervisor, Damir Filipović, for his availability and continuous support during these five years.
+This thesis greatly benefited from his countless valuable comments and flow of smart ideas.
 He guided me and I learned a lot from him about doing research.
@@
 During these years I had the chance to work with extremely nice and intelligent people.
-I am thankful to my colleagues and friends Antoine Didisheim and Coralie Jaunin for the countless chess games and `bridge tours` around the campus.
+I am thankful to my colleagues and friends Antoine Didisheim and Coralie Jaunin for the many chess games and `bridge tours` around the campus.
 
-I would like to thank my family: my parents Jacques and Sonia, my sister Isabelle and her husband Edward and my fiancée Cathy for their unconditional love and for bearing all my endless monologues about my latest algorithms.
+I would like to thank my family: my parents Jacques and Sonia, my sister Isabelle and her husband Edward, my nephew Dylan and my fiancée Cathy for their unconditional love and for bearing all my endless monologues about my latest models.
 I dedicate this thesis to them.

### 2022-06-03 main/ch1_introduction.tex
Add files via upload

\cleardoublepage
\chapter*{Introduction}
\markboth{Introduction}{Introduction}
\addcontentsline{toc}{chapter}{Introduction}

Improved computational power, the rise of Big Data and recent developments in machine learning have created new areas of research in finance. This thesis consists of three machine learning applications to risk management. The first two chapters use neural networks to mitigate default and systemic risk and the last chapter is a financial sentiment analysis using natural language processing.  \\

The first chapter builds on the works from \citet{DSW} and \citet{Duan2012}. Both models use a doubly stochastic argument to derive multi-period default probabilities. In particular, they estimate the intensities of two Poisson processes, one governing default and the other governing other exits. \citet{DSW} generates future random values for the covariates using a VAR process while \citet{Duan2012} relaxes this assumption and use forward intensities. The latter specifies the intensities as a linear function of state variables and uses maximum likelihood to estimate the parameters. The first chapter of this thesis extends existing literature by removing the assumption of linear intensities and uses artificial neural networks to estimate the intensities of the Poisson processes. Neural networks are well-suited in this framework because they allow an easy replication of the linear formulation in \citet{Duan2012}. Increasing the network's width and depth allows for non-linearities and out-of-sample Lorenz Curves show that the neural network's approach outperforms the linear assumption for every horizon. Finally, I show what are the most important predictors of default in the short and longer term. Interdepencies in a network are at the heart of systemic risk. The second chapter - connected to the first one as another application of neural networks to risk management - employs Granger causality to identify interconnections among a set of institutions. We build an information measure known as directed information (DI) capable of capturing causal relationships in both linear and non-linear systems. The output of this approach is a directed graph that visualizes the interconnections among a set of time series. Computing DI has both computational and sample complexity which makes it not suitable for inferring the causal structure of large networks. To overcome this problem, we develop a novel approach based on recurrent neural networks that reduces complexity of evaluating DI in high-dimensional settings. We show that our approach performs well both in a linear and non-linear simulated enviornments, then apply it to infer the causal relationships among US firms from 1990 to 2020. \\

The last chapter uses natural language processing to assess the predictive power of social media on stock returns.We scrape 90 millions messages from Stocktwits - a microblogging platform similar to Twitter but designed for finance professionals. One of the challenge in this context is to create a classifier that understands the vocabulary of the message posted by the users. After preprocessing steps, we use TFIDF vectorization to compute the importance of each word in the message. This transforms the messages from words to a vector of numbers which is now in the same dimension as the vocabulary. Then, we build two adversarial logistic regressions using the TFIDF vectors as features and the user-labels as targets. The first model predicts positive or not positive, the second model predicts negative or not negative. When the models agree, we get our label and when they disagree, we treat the tweet as neutral. This procedure allows us to create an artificial neutral class that absorbs all the tweets that do not convey financial information. Finally, with daily intervals, we aggregate tweets predicted sentiments per ticker to compute daily polarity time-series. We then use the daily volume of messages on a given firm to identify sudden peak of activity, indicating a firm event. Computing cumulative average abnormal return and cumulative abnormal polarity in a 41 days window centered at the identified event, we show that abnormal polarities have significant predictive power on the type of event.

### 2022-06-03 main/paper1.tex
Add files via upload

@@
 \end{tikzpicture}
-\caption{Example of the lifespan of a firm.
-Each dot corresponds to a time period where the econometrician gathers firm-specific (if available) and macroeconomic variables.
+\mycaption{Example of the lifespan of a firm}{Each dot corresponds to a time period where the econometrician gathers firm-specific (if available) and macroeconomic variables.
 $t^0_i$ is the entry time of the firm i and $\tau_{D_i}$ denotes the default time of the firm i.}
@@
 \end{tikzpicture}
-    \caption{Illustration of a neural network [5, 3].
-In this network, there are three input features, five neurons in the first hidden layer and 3 neurons in the second hidden layer.
+    \mycaption{Illustration of a neural network \text{[}5, 3\text{]}}{In this network, there are three input features, five neurons in the first hidden layer and 3 neurons in the second hidden layer.
 Each neuron is activated with a sigmoid function.}
@@
     \includegraphics[width=\textwidth]{def_surv_oe_plot.jpg}
-    \caption{Number of firms defaulted, exited for other reasons, surviving each year}
+    \mycaption{Number of firms defaulted, exited for other reasons, surviving each year}{}
     \label{fig:defsurv}
@@
     \end{tabular}
-    \caption{Number of firms in each category for each horizon of prediction $\tau$ which corresponds to the firm-month observations used in the likelihood for horizon $\tau$}
+    \mycaption{Number of firms in each category for each horizon of prediction $\tau$}{These correspond to the firm-month observations used in the likelihood for horizon $\tau$.}
     \label{tab:observ}
@@
     \end{tabular}
-    \caption{Mean of variables for surviving firms, defaulted firms and other exits. $\Delta$ are one period lagged differences.}
+    \mycaption{Mean of variables for surviving firms, defaulted firms and other exits}{$\Delta$ are one period lagged differences.}
     \label{tab:meanvar}
@@
     \includegraphics[scale=0.6]{corrmatrix.jpeg}
-    \caption{Correlation matrix for firm-specific and macroeconomic covariates}
+    \mycaption{Correlation matrix for firm-specific and macroeconomic covariates}{}
     \label{fig:corrmat}
@@
     \includegraphics[width=\textwidth]{duan_vs_NN.jpeg}
-    \caption{Gini coefficient : Linear vs Neural Networks}
+    \mycaption{Gini coefficient : Linear vs Neural Networks}{}
     \label{fig:ginicomp}
@@
         \end{subfigure}
-        \caption{Out-of-sample Lorenz Curves for each horizon. The orange line shows the performance of Duan's model and the red line shows the performance of the [5,3] neural network.} 
+        \mycaption{Out-of-sample Lorenz Curves for each horizon}{The orange line shows the performance of Duan's model and the red line shows the performance of the [5,3] neural network.} 
         \label{fig:perf}
@@
     \end{tabular}
-    \caption{Gini coefficients}
+    \mycaption{Gini coefficients}{}
     \label{tab:ginicoeff}
@@
     \includegraphics[width=\textwidth]{lor0duan.png}
-    \caption{Comparison with the benchmark model \citet{Duan2012}.
-The green line shows the performance of the [1] neural network (i.e a single hidden layer with a single neuron) with an exponential activation function.
+    \mycaption{Comparison with the benchmark model \citet{Duan2012}}{The green line shows the performance of the [1] neural network (i.e a single hidden layer with a single neuron) with an exponential activation function.
 The orange line shows the performance of Duan's model and the red line shows the performance of the [5,3] neural network.}
@@
     \includegraphics[width=\textwidth]{compgraphr.png}
-    \caption{Trained [5, 3] Neural Network for forward default intensity at horizon 0.
-Negative weights are drawn as blue lines connecting neurons where orange lines show positive weights.
+    \mycaption{Trained \text{[}5, 3\text{]} Neural Network for forward default intensity at horizon 0}{Negative weights are drawn as blue lines connecting neurons where orange lines show positive weights.
 The thicker the line, the higher the weight is in absolute value.}
@@
         \end{subfigure}
-        \caption{Sensitivities for each horizon.} 
+        \mycaption{Sensitivities for each horizon}{}
         \label{fig:sensi}

### 2022-06-03 main/paper2.tex
Add files via upload

@@
 }
-   \caption{DIG of the system in \eqref{model1}.}\label{fig:example_dig}
+   \mycaption{DIG of the system in \eqref{model1}}{}
+\label{fig:example_dig}
 \end{figure}
@@
 \includegraphics[scale=.4]{koopman.pdf}
-\caption{Koopman lifting technique compared to classical non-linear identification.}\label{fig:koopman}
+\mycaption{Koopman lifting technique compared to classical non-linear identification}{}
+\label{fig:koopman}
 \end{figure}
@@
   }
- \caption{Infer-DIG}\label{alg}
+ \mycaption{Infer-DIG}{}
+ \label{alg}
 \end{algorithm}
@@
     \includegraphics[scale=.2]{fig_matrix.pdf}
-    \caption{Structure of matrix $\textbf{A}$ in \eqref{eq:ex_lin} that is build using \eqref{eq:rules}.}
+    \mycaption{Structure of matrix $\textbf{A}$ in \eqref{eq:ex_lin} that is build using \eqref{eq:rules}.}{}
     \label{fig:matrix}
@@
         \end{subfigure}
-        \caption{Precision and recall curves in the linear framework for the \textit{True parents}, \textit{Most correlated}, \textit{Ideal portfolio}, and the \textit{RNN}, respectively.} 
+        \mycaption{Precision and recall curves in the linear framework.}{Precision and recall curves for the \textit{True parents}, \textit{Most correlated}, \textit{Ideal portfolio}, and the \textit{RNN}, respectively.}
         \label{perflin}
@@
 \includegraphics[width=\textwidth]{precrecall_15assets.jpg}
-\caption{Precision-Recall curves for the quadratic model.}\label{fig:perfnonlin}
+\mycaption{Precision-Recall curves for the quadratic model.}{}
+\label{fig:perfnonlin}
 \end{figure}
@@
 \end{tabular}
-\caption{Degree of Granger Causality (DGC) for each sub-graph. DGC is defined as the fraction of relationships in the network among all potential relationships.}
+\mycaption{Degree of Granger Causality (DGC) for each sub-graph.}{DGC is defined as the fraction of relationships in the network among all potential relationships.}
 \label{GrangerCausality}
@@
         \end{subfigure}
-        \caption{Empirical DIG for the periods 1990-1994 (top) and 1995-1999 (bottom). Interactive graphs can be found at https://marcaureledivernois.github.io/firm-network/} 
+        \mycaption{Empirical DIG for the periods 1990-1994 and 1995-1999.}{Empirical DIG for the periods 1990-1994 (top) and 1995-1999 (bottom).
+Interactive graphs can be found at https://marcaureledivernois.github.io/firm-network/}
         \label{empdig1}
@@
         \end{subfigure}
-        \caption{Empirical DIG for the periods 2000-2004 (top) and 2005-2009 (bottom). Interactive graphs can be found at https://marcaureledivernois.github.io/firm-network/} 
+        \mycaption{Empirical DIG for the periods 2000-2004 and 2005-2009.}{Empirical DIG for the periods 2000-2004 (top) and 2005-2009 (bottom).
+Interactive graphs can be found at https://marcaureledivernois.github.io/firm-network/}
         \label{empdig2}
@@
         \end{subfigure}
-        \caption{Empirical DIG for the periods 2010-2014 (top) and 2015-2019 (bottom). Interactive graphs can be found at https://marcaureledivernois.github.io/firm-network/} 
+        \mycaption{Empirical DIG for the periods 2010-2014 and 2015-2019.}{Empirical DIG for the periods 2010-2014 (top) and 2015-2019 (bottom).
+Interactive graphs can be found at https://marcaureledivernois.github.io/firm-network/}
         \label{empdig3}
@@
 \end{tabular}}
-\caption{Outdegrees (Out) ranked for each sub-graph. Outdegree is defined as the number of edges going out of a specific node.}
+\mycaption{Outdegrees ranked for each sub-graph.}{Outdegree (Out) is defined as the number of edges going out of a specific node.}
 \label{outdegree}
@@
 \end{tabular}}
-\caption{Indegrees (In) ranked for each sub-graph. Indegree is defined as the number of edges going to a specific node.}
+\mycaption{Indegrees ranked for each sub-graph.}{Indegree (In) is defined as the number of edges going to a specific node.}
 \label{indegree}

### 2022-06-03 main/paper3.tex
Add files via upload

@@
     \includegraphics[width=0.75\textwidth]{tweet_vol_daily.JPG}
-    \caption{Number of messages posted daily on StockTwits.}
+    \mycaption{Number of messages posted daily on StockTwits.}{}
     \label{fig:tweet_vol_daily}
@@
 \includegraphics[width=0.75\textwidth]{stocktwits_screeshot_3march2020.JPG}
-\caption{Screenshot of StockTwits as of 3rd March 2020.
-The first message is labeled as bullish by the user ``satkaru", the two next are unlabeled messages that will be classified using a machine learning classifier and the last message is labeled as bearish by the user ``Etrading''.}
+\mycaption{Screenshot of StockTwits as of 3rd March 2020.}{The first message is labeled as bullish by the user ``satkaru", the two next are unlabeled messages that will be classified using a machine learning classifier and the last message is labeled as bearish by the user ``Etrading''.}
 \label{screenshot}
@@
         \end{subfigure}
-        \caption{Number of user-labeled messages in each category. There are around 60 million unlabeled messages that we classify using the 30 million user-labeled bullish and bearish messages.} 
+        \mycaption{Number of user-labeled messages in each category.}{There are around 60 million unlabeled messages that we classify using the 30 million user-labeled bullish and bearish messages.} 
         \label{categrepartition-preclassif}
@@
 
-        \caption{User summary statistics.
-Top graph is a log-log histogram of the number of followers per user and the bottom graph shows the log-log histogram of the number of messages posted by users.}
+        \mycaption{User summary statistics.}{Top graph is a log-log histogram of the number of followers per user and the bottom graph shows the log-log histogram of the number of messages posted by users.}
         \label{fig:data-stocktwits1}
@@
         \end{subfigure}
-        \caption{Firm summary statistics.
-Top graph shows the top 30 most discussed tickers on the platform.
+        \mycaption{Firm summary statistics.}{Top graph shows the top 30 most discussed tickers on the platform.
 SPY is the ticker of the S\&P and AAPL is the ticker for Apple.
@@
    \vspace{0.1cm}
-    \caption{Bullish word cloud (top), bearish word cloud (bottom).
+    \mycaption{Bullish and bearish word clouds.}{Bullish word cloud (top), bearish word cloud (bottom).
 These correspond to the most frequent terms (up to 3-grams) in user-labeled bullish (bearish) messages relative to their total appearance.
@@
 \end{tabular*}
-\caption{Preprocessing of five sample messages.
-Preprocessing operations include: punctuation removal, lower casing, apostrophe handling, contraction form handling (i.e.
+\mycaption{Preprocessing of five sample messages.}{Preprocessing operations include: punctuation removal, lower casing, apostrophe handling, contraction form handling (i.e.
 ``won't" becomes ``will not"), tickers removal, users removal, URLs removal, parsing and a simple spell corrector dealing with more than two repeated characters (i.e.
@@
     \includegraphics[width=0.75\textwidth]{preclassif_acrosstime.JPG}
-    \caption{Proportions of user-labeled messages in each category: bullish (green), bearish (red), and unlabeled (gray). Proportions are aggregated monthly.}
+    \mycaption{Proportions of user-labeled messages in each category}{Proportions of user-labeled messages in each category: bullish (green), bearish (red), and unlabeled (gray).
+Proportions are aggregated monthly.}
     \label{fig:ts_neutral_vs_uncl}
@@
     \includegraphics[width=0.75\textwidth]{optimal thresholds.JPG}
-    \caption{Optimal classification thresholds.
-The green (red) line is the F1 score for the bullish versus non-bullish (bearish versus non-bearish) classification, respectively.
+    \mycaption{Optimal classification thresholds.}{The green (red) line is the F1 score for the bullish versus non-bullish (bearish versus non-bearish) classification, respectively.
 Circles indicate the maximal F1 scores, along with the corresponding classification score thresholds.}
@@
     \includegraphics[width=0.75\textwidth]{postclassif_acrosstime3.JPG}
-    \caption{Proportions of classified messages in each category: bullish (light green predicted, green user-labeled), bearish (light red predicted, red user-labeled), and neutral (gray).
+    \mycaption{Proportions of classified messages in each category}{Bullish (light green predicted, green user-labeled), bearish (light red predicted, red user-labeled), and neutral (gray).
 Proportions are aggregated monthly.}
@@
     \includegraphics[width=0.75\textwidth]{spy_vs_marketpol.JPG}
-    \caption{Market polarity on the y-axis versus the polarity of the SPY on the x-axis. The red line shows the linear regression line and coefficients.}
+    \mycaption{Market polarity versus the SPY polarity.}{Market polarity on the y-axis versus the polarity of the SPY on the x-axis. The red line shows the linear regression line and coefficients.}
     \label{fig:marketpolspy}
@@
 \end{tabular}
-\caption{Results from linear regressions of contemporaneous and next period stock returns on polarity.
+\mycaption{Linear regressions }{Results from linear regressions of contemporaneous and next period stock returns on polarity.
 Stock returns are trimmed at the 5\% percentile on both sides.
@@
         \end{subfigure}
-        \caption{Time series of daily polarity (red - left axis) and daily stock returns (blue - right axis) since 1st of January 2019 for the top 6 most discussed tickers.
+        \mycaption{Time series of daily polarity}{Time series of daily polarity  (red - left axis) and daily stock returns (blue - right axis) since 1st of January 2019 for the top 6 most discussed tickers.
 Pearson correlation between the two time series is shown in the title.}
@@
     \includegraphics[width=0.75\textwidth]{timeline.JPG}
-    \caption{Timeline of our event studies.
-$\tau$ is an event date, the event window has 41 business days centered around the event date and the estimation window is a one-year rolling period prior the event.}
+    \mycaption{Timeline of our event studies.}{$\tau$ is an event date, the event window has 41 business days centered around the event date and the estimation window is a one-year rolling period prior the event.}
     \label{fig:timeline}
@@
     \includegraphics[width=0.75\textwidth]{activity-volume.jpg}
-    \caption{Changes in weekly volume of transactions on the y-axis versus changes in message activity on the x-axis. Activity is measured in weekly messages posted per firm.}
+    \mycaption{Volume of transactions and message activity.}{Changes in weekly volume of transactions on the y-axis versus changes in message activity on the x-axis.
+Activity is measured in weekly messages posted per firm.}
     \label{fig:actvol}
@@
     \includegraphics[width=0.75\textwidth]{abnpol_distr_all.JPG}
-    \caption{Empirical distribution of abnormal polarities. Red dashed lines show the one-third and two-third percentiles.}
+    \mycaption{Empirical distribution of abnormal polarities.}{Red dashed lines show the one-third and two-third percentiles.}
     \label{fig:abnpol_all}
@@
     \includegraphics[width=0.75\textwidth]{activity_aapl.JPG}
-    \caption{Daily message volume for Apple.
-Events are days with an unusual high number of messages.
+    \mycaption{Daily message volume for Apple.}{Events are days with an unusual high number of messages.
 Green upper-triangles show bullish events, gray circles are neutral events and red down-triangles represent bearish events.}
@@
     \includegraphics[width=0.75\textwidth]{caar.JPG}
-    \caption{Cumulative average abnormal returns around identified events.
-CAAR related to bearish, neutral and bullish events are displayed with the red, gray and green line, respectively.
+    \mycaption{Cumulative average abnormal returns around identified events.}{CAAR related to bearish, neutral and bullish events are displayed with the red, gray and green line, respectively.
 Areas around lines show confidence intervals at the 95\% level.}
@@
     \includegraphics[width=0.75\textwidth]{caap.JPG}
-    \caption{Cumulative average abnormal polarity around identified events.
-CAAP related to bearish, neutral and bullish events are displayed with the red, gray and green line, respectively.
+    \mycaption{Cumulative average abnormal polarity around identified events.}{CAAP related to bearish, neutral and bullish events are displayed with the red, gray and green line, respectively.
 Areas around lines show confidence intervals at the 95\% level.}
@@
         \end{subfigure}
-        \caption{Distributions of CAP and CAR 5 days before an event, on event date and 5 days after an event.
+        \mycaption{Distributions of CAP and CAR}{Distributions of CAP and CAR 5 days before an event, on event date and 5 days after an event.
 The line inside a box shows the median while the edges of each box represent the 25\% and 75\% quantile of the distribution.
@@
 \end{tabular}
-\caption{Mann-Whitney U-test estimates for pairwise significant differences between distribution medians.
+\mycaption{Mann-Whitney U-test estimates}{Mann-Whitney U-test estimates for pairwise significant differences between distribution medians.
 Under the null hypothesis, the two samples represent two distributions with equal median values.
@@
     \includegraphics[width=0.75\textwidth]{cross_sectional_cap.jpg}
-    \caption{Cross-sectional statistics of $CAP^{(R)}$.
-The blue lines shows the cross-sectional mean of the CAP across the 19 tickers over the year.
+    \mycaption{Cross-sectional statistics of $CAP^{(R)}$.}{The blue lines shows the cross-sectional mean of the CAP across the 19 tickers over the year.
 The green and red lines show the up and down 99\% confidence band.
@@
     \includegraphics[width=0.75\textwidth]{cumulative_log_return.jpg}
-    \caption{Cumulative log returns of both portfolios (x=2.58) and the S\&P500.}
+    \mycaption{Cumulative log returns of both portfolios (x=2.58) and the S\&P500.}{}
     \label{fig:cumretpf}
@@
         \end{subfigure}
-        \caption{Long and Short portfolio for x=2.58.} 
+        \mycaption{Long and Short portfolios for x=2.58.}{}
         \label{fig:pf}

### 2022-06-03 tail/appendix.tex
Add files via upload

@@
 
-\chapter{Appendix to Chapter 1}
-\section{Relation between $g_{it}(\tau)$ and $\psi_{it}(\tau)$}
+\chapter{Appendix}
+
+\section{Appendix to Chapter 1}
+\subsection{Relation between $g_{it}(\tau)$ and $\psi_{it}(\tau)$}
 
@@
 
-\section{Computation of $\psi_{it}(\tau) \tau$}
+\subsection{Computation of $\psi_{it}(\tau) \tau$}
 
@@
 
-\section{Likelihood function}
+\subsection{Likelihood function}
 \label{appendix1}
@@
 
-\section{Distance-to-Default estimation}
+\subsection{Distance-to-Default estimation}
 \label{appendix2}
 
-\subsection{Distance-to-default}
+\subsubsection{Distance-to-default}
 
@@
 \includegraphics[width=6in]{DTD_Crosbie_Bohn.JPG} 
-\caption{Distance-to-default} 
+\mycaption{Distance-to-default}{} 
 \label{fig:DD}
@@
 
-\subsection{Variance restriction method to estimate DtD}
+\subsubsection{Variance restriction method to estimate DtD}
 The variance restriction method is used in \citet{DSW} to estimate the distance-to-default (DtD).
@@
 
-\subsection{Maximum likelihood estimation to estimate DtD}
+\subsubsection{Maximum likelihood estimation to estimate DtD}
 
@@
 
-\chapter{Appendix to Chapter 2}
-\section{Technical proofs}
-\subsection{Proof of Lemma \ref{lemma:small}}
+\section{Appendix to Chapter 2}
+\subsection{Technical proofs}
+\subsubsection{Proof of Lemma \ref{lemma:small}}
 From the definition of DIG, we know that for any $R_k\not\in\mathcal{PA}_j$, 
@@
 
-\subsection{Proof of Lemma \ref{lemma:lin}}
+\subsubsection{Proof of Lemma \ref{lemma:lin}}
 Consider the VAR model in \eqref{eq:linearsys}. 
@@
 
-\subsection{Proof of Lemma \ref{lemma:nonlinear}}
+\subsubsection{Proof of Lemma \ref{lemma:nonlinear}}
 The proof is similar to the linear version and uses the fact that exogenous noises $\{\varepsilon_{j,t}\}$ are independent. More precisely, we have  
@@
 
-\section{k-Nearest Neighbors estimator of mutual information}
+\subsection{k-Nearest Neighbors estimator of mutual information}
 Suppose that $N+M$ i.i.d. realizations $\{\textbf{x}_1, ..., \textbf{x}_{N+M}\}$ are available from $\PP_{X,Y,Z}$, where $\textbf{x}_i$ denotes the $i$-th realization of $(X,Y,Z)$.
@@
 
-\section{Koopman-based Lifting Method}
+\subsection{Koopman-based Lifting Method}
 Let $\textbf{X}_t:=\{X_{1,t}, ..., X_{m,t}\}$ denote a network of $m$ time series such that 
@@
 
-\section{Ideal portfolio}
+\subsection{Ideal portfolio}
 In this section, we show how the ideal portfolio is related to the coefficients of the linear system in \eqref{eq:linearsys}.  
@@
 \begin{comment}
-\section*{Appendix D : Ideal portfolio}
+\subsection*{Appendix D : Ideal portfolio}
 In this section, we present the relation between the ideal portfolio defined in Lemma \ref{lemma:lin} and maximally correlated portfolio. 
@@
 
-\chapter{Appendix to Chapter 3}
-\section{Number of messages}
+\section{Appendix to Chapter 3}
+\subsection{Number of messages}
 As the same information sometimes refer to several companies, users are allowed to identify more than one ticker per message.
@@
     \includegraphics[width=0.75\textwidth]{hist_tic.JPG}
-    \caption{Histogram of the number of tickers per message, across all messages referring to more than one ticker.
+    \mycaption{Number of tickers per message}{Histogram of the number of tickers per message, across all messages referring to more than one ticker.
 The maximum number of tickers per message amounts to 28 and corresponds to 11 messages in the sample.}
@@
     \includegraphics[width=0.75\textwidth]{uniq.JPG}
-    \caption{Total number of messages with double counting (red) compared to the total number of messages without double counting (blue). Numbers are aggregated daily.}
+    \mycaption{Number of messages}{Total number of messages with double counting (red) compared to the total number of messages without double counting (blue). Numbers are aggregated daily.}
     \label{fig:uniq}
@@
     \includegraphics[width=0.75\textwidth]{ratio.JPG}
-    \caption{Ratio between the number of messages with double counting and the number of messages without double counting. Numbers are aggreaged daily.}
+    \mycaption{Number of messages II}{Ratio between the number of messages with double counting and the number of messages without double counting. Numbers are aggreaged daily.}
     \label{fig:ratio}
@@
 
-\section{Coverage}
+\subsection{Coverage}
 
@@
     \includegraphics[width=0.75\textwidth]{trim.JPG}
-    \caption{Coverage as a function of the median threshold. A lower threshold increases the coverage at the expense of a bigger bias in the polarity.}
+    \mycaption{Coverage}{Coverage as a function of the median threshold. A lower threshold increases the coverage at the expense of a bigger bias in the polarity.}
     \label{fig:coverage}
@@
 \end{tabular}
-\caption{Coverage after the trimming process. List of tickers and corresponding market capitalization as of 31st of December 2019.}
+\mycaption{Coverage}{Coverage after the trimming process. List of tickers and corresponding market capitalization as of 31st of December 2019.}
 \label{tab:listcov}
@@
 \newpage
-\section{Tutorial for StockTwits messages extraction}
+\subsection{Tutorial for StockTwits messages extraction}
 We use CRSP to get stock prices of all US and Canadian listed firms from 1990 to 2020.
@@
 
-\section{Portfolio construction - various $x$}
+\subsection{Portfolio construction - various $x$}
 As the thresholds $U_t(x)$ and $L_t(x)$ are functions of the hyperparameter $x$, we provide for robustness check the results of our portfolio construction with $x$ = 1.96  (95\% confidence band) and $x$ = 2.81 (99.5\% confidence band).
-\subsection{x = 1.96}
+\subsubsection{x = 1.96}
 
@@
     \includegraphics[width=0.75\textwidth]{cumulative_log_return1.jpg}
-    \caption{Cumulative log returns of both portfolios (x=1.96) and the S\&P500.}
+    \mycaption{Cumulative log returns of both portfolios (x=1.96) and the S\&P500.}{}
     \label{fig:cumpf1}
@@
         \end{subfigure}
-        \caption{Long and Short portfolio for x = 1.96.} 
+        \mycaption{Long and Short portfolios for x = 1.96.}{}
         \label{fig:pf1}
@@
 \clearpage
-\subsection{x = 2.81}
+\subsubsection{x = 2.81}
 
@@
     \includegraphics[width=0.75\textwidth]{cumulative_log_return3.jpg}
-    \caption{Cumulative log returns of both portfolios (x=2.81) and the S\&P500.}
+    \mycaption{Cumulative log returns of both portfolios (x=2.81) and the S\&P500.}{}
     \label{fig:cumpf3}
@@
         \end{subfigure}
-        \caption{Long and Short portfolio for x=2.81.} 
+        \mycaption{Long and Short portfolio for x=2.81.}{}
         \label{fig:pf3}

### 2022-07-24 main/ch1_introduction.tex
Add files via upload

@@
 The first two chapters use neural networks to mitigate default and systemic risk and the last chapter is a financial sentiment analysis using natural language processing.
-\\
 
@@
 In particular, they estimate the intensities of two Poisson processes, one governing default and the other governing other exits.
-\citet{DSW} generates future random values for the covariates using a VAR process while \citet{Duan2012} relaxes this assumption and use forward intensities.
+\citet{DSW} generate future random values for the covariates using a VAR process while \citet{Duan2012} relax this assumption and use forward intensities.
 The latter specifies the intensities as a linear function of state variables and uses maximum likelihood to estimate the parameters.
@@
 
-The last chapter uses natural language processing to assess the predictive power of social media on stock returns.We scrape 90 millions messages from Stocktwits - a microblogging platform similar to Twitter but designed for finance professionals.
+\clearpage
+The last chapter uses natural language processing to assess the predictive power of social media on stock returns.
+We scrape a large sample of messages from Stocktwits, a microblogging platform similar to Twitter but designed for finance professionals.
 One of the challenge in this context is to create a classifier that understands the vocabulary of the message posted by the users.
@@
 This transforms the messages from words to a vector of numbers which is now in the same dimension as the vocabulary.
-Then, we build two adversarial logistic regressions using the TFIDF vectors as features and the user-labels as targets.
-The first model predicts positive or not positive, the second model predicts negative or not negative.
-When the models agree, we get our label and when they disagree, we treat the tweet as neutral.
+Next, we build two adversarial logistic regressions using the TFIDF vectors as features and the user-labels as targets.
+The first (second) classifier sets bullish (bearish) as positive and non-bullish (non-bearish) as negative class.
+When the models agree, the classification is trivial and when they disagree, we treat the tweet as neutral.
 This procedure allows us to create an artificial neutral class that absorbs all the tweets that do not convey financial information.
-Finally, with daily intervals, we aggregate tweets predicted sentiments per ticker to compute daily polarity time-series.
+Finally, with daily intervals, we aggregate the predicted sentiments per ticker to compute daily polarity time-series.
 We then use the daily volume of messages on a given firm to identify sudden peak of activity, indicating a firm event.
 Computing cumulative average abnormal return and cumulative abnormal polarity in a 41 days window centered at the identified event, we show that abnormal polarities have significant predictive power on the type of event.
+The performance of sentiment-sorted portfolios illustrates the economic relevance of our sentiment measure.

### 2022-07-24 main/paper1.tex
Add files via upload

@@
 \section{Introduction}
-\label{S:1}
+\label{S:1.1}
 The first default prediction models appeared forty years ago with the first generation model presented by \citet{altman}.
 This work led to the so-called Altman Z-score formula which uses accounting data to compute the default probability of a firm in the next two years.
-However, when used for financial firms, Altman's Z-score formula needs to be used with care  because, as I will discuss in this paper, financial firms have to be treated carefully due to their frequent use of off-balance sheet financing.
+However, when used for financial firms, Altman's Z-score formula needs to be used with care  because, as I will discuss in this chapter, financial firms have to be treated carefully due to their frequent use of off-balance sheet financing.
 Twenty years later, a second generation of reduced-form models used econometrical tools such as maximum likelihood, probit, and logit regressions.
@@
 One innovative recent development is the use of doubly stochastic Poisson intensity model combined with multiple logit regression to account for multi-period default probability estimation.
-This model has been proposed by \citet{DSW} in \textit{Multi-period corporate default prediction with stochastic covariates}.
-The main contribution over prior work was to exploit the time-series dynamics of the explanatory covariates in order to estimate the likelihood of default over several future periods.
+This model is proposed by \citet{DSW} in \textit{Multi-period corporate default prediction with stochastic covariates}.
+The main contribution over prior work is to exploit the time-series dynamics of the explanatory covariates in order to estimate the likelihood of default over several future periods.
 Their model employs firm-specific and macroeconomic data to create a Markov state vector $X_t$ in order to compute independent firm default intensities $\lambda(t)$ and other types of exit intensities $\phi(t)$.
-\citet{DSW} is the first model capable of multi-period default probability estimation using time dynamics of covariates $X_t$.
+The model proposed in \citet{DSW} is the first one capable of multi-period default probability estimation using time dynamics of covariates $X_t$.
 Applications of \citet{DSW} works are various.
 We can find them in credit rating by credit rating agencies, banks who want to calculate the minimal amount of capital to be held and other researches analyzing the link between macroeconomic cycles and firm's default probabilities.
-Covariates used by \citet{DSW} are firm's distance to default, firm's trailing one-year stock return, three-month Treasury bill rate and trailing one-year return on S\&P500.
-Estimating their model on US-listed Industrial firms between 1980 and 2004, they find that distance to default and the current state of the economy have a significant impact on default hazard rates.
+Covariates used by \citet{DSW} are firm's Distance-to-Default, firm's trailing one-year stock return, three-month Treasury bill rate and trailing one-year return on S\&P500.
+Estimating their model on US-listed industrial firms between 1980 and 2004, they find that Distance-to-Default and the current state of the economy have a significant impact on default hazard rates.
 
-The two papers the closest to my paper are the work from \citet{DSW} and \citet{Duan2012}.
+\clearpage
+The two papers the closest to this chapter are the work from \citet{DSW} and \citet{Duan2012}.
 The first model uses the doubly stochastic argument to derive multi-period default probabilities.
-To do so, it requires some strong assumptions (i.e Vectorial Autoregressive process) regarding the behavior of the time-series of covariates to generate future random values for the covariates.
-Obviously, if the process is misspecified, biases are introduced both in the forecasted covariates and in the future default probabilities.
-Five years later, \citet{Duan2012} shows that we can relax the VAR assumption with the use of forward intensities.
-This paper explains how we can reduce biases by projecting current event realizations on past data.
+To do so, it requires strong assumptions (e.g.
+vector autoregressive process) regarding the behavior of the time-series of covariates to generate future random values for the covariates.
+If the process is misspecified, biases are introduced both in the forecasted covariates and in the future default probabilities.
+Five years later, \citet{Duan2012} show that we can relax the VAR assumption with the use of forward intensities.
+This chapter explains how we can reduce biases by projecting current event realizations on past data.
 For convenience, they specify the intensities as a linear function of state variables.
 I wish to extend the latter by removing the assumption of linear intensities and use an artificial neural network to estimate the intensities of the Poisson processes governing both default and other exits.
-Many papers on estimating default probabilities with machine learning techniques have already been written.
-For instance, \citet{barboza2017machine} test several machine learning models (support vector machines, bagging, boosting, and random forest) to predict bankruptcy one year prior to the event.
+Machine learning techniques for default probabilities estimation are increasingly drawing attention.
+\citet{barboza2017machine} test several machine learning models (support vector machines, bagging, boosting, and random forest) to predict bankruptcy one year prior to the event.
 They report a substantial improvement in prediction accuracy using these machine learning techniques especially when, in addition to the original Altman’s Z-score variables, six complementary financial indicators are included.
-To the best of my knowledge, such types of models do not rely on theoretical foundations and this is where the forward intensity model can contribute to the literature.
-It is able to provide multi-period predictions and is supported by a solid mathematical and econometrical background.
+However, pure data-driven models often lacks economic relevance and this is where the forward intensity model can contribute to the literature.
+It is able to provide multi-period predictions while being supported by an economic and econometric background.
 
-In \citet{DSW}, one of the main assumption is that the covariates governing both default and other exits intensities follow a high-dimensional vector auto-regressive (VAR) process.
+In \citet{DSW}, one of the main assumption is that the covariates governing both default and other exits intensities follow a high-dimensional vector autoregressive (VAR) process.
 Using this type of process forces the model to greatly reduce either the number of firms in the sample or the number of state variables explaining firm attributes; if one does not restrict the number of firms or variables in the estimation, the dimension of the model will simply be too high and it will considerably increase computational time.
-A major step forward made in \citet{Duan2012} was to get rid off the VAR process in order to reduce computational time by using ``a new reduced-form approach based on a \textit{forward} intensity model''.
+A major step forward made in \citet{Duan2012} was to get rid off the VAR process in order to reduce computational time by using a new reduced-form approach based on a forward intensity model'.
 These forward intensities produce a term structure of bankruptcy probabilities without using any sort of high-dimensional process.
 Using this method allows the model to incorporate a lot more state variables or individual firms in the sample.
-Moreover, \citet{Duan2012} state that their model may also improve robustness to misspecification because in a VAR model, estimation of future values are highly sensitive to any biases.
-On the other side, the forward intensity model approach is a ``direct projection of past data on current realizations'' which does not involve random estimation of future values.
-\\
+Moreover, \citet{Duan2012} state that their model may also improve robustness to misspecification because in a VAR model, estimation of future values are highly sensitive to any bias.
+On the other side, the forward intensity model approach is a projection of past observations on current realizations, which does not involve random estimation of future values.
+
 Regarding covariates used, both \citet{Duan2012} and \citet{DSW} estimate their own Distance-to-Default (hereafter DtD).
-An important aspect to highlight is that distance to defaults specified in \citet{DSW} differ from those estimated in \citet{Duan2012} since the first one are estimated using the Variance Restriction method (see \textit{Measuring Distance-to-Default for Financial and Non-Financial Firms} \citep{Duan2012DTD}) and the second one using the transformed-data maximum likelihood estimation method to account for financial firms.
-The variance restriction method is a popular way to implement the \citet{Merton1974} model but fails to estimate properly the default point for financial firms.
-Following the KMV assumption (see \citet{KMV}), the default point in this method is specified as short term debt plus one half of long term debt and does not take into account other liabilities.
-However, it is well-known that financial firms such as banks specify a high portion of their debt as ``other liabilities".
-Hence, to include financial firms in the sample, the default point has to be adjusted to the sum of short term debt, one half of long term debt and a fraction $\delta$ of other liabilities.
+An important aspect to highlight is that Distances-to-Default specified in \citet{DSW} differ from those estimated in \citet{Duan2012} since the former estimates DtD using the variance restriction method (see \citep{Duan2012DTD}) while the latter uses the transformed-data maximum likelihood estimation method to account for financial firms.
+The variance restriction method is a popular way to implement the \citet{Merton1974} model but fails at estimating properly the default point for financial firms.
+Following the KMV assumption (see \citet{KMV}), the default point in this method is specified as short-term debt plus one half of long-term debt and does not take into account other liabilities.
+However, it is well-known that financial firms such as banks specify a high portion of their debt as other liabilities.
+Hence, to include financial firms in the sample, the default point has to be adjusted to the sum of short-term debt, one half of long-term debt and a fraction $\delta$ of other liabilities.
 The method used by \citet{Duan2012} is using a maximum likelihood estimation in order to estimate the unknown fraction $\delta$.
 The Appendix provides additional methodological information on the DtD estimation using the variance restriction method and the maximum likelihood estimation.
-\\
-To summarize, the forward intensity models require an assumption to link covariates to intensities.
-\citet{Duan2012} uses a linear assumption and a maximum likelihood to estimate those parameters.
-This paper contributes to previous literature by using neural networks to relax the linear assumption of forward intensities.
+
+To summarize, the forward intensity model requires an assumption to link covariates to intensities.
+\citet{Duan2012} use a linear assumption and a maximum likelihood to estimate those parameters.
+This chapter contributes to previous literature by using neural networks to relax the linear assumption of forward intensities.
 Neural networks allow the estimation of highly non-linear functions without specifying the form of the relationships.
-The organization of the paper is as follows.
-Section 2 sets up the reduced-form model of default, develops the likelihood function used as loss function later on and describes the neural network approach.
-Section 3 discusses summary statistics of the dataset used.
-Section 4 presents the results.
-Section 5 concludes.
-Appendices at the end of the paper contains several proofs and more details on the Distance-to-Default estimation.
+The remainder of this chapter is structured as follows.
+Section \ref{S:1.2} sets up the reduced-form model of default, develops the likelihood function used as loss function later on and describes the neural network approach.
+Section \ref{S:1.3} discusses summary statistics of the dataset used.
+Section \ref{sec4} presents the results.
+Section \ref{S.1.5} concludes.
+Appendix \ref{appendix1} at the end of the thesis contains several proofs and more details on the Distance-to-Default estimation.
 
 \section{Methodology}
-\label{S:2}
+\label{S:1.2}
 
@@
 Let us denote by $Z_{it}$ the set of firm-specific variables at time t for the firm i and $Y_t$ the set of macroeconomic variables at time t.
-The first time of entry of the firm i is denoted  $t_i^0$.
+Let  $t_i^0$ be the first time of entry of firm $i$ in the dataset.
 The econometrician's information set $\mathscr{F}_t$ at time t is thus
@@
 In Figure \ref{fig:lifespan}, each dot corresponds to a period.
-At each period, the econometrician gathers firm-specific variables (DTD, Cash/TA, NI/TA, ...) and macroeconomic variables (S\&P500 return, treasury rate).
-For a particular point in time t, the econometrician knows the whole time-serie of macroeconomic variables until t irrespective of $t_i^0$, and the time-series of firm-specific variables from $t_i^0$ to $\tau_{D_i}$ if $t_i^0 \leq t$ and $t \leq \tau_{D_i}$.
+At each period, the econometrician gathers firm-specific variables (DtD, Cash/Total Assets, Net Income/Total Assets, ...) and macroeconomic variables (S\&P500 return, treasury rate).
+For a particular point in time $t$, the econometrician knows the whole time-serie of macroeconomic variables until $t$ irrespective of $t_i^0$, and the time-series of firm-specific variables from $t_i^0$ to $\tau_{D_i}$ if $t_i^0 \leq t$ and $t \leq \tau_{D_i}$.
 Following Proposition 1 in \citet{DSW}, the conditional probability of default within $s$ years can be computed as 
@@
 \citet{Duan2012} propose a model to predict corporate defaults at multiple horizons by estimating these forward intensities via maximum likelihood.
-To do so, they use a linear assumption in the relationship between the variables and the forward intensities (i.e $f_{it}(\tau) = \exp(\alpha_0(\tau) + \alpha_1(\tau) x_{it,1} + \alpha_2(\tau) x_{it,2} + ...
+To do so, they use a linear assumption in the relationship between the variables and the forward intensities (i.e.
+$f_{it}(\tau) = \exp(\alpha_0(\tau) + \alpha_1(\tau) x_{it,1} + \alpha_2(\tau) x_{it,2} + ...
 + \alpha_k(\tau) x_{it,k}$).
@@
 Since we do not have the exact knowledge of $\lambda$ and $\phi$, \citet{Duan2012} propose to use forward intensity rates.
-However, we need to translate the default probability given in equation \ref{eq:3.6} (which depends on spot intensities) into a formula that depends on forward intensities.
-To do so, we first compute the probability of surviving as a function of forward combined intensity, which will be later used to compute the probability of default as a function of both combined and default forward intensities.
+However, the default probability given in equation \ref{eq:3.6} (which depends on spot intensities) needs to be translated into a formula that depends on forward intensities.
+To do so, we first compute the survival probability as a function of forward combined intensity, which will be later used to compute the probability of default as a function of both combined and default forward intensities.
 Let us denote by $F_{it}(\tau)$ the conditional distribution function of the combined (default and other exits) exit time evaluated at $t+\tau$.
@@
 \end{equation}
-Then, let us introduce the quantity $\psi_{it}(\tau)$ to be :
+Next, let us introduce the quantity $\psi_{it}(\tau)$ to be :
 \begin{equation} \label{eq:3.8}
@@
 We now need to compute $\psi_{it}(\tau)\cdot\tau$.
-At this point, \citet{Duan2012} makes the assumption that $\psi_{it}$ is differentiable and defines the forward combined exit intensity as
+At this point, \citet{Duan2012} make the assumption that $\psi_{it}$ is differentiable and define the forward combined exit intensity as
 \begin{equation} \label{eq:3.10}
@@
 \end{equation}
-Equation \ref{eq:3.10} comes from the definition of a hazard rate function. Referring  P.87 of \citet{book}, the definition of a hazard rate function is the following:
-%\newtheorem*{mydef}{Definition}
+Equation \ref{eq:3.10} comes from the definition of a hazard rate function. Referring \citet{book}, the definition of a hazard rate function is the following:
 \begin{definition}[Hazard rate]
-Let denote by $\tau$ a stopping time and $F(T) \equiv \mathbb{P}[\tau \leq T]$ its distribution function.
+Let $\tau$ be a stopping time and $F(T) \equiv \mathbb{P}[\tau \leq T]$ its distribution function.
 Assume that $F(T) < 1 \forall T$, and that F(T) has a density f(T).
-The hazard rate function h of $\tau$ is :
+The hazard rate function $h$ of $\tau$ is :
 \begin{equation*}
@@
 A hazard rate is the local arrival probability of a stopping time per time interval.
-Please note that under suitable regularity conditions, intensities and hazard rates are closely similar.
+Under suitable regularity conditions, intensities and hazard rates are closely similar.
 In particular, in our doubly-stochastic framework, hazard rates and intensities are equivalent.
-This is why in \citet{Duan2012}, we have $\lambda(t) = h(t)$ and the distinction between hazard rates and intensity is often not made.
+Hence, in \citet{Duan2012} and this chapter, $\lambda(t) = h(t)$ and the distinction between hazard rates and intensity is not made.
 \\
@@
 
-\noindent Then, we can compute the quantity $\psi_{it}(\tau) \tau$ that we were looking for as :
+\noindent Next, we can compute\footnote{Proofs of the following formulations can be found in the Appendix \ref{appendix1}.} the quantity $\psi_{it}(\tau) \tau$ that we were looking for as :
 \begin{equation}
@@
 
-\noindent For an interested reader, proofs of the above formulations can be found in the Appendix. Hence, the probability of surviving over [t,t+$\tau$] is given by
+\noindent Hence, the probability of surviving over [t,t+$\tau$] is given by :
 \begin{equation} \label{survprob}
@@
 
-The forward default intensity for horizon $\tau$ is defined as the limit for a small time step of the probability of defaulting in this small time step given that the firm has survived.
+The forward default intensity for horizon $\tau$ is defined as the limit for a small time step of the probability of defaulting in this small time step given that the firm survives until the considered horizon.
 The probability is Bayesian and the forward default intensity denoted $f_{it}(\tau)$ is the following :
+
 \begin{equation}
@@
 \end{equation}
+
 Hence, the probability of defaulting between t and t+$\tau$ is given by :
+
 \begin{equation}\label{probdef}
@@
 
-In this setup, the likelihood function depends on three types of probabilities (default, other exit and surviving) which themselves depend on the types of intensities (default and other exits).
+In this setup, the likelihood function depends on three types of probabilities (default, other exit and surviving) which themselves depend on two types of intensities (default and other exits).
 The negative log-likelihood function has to be adjusted to the neural network framework and can be used as objective function to be minimized as long as we feed mini-batches and not single data points (``online learning'').
-Mini-batch feeding is a common practice in machine learning papers and consists of splitting the available data in batches of fixed size.
-Then, each backward pass takes one batch to perform one gradient decent step.
+Mini-batch feeding is a common practice in machine learning and consists of splitting the available data in subsets of fixed size.
+Next, each backward pass takes one batch to perform gradient descent to update the parameters of the model.
 \\
@@
 The use of the $\tau$ index is to account for multiperiod prediction.
-When $\tau = 0$, the forward intensity model is computing spot intensities.
+When $\tau = 0$, the forward intensity model computes spot intensities.
 When we set $\tau = 1$, the forward intensity model produces estimates one step ahead, and so on so forth.
 I denote $X_{it} = (x_{it,1}, x_{it,2} ,...)$ the set of firm-specific and macroeconomic variables explaining both default and combined exit intensities.
-As specified in \citet{Duan2012} $f_{it}(\tau)$ and $g_{it}(\tau)$ are functions of $X_{it}$ and can be specified as any form of function as long as they satisfy the following constraints.
+As specified in \citet{Duan2012}, $f_{it}(\tau)$ and $g_{it}(\tau)$ are functions of $X_{it}$ and can be specified as any form of function as long as they satisfy the following constraints.
 Since combined exit intensity has to be greater or equal than default intensity, we need to make sure that the forms specified for $f_{it}(\tau)$ and $g_{it}(\tau)$ satisfy the following conditions : $f_{it}(\tau) \leq g_{it}(\tau)$, $f_{it}(\tau) > 0$, $g_{it}(\tau) > 0$.
-\\
 
@@
 $N^{(\lambda)}$ and $N^{(\mu)}$ stand for the output of the neural network for $f_{it}$ and $h_{it}$ respectively.
-The log-likelihood for prediction horizon $\tau$ is expressed as
+The log-likelihood for prediction horizon $\tau$ is expressed\footnote{Proof of the above formulation can be found in the Appendix \ref{appendix1}.} as
 \begin{equation} \label{decomp1}
@@
 
-Proof of the above formulation can be found in the Appendix.
 As in previous studies, the likelihood functions still exhibit the decomposable property which allows to estimate the model for each horizon of prediction independently.
 
-Since the intensities are directly driven by the covariates, \citet{Duan2012} requires an assumption on the mapping from the covariates to these intensities.
-In \citet{Duan2012} the mapping is made with a linear assumption, whereas in the framework of this paper, the mapping depends on the whole architecture of the neural network.
-When the neural network has only one hidden layer of one neuron coupled with an exponential activation function, the model boils down to \citet{Duan2012} as the intensities will be a linear combination of the covariates.
+Since the intensities are directly driven by the covariates, \citet{Duan2012} require an assumption on the mapping from the covariates to these intensities.
+In \citet{Duan2012} the mapping is made with a linear assumption, whereas in the framework of this chapter, the mapping depends on the whole architecture of the neural network.
+When the neural network has only one hidden layer of one neuron coupled with an exponential activation function, the model boils down to \citet{Duan2012} as the intensities would be a linear combination of the covariates.
 As the width and depth of the network increases, we depart more and more from the linear assumption and we allow more non-linearities to be incorporated in these intensities.
@@
 It learns how to compute the output by tuning weights in order to minimize a given loss function.
-A neural network is constructed by juxtaposing several ``hidden'' layers of neurons.
+A neural network is constructed by juxtaposing several hidden layers of neurons.
 The input of each layer is a data transformation of the output of the previous layer.
@@
 The forward pass takes as input a batch of data and computes the loss value, the backward pass then computes the gradient and adjust the weights of the network based on a learning rate hyperparameter.
-As an illustration, figure \ref{fig:net} shows how a neural network looks like with 2 hidden layers : 5 neurons in the first layer and 3 neuron in the second layer.
+As an illustration, Figure \ref{fig:net} shows a neural network with 2 hidden layers : 5 neurons in the first layer and 3 neurons in the second layer.
 In this example, 3 features (inputs) are fed to the network.
@@
 The outputs of the first layer are also weighted to produce the inputs of the second hidden layer.
-Non-linearity is introduced in each node with a non-linear activation function (i.e sigmoid).
+Non-linearity is introduced in each node by a non-linear activation function (e.g.
+sigmoid).
 Finally, the output of the second hidden layer is aggregated to produce the final output of the model.
-The neural networks in this paper are implemented in Python using the library TensorFlow.
-Approximately one hour of computing time is needed to fit all networks on a 32GB RAM quad core 2.7 GhZ computer, and GPU computing is not necessary as the networks are rather small.
+The neural networks in this chapter are implemented in Python using the library TensorFlow.
+Approximately one hour of computing time is needed to fit all networks on a 32GB RAM quad core 2.7 GHz computer.\footnote{GPU computing is not necessary in this context as the networks are small.}
 
@@
 \end{tikzpicture}
-    \mycaption{Illustration of a neural network \text{[}5, 3\text{]}}{In this network, there are three input features, five neurons in the first hidden layer and 3 neurons in the second hidden layer.
+    \mycaption{Illustration of a neural network \text{[}5, 3\text{]}}{In this network, there are three input features (green circles), five neurons in the first hidden layer and 3 neurons in the second hidden layer.
 Each neuron is activated with a sigmoid function.}
@@
 A deeper network allows for more non-linearities in the approximation of the function, at the expense of having more parameters to estimate.
-For instance, suppose that every observation comes with 12 input features (i.e : $x$ is a vector of shape 1x12), a [5, 3] (i.e 2 layers neural network with 5 neurons in the first hidden layer and 3 neurons in the second hidden layer) can be viewed as a function computing the output $N_{it}^{(\lambda)}$ in the following way :
+For instance, suppose that every observation comes with 12 input features (i.e.
+: $x$ is a vector of shape 1x12), a [5, 3] (i.e.
+2 layers neural network with 5 neurons in the first hidden layer and 3 neurons in the second hidden layer) can be viewed as a function computing the output $N_{it}^{(\lambda)}$ in the following way :
 
@@
 \noindent  with the activation functions being for instance the sigmoid function $\phi_1(x) = \phi_2(x) = \frac{1}{1+e^{-x}}$, $x$ being the data input, $w_1$, $w_2$ and $w$ weight matrices and $b_1$ and $b_2$ biases matrices.
-In this setup, the number of parameters to estimate is equal to 12x5 + 1x5 + 5x3 + 1x3 + 3x1 = 86.
+In this architecture, the number of parameters to estimate is equal to 12x5 + 1x5 + 5x3 + 1x3 + 3x1 = 86.
 \\
-Neural networks are also well-suited for this paper because they allow an easy replication of the benchmark model \citet{Duan2012}.
-More specifically, if the activation function $\phi(x)$ is chosen as being an exponential $\exp(x)$, and the network architecture is [1] (i.e a single hidden layer with a single neuron), the output $N_{it}^{(\lambda)}$ becomes
+Neural networks are also well-suited for this chapter because they allow an easy replication of the benchmark model \citet{Duan2012}.
+More specifically, if the activation function $\phi(x)$ is chosen as being an exponential $\exp(x)$, and the network architecture is [1] (i.e.
+a single hidden layer with a single neuron), the output $N_{it}^{(\lambda)}$ becomes the linear assumption
 
@@
 \section{Empirical section}
-\label{S:3}
+\label{S:1.3}
 
@@
 DLRSN contains the code that indicates the reason a company become inactive on the database.
-I consider firms with a DLRSN code 2 (bankruptcy) or 3 (liquidation) to be defaulted, any other DLRSN code as ``other exits" and no DLRSN code as surviving.
-For additional information on DLRSN and DLDTE codes, please refer to the Wharton WRDS documentation.
+I consider firms with a DLRSN code 2 (bankruptcy) or 3 (liquidation) to be defaulted, any other DLRSN code as other exits and no DLRSN code as surviving.
+For additional information on DLRSN and DLDTE codes, I refer to the Wharton WRDS documentation.
 I focus on the period from 1991 to 2018 to match the accounting data with the bankruptcy data.
 Using WRDS database, I have downloaded accounting information for every company that has been listed someday on either NYSE, AMEX or NASDAQ between 1991 and 2018.
-This means I have 27 years of data where firms entered and/or exited anywhere in this sample.
-Using this kind of sample brings a problem of cylindric data : firm's entering/exiting time obviously are not the same for each company.
-We can see the whole dataset as a 3D matrix with x-axis being features, y-axis being time, and z-axis being firm.
-We only need to fill the matrix with NaN for elements where the firm $i$ doesn't exist or already left at time $t$.
-Since neural networks need plenty of data points to be well trained, I chose not to remove firms even if it has a short lifespan.
-When a variable is completely missing for a firm, I drop the whole firm because the likelihood is not specified if a variable is fully missing.
+The dataset spans 27 years of data where firms entered and/or exited anywhere in this sample.
+Using this kind of sample brings a problem of cylindric data : firm's entering/exiting time are not the same for each company.
+The dataset is represented as a three dimensions matrix with the x-axis being features (i.e.
+variables), the y-axis being time, and the z-axis being firm.
+I fill the matrix with missing values for elements where the firm $i$ doesn't exist or already left at time $t$.
+Since neural networks need a lot of data points to be well trained, I chose not to remove firms even if it has a short lifespan.
+When a firm has a variable completely missing, I drop the whole firm because the likelihood is not specified if a variable is fully missing.
 However, when the variable is not fully missing but only some data points are not available, I use the last available information before the missing entry.
@@
 Finally, I standardize all variables by substracting the mean of the variable and dividing the result by its standard deviation.
-The test set is also standardized using the mean and standard deviation from the training set.
+The variables in the test set are also standardized using their respective mean and standard deviation from the training set.
 Table \ref{tab:observ} shows the number of firms in the three categories for each horizon of prediction $\tau$ which corresponds to the firm-month observations used in the likelihood for horizon $\tau$ (see equations \ref{small_lik_f} and \ref{small_lik_h}).
@@
     \centering
-    \includegraphics[width=\textwidth]{def_surv_oe_plot.jpg}
-    \mycaption{Number of firms defaulted, exited for other reasons, surviving each year}{}
+    \includegraphics[width=0.75\textwidth]{def_surv_oe_plot.jpg}
+    \mycaption{Defaults, exits for other reasons and survivals each year}{Top plot shows the number of defaulted firms, middle plot shows the number of exits for other reasons and bottom plot shows the number of firms surviving each year.}
     \label{fig:defsurv}
@@
     \hline\hline
-Horizon & Surviving & Defaults & Other exits \\
+Horizon & Survivals & Defaults & Other exits \\
     \hline
@@
 
-\citet{leippold} used a theoretical model to show that the most powerful default predictor must incorporate both macroeconomic and accounting data.
-For more transparency and to allow better comparison with previous literature, I chose to work with a similar set of features similar than my benchmark model \citet{Duan2012}.
+\citet{leippold} use a theoretical model to show that the most powerful default predictor must incorporate both macroeconomic and accounting data.
+For more transparency and to allow better comparison with previous literature, I choose to work with a similar set of features than the benchmark model \citet{Duan2012}.
 Firm-specific values are common to each firm, macroeconomics variables are function of market data.
@@
 \item Treasury : 3-month annualized US Treasury bill rate. I use this variable as the risk-free rate $r$ in the model. 
-\item DtD : the Distance-to-default is a volatility adjusted leverage measure for gauging how far is the firm from default.
+\item DtD : the Distance-to-Default is a volatility adjusted leverage measure for gauging how far is the firm from default.
 It was first introduced by \citet{Merton1974} model which treat firm's equity value as a call option on the underlying asset and the strike being the debt level.
 The DtD can be viewed as the number of standard deviations of annual asset growth by which the asset level exceeds the firm's liabilities.
-In this paper, the DtD is estimated using a maximum likelihood taking into account other liabilities of each firm to handle financial firm's bias.
+In this chapter, the DtD is estimated using a maximum likelihood taking into account other liabilities of each firm to handle financial firm's bias.
 It is well-known in the literature that DtD is a significant measure to estimate default probabilities but has to be used conjointly with other variables.
-See the Appendix for additional information regarding the estimation procedure of DtD.
+See the Appendix \ref{appendix1} for additional information regarding the estimation procedure of DtD.
 \item CASH/TA : ratio between cash and short-term investment to total assets. Both quantities are taken from the balance sheet of the company.
-\item NI/TA : ratio between net income to the total assets. A loss is registered as a negative net income.
-\item SIZE : logarithm of the ratio between a firm market equity value to the average market equity value of the whole S\&P500.
+\item NI/TA : ratio between net income to total assets. A loss is registered as a negative net income.
+\item SIZE : logarithm of the ratio between market equity value to the cross-sectional average market equity value of the stocks in the S\&P500.
 The market equity value is computed as the stock price multiplied by the number of outstanding shares at the end of each quarter.
@@
 This section depicts the summary statistics of the dataset.
-Table \ref{tab:meanvar} shows a summary of the means of variables for each three categories of firm.
+Table \ref{tab:meanvar} shows a summary of the mean of each variable for the three categories of firm.
 Please note that we have to be very careful when comparing two means in this table.
-Comparing means needs to be done with confidence intervals and tests of significance which involves standard deviations.
+Comparing means needs to be done with confidence intervals and significance tests which involves standard deviations.
 This table is presented to give a rough idea without performing any statistical inference.
@@
     \end{tabular}
-    \mycaption{Mean of variables for surviving firms, defaulted firms and other exits}{$\Delta$ are one period lagged differences.}
+    \mycaption{Mean of variables for surviving firms, defaulted firms and other exits}{The prefix $\Delta$ stands for one period lagged differences.}
     \label{tab:meanvar}
@@
 
-\begin{figure}[]
+\begin{figure}
     \centering
     \includegraphics[scale=0.6]{corrmatrix.jpeg}
-    \mycaption{Correlation matrix for firm-specific and macroeconomic covariates}{}
+    \mycaption{Correlation matrix for firm-specific and macroeconomic covariates}{Red values represent positive correlation coefficients, blue values shows negative correlation coefficients.}
     \label{fig:corrmat}
@@
 As every neural network, we need to chose the optimal hyperparameters to achieve the highest performance.
-In my setup, this consists mainly of choosing the architecture of the model (i.e the number of neurons and layers).
+In my setup, this consists mainly of choosing the architecture of the model (i.e.
+the number of neurons and layers).
 To do so, there is currently no other better method than trial and error.
-However, I need to be careful to not chose hyperparameters that overfit the test set.
+However, hyperparameters need to be carefully chosen to not overfit the test set.
 To avoid any overfitting, I perform a 5-fold cross validation for each horizon of prediction.
@@
 The Lorenz curve plots on the x-axis the cumulative percentage of observations against the fraction of defaults on the y-axis.
-These curves are often used in the literature for default prediction (for instance \citet{leippold}, \citet{Duan2012}) and I have seen very similar plots under two other names : power curves and cumulative accuracy profiles.
-They are different from ROC curves and precision-recall curves because they don't rely on thresholds to discriminate of true positives against false positives.
-I believe that they are particularly well-suited as performance measure in this model because of the multiperiod framework involved.
-The idea is that if the model is outputting a false positive for an horizon x but the true positive is horizon x+1, the model should not be ``too much" penalized.
+These curves\footnote{Similar plots exist also under different names (e.g.
+power curves, cumulative accuracy profiles).} are often used in the literature for default prediction (for instance \citet{leippold}, \citet{Duan2012}).
+They are different from ROC curves and precision-recall curves because they don't rely on thresholds to discriminate between true positives and false positives.
+Hence, they are particularly well-suited as performance measure in this model because of the multiperiod framework involved.
+The idea is that if the model is outputting a false positive for an horizon $\tau$ but the true positive is horizon $\tau+1$, the model should not be too much penalized.
 The ROC and precision-recall curves would treat this as a false positive even though the prediction was not that far from the target.
@@
 We can easily see whether the model is outputting high probabilities for defaulted firms and small probabilities for surviving firms.
-\\
-Finally, the Gini coefficient is used as my scalar summary statistic to compare models.
+
+Finally, the Gini coefficient is used as the scalar summary statistic to compare models.
 It measures the degree of inequality of the Lorenz curve.
-A perfect model has a Gini coefficient close to 1 (depending on the fraction of defaults in the dataset) and a poor model has a Gini coefficient of 0 (perfect equality).
+A perfect model has a Gini coefficient close to 1 and a poor model has a Gini coefficient of 0 (perfect equality).
 
@@
 This is usually known in the machine learning literature as the bias-variance tradeoff.
-Figure \ref{fig:ginicomp} shows the average Gini coefficient compute on the validation test across all folds of the cross validation for each horizon.
+Figure \ref{fig:ginicomp} shows the average Gini coefficient computed on the validation test across all folds of the cross validation for each horizon.
 I compare the scores obtained with different network architectures with those using the linear assumption.
-In this setup, one can interpret ``architecture'' as ``non-linearity degree'' as higher architecture involves more weights in the output function to be estimated.
+In this setup, one can interpret the architecture model as the non-linearity degree because higher architecture involves more weights in the output function to be estimated.
 By disantangling the spaghettis, we can clearly see the bias-variance tradeoff.
@@
 Finally, increasing the depth again to [10, 5] decreases Gini coefficient at all horizons, suggesting a severe overfitting of the forward intensities function.
-For next sections, I will now only show out-of-sample results of the [5, 3] architecture using the test set preliminary set on the side.
-\\
+For the following sections, I will now only show out-of-sample results of the [5, 3] architecture using the test set.
 
@@
     \centering
-    \includegraphics[width=\textwidth]{duan_vs_NN.jpeg}
-    \mycaption{Gini coefficient : Linear vs Neural Networks}{}
+    \includegraphics[width=0.75\textwidth]{duan_vs_NN.jpeg}
+    \mycaption{Gini coefficient : Linear vs Neural Networks}{Yellow line shows the Gini coefficient of the linear model for every horizon.
+The remaining lines show the Gini coefficient for different neural network architectures.
+Green line is the best performing model as it exhibits the highest Gini for every horizon.}
     \label{fig:ginicomp}
@@
 
-Figure \ref{fig:perf} shows Lorenz curves for horizon 0, 3, 6, 12, 24 and 35.
-Please note that all results generalize well to all others horizons.
+Figure \ref{fig:perf} shows Lorenz curves for horizon 0, 3, 6, 12, 24 and 35 but all results generalize well to all others horizons.
 The curves are completely out-of-sample since they are computed on the test set that the model has never seen before.
-Table \ref{fig:ginicomp} shows Gini coefficients for both the linear assumption \citet{Duan2012} and for the [5, 3] for the same horizons.
+Table \ref{tab:ginicoeff} shows Gini coefficients for both the linear assumption \citet{Duan2012} and for the [5, 3] for the same horizons.
 Overall, as expected the neural network clearly outperforms the linear assumption, suggesting that the linear assumption from \citet{Duan2012} can be greatly improved by adding non-linearities in the specification of the intensities.
 Unfortunately, it is difficult to tell which kind of non-linearities should be taken into account since neural networks are often seen as a black box.
-However, I will still try to answer this question by looking how the average intensity outputted by the model changes when we change a feature ceteris paribus (see section dedicated to ``sensitivities").
-Another attempt at answering this question is described in the section ``computational graph'', where I deep dive into the weights of the network to understand how the output is computed.
+However, I will still try to answer this question by looking how the average intensity outputted by the model changes when we change a feature ceteris paribus (see Section \ref{sec4-4} dedicated to sensitivities).
+Another attempt at answering this question is described in Section \ref{sec4-3}, where I dive into the weights of the network to understand how the output is computed.
 
-For comparison purposes, figure \ref{fig:lor0duan} shows the Lorenz curves for both linear assumption, its replication in the neural network framework, and the [5, 3] model.
-The linear assumption and its replication have similar performance, suggesting again that the neural network ``NN+exp+[1]" is able to easily replicate the linear framework depicted in \citet{Duan2012}.
+For comparison purposes, Figure \ref{fig:lor0duan} exhibits the Lorenz curves of the linear assumption, its replication in the neural network framework, and the [5, 3] model.
+The linear assumption and its replication have similar performance, showing that the neural network ``NN+exp+[1]" is indeed able to replicate the linear framework depicted in \citet{Duan2012}.
 
@@
     \end{tabular}
-    \mycaption{Gini coefficients}{}
+    \mycaption{Gini coefficients}{Comparison of the out-of-sample Gini coefficients associated to the [5,3 ] neural network and the linear assumption.}
     \label{tab:ginicoeff}
@@
     \centering
-    \includegraphics[width=\textwidth]{lor0duan.png}
-    \mycaption{Comparison with the benchmark model \citet{Duan2012}}{The green line shows the performance of the [1] neural network (i.e a single hidden layer with a single neuron) with an exponential activation function.
+    \includegraphics[width=0.75\textwidth]{lor0duan.png}
+    \mycaption{Comparison with the benchmark model \citet{Duan2012}}{The green line shows the performance of the [1] neural network (i.e.
+a single hidden layer with a single neuron) with an exponential activation function.
 The orange line shows the performance of Duan's model and the red line shows the performance of the [5,3] neural network.}
@@
 
+\clearpage
 \subsection{Computational graph}
@@
     \centering
-    \includegraphics[width=\textwidth]{compgraphr.png}
+    \includegraphics[width=0.75\textwidth]{compgraphr.png}
     \mycaption{Trained \text{[}5, 3\text{]} Neural Network for forward default intensity at horizon 0}{Negative weights are drawn as blue lines connecting neurons where orange lines show positive weights.
@@
     \item Compute the forward default intensity for all the observations in the test set and average the result.
-Then, keeping everything else equal, change the value of one feature by an absolute value and feed the ``updated" observations to the network as new input and compute the new average forward default intensity.
-    \item Repeat step 1 for all absolute values in some interval
-    \item Plot the average intensity against the absolute change
+Keeping everything else equal, change the value of one feature by an absolute value and feed the updated observations to the network as new input and compute the new average forward default intensity.
+    \item Repeat step 1 for all absolute values in some interval.
+    \item Plot the average intensity against the absolute change.
     \item Repeat step 1-3 for all 11 other features.
@@
 
-Keep in mind that there is a non-negativity constraint on forward intensities.
+As explained in Section \ref{SSS:2-1-2}, I impose a non-negativity constraint on forward intensities.
 A decreasing relationship means that an increase of the associated variable decreases the probability of default of the firm.
@@
 We should expect CASH/TA, size, DtD, NI/TA, Market-to-book ratio and all their lagged differences ($\Delta$) to be decreasing.
-\\
 
@@
 The model also predicts that firms with low NI/TA or decreasing NI/TA should have higher probability of default.
-Finally, forward intensities in DtD should be decreasing for all horizons to reflect that a higher distance to default makes the firm less likely to default.
-At horizon 0, a negative change in distance-to-default has a substantially greater effect on forward intensities than any other variables.
+Finally, forward intensities in DtD should be decreasing for all horizons to reflect that a higher Distance-to-Default makes the firm less likely to default.
+At horizon 0, a negative change in Distance-to-Default has a substantially greater effect on forward intensities than any other variables.
 This result is consistent with \cite{DSW}.
@@
         \end{subfigure}
-        \mycaption{Sensitivities for each horizon}{}
+        \mycaption{Sensitivities for each horizon}{These plots show the average default forward intensity against a change in the specified variable, ceteris paribus}
         \label{fig:sensi}
@@
 \section{Conclusion}
-\label{S.5}
+\label{S.1.5}
 
@@
 The econometric model governing the forward intensity written by \citet{Duan2012} has been adapted to this new framework to allow the use of neural networks.
-Neural networks are also well-suited for this paper because they allow an easy replication of the benchmark model \citet{Duan2012}.
-More specifically, if the network architecture is [1] (i.e : one layer and one neuron) and the activation function is chosen as being an exponential $\exp(x)$, the neural network boils down to a logit regression.
-The dataset used in this paper is the same as previous literature to allow easier comparison.
+Neural networks are also well-suited for this chapter because they allow an easy replication of the benchmark model \citet{Duan2012}.
+More specifically, if the network architecture is [1] (i.e.
+: one layer and one neuron) and the activation function is chosen as being an exponential, the neural network boils down to a logit regression.
+The dataset used in this chapter is the same as previous literature to allow easier comparison.
 It consists of 5 firm-specific variables computed from accounting data and 2 macroeconomic variables to control for the health of the economy.
-I also account for momentum of these variables by feeding the model the one lagged differences of each variable.
+I also account for momentum of these variables by feeding the model the one-lagged differences of each variable.
 Looking at summary statistics only, the average defaulted firm is small, has low cash, low market-to-book ratio, low Distance-to-Default and has large and negative profits.
-To measure the discriminatory power of the models, I follow previous literature and use Lorenz curves (also known as ``Cumulative accuracy profile" or ``Power curves").
+To measure the discriminatory power of the models, I follow previous literature and use Lorenz curves (also known as ``cumulative accuracy profile" or ``power curves").
 The idea behind Lorenz curves is to order default probabilities and look how they are distributed across defaulted and non-defaulted firms.
 The average Gini coefficient across all folds of the cross-validation is used as comparison tool to gauge the accuracy of the model.
-Results show that the architecture [5, 3] (i.e 2 layers with 5 neurons in the first hidden layer and 3 neurons in the second hidden layer) seems to outperform others architecture.
-In this setup, one can interpret ``architecture'' as ``non-linearity degree'' as higher architecture involves more weights in the output function to be estimated.
+Results show that the architecture [5, 3] (i.e.
+2 layers with 5 neurons in the first hidden layer and 3 neurons in the second hidden layer) seems to outperform others architecture.
+In this setup, one can interpret the architecture as the non-linearity degree as higher architecture involves more weights in the output function to be estimated.
 Out-of-sample Lorenz Curves and Gini coefficients show that the neural network approach outperforms the linear assumption for every horizon, suggesting the presence of non-linearities in forward default intensities.
-Finally, even if neural networks are known as being black boxes, one can try to see how the model reacts to a change of input variables.
+Finally, even if neural networks are known to be black boxes, I show how the model reacts to a change of input variables.
 Most sensitivities plots show non-linear relationships, which is not surprising given the nature of the neural network specification.
-It appears that the most important predictors of default in this model are in the short term Market-to-book ratio, Distance-to-default and NI/TA, and in the long run NI/TA, CASH/TA, $\Delta$CASH/TA and Distance-to-default.
+It appears that the most important predictors of default in this model are in the short term Market-to-book ratio, Distance-to-Default and NI/TA, and in the long run NI/TA, CASH/TA, $\Delta$CASH/TA and Distance-to-Default.
 Further works could involve more variables in the estimation.

### 2022-07-24 main/paper2.tex
Add files via upload

@@
 
-\section{Introduction}\label{S:1}
+\section{Introduction}\label{S:2.1}
 
@@
 In the great majority of practical applications, Granger-causality has been studied in the context of Vector Autoregressive (VAR) models. 
-For instance, \citet{billio2012econometric} proposes several measures  based on Granger-causality to capture the connections between the monthly returns of different financial institutions. 
-It uses principle component analysis and ``pairwise'' Granger-causality tests to identify the causal networks. 
+For instance, \citet{billio2012econometric} propose several measures  based on Granger-causality to capture the connections between the monthly returns of different financial institutions. 
+It uses principle component analysis and pairwise Granger-causality tests to identify the causal networks. 
 Other related works are \citet{diebold2014network} and \citet{barigozzi2016network} in which the authors propose connectedness measures based on generalized variance decomposition.
@@
 
-Contributions of this paper are both in network identification literature as well in finance.
+\clearpage
+Contributions of this chapter are both in network identification literature and in finance.
 %This paper adds contribution in the network identification literature as well as in finance literature. 
-Our contribution to network identification are as follows:
+Our contributions to network identification are as follows:
 \begin{itemize}
     \item We use an information-theoretic measure known as directed information (DI) to infer the Granger-causalities among a set of time series. 
-This measure is non-parametric, i.e., it does not depend on the underlying model of the dynamics and it is capable of capturing causal relationships in both linear and nonlinear systems.
+This measure is non-parametric (i.e.
+it does not depend on the underlying model of the dynamics) and it is capable of capturing causal relationships in both linear and nonlinear systems.
 The output of this approach is a directed graph known as Directed Information Graph (DIG) that visualizes the interconnections among a set of time series such as stock returns.
 
-\item Computing DI has both high computational and sample complexity which makes it not suitable for inferring the causal structure of large networks. 
+\item Computing DI has high computational and sample complexity which makes it not suitable for inferring the causal structure of large networks. 
 To overcome this problem, we develop a novel approach based on Recurrent Neural Networks (RNNs) that reduces the complexity of evaluating DIs in high-dimensional settings. 
@@
 This is due to information embedded in several economic factors already included in the market index.
-They show that only the lagged value of the market index is enough to predict the forthcoming value of the index.
 
@@
 In recent years, several approaches have been developed to generalize the applicability of Granger-causality to non-linear and large dynamics. 
-To mention a few, \citet{psaradakis2005markov} that introduces different terminologies for causality based on Granger’s ideas and provide a set of parametric non-causality constraints in the context of Markov switching VAR models.
-In a similar context, \citet{bianchi2019modeling} investigates time-varying systemic risk based on a range of multi-factor asset pricing models and develops a Markov Chain Monte Carlo (MCMC) scheme to infer their model parameters and consequently obtain their corresponding networks.
-Another attempt is \citet{bonaccolto2019estimation} in which the authors explore quantile-based methods of Granger causality.
+To mention a few, \citet{psaradakis2005markov} introduce different terminologies for causality based on Granger’s ideas and provide a set of parametric non-causality constraints in the context of Markov switching VAR models.
+In a similar context, \citet{bianchi2019modeling} investigate time-varying systemic risk based on a range of multi-factor asset pricing models and develop a Markov Chain Monte Carlo (MCMC) scheme to infer their model parameters and consequently obtain their corresponding networks.
+Another attempt is \citet{bonaccolto2019estimation} in which the authors explore quantile-based methods of Granger-causality.
 %This work focuses on the linear factor model augmented with network dependence introduced by \citet{billio2018networks} and develops a method to identify causality among quantiles of the time series.
 This method is consistent with \citet{hong2009granger} and \citet{corsi2018measuring} that focus on causality among tail events. 
-These methods are suitable for capturing causal relationships that are not in the center of their distributions, or in the mean but they are in the tails of their distributions.
+These methods are suitable for capturing causal relationships that are not in the center of their distributions, or in the mean but in the tails of their distributions.
 It is important to emphasize that our proposed approach using DI is also capable of capturing such causal relations.
@@
 Often, the problem of network identification in high dimensional settings requires more considerations and even its own techniques. 
-For instance, \citet{billio2019bayesian} proposes a Bayesian non-parametric Lasso prior (BNP-Lasso) for high-dimensional VAR models that can improve efficiency and accuracy. 
+For instance, \citet{billio2019bayesian} propose a Bayesian non-parametric Lasso prior (BNP-Lasso) for high-dimensional VAR models that can improve efficiency and accuracy. 
 In order to overcome overparametrization and overfitting issues in large VAR models, BNP-Lasso clusters the VAR coefficients into groups and shrinks the coefficients of each group toward a common location.
@@
 However, this method is limited to linear models with Gaussian innovations.
-To overcome this limitation, \citet{kalli2018bayesian} proposes a Bayesian non-parametric approach that allows for nonlinearity in the conditional mean, heteroskedasticity in the conditional variance, and non-Gaussian innovations.
+To overcome this limitation, \citet{kalli2018bayesian} propose a Bayesian non-parametric approach that allows for nonlinearity in the conditional mean, heteroskedasticity in the conditional variance, and non-Gaussian innovations.
 However, unlike the BNP-Lasso, it does not allow sparsity in the model.
-\citet{petrova2019quasi} proposes yet another non-parametric, quasi-Bayesian likelihood estimation methodology for high dimensional setting with time varying parameters. 
+\citet{petrova2019quasi} proposes yet another non-parametric, quasi-Bayesian likelihood estimation methodology for high dimensional setting with time-varying parameters. 
 %This method is also non-parametric. 
@@
 
-The rest of the paper is organized as follows. 
-Section \ref{Sec1} reviews the notion of Granger causality and formally introduce directed information graphs which is suitable for linear and nonlinear systems. 
+The rest of the chapter is organized as follows. 
+Section \ref{Sec1} reviews the notion of Granger-causality and formally introduces directed information graphs which are suitable for linear and nonlinear systems. 
 In Section \ref{sec:method}, we introduce a novel approach for inferring the Granger-causal network of high dimensional nonlinear systems. 
-Finally, in Section \ref{sec:exp}, we apply our method to learn the causal network of both synthetic and real-world dataset.
-For the real-world experiment, we used the daily stock prices of major US firms.
+Finally, in Section \ref{sec:exp}, we apply our method to learn the causal network of both synthetic and real-world dataset. For the real-world experiment, we use the daily stock prices of US firms.
 
@@
 
-In this section, we present a statistical approach to learn the causal interconnections in a dynamical systems based on Granger causality \citet{granger}. 
+In this section, we present a statistical approach to learn the causal interconnections in a dynamical systems based on Granger-causality (see \citet{granger}). 
 We begin by introducing some notations.
@@
 Researchers from different fields have developed various frameworks and graphical models to capture and represent interconnections among variables or processes.
-One of the most popular and widely used frameworks in economics is the notion of Granger causality.
-The basic idea in this framework was originally introduced by Wiener \citet{wiener1956theory}, and later formalized by Granger \citet{granger}. 
+One of the most popular and widely used frameworks in economics is the notion of Granger-causality.
+The basic idea in this framework was originally introduced by \citet{wiener1956theory}, and later formalized by \citet{granger}. 
 The idea is as follows: ``we say that $X$ is causing $Y$ if we are better able to predict the future of $Y$ using all available information than if the information apart from the past of $X$ had been used."
@@
 Clearly, the performance of the second predictor is bounded by the first predictor.
-If they have the same performance, then we say $X$ does not Granger cause $Y$.
+If they have the same performance, then we say $X$ does not Granger-cause $Y$.
 It is important to emphasize that this formulation is only applicable in linear systems.
 
-To go beyond linear systems, works such as \citet{directed} and \citet{massey} use information-theoretical measures and generalize Granger causality. 
-In this work, we introduce and apply directed information (DI) \citet{directed}, an information-theoretical tools to measure interconnections among firms.
-DI has been used in many applications to infer causal relationships.
-For example, \citet{spike} and \citet{di12} used it for analyzing neuroscience data and \citet{acc2014}  and \citet{etesami2018econometric} applied to market data.
+To go beyond linear systems, works such as \citet{directed} and \citet{massey} use information-theoretical measures and generalize Granger-causality. 
+In this work, we introduce and apply directed information (see \citet{directed}), an information-theoretical tools to measure interconnections among firms.
+Directed information (DI) has been used in many applications to infer causal relationships.
+For example, \citet{spike} and \citet{di12} use it for analyzing neuroscience data and \citet{acc2014}  and \citet{etesami2018econometric} apply to market data.
 
-In order to visualize the inferred interconnections among time series using DI, directed information graphs (DIGs) have been developed \citet{directed}.  
+Directed information graphs (DIGs) have been developed to visualize the inferred interconnections among time series (see \citet{directed}).  
 DIGs are a type of graphical models in which nodes represent time series and arrows indicate the direction of causation. We use DIG to represent the causal network among the covered firms.
@@
 
-In the rest of this section, we describe how the DI can capture the interconnections in causal\footnote{In causal systems, given the full past of the system, the present of the processes become statistically independent.
-In other words, there are no simulations relationships between the time series.} dynamical systems (linear or non-linear) and formally define DIGs.
+In the rest of this section, we describe how the DI can capture the interconnections in causal dynamical systems (linear or non-linear) and formally define DIGs.
 
@@
 
-At time $t$ for an outcome $y_t\in \mathcal{Y}$, the difference between the log-losses of the two predictors compares their performances.  This difference is also called \textit{regret},
+At time $t$ for an outcome $y_t\in \mathcal{Y}$, the difference between the log-losses of the two predictors compares their performances.  This difference is also called regret, denoted $r_t$ :
 \begin{align}
-r_t&:=-\log{p(Y_t=y_t|Y^{t-1},Z^{t-1})}-\big(-\log {p(Y_t=y_t|Y^{t-1},Z^{t-1},X^{t-1})}\big)\\
+r_t:&=-\log{p(Y_t=y_t|Y^{t-1},Z^{t-1})}-\big(-\log {p(Y_t=y_t|Y^{t-1},Z^{t-1},X^{t-1})}\big) \nonumber \\
 &=\log \frac{p(Y_t=y_t|Y^{t-1},Z^{t-1},X^{t-1})}{p(Y_t=y_t|Y^{t-1},Z^{t-1})}.
@@
 \end{align}
-The average regret in \eqref{eq:ave} is called \textit{directed information} (DI) and will be our measure of causation in this work.
+The average regret in \eqref{eq:ave} is called directed information (DI) and will be our measure of causation in this work.
 This measure is always positive and if it is zero, it indicates that the history of time series $X$ contains no significant information that would help in predicting the future of time series $Y$ given the history of $Y$ and $Z$.
@@
 Similarly, the children set of node $R_j$ is given by $\mathcal{CH}_j:=\{R_k: d_{k,j}=1\}$.
-Next example demonstrates the DIG of a simple linear system.
+Example \ref{r1} demonstrates the DIG of a simple linear system.
 \begin{example}\label{r1}
@@
 where $N_X$, $N_Y$, and $N_Z$ are three independent stationary Gaussian processes with zero mean and a diagonal covariance matrix (1, 0.9, 1). 
-Since the dynamic is linear and the exogenous noises are Gaussian, we can compute the DIs using the following expression\footnote{Equation \eqref{simueq} does not hold in general setting.}\citet{acc2014}.
+Since the dynamic is linear and the exogenous noises are Gaussian, we can compute the DIs using the following expression\footnote{Equation \eqref{simueq} does not hold in more general settings.}(see \citet{acc2014}).
 \begin{equation}\label{simueq}
@@
 \end{equation}
-where $|\Sigma_{Y_{t-1},Y_{t},X_{t-1}}|$ denotes the determinant of the covariance matrix of $\{Y_{t-1},Y_{t},X_{t-1}\}$. Using \eqref{simueq}, we computed the DIs of this system,
-\begin{figure}
+where $|\Sigma_{Y_{t-1},Y_{t},X_{t-1}}|$ denotes the determinant of the covariance matrix of $\{Y_{t-1},Y_{t},X_{t-1}\}$. Using \eqref{simueq}, we compute the DIs of this system,
+\begin{small}
+\begin{align}  \label{di_xyz}
+&I(Y\rightarrow X||Z)=0,\ \ I(Z\rightarrow X||Y)=0,\ \ I(Z\rightarrow Y||X)=0,\ \ I(Z\rightarrow Z||X,Y)=0,\\
+&I(X\rightarrow Z||Y)=0,\ \ \ \ I(X\rightarrow Y||Z)\approx 0.1, \ \ \ I(Y\rightarrow Z||X)\approx 0.03. \nonumber
+\end{align}
+\end{small}
+Figure \ref{fig:example_dig} illustrates the DIG of the system \eqref{model1}. In this example, $\mathcal{PA}_Z=\{Y\}$ and $\mathcal{CH}_Z=\{\}$.
+\begin{figure}[h]
 \hspace{4cm}
@@
 }
-   \mycaption{DIG of the system in \eqref{model1}}{}
+\mycaption{DIG estimated from the directed informations \eqref{di_xyz}}{X influences Y, Y influences Z. This DIG correctly estimates the dynamics of the system in \eqref{model1}.}
 \label{fig:example_dig}
 \end{figure}
-\begin{small}
-\begin{align*}
-&I(Y\rightarrow X||Z)=0,\ \ I(Z\rightarrow X||Y)=0,\ \ I(Z\rightarrow Y||X)=0,\ \ I(Z\rightarrow Z||X,Y)=0,\\
-&I(X\rightarrow Z||Y)=0,\ \ \ \ I(X\rightarrow Y||Z)\approx 0.1, \ \ \ I(Y\rightarrow Z||X)\approx 0.03.
-\end{align*}
-\end{small}
-Figure \ref{fig:example_dig} illustrates the DIG of this system. In this example, $\mathcal{PA}_Z=\{Y\}$ and $\mathcal{CH}_Z=\{\}$.
 \end{example}
@@
 
+\clearpage
+
 \begin{remark}
-
 A causal model allows a factorization of the joint density function in some specific ways.
@@
 Such factorization is called a generative model. 
-
 \end{remark}
@@
 On the other hand, estimating DI  requires estimating all the expectation terms in (\ref{cdif}).
-In information theory this expectation is known as \textit{conditional mutual information}\footnote{For more details see \citet{cover2012elements}.}, i.e., 
+In information theory this expectation is known as \textit{conditional mutual information}\footnote{See \citet{cover2012elements} for more details.}, i.e., 
 \begin{align}
@@
 samples of the time series such as plug-in empirical estimator and k-nearest neighbor estimator.
-For an overview of such estimators see the Appendix and the articles in \citet{paninski2003estimation}, \citet{noshad2019scalable}, and \citet{jiao2013universal}.
+For an overview of such estimators see Appendix \ref{appendix2} and the articles in \citet{paninski2003estimation}, \citet{noshad2019scalable}, and \citet{jiao2013universal}.
 
@@
 %More precisely, we estimated each of the above conditional mutual information using the $k$-nearest method of \citet{sricharan2011k}. 
-For the sake of completeness, we describe the steps of this method in Appendix.
+For the sake of completeness, we describe the steps of this method in Appendix \ref{appendix2}.
 
-Side information can also help to infer the DIG of a dynamical system without directly estimating the DIs but instead it provides an alternative approach to identify the DIG. 
+Side information can also help to infer the DIG of a dynamical system without directly estimating the DIs but instead providing an alternative approach to identify the DIG. 
 For example, if it is given that the underlying dynamic is linear, i.e., $\textbf{X}_t = \textbf{AX}_{t-1} + \textbf{N}_t$, then it has been shown in \citet{acc2014} that the support\footnote{The support of a matrix $\textbf{B}=[b_{i,j}]$ is a binary matrix of the same dimension as $\textbf{B}$ such that its entry $(i,j)$ is one if and only if $b_{i,j}\neq0$.} of the coefficient matrix $\textbf{A}$ is equal to the adjacency matrix of its corresponding DIG.
@@
 The main reason is that without any side information, estimating the DI has high computational and sample complexity. 
-Furthermore, the estimating complexity of DI increase with the dimension of the network.
+Furthermore, the estimating complexity of DI increases with the dimension of the network.
 This is due to the fact that the DI in \eqref{cdif} measures the influence from $R_i$ to $R_j$ by conditioning on the information from the remaining network $\mathcal{R}_{-\{i,j\}}$. 
@@
 
-The authors in \citet{quinn2015bounded} presented a more generalized version of the above approximation in which they identify the optimal connected bounded in-degree\footnote{Connected bounded in-degree graphs with bound $k$ are connected directed graphs in which each node has at most $k$ number of parents.} approximations.
+The authors in \citet{quinn2015bounded} present a more generalized version of the above approximation in which they identify the optimal connected bounded in-degree\footnote{Connected bounded in-degree graphs with bound $k$ are connected directed graphs in which each node has at most $k$ number of parents.} approximations.
 This method reduces the size of the conditioning set in \eqref{cdif} to some constant value (bound of the in-degrees) which is independent of the network size. 
@@
 
-In order to present our method, we need the following preliminary result that characterizes an important property of DI in \eqref{cdif}. All the proofs are presented in Appendix A.
+In order to present our method, we need the following preliminary result that characterizes an important property of DI in \eqref{cdif}. All the proofs are presented in the Appendix \ref{appendix2}.
 
@@
 %Afterward, we propose two approaches for two sub-classes of \eqref{nonlinear} to obtain this times series.
+\clearpage
 \begin{lemma}\label{lemma:nonlinear}
@@
 
-\paragraph{Koopman-based lifting technique}
+\subsubsection{Koopman-based lifting technique}
 Consider a particular sub-class of the non-linear system in \eqref{nonlinear} whose dynamic is defined by
@@
 However, in this formulation, the weights $\{w_{,k}\}$ are unknown. 
-An approach to obtain the weights is a non-linear filtering technique known as Koopman-based lifting \citet{koopman1931hamiltonian}. 
+An approach to obtain the weights is a non-linear filtering technique known as Koopman-based lifting (see \citet{koopman1931hamiltonian}). 
 This technique takes observational data and a set of library functions as inputs and obtains the unknown coefficients $\{w_{,k}\}$.
-The main steps of this technique are; transforming the data (lifting the data),  applying a linear identification on the lifted data, and finally applying another transformation to bring down the results into the original vector field.
+The main steps of this technique are transforming the data (lifting the data),  applying a linear identification on the lifted data, and finally applying another transformation to bring down the results into the original vector field.
 Figure \ref{fig:koopman} illustrates the main steps.
-For more details see Appendix and \citet{mauroy2019koopman}.
-\begin{figure}
+For more details see Appendix \ref{appendix2} and \citet{mauroy2019koopman}.
+\begin{figure}[h]
 \centering
 \includegraphics[scale=.4]{koopman.pdf}
-\mycaption{Koopman lifting technique compared to classical non-linear identification}{}
+\mycaption{Koopman lifting technique compared to classical non-linear identification}{Starting from the bottom left plot, the main steps of this technique are transforming the data (lifting the data),  applying a linear identification on the lifted data, and finally applying another transformation to bring down the results into the original vector field.}
 \label{fig:koopman}
@@
 First, the Koopman's performance depends on the choice of the library functions and second, it often fails to estimate the real time series $Q$.
-More precisely, this technique involves the computation of matrix $\textbf{L}:=log(\textbf{P}_x^{\dagger}\textbf{P}_y)/T_s$, where $\textbf{P}_x$ and $\textbf{P}_y$ are estimated from the observational data\footnote{See Appendix for more details.}.
+More precisely, this technique involves the computation of matrix $\textbf{L}:=log(\textbf{P}_x^{\dagger}\textbf{P}_y)/T_s$, where $\textbf{P}_x$ and $\textbf{P}_y$ are estimated from the observational data\footnote{See Appendix \ref{appendix2} for more details.}.
 Matrix $\textbf{P}_x^{\dagger}$ denotes the pseudo-inverse, and the function $\log(\cdot)$ denotes the (principal) matrix logarithm.
@@
 
-\paragraph{RNNs method}
- Recurrent neural networks are a specific class of Neural Networks well suited to learn time series. 
+\subsubsection{RNNs method}
+ Recurrent neural networks are a specific class of neural networks well suited to learn time series. 
  They are distinguished by their memory as they are able to remember information from prior inputs to influence their current outputs.
@@
 
- \begin{algorithm}[t]
+ Given the aforementioned result, we train a RNN using the observational data to estimate the time series $Q$ defined in Lemma \ref{lemma:nonlinear}.
+ More precisely, our RNN maps $\mathcal{X}_{-\{i\}}^{t-1}$ as the inputs to $X_{j,t}$ as the output. 
+ Let $R_j(\mathcal{X}_{-\{i\}}^{t-1};\Theta^*)$ denotes the trained RNN with parameters $\Theta^*$. 
+ In this case, the time series $Q$ can be written as  $Q_{t-1}=R_{j}(\mathcal{X}_{-\{i\}}^{t-1};\Theta^*)$.
+ Finally, we use \eqref{eq:non-q} to detect whether $X_i$ has influence on $X_j$ or not. 
+ Algorithm \ref{alg}  summarizes the steps of our RNN method.
+
+ \begin{algorithm}[h]
  \textbf{Input:} Observational data of $m$ time series up to time $T$, $\mathcal{X}^T$, Threshold $\alpha>0$\;
@@
  \For{ $i,j=1,...,m$}{
-  Train an RNN $R_j(\cdot;\Theta^*)$ that maps $\mathcal{X}_{-\{i\}}^{t-1}$ to $X_{j,t}$\; 
+  Train a RNN $R_j(\cdot;\Theta^*)$ that maps $\mathcal{X}_{-\{i\}}^{t-1}$ to $X_{j,t}$\; 
   Define $Q_{t-1}=R_{j}(\mathcal{X}_{-\{i\}}^{t-1};\Theta^*)$\;
@@
 \end{algorithm}
- Given the aforementioned result, we train an RNN using the observational data to estimate the time series $Q$ defined in Lemma \ref{lemma:nonlinear}.
- More precisely, our RNN maps $\mathcal{X}_{-\{i\}}^{t-1}$ as the inputs to $X_{j,t}$ as the output. 
- Let $R_j(\mathcal{X}_{-\{i\}}^{t-1};\Theta^*)$ denotes the trained RNN with parameters $\Theta^*$. 
- In this case, the time series $Q$ can be written as  $Q_{t-1}=R_{j}(\mathcal{X}_{-\{i\}}^{t-1};\Theta^*)$.
- Finally, we use \eqref{eq:non-q} to detect whether $X_i$ has influence on $X_j$ or not. 
- Algorithm Infer-DIG in \ref{alg} summarizes the steps of our RNN method.
 
@@
  \subsection{Linear Gaussian Framework} \label{linfr}
- In this experiment, we consider a linear system, a VAR(1) model whose dynamic is given by 
+ In this experiment, we consider a linear system, a VAR(1) model whose dynamics are given by 
 \begin{equation}\label{eq:ex_lin}
@@
 where $1_{a>b}$ denotes the indicator function which is equal to 1 when $a>b$ and $0$ otherwise and $\underline{\epsilon}$ and $\bar{\epsilon}$ are thresholds to define non-zero entries in the upper-left and the lower part of $\textbf{A}$, respectively.
+
 Figure \ref{fig:matrix} illustrates the structure of the resulting $\textbf{A}$.
@@
 In our experiment, we select $(s,m)=(85,100)$ and $(\underline{\epsilon},\bar{\epsilon})=(0.4,0.7)$. 
-%further results are robust to changes of parameters.
+Finally, to guarantee the stability of the time series, we rescale\footnote{Formally, we use $\textbf{A}/(\rho(\textbf{A}) + \epsilon)$, where $0<\epsilon$.} $\textbf{A}$ such that its spectral radius is strictly less than one, i.e.,  $\rho(\textbf{A})< 1$.
+Once the matrix $\textbf{A}$ is defined, we simulate the time series using \eqref{eq:ex_lin} for a period of $T = 30000$ and use the resulting data for our estimations.
+
 \begin{figure}
     \centering
-    \includegraphics[scale=.2]{fig_matrix.pdf}
-    \mycaption{Structure of matrix $\textbf{A}$ in \eqref{eq:ex_lin} that is build using \eqref{eq:rules}.}{}
+    \includegraphics[scale=0.2]{fig_matrix.pdf}
+    \mycaption{Adjency matrix A for the linear Gaussian framework}{Structure of matrix A in \eqref{eq:ex_lin} that is built using \eqref{eq:rules}.}
     \label{fig:matrix}
 \end{figure}
-Finally, to guarantee the stability of the time series, we rescale\footnote{Formally, we use $\textbf{A}/(\rho(\textbf{A}) + \epsilon)$, where $0<\epsilon$.} $\textbf{A}$ such that its spectral radius is strictly less than one, i.e.,  $\rho(\textbf{A})< 1$.
-Once the matrix $\textbf{A}$ is defined, we simulate the time series using \eqref{eq:ex_lin} for a period of $T = 30000$ and use the resulting data for our estimations.
 
@@
 \end{align}
-where $\widehat{I}(\cdot\rightarrow\cdot||\cdot)$ denotes the estimated DI from the data. 
+where $\widehat{I}(X_i\rightarrow X_j||\mathcal{C}_j)$ denotes the estimated DI from $X_i$ to $X_j$ given the conditioning set $\mathcal{C}_j$. 
 In order to compare the performances of the aforementioned four approaches, we use the precision and recall measure between the true DIG (obtained from $\textbf{A}$) and their estimated DIGs. 
@@
 
-\begin{figure*}
+\begin{figure*}[h]
         \centering
@@
         \end{subfigure}
-        \mycaption{Precision and recall curves in the linear framework.}{Precision and recall curves for the \textit{True parents}, \textit{Most correlated}, \textit{Ideal portfolio}, and the \textit{RNN}, respectively.}
+        \mycaption{Precision and recall curves in the linear framework.}{Precision and recall curves for the \textit{true parents}, \textit{most correlated}, \textit{ideal portfolio}, and the \textit{RNN}, respectively.}
         \label{perflin}
@@
 On the other hand, we observe high accuracy from the \textit{RNN} approach which is a striking result.
-This result is an evidence that an RNN is capable of estimating the ideal portfolio, i.e., the time series $Q$ in Lemma \ref{lemma:lin} without any side information about the underlying model.
+This result is an evidence that a RNN is capable of estimating the ideal portfolio, i.e., the time series $Q$ in Lemma \ref{lemma:lin} without any side information about the underlying model.
 
@@
 Similar to the linear setting, we use the \textit{true parents} as a benchmark since it has the ideal performance. 
-%Conditioning on the true parents obviously performs the best as all the information is included in the Directed Information. 
-%However, it is also the heaviest in terms of computational time. 
 It is however important to emphasize that this conditioning approach has higher complexity compared to the others. 
@@
 With this method, we could slightly reduce the estimation complexity of the DIs compared to the  \textit{true parent} approach but this comes with the price of losing the performance.
-clearly, the performance of the \textit{most correlated} approach can be improved by increasing $k$ but this will increase the complexity.
+Clearly, the performance of the \textit{most correlated} approach can be improved by increasing $k$ but this will increase the complexity.
 
@@
 Finally, as it is shown in Figure \ref{fig:perfnonlin}, the \textit{RNN} approach outperforms the \textit{most correlated} and the \textit{ideal portfolio} approaches and it shows close performance to the \textit{true parents} but with the size of the conditioning set equal to one.
-This result once more fortifies our claim that with an RNN we can summarize the information of the network into one time series and use it for detecting the causal relationships. 
+This result once more fortifies our claim that with a RNN we can summarize the information of the network into one time series and use it for detecting the causal relationships. 
 This claim is due to Lemma \ref{lemma:nonlinear} and the universal approximation theorem which states that a neural network with enough hidden layers can approximate any non-linear function  \citet{HORNIK1989359}.
-The slight difference between the performance of the \textit{RNN} and the \textit{true Parents} is because of the estimation error in the recurrent neural network.
+The slight difference between the performance of the \textit{RNN} and the \textit{true parents} is because of the estimation error in the recurrent neural network.
 
@@
 
-\begin{figure}
+\begin{figure}[h]
 \centering
-\includegraphics[width=\textwidth]{precrecall_15assets.jpg}
-\mycaption{Precision-Recall curves for the quadratic model.}{}
+\includegraphics[width=0.75\textwidth]{precrecall_15assets.jpg}
+\mycaption{Precision-recall curves for the quadratic model}{Blue, green, yellow and red lines show the precision-recall curve of the \textit{true parents}, \textit{most correlated}, \textit{ideal portfolio}, and the \textit{RNN}, respectively.}
 \label{fig:perfnonlin}
@@
 
-This section describes how to apply our approach to empirical data and obtain the DIG of some US firms. 
+This section describes how to apply our approach to empirical data and obtain the DIG of US firms. 
 We extracted the daily stock prices and the daily US Treasury rate as risk free returns from the CRSP database from 1990 to 2020. 
@@
 
-Table \ref{GrangerCausality} shows the Degree of Granger Causality (DGC) defined as the fraction of relationships in the network among all potential relationships. Formally,
+Table \ref{GrangerCausality} shows the Degree of Granger-Causality (DGC) defined as the fraction of relationships in the network among all potential relationships. Formally,
 \begin{equation}
@@
 \begin{table}[]
+\centering
 \begin{tabular}{cccccc}
@@
 
-In this work, we introduce an information-theoretic measure known as directed information that is capable of capturing nonlinear Granger-causality in an interactive system. 
+In this chapter, we introduce an information-theoretic measure known as directed information that is capable of capturing nonlinear Granger-causality in an interactive system. 
 We develop a novel algorithm based on recurrent neural network utilized with directed information. 

### 2022-07-24 main/paper3.tex
Add files via upload

--- 2022-05-24 main/paper3.tex
@@
 \section{Introduction}
-\label{S:1}
+\label{S:3.1}
 
-Recent developments in artificial intelligence and the growing amount of alternative data have created new areas of research in finance.
-In particular, often coming from news, social media or annual reports, textual data is increasingly used in the literature.
-\citet{nikfarjam2010text} documents recent text mining approaches for stock market prediction.
-The pioneer work from \citet{antweiler2004all} computes a bullishness measure out of 1.5 million messages posted on Yahoo!
+Can the stock market be predicted by analyzing social media?
+Recent developments in machine learning and the growing quantities of available text data from online news, social media and annual reports have triggered intensive research in finance.
+In their pioneering paper, \citet{antweiler2004all} compute a bullishness measure out of 1.5 million messages posted on Yahoo!
 Finance and Raging Bull and finds that stock messages help predict market volatility.
 Their results clearly reject the hypothesis that all that talk is just noise.
-They show that there is financially relevant information present.
-Growing from that, there are now two main sides of sentiment analysis literature studying the forecasting abilities of natural language processing (NLP hereafter): text mining of annual reports and sentiment analysis using social media.
-On one side, \citet{shirata2011extracting} and \citet{cecchini2010making} report that extracting phrases from annual reports may be an effective predictor of corporate bankruptcy.
-Similarly, \citet{loughran2011barron} identify phrases that might be red flags indicating questionable behavior.
-On the other side, \citet{renault2017intraday} builds an intraday investor sentiment indicator using messages and finds that the change in investor sentiment of the first half-hour of a trading day helps forecast the last half-hour market return of that trading day.
+They show that there is financially relevant information present in social media.
+In a similar vein, \citet{tet_07} constructs a measure of media pessimism from a Wall Street Journal column and finds that it predicts downward pressure on market prices.
 
-We conjecture that our investor sentiment measure can serve as a proxy of unobservable firm fundamentals.
-For instance, misaligned managerial and shareholder incentives are not easy to observe quantitatively in firm's annual reports (see \citet{nikolov2014agency}) but analysts may talk about them freely in bearish messages.
-This flow of data can be used throughout a period to improve forecasts (see \textit{nowcasting} in \citet{challet2013predicting}).
-\citet{tetlock2008more} uses Wall Street Journal stories to examine whether the usage of language is able to predict individual firms’ accounting earnings and stock returns.
-They find that some aspects of firms' fundamentals are hard-to-quantify, but investors may use linguistic media content to capture information and incorporate it into stock prices.
+Most of the previous financial studies of social media rely on pre-defined or manually annotated sentiment dictionaries.
+Such approaches are limited in various ways.
+How to create a sentiment classifier that understands the vocabulary of the messages posted by the investors?
+For instance, ``bull" is an animal in everyday language but it is someone optimistic in the financial jargon.
+\citet{loughran2011liability} create a word list, which helps classify tone in a financial document.
+However, this might not be sufficient in the context of social media because messages posted present many typos, abbreviations and slang, so one needs to have an additional layer of data preprocessing.
+For instance, the word ``goooooood" would not be recognized by the model if it is not corrected into ``good" first.
+On the other hand, manually annotating and validating dictionaries is not a scalable approach to handling social media content.
 
-Most papers studying social media's predictive power use Twitter as their primary source of data.
-Twitter has the advantage of being used by a wide range of people across the world and a few influencers can attract attention of many investors.
-In 2013, Carl Icahn tweeted that following a meeting with Tim Cook (Apple CEO), he bought a large position in Apple and believed that the company is extremely undervalued.
+Our paper overcomes these limitations.
+We develop a machine learning algorithm to classify the sentiment of a large sample of StockTwits messages as bullish, bearish, or neutral.
+The sample consists of all messages referring to US and Canadian stocks, including ETFs and other types of securities available on CRSP/Compustat, from January 2010 to March 2020.
+We train our machine learning classifier on the set of all user sentiment-labeled messages, which constitute about one third of the sample.
+We then classify the sentiment of all remaining messages.
+Our method scales and performs very well.
+It achieves an out-of-sample accuracy of 85.9\%, which compares well to the anecdotal 80 to 85\% probability that human annotators agree on the sentiment of a document, see, e.g., \citet{wilson2005recognizing} and \citet{chen2020large}.
+As a side product, we generate a vocabulary of one million investor sentiment-labeled terms consisting of up to three words that frequently appear in StockTwits messages.
+
+We then construct a stock-aggregate daily sentiment polarity measure and relate it to daily stock returns.
+We find that polarity is positively associated with contemporaneous returns.
+However, unconditionally, polarity cannot predict next-day returns, which is in line with the efficient market hypothesis (EMH).
+We then conduct an event study.
+We define events as days of sudden peaks of message volume of individual tickers.
+We classify events as bullish, bearish, or neutral depending on the prevailing polarities.
+We find that bullish (bearish) events are strongly associated with large positive (negative) abnormal returns.
+Cumulative abnormal returns over the preceding 20 days of an event have no predictive power on the type of event.
+Returns normalize immediately after the jump on the event date, which again is in line with the EMH.
+In contrast, remarkably, we find that cumulative abnormal polarity has statistically significant predictive power on the type of event.
+We assess the economic relevance of our findings with the performance of cumulative abnormal polarity ranked portfolios.
+We find that for appropriate choices of thresholds, cumulative abnormal polarities provide valuable signals for stock market investments.
+
+As a technical byproduct, we develop a sentiment classifier of micro-blogs for imbalanced data.
+This addresses the stylized fact that bloggers post more bullish than bearish-labeled messages.
+In our sample, the ratio is five to one.
+On the other hand, we find that not all messages carry a substantial stock market relevant sentiment.
+Rather than re-sampling from the underrepresented bearish class, we thus introduce an auxiliary neutral class.
+We then run two independent binary classifiers.
+The first (second) classifies messages as bullish versus non-bullish (bearish versus non-bearish).
+We aggregate the two binary outcomes and classify a message as bullish (bearish) for the concordant combination bullish/non-bearish (non-bullish/bearish), and neutral otherwise.
+This approach is very simple and efficient, and eliminates the class imbalance bias at the same time.
+It builds on any traditional binary classifiers.
+We use logistic regression on Term Frequency-Inverse Document Frequency (TFIDF)-vectorized messages.
+TFIDF is a weighting scheme gauging the importance of a word in a document.
+
+Our paper contributes to the growing literature on machine learning classification of social media and its interaction with the stock market.
+Most previous financial studies use Twitter as their primary source of data.
+Twitter has the advantage of being used by a wide range of people across the world and a few influencers can attract the attention of many investors.
+In 2013, following a meeting with Tim Cook (Apple CEO), Carl Icahn tweeted that he bought a large position in Apple and believed that the company is extremely undervalued.
 This bullish tweet caused the market capitalization of Apple to jump by \$12 billion.
-In 2019, JPMorgan has created the \textit{Volfefe Index} to track Donald Trump's tweets impact on the stock market.
+In 2019, JPMorgan has created the Volfefe Index to track Donald Trump's tweets impact on the stock market.
 However, it is more difficult to disentangle noise from relevant tweets in Twitter than in other more focused social media.
 Results from \citet{Ghoshal} show that StockTwits is significantly more informative than Twitter data.
-This is not surprising as StockTwits is a finance-only social media whereas Twitter also captures irrelevant opinions on a wide range of non-finance related matters.
+This is not surprising as StockTwits is a finance-focused platform whereas Twitter also captures irrelevant opinions on a wide range of non-finance related matters.
 
-To the best of our knowledge, our paper is the first that compares predictive power of messages in general and around specific events.
-\citet{ranco2015effects} is likely the closest paper to our study.
-However, the finance-tailored data we use allow us to get higher contemporaneous correlations between stock returns and polarity.
-We also use much more messages (90 million versus 1 million) during a longer period (10 years versus 13 months).
-We also add contribution by looking at the predictive power of cumulative average abnormal polarity (CAAP hereafter).
-Our paper also contributes to the Efficient Market Hypothesis (EMH hereafter) literature by gauging how cumulative average abnormal returns (CAAR hereafter) and CAAP behave around sudden peaks of message activity.
-We collect 90 million messages published on StockTwits from mid-2010 to the end of March 2020.
-On this microblogging platform, users are invited to identify firms with their ticker when they share opinions.
-Another useful feature on the platform is the possibility to explicitly label their own messages as \textit{bearish} or \textit{bullish} when users post them.
-We believe that messages on the platform are reliable for several reasons.
-First, users have incentives to publish valuable information in order to maintain or increase mentions and/or retweets and thus have a greater share of voice in the forum (\citet{sprenger2014tweets}).
-In addition, market manipulations happen rarely because malicious users have incentives to post fake news only if they previously traded in the same direction than the news they are creating, which will only benefit them if they already have influence on the platform.
-Finally, SEC closely monitors large influencers to prevent any market abuse.
+Our paper is the first that analyzes the predictive power of StockTwits messages on stock returns unconditionally and around specific events.
+\citet{renault2017intraday} builds an intraday investor sentiment indicator using messages and finds that the change in investor sentiment of the first half-hour of a trading day helps forecast the last half-hour market return of that trading day.
+However, his classifier is based on a dictionary consisting of 8 thousand manually validated and modified terms, which limits its scalability.
+\citet{renault2020sentiment} uses larger data sets and compares various classifiers, including machine learning.
 
-The challenge in this context is to create a classifier that understands the vocabulary of the messages posted by the investors.
-For instance, ``bull" is an animal in everyday language but it is someone optimistic in the financial jargon.
-Work has already been done in this direction: \citet{loughran2011liability} creates a word list, which helps classify tone in a financial document.
-However, this might not be sufficient in the context of social media because messages posted present many typos, abbreviations and slang, so one needs to have an additional layer of data preprocessing.
-For instance, the word ``goooooood" would not be recognized by the model if it is not corrected into ``good" first.
+Our approach is in some parts similar to \citet{ranco2015effects}, who also study the relation of micro-blog sentiments with stock returns.
+However, they use Twitter data, whereas the finance-tailored StockTwits data we use results in higher contemporaneous correlations between stock returns and polarity.
+They manually annotate 100 thousand tweets, which limits the scalability of their approach.
+Our sample is much larger (90 million versus 1 million messages) and covers a longer period (10 years versus 13 months).
 
-In this paper, we use a logistic regression on Term Frequency-Inverse Document Frequency (TFIDF hereafter) vectorized labeled messages to classify unlabeled messages in either bearish, bullish or neutral class.
-TFIDF is a weighting scheme gauging the importance of a word in a document (see \citet{erdemlioglu2017market}).
-As users are prone to post more bullish messages than bearish messages, a good classifier needs to take into account the unbalanced data.
-Without resampling, the classifier outputs classification scores biased towards the over-represented class.
-Because we are creating an artificial neutral class, we chose to not resample the data but rather optimally select classification score thresholds by maximizing F1 scores of two distinct classification algorithms: bullish versus non-bullish and bearish versus non-bearish.
-Then, we aggregate messages daily for every firm to compute sentiment polarity time series for individual firms and for the whole economy.
-%\todo{we should find a concise definition of polarity to be put in the abstract and introduction} \TODO{agree.
-what about something like : degree of bullishness observed on a given firm on a given day}
-%Similarly, \citet{pang2002thumbs} suggests that ``there are certain words people tend to use to express strong sentiments, so that it might suffice to simply produce a list of such words by introspection and rely on them alone to classify the texts."
+\citet{ke_kel_xiu_20} extract sentiment from news articles on the Dow Jones Newswires.
+They train a  sentiment score directly on returns.
+In contrast, we use user sentiment-labeled StockTwits messages as training and validation set for our sentiment classifier.
 
-We then use the daily volume of messages on a given firm to identify sudden peak of activity, indicating a firm event.
-Using abnormal polarity on each event date, we are able to classify events into three classes: bullish, neutral and bearish.
-We then compute cumulative average abnormal return and cumulative average abnormal polarity in a 41 days window centered at the identified event.
-We show that abnormal polarities have significantly higher predictive power than abnormal returns.
-On average, changes of polarity are associated with changes of contemporaneous return of the same sign; but this result does not hold against next-day return.
-However, when we focus on specific events, polarity has strong predictive power.
-It is also interesting to note that polarity tends to be biased towards recent past events.
-Finally, as robustness check, our event study on CAAR is similar to previous literature and is consistent with Fama’s theory.
+Our paper also contributes to the EMH literature by gauging how cumulative average abnormal returns and abnormal polarities behave around sudden peaks of message activity.
 
 The remainder of the paper is structured as follows.
-Section \ref{S:data} presents the data.
-Section \ref{S:classification} develops the NLP logistic classifier on TFIDF vectorization.
-Section \ref{S:polarity} explains our polarity measure.
-Section \ref{S:event studies} contains the results of the event studies.
-Section \ref{S:portfolios} shows the portfolio construction and Section \ref{S:conclusion} concludes.
+Section \ref{S:data} discusses StockTwits and stock market data.
+Section \ref{S:classification} develops our sentiment classifier based on TFIDF vectorization.
+Section \ref{S:polarity} introduces the sentiment polarity measure and relates it to stock returns.
+Section \ref{S:event studies} contains the event study.
+Section \ref{S:portfolios} discusses the sentiment-sorted portfolio performance.
+Section \ref{S:conclusion} concludes.
+The appendix contains additional statistics and background material.
 
-\section{Data}
+\section{StockTwits and Stock Market Data}
 \label{S:data}
 
-Data is coming from two sources: Compustat/CRSP for the stock prices, StockTwits for the investor sentiment.
-From the CRSP database, we extract daily stock prices (closing prices), daily volume of transactions and number of shares outstanding from 2010 to 2020 for all US and Canadian firms.
-Stock prices and number of shares are adjusted to account for any distribution (i.e.
-dividends, stock splits) so that a comparison can be made on an equivalent basis before and after the distribution.
+StockTwits is a large social media platform similar to Twitter but designed for investors and traders.
+Users register online and can post messages about any listed stock through the prefix \$ followed by the ticker of the stock.
+StockTwits was created in 2008 as an app built on the Twitter's API and later detached from Twitter to build a standalone social network.
+As of April 2019, it had over two million registered users and the number of daily posted messages has been growing exponentially, see Figure~\ref{fig:tweet_vol_daily}.
 
-StockTwits is a large social network similar to Twitter but designed for investors and traders.
-Users register online and can post messages about any listed firm through the prefix \$ followed by the ticker of the firm.
-StockTwits was created in 2008 as an app built on the Twitter's API and later detached from Twitter to build a standalone social network.
-As of April 2019, it has over two million registered users and the number of messages posted is growing exponentially (see Figure \ref{fig:tweet_vol_daily}).
-StockTwits describes itself as ``the voice of social finance and the best way to find out what is happening right now in the markets and stocks you care about".
-In practice, it is effectively used by finance professionals to express their opinions on individual firms and the market as a whole.
-Since mid-2010, users also have the possibility to label their own messages as either \textit{bullish} or \textit{bearish}.
-This feature is very useful for researchers as even though many messages are unlabeled, it allows for sentiment classification using NLP techniques.
-The reasons of using StockTwits and not another social media is threefold.
-First, one of the main challenge in Natural Language Processing is the creation of an appropriate labeled vocabulary.
-\citet{loughran2011barron}) shows that it is essential to have a specific vocabulary to interpret finance documents (i.e many words have a different meaning in finance than in traditional English (e.g “bear trap”)).
-In addition to that, social media slang is an additional layer of language complexity.
-To this extent, the functionality to self-tag \textit{bullish} and \textit{bearish} messages that StockTwits implemented in 2012 is very valuable as it allows the creation of a specific labeled vocabulary out of labeled messages.
-We are not aware of any other social media platform in finance offering this functionality.
-Second, StockTwits is less noisy than Twitter because messages focus on finance and economics matters only.
-Third, extracting data out of StockTwits is easy because of its API.
-StockTwits' API is designed to query the database to download messages via JSON requests.
-Using a Python script, we extract all messages since the 10th of July 2009 until the 31st of March 2020 for a preset list of tickers coming from the Compustat database corresponding to all US and Canadian firms.
-This results in 90 million messages, which we download and store as JSON files.
-As one message may refer to several tickers, we consider a message with two or more tickers as one message for every ticker identified.
-We refer to the appendix for more information about this process.
+\clearpage
 
- \begin{figure}
+\begin{figure}[h]
     \centering
     \includegraphics[width=0.75\textwidth]{tweet_vol_daily.JPG}
-    \caption{Number of messages posted daily on StockTwits.}
+    \mycaption{Number of messages posted daily on StockTwits}{Message volume posted on StockTwits. Numbers are aggregated daily.}
     \label{fig:tweet_vol_daily}
@@
 
-Every message comes with the following eight features: (1) the ticker discussed (2) the exact timestamp of the message, (3) a unique message identifier, (4) the body of the message, (5) the sentiment label (bearish, bullish, or none) entered by the user, (6) a unique identifier of the user who sent the message, (7) the number of messages published by the user who sent the message, and (8) the number of followers of the user who sent the message.
-Figure~\ref{screenshot} shows a screenshot of the StockTwits website as of 3rd March 2020, for a query on the AAPL\footnote{AAPL is the ticker for Apple.} ticker.
-The first message is labeled as bullish by the user ``satkaru", the two next are unlabeled messages that will be classified using a machine learning classifier, and the last message is labeled as bearish by the user ``Etrading''.
+StockTwits describes itself as ``the voice of social finance and the best way to find out what is happening right now in the markets and stocks you care about''.
+In practice, it is effectively used by finance professionals to express their opinions on individual stocks and the market as a whole.
+Importantly, users have the option to label their posted messages as either bullish or bearish.\footnote{This optional label was effectively available as of mid-2010.} This feature is key for our approach, as it allows for sentiment classification of all messages using machine learning trained on the user-labeled messages.\footnote{One third of the messages in our sample have a user labeled sentiment.}
+
+The reasons for using StockTwits and not other social media data for financial studies are at least threefold.
+First, a major challenge in applying natural language processing is the creation of an appropriate labeled vocabulary.
+\citet{loughran2011barron} show that it is essential to have a specific vocabulary to interpret finance documents (i.e., many words have a different meaning in finance than in traditional English, such as “bear trap”).
+In addition to that, social media slang is an additional layer of language complexity.
+To this extent, the functionality to self-tag bullish and bearish messages that StockTwits is extremely valuable as it allows the creation of a specific labeled vocabulary out of labeled messages.
+We are not aware of any other social media platform in finance offering this functionality.
+Second, text data from StockTwits is more reliable and less noisy than from general purpose platforms, such as Twitter, because messages focus on finance and economics matters only.
+Micro-bloggers have incentives to post valuable information in order to maintain or increase mentions and retweets, and thus have a greater share of voice in the forum (\citet{sprenger2014tweets}).
+On the other hand, StockTwits messages might be biased and subject to malicious users that try to manipulate the market.
+However, market manipulations likely happen only rarely as the SEC closely monitors potential influencers to prevent any market abuse.
+Third, extracting data from StockTwits is easy because of its API.
+StockTwits' API is designed to query the database to download messages via JSON requests.
+We provide a short tutorial in Appendix~\ref{app_tut}.
+
+We use stock market data from CRSP/Compustat.
+We extract daily closing prices, daily volume of transactions and number of shares outstanding for all US and Canadian stocks, as well as ETFs and some other types of securities, from January 2010 to March 2020.
+Stock prices and number of shares are adjusted to account for any distribution (i.e., dividends, stock splits) so that a comparison can be made on an equivalent basis before and after the distribution.
+We use as risk-free rate the 3-month US T-bill rate, converted into daily risk-free returns.
+We henceforth refer to daily stock excess returns over risk-free simply as returns.
+Using a Python script, we then extract all messages from StockTwits for the list of tickers corresponding to the sample of US and Canadian stocks.
+This results in 90 million messages, which we download and store as JSON files.\footnote{A message may refer to multiple tickers.
+We count any such message towards any ticker that it refers to.
+We give more information about this double counting in Appendix \ref{app_counting}.} Overall, our sample covers 8843 tickers, whereof 75\% refer to ordinary common share, 15\% to ETFs, and the remaining 10\% to other types of securities.
+Henceforth, we interchangeably refer to any of these securities as either a stock or a ticker.
+
+Every StockTwits message includes eight features: (1) the reference ticker(s), (2) a timestamp, (3) a unique message identifier, (4) the body of the message, (5) the sentiment label (bearish, bullish, or none) entered by the user, (6) a unique identifier of the user who posted the message, (7) the number of messages published by the user who posted the message, and (8) the number of followers of the user who posted the message.
+Our sentiment analyses builds on the first five features.
+The last three provide additional information on the network structure, which we briefly discuss in Appendix \ref{app_userstats}.
+
+Figure~\ref{screenshot} shows a screenshot of the StockTwits website as of 3rd March 2020, for a query of AAPL, which is the ticker for Apple.
+The first message is labeled as bullish by the user ``satkaru", the two next are unlabeled messages that will be classified by our machine learning algorithm, and the last message is labeled as bearish by the user ``Etrading''.
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.75\textwidth]{stocktwits_screeshot_3march2020.JPG}
+\mycaption{Screenshot of messages posted on StockTwits}{Screenshot is from the 3rd of March 2020 for a query on AAPL, the ticker of Apple.}
+\label{screenshot}
+\end{figure}
+
+Left plot of Figure \ref{fig:data-stocktwits2} shows the top 30 most discussed tickers on StockTwits.
+SPY, a large ETF that tracks the S\&P 500 stock market index, is the most discussed ticker, followed by Apple and other big tickers.
+The messages about the 15 (30) biggest tickers represent 20\% (25\%) of the total number of messages, which indicates that users talk about a wide panel of tickers and not only big firms.
+Right graph shows a histogram of the number of messages per ticker.
+The x-axis is log-scaled because due to extreme values the distribution is highly skewed.
 
@@
 \centering
-\includegraphics[width=0.75\textwidth]{stocktwits_screeshot_3march2020.JPG}
-\caption{Screenshot of StockTwits as of 3rd March 2020.
-The first message is labeled as bullish by the user ``satkaru", the two next are unlabeled messages that will be classified using a machine learning classifier and the last message is labeled as bearish by the user ``Etrading''.}
-\label{screenshot}
-\end{figure*}
-
-Figure \ref{categrepartition-preclassif} shows the distribution of user-labeled and unlabeled messages.
-Overall, around 30 million messages are user-labeled and 60 million messages are unlabeled.
-Among the user-labeled messages we find five times more bullish than bearish ones.
-This ratio indicates that investors are on average optimistic about the market, which is consistent with findings in the literature, e.g., \citet{renault2017intraday}.
-Such class imbalance is a well-known issue in machine learning classification, which we address below.
-We will classify the unlabeled messages using a machine learning algorithm trained on the set of user-labeled messages.
-Since not every message contains substantial information, we believe that the sentiment classification should not be a bullish/bearish dichotomy.
-Hence, we allow for a neutral class to account for messages that do not take a clear stand.
-
-\begin{figure*}[h]
-\centering
-        \begin{subfigure}[b]{0.75\textwidth}   
-            \centering 
-            \includegraphics[width=\textwidth]{preclassif.JPG}
-        \end{subfigure}
-        \caption{Number of user-labeled messages in each category. There are around 60 million unlabeled messages that we classify using the 30 million user-labeled bullish and bearish messages.} 
-        \label{categrepartition-preclassif}
-\end{figure*}
-
-Figure \ref{fig:data-stocktwits1} shows a log-log histogram of the number of followers per user and a log-log histogram of the number of messages posted by users.
-As expected, there are few users with many followers (they can be seen as ``influencers"), and many users with a few followers.
-In addition, most users seem to post on average between 10 and 400 messages and a few post a lot more.
-
-\begin{figure*}[]
-\centering
-        \begin{subfigure}{0.75\textwidth}   
-            \centering 
-            \includegraphics[width=\textwidth]{Followers.JPG}
-        \end{subfigure}
-        \quad
-        \begin{subfigure}{0.75\textwidth}
-            \centering
-            \includegraphics[width=\textwidth]{Tweets.jpg}
-        \end{subfigure}
-
-        \caption{User summary statistics.
-Top graph is a log-log histogram of the number of followers per user and the bottom graph shows the log-log histogram of the number of messages posted by users.}
-        \label{fig:data-stocktwits1}
-\end{figure*}
-
-Figure \ref{fig:data-stocktwits2} shows the top 30 most discussed tickers on the platform.
-Of all tickers extracted, around 75\% are ordinary common share, 15\% are ETFs and the remaining 10\% are other types of securities.
-Not surprisingly, the S\&P index is the most discussed, followed by Apple and other big tickers.
-The messages about the 15 (30) biggest tickers represent 20\% (25\%) of the total number of messages, which indicates that users talk about a wide panel of tickers and not only big firms.
-The bottom graph shows a histogram of the number of messages per ticker.
-The x-axis is log-scaled because of extreme values, the distribution is highly skewed.
-
-\begin{figure*}[]
-\centering
-\begin{subfigure}{0.75\textwidth}  
+\begin{subfigure}{0.49\textwidth}  
             \centering 
@@
         \end{subfigure}
-        \quad
-        \begin{subfigure}{0.75\textwidth}   
+        \begin{subfigure}{0.49\textwidth}   
             \centering 
@@
         \end{subfigure}
-        \caption{Firm summary statistics.
-Top graph shows the top 30 most discussed tickers on the platform.
-SPY is the ticker of the S\&P and AAPL is the ticker for Apple.
-Bottom graph shows the distribution of the number of messages across tickers.}
+        \mycaption{Ticker summary statistics}{
+        Left graph shows the top 30 most discussed tickers on StockTwits.
+SPY is the ticker of a large ETF tracking the S\&P 500 and AAPL is the ticker for Apple.
+Right graph shows the distribution of the number of messages across tickers.}
         \label{fig:data-stocktwits2}
@@
 
-Messages contain qualitative information, which needs to be transformed into quantitative data for the computer to understand.
-Thereto, we first apply the following preprocessing operations: an apostrophe handler, a contraction form handler (i.e.
-``aren't" becomes ``are not"), tickers removal, stop words (i.e.
-``a", ``the", ``of") removal\footnote{We follow \citet{renault2020sentiment} and \citet{saif2014stopwords} and use a restrictive list of stopwords to avoid accuracy decrease.}, users removal, lemmatization, URLs removal and a simple spell corrector dealing with more than two repeated characters (i.e.
-``soooo goooood" becomes ``soo good").
+Text messages need to be transformed into a quantitative vector to be fed into our machine learning classifier, which in turn computes a sentiment score.
+This transformation consists of several steps.
+First, we apply some preprocessing operations to the text messages: an apostrophe handler, a contraction form handler (e.g., ``aren't" becomes ``are not"), tickers removal, stop words removal (e.g., ``a", ``the", ``of")\footnote{We follow \citet{renault2020sentiment} and \citet{saif2014stopwords} and use a restrictive list of stopwords to avoid accuracy decrease.}, users removal, lemmatization, URLs removal and a simple spell corrector dealing with more than two repeated characters (e.g., ``soooo goooood" becomes ``soo good").
 Table \ref{tab:preprocess} shows five examples of messages before and after preprocessing.
 
-One of the first steps in NLP is the tokenization : the way of slicing a piece of text in smaller units called tokens or terms.
-In financial lingo, some words only have meaning when associated with other words (i.e ``bad apple" or ``bear flag").
-N-gram models allow accounting for words frequently occurring together with other words.
-The main hyperparameter in N-gram models is the size of the group of words considered : unigram is a term with only one word, bigram is a term with two consecutive words, etc.
-Bigger N-grams models increase dramatically the size of the vocabulary (i.e : the collection of all terms considered).
-We are able to control the size of the vocabulary by tuning a hyperparameter keeping only a given number of most frequent terms.
-We chose to work with a vocabulary consisting of the one million most frequent unigrams, bigrams or trigrams.
-
-Figure \ref{wordclouds} represents the bullish and bearish word clouds.
-These correspond to the most frequent terms (up to 3-grams) in user-labeled bullish (bearish) messages relative to their total appearance.
-The size of the terms represent their importance in the cloud.
-In the bullish cloud, we see words such as ``bullish divergence", ``room to grow", ``lot potential" which are clearly bullish signals.
-In the bearish cloud, we find words such as ``recent resistance", ``short setup", ``bad apple" which indeed indicate bearish signals.
-These findings are reassuring in the sense that the content of the messages on the platform are consistent with their labels.
-The term ``aldox" in the bullish cloud caught our attention.
-After some research, it is an abbreviation for Aldoxorubicin, a drug against tumors and is associated with pharmaceutical messages where investors were very enthusiastic about it (example of related message: ``aldox is on the slide.
-have great faith this is truly world change").
-That is why the term is appearing almost exclusively in bullish messages, hence in the bullish cloud.
-
-On another note, most of the times messages are written by humans.
-It is however possible that some messages are generated by a robot (for instance to spread news or articles).
-In such cases, we define a neutral class that would absorb these messages if no substantial information is embedded in these messages.
-The term "long position open" present in the bearish cloud is an anomaly due to bearish user-labeled messages of intraday alerts such as “sell \$labd close labd long position.
-open labd short position.
-time: 14:53 ny price: \$13.64 zquant intraday alerts”.
-However, such anomalies are not an issue, a message “long position open” has a bullish probability of 0.91 and gets classify as bullish as it should.
-This is the case because when classifying a message, the score of the unigram “long” is much stronger than the score of the trigram “long position open”.
-From a linguistic point of view our approach is brute force.
-However, in turn, it works at a massive scale.
-Overall, the data quality is very good.
-
-\begin{figure}
-\begin{center}
-    \includegraphics[width=0.75\textwidth]{Positive rate words.jpg}
-
-    \vspace{1cm}
-
-    \includegraphics[width=0.75\textwidth]{Negative rate words.jpg}
-   \vspace{0.1cm}
-    \caption{Bullish word cloud (top), bearish word cloud (bottom).
-These correspond to the most frequent terms (up to 3-grams) in user-labeled bullish (bearish) messages relative to their total appearance.
-The size of the terms represent their importance in the cloud.}
-    \label{wordclouds}
-  \end{center}
-\end{figure}
-
-\newpage
-\begin{table}[H]
-\begin{tabular*}{\textwidth}{l@{\extracolsep{\fill}}ll}
-\hline \multicolumn{2}{l}{Before preprocessing}                         \\   \hline \hline                                                                          
+\begin{table}[h]
+\centering
+\begin{tabular}{ll} \hline
+\multicolumn{2}{l}{Before preprocessing}                                   \\ \hline  
 (1) & @CassandraTwit \$uvxy contango 3.5\%...still long. goooooood        \\
 (2) & \$FRPT Take profits while you still can.                            \\
-(3) &  \$UVXY \$tvix go time boys and girls.
-Holding overnight again                                                                                                                                        \\
+(3) & \$UVXY \$tvix go time boys and girls. Holding overnight again       \\
 (4) & \$dnr Nice upgrade as company goes into its quiet period!           \\
- (5) & \$SPY market won't reverse again towards closing. Get put options. \\ 
- \hline                                                                                          
-\end{tabular*}
-\end{table}
-
-\begin{table}[H]
-\begin{tabular*}{\textwidth}{l@{\extracolsep{\fill}}ll}
-\hline \multicolumn{2}{l}{After preprocessing}               \\   \hline \hline                                                                       
+(5) & \$SPY market won't reverse again towards closing. Get put options.  \\ \hline   \hline
+\multicolumn{2}{l}{After preprocessing}                                   \\ \hline 
 (1) & contango still long good                                            \\
@@
 (5) & market will not reverse again towards closing get put options      \\ \hline
-\end{tabular*}
-\caption{Preprocessing of five sample messages.
-Preprocessing operations include: punctuation removal, lower casing, apostrophe handling, contraction form handling (i.e.
-``won't" becomes ``will not"), tickers removal, users removal, URLs removal, parsing and a simple spell corrector dealing with more than two repeated characters (i.e.
-``goooood" becomes ``good")}
+\end{tabular}
+\mycaption{Preprocessing of five sample messages}{Preprocessing operations include: punctuation removal, lower casing, apostrophe handling, contraction form handling (i.e., ``won't" becomes ``will not"), tickers removal, users removal, URLs removal, parsing and a simple spell corrector dealing with more than two repeated characters (i.e., ``goooood" becomes ``good")}
 \label{tab:preprocess}
@@
 
-\section{Sentiment classification}\label{S:classification}
-
-Since mid-2010, StockTwits users have the choice to label their own messages as either \textit{bearish} or \textit{bullish} or to leave it unlabeled.
-Unlabeled messages are tricky to deal with because the user either deliberately chose to leave the message neutral by not labeling it or forgot to click on a label.
-Figure \ref{fig:ts_neutral_vs_uncl} shows the proportions of user-labeled messages in each category across time.
-In the early years of the platform, most messages are unlabeled, presumably because users were not familiar with the sentiment label yet.
-Albeit the proportion of unlabeled messages monotonically declines over the years, almost 60\% of the more recent messages are still unlabeled.
-We believe that by far not all unlabeled messages reflect neutral opinions.
-Figure \ref{fig:postclassif_acrosstime} shows that a lot of unlabeled messages get classified in either bullish or bearish messages, hence these unlabeled messages had indeed valuable information that we are now able to capture.
+The next step is tokenization: the slicing of a text message into smaller units called terms or tokens.
+In financial lingo, some words only have meaning when associated with other words (i.e., ``bad apple" or ``bear flag").
+$N$-gram models allow accounting for words frequently occurring together with other words.
+The main hyperparameter in an $N$-gram model is the number $N$ of words that form a term: a unigram is a term with only one word, a bigram is a term with two consecutive words, etc.
+Larger $N$-gram models increase dramatically the size of the vocabulary (i.e., the collection of all terms considered).
+We select $N=1,2,3$ and truncate the resulting vocabulary such that it consists of the one million most frequent terms in the union of all unigrams, bigrams and trigrams.
 
@@
 \centering
-    \includegraphics[width=0.75\textwidth]{preclassif_acrosstime.JPG}
-    \caption{Proportions of user-labeled messages in each category: bullish (green), bearish (red), and unlabeled (gray). Proportions are aggregated monthly.}
-    \label{fig:ts_neutral_vs_uncl}
+\begin{subfigure}{0.49\textwidth}
+\centering
+\includegraphics[width=1\textwidth]{Positive rate words.jpg}
+\end{subfigure}
+\begin{subfigure}{0.49\textwidth}
+\centering
+        \includegraphics[width=1\textwidth]{Negative rate words.jpg}
+\end{subfigure}
+    \mycaption{Word clouds}{Bullish word cloud (left), bearish word cloud (right).
+These correspond to the most frequent terms (up to trigrams) in user-labeled bullish (bearish) messages relative to their total appearance.
+The size of the terms represents their importance in the cloud.}
+    \label{wordclouds}
 \end{figure}
 
-As one of the goal of this paper is to build an accurate time-series sentiment measure for individual firms, it motivates the use of Natural Language Processing to classify all unlabeled messages in either bullish, neutral or bearish class.
+Figure \ref{wordclouds} represents the bullish and bearish word clouds.
+These represent the most frequent terms in all user-labeled bullish and bearish messages relative to their total appearance, respectively.
+The size of the terms represents their relative weight in the cloud.
+In the bullish cloud, we see terms such as ``bullish divergence", ``room to grow", ``lot potential" which we can clearly interpret as bullish signals.
+In the bearish cloud, we find terms such as ``recent resistance", ``short setup", ``bad apple" which again we can directly interpret as bearish signals.
+These findings are reassuring in the sense that the content of the messages on StockTwits are consistent with their labels.
+We checked for anomalies at random, but did not find significant issues.
+Appendix \ref{app_ano} discusses two such anomalies.
 
-To classify unlabeled messages, we use a logistic regression of the labels on TFIDF transformed user-labeled messages, as in \citet{nlp1} and \citet{classifq}.
-TFIDF stands for Term Frequency-Inverse Document Frequency and is a widely used method to transform a text, in our case a message $m$, into a numerical vector, $TFIDF_m$.
+%\begin{figure}[h]
+%\begin{center}
+%    \includegraphics[width=0.5\textwidth]{Positive rate words.jpg}
+
+%    \vspace{1cm}
+
+%    \includegraphics[width=0.5\textwidth]{Negative rate words.jpg}
+%   \vspace{0.1cm}
+%    \caption{Bullish word cloud (top), bearish word cloud (bottom).
+These correspond to the most frequent terms (up to trigrams) in user-labeled bullish (bearish) messages relative to their total appearance.
+The size of the terms represent their importance in the cloud.}
+%    \label{wordclouds}
+%  \end{center}
+%\end{figure}
+
+\section{Sentiment Classification}\label{S:classification}
+
+Left plot of Figure \ref{fig:classif_time} shows the proportions of user sentiment-labeled messages across time.
+In the early years of the platform, most messages were unlabeled, presumably because users were not familiar with the sentiment label yet.
+Albeit the proportion of unlabeled messages monotonically declines over the years, almost 60\% of the more recent messages are still unlabeled.
+Overall, around 30 million messages are user-labeled and 60 million messages are unlabeled.
+We conjecture that by far not all unlabeled messages reflect market neutral opinions.
+Indeed, the right plot of Figure \ref{fig:classif_time} reveals that a substantial part of user-unlabeled messages is machine learning classified as bullish or bearish.
+Hence these user-unlabeled messages contain indeed market relevant information, which we are able to capture by our algorithm.
+
+%\begin{figure}[h]
+%    \centering
+%    \includegraphics[width=0.5\textwidth]{preclassif_acrosstime.JPG}
+%    \caption{Proportions of user-labeled messages: bullish (green), bearish (red), and unlabeled (gray) across time. Proportions are aggregated monthly.}
+%    \label{fig:ts_neutral_vs_uncl}
+%\end{figure}
+
+\begin{figure}[h]
+\centering
+\begin{subfigure}{0.49\textwidth}
+\centering
+\includegraphics[width=1\textwidth]{preclassif_acrosstime.JPG}
+\end{subfigure}
+\begin{subfigure}{0.49\textwidth}
+\centering
+        \includegraphics[width=1\textwidth]{postclassif_acrosstime3.JPG}
+\end{subfigure}
+
+    \mycaption{Message classification}{Left plot shows the proportions of user-labeled messages: bullish (green), bearish (red), and unlabeled (gray) across time.
+Right plot shows the proportions of machine learning classified messages: bullish (light green predicted, green user-labeled), bearish (light red predicted, red user-labeled), and neutral (gray) across time.
+Proportions are aggregated monthly.}
+    \label{fig:classif_time}
+
+\end{figure}
+
+Among the user-labeled messages we find five times more bullish than bearish ones.
+This ratio indicates that investors are on average optimistic about the market, which is consistent with findings in the literature, e.g., \citet{renault2017intraday}.
+Such an imbalance is a well-known issue in machine learning classification as it creates a bias towards the over-represented class, see \citet{chawla2004special}.
+There are various ways to tackle class imbalance.
+An all-purpose standard approach in machine learning is to over-sample the minority class, which consists of randomly re-sampling from the minority class and thus artificially re-balance the class sizes in the data.
+We use a different approach, which is tailored for our setup.
+As not every message carries a substantial stock market relevant sentiment, we deviate from the bullish--bearish dichotomy.
+Instead, we create an auxiliary neutral sentiment class to account of messages that do not take a clear stand.
+See Appendix \ref{app_exampleclass} for examples of such neutral messages.
+
+We then run two independent binary classifiers.
+The first (second) classifies messages as bullish versus non-bullish (bearish versus non-bearish).
+We aggregate the two binary outcomes and classify a message as bullish (bearish) for the concordant combination bullish/non-bearish (non-bullish/bearish), and neutral otherwise.
+This approach is very simple and efficient, and eliminates the class imbalance bias at the same time.
+It builds on any traditional binary classifiers.
+We use logistic regression on Term Frequency-Inverse Document Frequency (TFIDF)-vectorized messages, as in, e.g., \citet{nlp1}, \citet{classifq}, \citet{erdemlioglu2017market}.
+TFIDF is a widely used method to transform a text, in our case a message $m$, into a numerical vector, $TFIDF_m$.
 The dimension of this vector is equal to the size of the vocabulary (the collection of all terms across all messages).
@@
 The first factor measures how frequently term $t$ appears in the message,
-\[ TF_{m,t} =\dfrac{\sum_{i=1}^{N_m} \textbf{1}_{t=t_{m,i}}}{N_m},\]
+\begin{equation}
+TF_{m,t} =\dfrac{\sum_{i=1}^{N_m} \textbf{1}_{t=t_{m,i}}}{N_m},
+\end{equation}
 where $N_m$ denotes the number of terms $t_{m,i}$ in message $m$. The second factor measures how important term $t$ is to the message,
-\[ IDF_t = \log \big(\dfrac{V}{\sum_{j=1}^{V} \textbf{1}_{t \in m_j}}\big),\]
+\begin{equation}
+IDF_t = \log \big(\dfrac{V}{\sum_{j=1}^{V} \textbf{1}_{t \in m_j}}\big),
+\end{equation}
 where $V$ denotes the total number of messages $m_j$. A term $t$ appearing in many documents (such as ``the", ``is", ``of") is likely to have low information content, hence a low $IDF_t$.
 
-As seen in Figure \ref{categrepartition-preclassif}, user-labeled messages exhibit five times more bullish messages than bearish messages.
-Such an imbalance is a well-known issue in machine learning classification and needs to be dealt with to avoid biases towards the over-represented class (see \citet{chawla2004special}).
-To tackle class imbalance, the most common technique is to randomly oversample the minority class, which consists of repeating some samples of the minority class and balance the number of samples between classes in the data.
-We follow a different approach.
-In addition to bullish and bearish classes, we define an artificial neutral sentiment class as it is possible that some users are not expressing any opinion and sometimes post finance-irrelevant messages.
-To do so, we optimally select classification score thresholds and eliminate the class imbalance bias at the same time.
+We use 80\% of the user sentiment-labeled messages as a training set and keep 20\% as a test set, then we run two binary classifiers.
+The first (second) classifier sets bullish (bearish) as positive and non-bullish (non-bearish) as negative class.
+Every message then classifies into one of the following combinations: (non-bullish, bearish), (bullish, bearish), (non-bullish, non-bearish), (bullish, non-bearish).
+For the first and last combinations, the two algorithms agree and the final classification is defined to be bearish (non-bullish, bearish) or bullish (bullish, non-bearish), respectively.
+For the two middle combinations, (bullish, bearish) and (non-bullish, non-bearish), the two algorithms disagree, so that the final classification is defined to be neutral.
+Formally, every message $m$ is mapped onto either
+\begin{equation}
+     m\mapsto \begin{cases} \text{(non-bullish, bearish)} & =: \text{bearish} \\ \text{(bullish, bearish)}& =: \text{neutral} \\ \text{(non-bullish, non-bearish)}& =: \text{neutral} \\ \text{(bullish, non-bearish)} &=: \text{bullish.} \end{cases}
+\end{equation}
 
-Performance measures widely used in machine learning classification are precision, recall, F1 score and accuracy.
-The first step is to define a class as the positive class.
-Instances (messages) are then divided into true positives TP (predicted positive, actual positive), false positives FP (predicted positive, actual negative), true negatives TN (predicted negative, actual negative), and false negatives FN (predicted negative, actual positive).
-Precision $PRE=\dfrac{TP}{TP+FP}$ is the proportion of true positives among the predicted positives.
-Recall $REC = \dfrac{TP}{TP+FN}$ is the proportion of true positives among the actual positives.
-The precision-recall trade-off is captured by the F1 score, $\dfrac{2 \cdot PRE \cdot REC}{PRE + REC}$, the harmonic mean of precision and recall.
-Accuracy $ACC = \dfrac{TP+TN}{TP+TN+FP+FN}$ is the fraction of correct predictions regardless of the label.
-
-We use 80\% of the user-labeled (bearish and bullish) messages as a training set and keep 20\% as a test set, then we run two binary classifiers.
-The first (second) classifier sets bullish (bearish) as positive and non-bullish (non-bearish) as negative class.
-Every message then falls into the following set of labels: \{(non-bullish, bearish), (bullish, bearish), (non-bullish, non-bearish), (bullish, non-bearish)\}.
-For the two outer cases the two algorithms agree and the final classification is defined to be bearish (non-bullish, bearish) or bullish (bullish, non-bearish), respectively.
-For the two inner cases, (bullish, bearish) and (non-bullish, non-bearish), the two algorithms disagree, so that the final classification is defined to be neutral.
-Formally, every message $m$ is then mapped on either
-
-\[ m\mapsto \begin{cases} \text{(non-bullish, bearish)} & =: \text{bearish} \\ \text{(bullish, bearish)}& =: \text{neutral} \\ \text{(non-bullish, non-bearish)}& =: \text{neutral} \\ \text{(bullish, non-bearish)} &=: \text{bullish.} \end{cases}\]
-
-Precision, recall, and F1 scores of the two algorithms differ because they depend on which class is defined as the positive one.
-To select optimal classification thresholds, we maximize the F1 score of either algorithm.
-The green (red) line in Figure \ref{fig:optimal_thresholds} is the F1 score for the bullish versus non-bullish (bearish versus non-bearish) classification, respectively.
-Circles indicate the maximal F1 scores, along with the corresponding classification score thresholds, 0.50 and 0.72, respectively.
+To select optimal classification thresholds, we maximize the F1 scores.
+The F1 scores of the two binary classifiers differ because they depend on which class is defined as the positive one.
+We recap the definition of the F1 score in Appendix \ref{app_class}.
+Figure~\ref{fig:optimal_thresholds} shows the F1 scores as functions of the threshold.
+Circles indicate the maximal F1 scores, along with the corresponding optimal thresholds, 0.50 and 0.72, respectively.
 
@@
     \includegraphics[width=0.75\textwidth]{optimal thresholds.JPG}
-    \caption{Optimal classification thresholds.
-The green (red) line is the F1 score for the bullish versus non-bullish (bearish versus non-bearish) classification, respectively.
-Circles indicate the maximal F1 scores, along with the corresponding classification score thresholds.}
+    \mycaption{Optimal classification thresholds}{The green (red) line is the F1 score for the bullish versus non-bullish (bearish versus non-bearish) classifier as a function of the threshold.
+Circles indicate the maximal F1 scores.}
     \label{fig:optimal_thresholds}
@@
 
-If the classification score of a message is bigger (smaller) than 0.72 (0.50), then both classifiers agree on the sentiment and the message is classified as bullish (bearish), respectively.
-If the classification score is between 0.50 and 0.72, the classifiers disagree, (bullish, bearish), and we consider the message as neutral.
-Finally, we overwrite the sentiment of a message predicted by the classifier by the user-labeled sentiment whenever the latter is available.
-Research in sentiment classification shows that human annotators tend to agree about 80 to 85\% of the time when evaluating the sentiment of a document (see e.g.
-\citet{wilson2005recognizing} and \citet{chen2020large}).
-This usually represents the accuracy that a sentiment classifier should meet or beat.
-The accuracy in the test set of our combined classifier is 86\%.
+If the sentiment score of a message is bigger (smaller) than 0.72 (0.50), respectively, then both classifiers agree and the sentiment of the message is classified as bullish (bearish), respectively.
+If the sentiment score is between 0.50 and 0.72, the classifiers disagree, (bullish, bearish), and we consider the message as neutral.
+Finally, we overwrite the predicted sentiment of any message by the user-labeled sentiment whenever the latter is available.
+Research in sentiment classification shows that human annotators tend to agree about 80 to 85\% of the time when evaluating the sentiment of a document (see, e.g., \citet{wilson2005recognizing} and \citet{chen2020large}).
+This is a benchmark for the accuracy that a sentiment classifier should meet or beat.
+The out-of-sample accuracy of our combined classifier is 85.9\%.
+Appendix~\ref{app_class} provides in-sample and out-of-sample confusion matrices for our combined classifier.
 
-Figure \ref{fig:postclassif_acrosstime} shows the proportions of classified messages in each category.
-Percentages of bearish (sum of labeled and classified as bearish) and bullish (sum of labeled and classified as bearish) messages are stable over time, suggesting that our classification method is robust.
-Even if most messages were not user-labeled in the early years of the platform, as seen in Figure \ref{fig:ts_neutral_vs_uncl}, we are now able to classify the sentiment of most messages posted in this period.
-Consistent with the over-representation of bullish messages observed in the user-labeled messages in Figure \ref{categrepartition-preclassif}, there are many more messages classified as bullish than bearish.
-Typical messages classified as bullish are messages such as “buy buy” or “hope the pump come soon” whereas typical bearish messages are messages such as “sell everything” or “start short position here”.
-Neutral messages are either empty, irrelevant to finance (e.g.
-“political posturing friend”\footnote{This is a reply to the following message : "honestly, how dumb can you be to believe that china was going to buy significant amount of agricultural products after the breakdown in trade talks.
-even if they buy it will be just a little bit and not significant"}) or ambiguous (e.g.
-“lol wow”).
+Right plot of Figure \ref{fig:classif_time} shows the proportions of our machine learning classifications across time.
+Percentages of bearish (user-labeled and classified as bearish) and bullish (user-labeled and classified as bearish) messages are stable over time, suggesting that our classification method is robust.
+Even if most messages were not user-labeled in the early years of the platform, as seen in the left plot of Figure \ref{fig:classif_time}, we are now able to classify the sentiment of all messages in the sample, including a neutral class.
+Consistent with the over-representation of bullish messages observed in the user-labeled messages, there are substantially more messages classified as bullish than bearish.
+Examples of classified messages are given in Appendix \ref{app_exampleclass}.
 
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.75\textwidth]{postclassif_acrosstime3.JPG}
-    \caption{Proportions of classified messages in each category: bullish (light green predicted, green user-labeled), bearish (light red predicted, red user-labeled), and neutral (gray).
+%\begin{figure}[h]
+%    \centering
+%    \includegraphics[width=0.5\textwidth]{postclassif_acrosstime3.JPG}
+%    \caption{Proportions of machine learning classified messages: bullish (light green predicted, green user-labeled), bearish (light red predicted, red user-labeled), and neutral (gray) across time.
 Proportions are aggregated monthly.}
-    \label{fig:postclassif_acrosstime}
-\end{figure}
+%    \label{fig:postclassif_acrosstime}
+%\end{figure}
 
+%\clearpage
 \section{Polarity}\label{S:polarity}
 
-To build sentiment measures for individual firms and the whole market on a given day, we count the sentiments in the messages.
-To match the timeline on which close-to-close stock returns are computed and to avoid forward-looking bias, we aggregate messages on a close-to-close manner.
-That is, polarity on day $t$ is computed with messages posted from 4:00 pm on day $t-1$ to 4:00 pm on day $t$.
-As stock returns are not computed outside of business days, we shift messages posted outside of business days to the next business day available, then remove any non-business day from our sample.
-We denote by $C_{i,t,j}$ the sentiment of the $j$th message about firm $i$ on day $t$.
-It is set to 1, 0, or -1 for bullish, neutral, or bearish, respectively.
-We follow \citet{ranco2015effects} and define the polarity of firm $i$ as
-\[    P_{i,t} = \dfrac{\sum_{j=1}^{V_{i,t}} (\textbf{1}_{C_{i,t,j} = 1} - \textbf{1}_{C_{i,t,j} = -1})}{\sum_{j=1}^{V_{i,t}} (\textbf{1}_{C_{i,t,j} = 1} + \textbf{1}_{C_{i,t,j} = -1})},\]
-where $V_{i,t}$ denotes the number of messages about firm $i$ on day $t$.\footnote{If $V_{i,t}=0$ then we set $P_{i,t} =0$.}
+We next aggregate message sentiments on a daily ticker-level and across the market.
+Thereto, we denote by $C_{i,t,j}=1,0,-1$ for bullish, neutral, bearish, respectively, the sentiment of the $j$th message about ticker $i$ on day $t$.\footnote{We follow the close-to-close convention.
+First, we remove all non-business days from the sample, whereby messages posted on non-business days count towards the next business day.
+``Day $t$'' then stands for the time interval from 4pm on the previous business day $t-1$ to 4pm on business day $t$.
+This convention is consistent with the stock return data, which are close-to-close, and thus avoids any look-ahead bias of our sentiment polarity.} We follow \citet{ranco2015effects} and define the sentiment polarity of ticker $i$ on day $t$ as
+\begin{equation}
+    P_{i,t} = \dfrac{\sum_{j=1}^{V_{i,t}} (\textbf{1}_{C_{i,t,j} = 1} - \textbf{1}_{C_{i,t,j} = -1})}{\sum_{j=1}^{V_{i,t}} (\textbf{1}_{C_{i,t,j} = 1} + \textbf{1}_{C_{i,t,j} = -1})},
+\end{equation}
+where $V_{i,t}$ denotes the number of messages about ticker $i$ on day $t$.\footnote{If $V_{i,t}=0$ then we set $P_{i,t} =0$.}
 
-As an aggregate measure, we define the polarity for the whole market as a weighted average over all firms 
-\begin{equation*}
+As an aggregate, we define the market polarity as a weighted average over all tickers 
+\begin{equation}
 \label{marketpoleq}
     P_t^M = \dfrac{\sum_{i} V_{i,t} \cdot P_{i,t}}{V^M_t} ,
-\end{equation*}
+\end{equation}
 where $V^M_t = \sum_{i} V_{i,t}$ denotes the number of messages on day $t$.
-
-Figure \ref{fig:marketpolspy} shows a scatter plot of the market polarity $P_{t}^M$ versus the polarity of the SPY\footnote{SPY is an ETF tracking the S\&P500 return.
-It is the largest ETF in the world.}.
-We do not expect the market polarity to be the SPY polarity because the stock universe is not the same (market polarity contains stocks that are not necessarily in the S\&P500 and vice versa).
-The slope coefficient of the regression line is statistically significantly positive and the contemporaneous Pearson correlation coefficient is 0.53, suggesting that the market polarity is an accurate measure of the aggregated sentiment of the market.
-Also, consistent with Figure \ref{fig:postclassif_acrosstime}, SPY and market polarities are bullish-biased.
+Figure \ref{fig:marketpolspy} shows a scatter plot of the market polarity $P_{t}^M$ versus the polarity of SPY.
+The slope coefficient of the regression line is statistically significantly positive and the contemporaneous Pearson correlation coefficient is 0.53, suggesting that the market polarity is an accurate measure of the aggregated sentiment of the market.\footnote{We do not expect $P^M_t$ to be equal to the SPY polarity because the underlying sets of stocks differ: market polarity contains stocks that are not in the S\&P500 and vice versa.} Also, consistent with Figure \ref{fig:classif_time}, SPY and market polarities are bullish-biased.
 
@@
     \includegraphics[width=0.75\textwidth]{spy_vs_marketpol.JPG}
-    \caption{Market polarity on the y-axis versus the polarity of the SPY on the x-axis. The red line shows the linear regression line and coefficients.}
+    \mycaption{Market polarity versus SPY polarity}{The red line shows the linear regression line and coefficients.}
     \label{fig:marketpolspy}
@@
 
-At this point, for stationary reasons\footnote{Computing the polarity ratio with few daily observations would lead to a spiky time series.}, we compute the median of daily message volume for each ticker and exclude from our sample tickers that have less than a median of 50 daily messages.
-Our final sample contains 19 tickers.
-We refer to the appendix for the list of tickers covered as well as more information about the trimming process.
+For the following time-series analysis and event study we restrict our sample.
+There are two reasons for doing so.
+First, we keep computational cost at a reasonable level.
+Second, and more importantly, the time series of ticker-individual polarities exhibit spikes and are too noisy if the daily message volumes $V_{i,t}$ are too small.
+We therefore exclude from now on tickers whose median of daily message volume is less than 50.
+Our reduced sample contains 19 tickers, covering a range of security types, sectors, and market capitalization.
+We refer to Appendix \ref{app_coverage} for more details on the coverage, including summary statistics and the list of tickers covered.
 
-To understand how polarity is related to investor sentiment, we run two linear regressions of contemporaneous daily returns on polarity and 5-days Cumulative Abnormal Polarity :
-\begin{align*}
-    R_{i,t} &= \alpha + \beta \cdot P_{i,t} + \epsilon_{i,t}, \\
-    R_{i,t} &= \alpha + \beta \cdot CAP_{i,5} + \epsilon_{i,t}. 
-\end{align*}
-Table \ref{regpolret} shows that $\beta$ is positive and significant for both regressions.
-This indicates that polarity is a good proxy for the sentiment of investors.
-Further supporting evidence is given by the correlation between polarity and contemporaneous stock returns at the firm level.
-Figure \ref{fig:ts} shows the time series during 2019 for the top 6 most discussed tickers.
-In our entire panel of firms, correlations are always positive and range between 0.1 and 0.3.
+To understand how our polarity measure is related to investor sentiment, we run linear regressions of contemporaneous and next-day returns on polarity:
+\begin{align}
+    R_{i,t} &= \alpha^{cont} + \beta^{cont} \cdot P_{i,t} + \epsilon^{cont}_{i,t}, \\
+    R_{i,t+1} &= \alpha^{next} + \beta^{next} \cdot P_{i,t} + \epsilon^{next}_{i,t}.
+\end{align}
+Table \ref{regpolret} shows that the regression coefficient is positive and significant for contemporaneous returns.
+This indicates that polarity is a good contemporaneous proxy for the sentiment of investors.
+Further supporting evidence is given by the correlation between polarity and contemporaneous returns at the ticker level.
 
-\begin{table}[]
+\begin{table}[h]
 \centering
 \begin{tabular}{l|c|c|c|c|c}
-                     & $R_{i,t}$ & $R_{i,t}$   & $R_{i,t+1}$ & $R_{i,t+1}$        \\ \toprule 
-Constant             & -0.0047*** & -7.92e-05   & -0.0002    & 7e-06         \\
-                     & (0.000)    & (9.81e-05)  & (0.000)    & (9e-05)       \\
-$P_{i,t}$           & 0.009***   &               & 0.0003     &               \\
-                     & (0.000)    &               & (0.000)    &               \\
-$CAP_{i,5}$         &            & 0.0007***     &            & -0.0002       \\
-                     &            & (0.000)       &            & (0.000)       \\ \bottomrule
-$R^2$                & 0.012      & 0.000         & 0.000      & 0.000              \\
-No. Obs.             & 34100      & 34024         & 34100      & 34024             
+                     & $R_{i,t}$  & $R_{i,t+1}$        \\ \hline   
+Constant             & -0.0047***  & -0.0002      \\
+                     & (0.000)     & (0.000)      \\
+$P_{i,t}$           & 0.009***      & 0.0003       \\
+                     & (0.000)       & (0.000)      \\ \hline
+$R^2$                & 0.012         & 0.000             \\
+No. Obs.             & 34100         & 34100            
 \end{tabular}
-\caption{Results from linear regressions of contemporaneous and next period stock returns on polarity.
+\mycaption{Regressions of returns on polarity}{Results from linear regressions of contemporaneous and next-day stock returns on polarity.
 Stock returns are trimmed at the 5\% percentile on both sides.
@@
 
-\begin{figure*}[]
+Figure \ref{fig:ts} shows the time series during 2019 for the top 6 most discussed tickers.
+In our entire sample of tickers, correlations are always positive and range between 0.1 and 0.3.
+In contrast, regressing next-day returns reveals that polarity has no predictive power for next-day stock returns unconditionally.
+In the following section we show how polarity has predictive power around specific events.
+
+\begin{figure*}[h]
         \centering
-        \begin{subfigure}[b]{0.48\textwidth}
+        \begin{subfigure}[b]{0.425\textwidth}
             \centering
@@
         \quad
-        \begin{subfigure}[b]{0.48\textwidth}  
+        \begin{subfigure}[b]{0.425\textwidth}  
             \centering 
@@
         \vskip\baselineskip
-        \begin{subfigure}[b]{0.48\textwidth}   
+        \begin{subfigure}[b]{0.425\textwidth}   
             \centering 
@@
         \quad
-        \begin{subfigure}[b]{0.48\textwidth}   
+        \begin{subfigure}[b]{0.425\textwidth}   
             \centering 
@@
                 \vskip\baselineskip
-        \begin{subfigure}[b]{0.48\textwidth}   
+        \begin{subfigure}[b]{0.425\textwidth}   
             \centering 
@@
         \end{subfigure}
-        \caption{Time series of daily polarity (red - left axis) and daily stock returns (blue - right axis) since 1st of January 2019 for the top 6 most discussed tickers.
+        \mycaption{Correlation between return and polarity}{Time series of daily polarity (red - left axis) and daily stock returns (blue - right axis) since 1st of January 2019 for the top 6 most discussed tickers.
 Pearson correlation between the two time series is shown in the title.}
@@
 
-We also run two linear regressions of next day returns on polarity and 5-days Cumulative Abnormal Polarity :
-\begin{align*}
-    R_{i,t+1} &= \alpha + \beta \cdot P_{i,t} + \epsilon_{i,t}, \\
-    R_{i,t+1} &= \alpha + \beta \cdot CAP_{i,5} + \epsilon_{i,t}. \\
-\end{align*}
-Table \ref{regpolret} reveals that polarity has no predictive power for next day stock returns unconditionally.
-However, the next section depicts how polarity still has embedded information around specific events.
+\clearpage\section{Event Study}\label{S:event studies}
 
-\section{Event studies}\label{S:event studies}
+Event studies constitute a statistical method widely used in financial econometrics, see, e.g., \citet{mackinlay1997event}.
+In general, they are used to measure the effect of events on the market value of stocks.
+Well-known applications of event studies include the testing of various forms of the efficient market hypothesis (EMH) (see \citet{fama1969} and \citet{fama1991efficient}).
 
-Event studies constitute a statistical method widely used in financial econometrics.
-In general, they are used to measure the effect of events on the market value of firms.
-Well known applications of event studies include the testing of various forms of the efficient market hypothesis (EMH) (see \citet{fama1969} and \citet{fama1991efficient}).
-What's more, as described in \citet{mackinlay1997event}, event studies can also be applied with little modification to other variables than stock returns.
-
-To design an event study, the first step is to define events of interest and the \textit{event window} over which a variable will be examined.
-Adhering to common practice, we choose the event window expanded from 20 business days before the event to 20 business days after the event.
-The \textit{estimation window} is used to estimate the parameters of the market model.
-A common choice is a one-year window ending 20 days before the event.
-Figure \ref{fig:timeline} shows the corresponding timeline.
+\subsection{Events}
+We define events as days with an unusual large number of messages for individual tickers.
+We conjecture that a sudden peak in StockTwits message volume indicates that an important corporate or stock market event is happening on the day of the peak.
+Figure \ref{fig:actvol} shows that increases (decreases) in message volumes are positively associated with increases (decreases) in contemporaneous weekly stock transaction volumes.
+These co-movements indicate that investors who post messages about stocks also trade them accordingly, adjusting their portfolios.
+This suggests that message volume peaks are a good proxy for corporate and stock market events.
 
@@
     \centering
-    \includegraphics[width=0.75\textwidth]{timeline.JPG}
-    \caption{Timeline of our event studies.
-$\tau$ is an event date, the event window has 41 business days centered around the event date and the estimation window is a one-year rolling period prior the event.}
-    \label{fig:timeline}
-\end{figure}
-
-Then, event studies require the distinction between a normal and an abnormal measure.
-The normal measure is defined as the expected measure conditioning on the event not taking place.
-In other words, we need an expectation of the variable of interest that would have been measured if the event never existed.
-The abnormal measure is defined as the measure minus the normal measure.
-Formally, let $X_{i,t}$ be a measure of interest (i.e.
-stock return, polarity, ...), $AX_{i,t}$ be the abnormal measure and $E(X_{i,t}|Z_{i,t})$ the normal measure with $Z_{i,t}$ being the conditional information for the normal measure model.
-We have:
-\[     AX_{i,t} = X_{i,t} - E(X_{i,t}|Z_{i,t}).\]
-
-The normal measure $E(X_{i,t}|Z_{i,t})$ is usually computed using either a constant mean model or a market model.
-Constant mean models use $E(X_{i,t}|Z_{i,t}) = \mu$ where $\mu$ is the mean of the measure during the estimation window.
-We chose to work with a market model but all following results hold with the constant mean model as well.
-As stated in \citet{mackinlay1997event}, the variance of the abnormal measure is not reduced a lot by choosing a more sophisticated model, hence the event study is not sensitive to the choice of the normal model.
-The market model we are using is defined as the following: 
-\begin{equation*}\label{eq_mm}
-    X_{i,t} = \alpha_i + \beta_i \cdot X_{t}^M + \epsilon_{i,t},
-\end{equation*}
-with $E(\epsilon_{i,t}) = 0$, $V(\epsilon_{i,t}) = \sigma_{\epsilon}^2$ and $X_{t}^M$ the measure of the whole market.
-We estimate the parameters of the market model in a one-year rolling estimation window.
-To avoid overlaps between estimation windows and events, we remove any event day from the estimation windows.
-This ensures that large event returns do not influence the parameters of the normal measure.
-
-\subsection{Events}
-We define an event as an unusual high number of daily messages for a particular firm.
-We conjecture that a sudden peak in StockTwits message volume indicates that an important firm event is happening on the day of the peak.
-As a robustness check, we show in Figure \ref{fig:actvol} that increases (decreases) in number of message volume are positively associated with increases (decreases) in contemporaneous weekly volume of stock transactions.
-These co-movements mean that investors not only post about these stocks but also trade them simultaneously, adjusting their portfolios.
-This indicates that the message volume peaks is a good proxy to identify event dates as there is no lag between message activity and trades.
-
-\begin{figure}
-    \centering
     \includegraphics[width=0.75\textwidth]{activity-volume.jpg}
-    \caption{Changes in weekly volume of transactions on the y-axis versus changes in message activity on the x-axis. Activity is measured in weekly messages posted per firm.}
+    \mycaption{Message activity and transaction volume}{Changes in weekly volume of transactions on the y-axis versus changes in message activity on the x-axis.
+Activity is measured in weekly messages posted per ticker.}
     \label{fig:actvol}
@@
 
-To measure unusual activity peaks, we use the market model (\ref{eq_mm}) on a one-year rolling estimation window and regress the daily relative change of message volume of individual firms $\dfrac{\Delta V_{i,t}}{V_{i,t-1}}$ on the daily relative change of total message volume $\dfrac{\Delta V_t^M}{V_{t-1}^M}$.
-Formally,
-\[     \dfrac{\Delta V_{i,t}}{V_{i,t-1}} = \alpha_i^V + \beta_i^V \cdot \dfrac{\Delta V_t^M}{V_{t-1}^M} + \epsilon_{i,t}.\]
-We can then compute the abnormal volume of messages for firm i as : 
-\[     AV_{i,t} = \dfrac{\Delta V_{i,t}}{V_{i,t-1}} - \hat{\alpha}_i^V - \hat{\beta}_i^V \cdot \dfrac{\Delta V_t^M}{V_{t-1}^M}. \]
-We define an event for ticker $i$ as a day $t$ where the standardized abnormal volume exceeds 2, 
-\[   \dfrac{AV_{i,t} - \mu_{AV_i}}{\sigma_{AV_i}} > 2.\]
+To measure unusual activity peaks, we use as benchmark model a one-year rolling window regression of daily relative message volume changes of ticker $i$ on daily relative total message volume changes.\footnote{Accounting for the lead time of the one-year rolling estimation window, the event study effectively applies to the shorter period from January 2011 to March 2020.} Formally,
+\begin{equation}
+\dfrac{\Delta V_{i,t}}{V_{i,t-1}} = \alpha_i^V + \beta_i^V \cdot \dfrac{\Delta V_t^M}{V_{t-1}^M} + \epsilon_{i,t},    
+\end{equation} 
+which gives the one-year rolling estimates $\hat{\alpha}_i^V$ and $\hat{\beta}_i^V$. We then define the abnormal message volume changes for ticker $i$ on day $t$ as 
+\begin{equation}
+    AV_{i,t} = \dfrac{\Delta V_{i,t}}{V_{i,t-1}} - \left(\hat{\alpha}_i^V + \hat{\beta}_i^V \cdot \dfrac{\Delta V_t^M}{V_{t-1}^M}\right).
+\end{equation}
+We define an event for ticker $i$ as any day $t$ where the standardized abnormal volume exceeds two, 
+\begin{equation}
+    \dfrac{AV_{i,t} - \hat{\mu}_{AV_i}}{\hat{\sigma}_{AV_i}} > 2,
+\end{equation}
+where $\hat{\mu}_{AV_i}$ and $\hat{\sigma}_{AV_i}$ denote the one-year rolling empirical mean and standard deviation.
 
-%To measure unusual activity peaks, we use the time series of individual firms message volume and identify days where the volume exceeds five times its rolling one-month median.
-Formally, the event dates for firm $i$ are the days $e$ such that:
-%\begin{equation}
-%    e = \{d : V_d > 5 \cdot V_m(d) \}
-%\end{equation}
-%where $V_d$ is the message volume on day d for firm i and $V_m(d)$ is the rolling one-month volume median at day d for firm i.
-
-Then, we define the type of the event as either bullish, neutral or bearish.
+Next we define the type of the event as either bullish, neutral or bearish.
 We use the abnormal polarity $AP_{i,t}$ of the event date to assess how on average investors perceive the event.
-Figure \ref{fig:abnpol_all} shows the distribution of abnormal polarities.
+Figure \ref{fig:abnpol_all} shows the distribution of abnormal polarities on event dates.
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.75\textwidth]{abnpol_distr_all.JPG}
+    \mycaption{Empirical distribution of abnormal polarities on event dates}{Red dashed lines show the one-third and two-third percentiles.}
+    \label{fig:abnpol_all}
+\end{figure}
 We chose to use the one-third (-0.03) and two-third percentile (0.07) of the distribution of abnormal polarities as thresholds for the type of the event.
-Conditionally on the existence of an event for firm $i$ at day $t$, we have:
-
-\[     Type_{i,t}= \begin{cases}
+We define the type of the event for ticker $i$ at $t$ as
+\begin{equation}
+Type_{i,t}= \begin{cases}
       Bullish  & \text{if}\ \ AP_{i,t}> 0.07, \\
@@
       Bearish  & \text{if}\ \ AP_{i,t} < -0.03. \\
-\end{cases} \]
+\end{cases}    
+\end{equation}
+
+Overall, across 19 tickers, we identify 1131 events, whereof 454 bullish, 294 neutral, and 383 bearish types.
+This coverage is on par with previous studies (e.g., \citet{mackinlay1997event} analyze 30 stocks and 600 events).
+Figure \ref{fig:events} shows the aggregate events and their types across time.
+The count of events looks stationary over time, apart from a build up phase of the platform in the early part until 2014.
+The distribution of event types is also balanced across time.
 
@@
     \centering
-    \includegraphics[width=0.75\textwidth]{abnpol_distr_all.JPG}
-    \caption{Empirical distribution of abnormal polarities. Red dashed lines show the one-third and two-third percentiles.}
-    \label{fig:abnpol_all}
+    \includegraphics[width=0.75\textwidth]{events_across_time.JPG}
+    \mycaption{Number of events of each type across time}{Green area shows bullish events, gray area shows neutral events and red area shows bearish events. Numbers are aggregated monthly.}
+    \label{fig:events}
 \end{figure}
 
-As an illustration, Figure \ref{fig:activityaapl} shows for Apple the time-serie of message volume and the bullish, neutral and bearish events identified as green up-triangles, gray circles and red down-triangles, respectively.
-Between 2011 and 2020, our algorithm identified 73 events for Apple.
-Interestingly, our methodology allows us to capture more than earning announcements : about half of these events correspond to earning announcements, but some also correspond to Apple \textit{Keynotes}\footnote{Keynotes are presentations that Apple gives to the press, often presenting new products.} or even CEO letters addressed to investors.
-Across 19 tickers, we identify 1131 events distributed across the three categories : 454 bullish events, 294 neutral events and 383 bearish events.
-This coverage is on par with previous studies (i.e: \citet{mackinlay1997event} has 30 firms and 600 events) Figure \ref{fig:events} shows the identified events and their types across the years.
-%Table \ref{tab:aaplevent} shows all events for Apple and their associated types.
-
-\begin{figure}[]
+As an illustration, Figure \ref{fig:activityaapl} shows for Apple the time-series of message volume and the the corresponding events.
+Between January 2011 and March 2020, our algorithm identified 73 events for Apple.
+\begin{figure}[h!]
     \centering
     \includegraphics[width=0.75\textwidth]{activity_aapl.JPG}
-    \caption{Daily message volume for Apple.
-Events are days with an unusual high number of messages.
+    \mycaption{Daily message volume for Apple}{Events are days with an unusual high number of messages.
 Green upper-triangles show bullish events, gray circles are neutral events and red down-triangles represent bearish events.}
@@
 \end{figure}
+What are these events?
+Remarkably, we capture a variety of corporate events and disclosures.
+Earning announcements constitute about half of the events.
+Other events include Apple Keynotes (presentations that Apple gives to the press, often presenting new products), or CEO letters addressed to investors.
+Table \ref{tab:event_aapl} lists a few examples.
 
-%\begin{table}[]
-%\caption{Identified event dates and associated description for Apple between %2013 and 2020.}
-%\centering
-%\begin{tabular}{|l|l|}
-%\toprule
-%Date         & Description\\
-%\bottomrule
-%2013-10-28  & Earnings announcement\\
-%2014-04-23  & Earnings announcement\\
-%2014-09-09  & Keynote \\
-%2015-10-27  & Earnings announcement\\
-%2016-01-26  & Earnings announcement\\
-%2016-07-26  & Earnings announcement\\
-%2016-09-07  & Keynote \\
-%2016-10-25  & Earnings announcement\\
-%2017-01-31  & Earnings announcement\\
-%2017-05-02  & Earnings announcement \\
-%2017-08-01  & Earnings announcement \\
-%2017-08-02  & Earnings announcement \\
-%2017-09-12  & Keynote \\
-%2017-11-02  & Earnings announcement\\
-%2018-05-01  & Earnings announcement\\
-%2018-07-31  & Earnings announcement\\
-%2018-08-01  & First company to reach \$1 trillion\\
-%2018-11-01  & Earnings announcement\\
-%2019-01-02  & CEO Letter to investors\\
-%2019-01-29  & Earnings announcement\\
-%2019-03-25  & Keynote \\
-%2019-04-30  & Earnings announcement\\
-%2019-07-30  & Earnings announcement\\
-%2019-09-10  & Keynote\\
-%\bottomrule
-%\end{tabular}
-%\label{tab:aaplevent}
-%\end{table}
+\begin{table}[h]
+\centering
+\begin{tabular}{l|l|c}
+Date         & Description & Type\\
+\hline  
+2012-04-24  & Earnings announcement & Bullish\\
+2012-09-12  & Presents iPhone 5 & Neutral \\
+2014-09-09  & Presents Apple Watch & Neutral \\
+2017-05-02  & Announces drop in iPhone sales & Bearish \\
+2017-08-31  & Earnings announcement & Bullish \\
+2017-09-12  & Presents iPhone X & Bearish \\
+2019-01-02  & CEO Letter to investors & Bearish \\
+2019-09-10  & Presents iPhone 11 & Neutral \\
+2019-10-30  & Earnings announcement & Bullish \\
+\end{tabular}
+\mycaption{Selected events and associated description and types for Apple}{This list is for illustration and non-exhaustive (9 out of 73).}
+\label{tab:event_aapl}
+\end{table}
 
-\begin{figure}[]
-    \centering
-    \includegraphics[width=0.75\textwidth]{events_across_time.JPG}
-    \caption{Number of events in each category across time. Numbers are aggregated monthly.}
-    \label{fig:events}
-\end{figure}
+%\clearpage
+\subsection{Abnormal Stock Returns}
+How do stock returns behave around events?
+Similar to the relative message volume changes, we use as benchmark model a one-year rolling window regression of the daily returns of ticker $i$ on the daily market returns, $R^M_{t}$, i.e., daily excess returns of the S\&P500,
+\begin{equation}
+    R_{i,t}= \alpha_i^R  + \beta_i^R \cdot R_t^M  + \epsilon_{i,t},
+\end{equation}
+which gives the one-year rolling estimates $\hat{\alpha}_i^R$ and $\hat{\beta}_i^R$. This implies the abnormal returns 
+\begin{equation}
+    AR_{i,t} = R_{i,t} - \left(\hat{\alpha_i}^R + \hat{\beta_i}^R \cdot R_t^M\right).
+\end{equation}
+We define the cumulative abnormal returns (CAR) around a ticker $i$ event $\tau$ as 
+\begin{equation}
+    CAR_i(\tau,t) = \sum_{s=- 20}^{t} AR_{i,\tau + s},
+\end{equation}
+and the cumulative average abnormal returns (CAAR) across all $N=1131$ events as 
+\begin{equation}
+         CAAR(t) = \dfrac{1}{N} \sum_{j=1}^N CAR_{i_j}(\tau_j,t).
+\end{equation}
 
-\subsection{CAAR and CAAP}
-We estimate the parameters for the market models using OLS, as it is a consistent and efficient estimator under general conditions.\footnote{An interested reader can refer to \citet{mackinlay1997event} for more details}
-
-\subsubsection{Abnormal Returns}
-Given the market models parameters $\hat{\alpha_i}^R$ and $\hat{\beta_i}^R$ estimated in \eqref{eq_mm}, the abnormal return is
-\[     AR_{i,t} = R_{i,t} - \hat{\alpha_i}^R - \hat{\beta_i}^R \cdot R_t^M,\]
-with $R_t^M$ being the market return. Then, we can compute the cumulative abnormal returns (CAR) around a firm $i$ event $\tau$ as:
-\[    CAR_i(\tau,t) = \sum_{s=- 20}^{t} AR_{i,\tau + s},\]
-and finally get the cumulative average abnormal returns (CAAR) across the N events as:
-\[     CAAR(t) = \dfrac{1}{N} \sum_{j=1}^N CAR_{i_j}(\tau_j,t).\]
-Variance of CAR is computed following \citet{mackinlay1997event} as :
-\[     var(CAR(\tau,t)) = \dfrac{1}{N^2} \sum_{i=1}^N (CAR_i(\tau,t) - CAAR(t))^2.\]
-
-Figure \ref{fig:ES_CAAR_MM} shows the CAAR around the events identified.
+Left plot of Figure \ref{fig:caarcaap} shows the CAAR around the events.
 This plot is consistent with \citet{mackinlay1997event}.
-It shows that CAAR related to bearish (bullish) events displays a downward (upward) jump at the event date respectively, and then the jumps are followed by a stable CAAR during the 20 days period after an event.
-Interestingly, there is a systematic (small) shift in the CAAR already 1 day before an event.
+It shows that CAAR related to bearish (bullish) events exhibits a significant downward (upward) jump at the event date, respectively.
+These jumps are followed by a flat CAAR during the 20 days after the event.
+Interestingly, there is a systematic shift in the CAAR already 1 day before the event.
+However, this shift is relatively small compared to the jump on the event day: one day before the event, the bullish (bearish) CAAR equals 0.019 (-0.023).
 The CAAR related to the neutral events exhibits a slight upward shift around the event date but it fades away after a few days.
 The CAAR related to bearish events shifts already a few days before the event but this shift is not statistically significant.
-Additionally, as we see in the top-left plot of Figure \ref{fig:boxplotcaar}, box plots are not shifted, indicating that the conditional distributions of the CAR are not statistically different from each other.
-Mann-Whitney U-test (Table \ref{utest_res}) shows that 5 days before an event, the median of the CAR distribution of the bullish events is neither statistically different from the median of the neutral nor the bearish events.
+This is in line with Figure \ref{fig:boxplotcaar}, which shows that the CAR distributions prior to the events are not significantly different from zero.
+This is confirmed by the Mann-Whitney U-tests shown in Table \ref{utest_res}.
+CAR has no predictive power on the type of the event: five days before an event, the median of the CAR distribution of the bullish events is not statistically different from the median of the neutral events.
+Same holds for the bearish events.
 
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.75\textwidth]{caar.JPG}
-    \caption{Cumulative average abnormal returns around identified events.
+%\begin{figure}[h]
+%    \centering
+%    \includegraphics[width=0.5\textwidth]{caar.JPG}
+%    \caption{Cumulative average abnormal returns around identified events.
 CAAR related to bearish, neutral and bullish events are displayed with the red, gray and green line, respectively.
 Areas around lines show confidence intervals at the 95\% level.}
-    \label{fig:ES_CAAR_MM}
-\end{figure}
+%    \label{fig:ES_CAAR_MM}
+%\end{figure}
 
-\subsubsection{Abnormal Polarity}
-Similar to the subsection above, given the market models parameters $\hat{\alpha_i}^P$ and $\hat{\beta_i}^P$ estimated in \eqref{eq_mm}, the abnormal polarity is
-\[    AP_{i,t} = P_{i,t} - \hat{\alpha_i}^P - \hat{\beta_i}^P \cdot P_t^M,\]
-with $P_t^M$ being the market polarity computed in \eqref{marketpoleq}.
-Then, we can compute the cumulative abnormal polarity (CAP) and cumulative average abnormal polarity (CAAP) from $\tau - 20$ to $t \le \tau + 20$ as:
+%\clearpage
+\subsection{Abnormal Polarity}
+How does sentiment polarity behave around events?
+Similar to the above, we use as benchmark model a one-year rolling window regression of the the daily polarity of ticker $i$ on the daily market polarity defined in \eqref{marketpoleq},
+\begin{equation}
+    P_{i,t}= \alpha_i^P  + \beta_i^P \cdot P_t^M  + \epsilon_{i,t},
+\end{equation}
+which gives the one-year rolling estimates $\hat{\alpha}_i^P$ and $\hat{\beta}_i^P$. This implies the abnormal polarity 
+\begin{equation}
+      AP_{i,t} = P_{i,t} - \left(\hat{\alpha_i}^P + \hat{\beta_i}^P \cdot P_t^M\right).
+\end{equation}
+We define the cumulative abnormal polarity (CAP) around a ticker $i$ event $\tau$ as 
+\begin{equation}\label{eq_CAPtaut}
+    CAP_i(\tau,t) = \sum_{s=- 20}^{t} AP_{i,\tau + s},
+\end{equation}     
+and the cumulative average abnormal polarities (CAAP) across all $N=1131$ events as 
+\begin{equation}
+       CAAP(t) = \dfrac{1}{N} \sum_{j=1}^N CAP_{i_j}(\tau_j,t).
+\end{equation}
 
-\begin{align*}
-    CAP_i(\tau,t) &= \sum_{s=- 20}^{t} AP_{i,s+\tau}, \\
-    CAAP(t) &= \dfrac{1}{N} \sum_{j=1}^N CAP_{i_j}(\tau_j,t).
-\end{align*}
+Right plot of Figure \ref{fig:caarcaap} shows the CAAP around the events.
+There are two main findings.
+First, in contrast to CAAR, the CAAP for bullish and bearish events is not constant after the event date, suggesting that users' sentiments about stocks tend to be biased towards recent past events.
+A possible explanation is that users might still post bullish (bearish) messages about a bullish (bearish) event during several days after the event.
+This is in contrast to the returns that immediately normalize after the event.
+Second, and more interestingly, the CAAP for bullish and bearish events shifts several days earlier than the CAAR.
+This indicates that investors are on average able to anticipate the type of an event in the near future.
+However, this sentiment only manifests through the social media, but not through abnormal returns.
 
-Figure \ref{fig:ES_cum_abn_pol_MM} shows the CAAP around the events identified.
-The main findings about polarity are twofold.
-First, conversely to CAAR, CAAP for bullish and bearish events are not constant after the event date, suggesting that users' sentiment about a firm tend to be biased towards recent past events.
-This can be explained by the fact that users might still post bullish or bearish messages about an event even several days after the event happening, even though the return had already adjusted.
-Second, and more interestingly, it looks like the CAAP for bullish and bearish events shift several days earlier than the CAAR.
-This would indicate that the users are on average able to anticipate a future bullish (bearish) event and post positively (negatively) about the firm days before the rise in CAAR.
-Figure \ref{fig:boxplotcaar} illustrates this striking result with box plots (see \citet{boxplot1} and \citet{boxplot2}) showing the distribution of the CAR and CAP.
-The line inside a box shows the median while the edges of each box represent the 25\% and 75\% quantile of the distribution.
-From above the edges of a box, a distance of 1.5 times the interquartile range is measured and a whisker is drawn up to the largest and lowest observed point from the data that falls within this distance.\footnote{Interquartile range is equal to the third quartile minus the first quartile.} The three plots on the left (right) show the boxplots for the CAR (CAP) 5 days before an event, on event date, and 5 days after the event, respectively.
-To show statistical significance, we use the Mann-Whitney U-test (see \citet{utest} and \citet{sheskin}) on every plot to assess whether the three samples (bullish, neutral and bearish) represent populations with different median values.\footnote{This interpretation only holds under stringent assumptions on the populations, namely that the two population distributions are equal up to a shift.}  Under the null hypothesis, the three samples represent distributions with equal medians.
- Let $\theta_i$ be the median of the distribution i.
-Formally, we test $H_0 : \theta_{bullish} = \theta_{neutral}$ against $H_1 : \theta_{bullish} > \theta_{neutral}$ and $H_0 : \theta_{neutral} = \theta_{bearish}$ against $H_1 : \theta_{neutral} > \theta_{bearish}$ 5 days before an event, on event date and 5 days after an event.
-We define U as the Mann-Whitney test statistic, Z as the normal approximation of the Mann-Whitney test statistic for large sample sizes, $n_1$ and $n_2$ as the sample sizes.
-We refer to \citet{sheskin} for the test statistic computation.
-Table \ref{utest_res} shows U-test estimates for pairwise comparisons.
-The null is rejected in every case except for CAR at $t-5$.
-5 days before the event, the boxes of the CAR between bullish, neutral and bearish events are on the same level, suggesting no predictive power of abnormal returns, consistent with the EMH.
-However, the boxes showing the CAP 5 days before the event are already shifted, meaning that conditionally on an event happening, investors are able on average to anticipate correctly the type of this event.
-On the event date, the boxes of the CAR shift as the abnormal returns jump for both bullish and bearish events, consistent again with the EMH.
-Finally, 5 days after the event, the distributions of the CAR are very similar to the distributions on the event dates.
-Again, this is consistent with the EMH as the abnormal returns adjusted quickly on event date and all information is now embedded in the prices.
-The distribution of the CAP 5 days after the event continued to shift compared to the event date, as people continue to post about recent past events.
-% \TODO{maybe this part of the short story about EMH, which we should provide earlier, at the beginning of the event study part, and then refer to it in the sequel of our text ..}
-
-\begin{figure}[]
-    \centering
-    \includegraphics[width=0.75\textwidth]{caap.JPG}
-    \caption{Cumulative average abnormal polarity around identified events.
+%\begin{figure}[h]
+%    \centering
+%    \includegraphics[width=0.5\textwidth]{caap.JPG}
+%    \caption{Cumulative average abnormal polarity around identified events.
 CAAP related to bearish, neutral and bullish events are displayed with the red, gray and green line, respectively.
 Areas around lines show confidence intervals at the 95\% level.}
-    \label{fig:ES_cum_abn_pol_MM}
-\end{figure}
+%    \label{fig:ES_cum_abn_pol_MM}
+%\end{figure}
 
-\begin{figure*}[]
+\begin{figure*}[h!]
 \centering
-        \begin{subfigure}[b]{0.42\textwidth}
+        \begin{subfigure}{0.49\textwidth}   
+            \centering 
+            \includegraphics[width=\textwidth]{caar.jpg}
+        \end{subfigure}
+        \begin{subfigure}{0.49\textwidth}   
+            \centering 
+            \includegraphics[width=\textwidth]{caap.jpg}
+        \end{subfigure}
+        \mycaption{CAAR and CAAP around identified events}{Cumulative average abnormal returns (left plot) and cumulative average abnormal polarity (right plot) around identified events.
+CAAR and CAAP related to bearish, neutral and bullish events are displayed with the red, gray and green line, respectively.
+Areas around lines show confidence intervals at the 95\% level.}
+        \label{fig:caarcaap}
+\end{figure*}
+
+Figure \ref{fig:boxplotcaar} illustrates this striking finding with box plots (see \citet{boxplot1} and \citet{boxplot2}) showing the distributions of the CAR and CAP, for all three event types, 5 days before the event, at the event date, and 5 days after the event, respectively.
+
+\begin{figure*}[h!]
+        \centering
+        \begin{subfigure}[b]{0.47\textwidth}
             \centering
@@
         \quad
-        \begin{subfigure}[b]{0.42\textwidth}  
+        \begin{subfigure}[b]{0.47\textwidth}  
             \centering 
@@
         \vskip\baselineskip
-        \begin{subfigure}[b]{0.42\textwidth}   
+        \begin{subfigure}[b]{0.47\textwidth}   
             \centering 
@@
         \quad
-        \begin{subfigure}[b]{0.42\textwidth}   
+        \begin{subfigure}[b]{0.47\textwidth}   
             \centering 
@@
                 \vskip\baselineskip
-        \begin{subfigure}[b]{0.42\textwidth}   
+        \begin{subfigure}[b]{0.47\textwidth}   
             \centering 
@@
         \end{subfigure}
-        \caption{Distributions of CAP and CAR 5 days before an event, on event date and 5 days after an event.
-The line inside a box shows the median while the edges of each box represent the 25\% and 75\% quantile of the distribution.
-From the edges of a box, a distance of 1.5 times the interquartile range is measured and a whisker is drawn up to the largest and lowest observed point from the data that falls within this distance.}
+        \mycaption{Distributions of CAP and CAR around events}{The line inside a box shows the median while the edges of each box represent the 25\% and 75\% quantile of the distribution.
+From above the edges of a box, a distance of 1.5 times the interquartile range is measured and a whisker is drawn up to the largest and lowest observed point from the data that falls within this distance.
+Interquartile range is equal to the third quartile minus the first quartile.}
         \label{fig:boxplotcaar}
@@
 
+To check statistical significance, we use the Mann-Whitney U-test (see \citet{utest} and \citet{sheskin}) to test whether the three samples (bullish, neutral and bearish) represent populations with different median values.\footnote{This interpretation only holds under stringent assumptions on the populations, namely that the two population distributions are equal up to a shift.
+Under the null hypothesis, the three samples represent distributions with equal medians.
+Let $\theta_i$ be the median of the distribution i.
+Formally, we test $H_0 : \theta_{bullish} = \theta_{neutral}$ against $H_1 : \theta_{bullish} > \theta_{neutral}$ and $H_0 : \theta_{neutral} = \theta_{bearish}$ against $H_1 : \theta_{neutral} > \theta_{bearish}$ 5 days before an event, on event date and 5 days after an event.
+We define U as the Mann-Whitney test statistic, Z as the normal approximation of the Mann-Whitney test statistic for large sample sizes, $n_1$ and $n_2$ as the sample sizes.
+We refer to \citet{sheskin} for the test statistic computation.}
+Table \ref{utest_res} shows U-test statistics for pairwise comparisons.
+The null is rejected in every case except for CAR at $\tau-5$.
+That is, 5 days before the event, CAR has no predictive power on the type of event.
+This is consistent with the EMH.
+In contrast, 5 days before the event, CAP can predict the type of event.
+At the event date, the medians of the CAR shift as the abnormal returns jump for both bullish and bearish events.
+Also this is consistent with the EMH.
+Finally, 5 days after the event, the distributions of the CAR are very similar to the ones at the event date.
+Again, this is consistent with the EMH, as all new information is instantaneously embedded into the prices and the returns normalize after the event, immediately.
+The medians of the CAP 5 days after the event exhibit an extended shift compared to the ones at the event date, as investors continue to post about recent past events.
+
 % Please add the following required packages to your document preamble:
 % \usepackage{multirow}
-\begin{table}[]
+\begin{table} 
 \centering
@@
 \end{tabular}
-\caption{Mann-Whitney U-test estimates for pairwise significant differences between distribution medians.
+\mycaption{Mann-Whitney U-test statistics}{Mann-Whitney U-test statistics for pairwise significant differences between distribution medians.
 Under the null hypothesis, the two samples represent two distributions with equal median values.
@@
 
-\section{Portfolios} \label{S:portfolios}
+\clearpage\section{Sentiment-Sorted Portfolios} \label{S:portfolios}
 
-For the construction of a portfolio, in practice, we cannot condition on an event happening over the holding period.
-But we can repeat an experiment many times by supposing there is an event during the holding period and invest according to this hypothesis (long or short according to CAP).
-We would hope that repeating this experiment many times, we will be correct a few times and make large gain (or loss).
-
-Our further intuition is that the larger the magnitude of $\mid$CAP$\mid$, the more likely there will be an event in the holding period.
-We have no statistical evidence for this, but intuitively, it makes sense to assume that the magnitude $\mid$CAP$\mid$ is a  “proxy” for predicting whether an event will happen or not.
-In that sense, we should better calibrate the thresholds such that the portfolio is zero quite often.
-Only if we have reason to believe that there will be an event (large magnitude $\mid$CAP$\mid$) then we invest, according to the sign of CAP.
-Otherwise we stay out of the market.
-
-The data consists of ticker-date instances $(i,t)$, for the universe of tickers $i=1,\dots,19$ and dates $t$ ranging through all business days of the sample period, excluding the first 14 days (for the CAP) and the last day (for the last holding period).
-Every ticker-date instance $(i,t)$ comes with the feature $CAP_{i,t}=\sum_{s=t-14}^t AP_{i,s}$, which is the running CAP over the last 14 days plus current day $t$ (we do end of day $t$ rebalancing).
-Note this is different from the definition of ``$CAP_i(\tau,t)$'' above.
-
-   % \item label $L_{it}$, taking 4 categorical values:
-   % \[ L_{i,t}=\begin{cases} \text{bullish},&\text{if there is a %bullish event of ticker $i$ during $t+1,\dots,t+5$,}\\ %\text{neutral},&\text{if there is a neutral event of ticker $i$ %during $t+1,\dots,t+5$,}\\ \text{no},&\text{if there is no event %of ticker $i$ during $t+1,\dots,t+5$,}\\ \text{bearish},&\text{if %there is a bearish event of ticker $i$ during %$t+1,\dots,t+5$}\end{cases}\]
-
-\subsection{CAP reset after an event}
-
-As CAP continues to shift after an event, for our portfolio construction, we chose to reset it to 0 after every event.
-This ensures that we are not exposed to short-term reversals.
-Let $CAP_{i,t}^{(R)}$ be the feature time-serie of the CAP reset to 0 after every event.
+We assess the economic relevance of the sentiment polarity and construct sorted portfolios. Thereto, we define for every ticker $i$ and day $t$ 
+\begin{equation}
+     CAP_{i,t}=\sum_{s=t-14}^t AP_{i,s},
+\end{equation}
+which is the running CAP over the last 14 days plus the current day $t$ (we rebalance the portfolio at the close on day $t$).\footnote{As above, we work here with the the restricted sample of 19 tickers and dates $t$ ranging through all business days of the sample period, excluding the first 14 days (for the CAP) and the last day (for the last portfolio holding period).} Note the difference to \eqref{eq_CAPtaut}.
+While we cannot predict the arrival of an event, we assume that the more $CAP_{i,t}$ deviates from zero the more likely there will be a an event on the next day.
+We will thus use $CAP_{i,t}$ as a baseline signal for market timing.
+However, as we have seen above, CAP continues to shift after an event.
+To avoid exposures to short-term reversals, we thus reset the running CAP after every event.
 Formally, let $\tau_{i,t}\le t$ denote the most recent past event date by $t$ of ticker $i$.
-Then we define
-\[ CAP_{i,t}^{(R)} = \sum_{s=\max\{t-14,\tau_{i,t}+1\}}^t AP_{i,s} = \begin{cases} CAP_{i,t},&\text{if $\tau_{i,t}<t-14$,} \\ \sum_{s= \tau_{i,t}+1}^t AP_{i,s},&\text{if $t-14\le \tau_{i,t}<t$,}\\
-0,&\text{if $\tau_{i,t}=t$,}\end{cases}\]
+Then we define the reset CAP
+\begin{equation}
+    CAP_{i,t}^{(R)} = \sum_{s=\max\{t-14,\tau_{i,t}+1\}}^t AP_{i,s} = \begin{cases} CAP_{i,t},&\text{if $\tau_{i,t}<t-14$,} \\ \sum_{s= \tau_{i,t}+1}^t AP_{i,s},&\text{if $t-14\le \tau_{i,t}<t$,}\\
+0,&\text{if $\tau_{i,t}=t$,}\end{cases}
+\end{equation}
 where we used the convention that $\sum_{s=t+1}^t \cdot =0$.
 
-\subsection{Cross-sectional thresholds}
+We also define time-varying thresholds on the reset CAP for market timing.
+For every day $t$, we estimate the mean $\mu_t$ and standard deviation $\sigma_t$ of $CAP_{i,t}^{(R)}$ across the 19 tickers $i$.
+For a fixed multiplier $x$, we define $U_t(x) = \mu_t + x \cdot \sigma_t$ the upper threshold, and $L_t(x) = \mu_t - x \cdot \sigma_t$ the lower threshold.
+Figure \ref{fig:cross-sec stats} shows the time series of the cross-sectional mean $\mu_t$ and the 99\% confidence interval, $L_t(x)$ and $U_t(x)$ for $x = 2.58$.
+As a robustness check of our approach, we observe that the mean is well centered at zero.
+We also see a regime change in the early 2015.
+In the first regime the standard deviation is much larger (and more volatile) than in the second regime.\footnote{We could not find an exogenous cause for this regime change.} Appendix~\ref{app_CAPport} contains the results for the 95\% ($x = 1.96$) and 99.5\% ($x = 2.81$) confidence intervals.
 
-We use time-varying thresholds to build a long and a short portfolio.
-For every day, we compute the cross-sectional mean and standard deviations of  $CAP_{i,t}^{(R)}$ across the 19 tickers.
-We will use the mean $\pm$ $x$ standard deviations as and up and down thresholds in the portfolio construction, with $x$ as hyperparameter.
-We define $U_t(x) = \mu_t + x \cdot \sigma_t$ the upper threshold, and $L_t(x) = \mu_t - x \cdot \sigma_t$ the lower threshold, with $\mu_t$ and $\sigma_t$ being the cross-sectional mean and standard deviations, respectively.
-Figure \ref{fig:cross-sec stats} shows the cross-sectional mean and its 99\% confidence interval ($x$ = 2.58).
-The mean is well centered at zero, which speaks for our method of computing CAP.
-It also shows that there are essentially two regimes, with a switch in early 2015.
-In the first regime the standard deviation is much larger (and more volatile) than in the second regime.
-The Appendix contains the results for 95\% ($x$ = 1.96) and 99.5\% confidence intervals ($x$ = 2.81).
-
-\begin{figure}
+\begin{figure}[h]
     \centering
     \includegraphics[width=0.75\textwidth]{cross_sectional_cap.jpg}
-    \caption{Cross-sectional statistics of $CAP^{(R)}$.
-The blue lines shows the cross-sectional mean of the CAP across the 19 tickers over the year.
-The green and red lines show the up and down 99\% confidence band.
-We use them as up and down thresholds for the portfolio construction.
-}
+    \mycaption{Cross-sectional statistics of $CAP_{i,t}^{(R)}$}{The blue line shows the daily cross-sectional mean, the green (red) line shows the cross-sectional mean plus (minus) 2.58 standard deviations, respectively.}
     \label{fig:cross-sec stats}
@@
 
-%\subsection{Long/not-long classification}
-%
-%We want to learn the binary label
-%
-%\[ L^{long}_{i,t} = \begin{cases} \text{long}, &\text{if %$L_{i,t}=\text{bullish}$,}\\\text{not-long}, &\text{if %$L_{i,t}\in\{\text{neutral},\text{no},\text{bearish}\}$,} \end{cases}
-%\]
-%Thereto, we choose a threshold $c^{long}$ and predict $\hat %L^{long}_{i,t}=\text{long}$ if $CAP_{i,t}^{(R)}>c^{long}$.
-%
-%Doing so gives a $2\times 2$ confusion matrix 
-%
-%\begin{table}[H]
-%\centering
-%\begin{tabular}{c|c|c|}
-%predicted label & \multicolumn{2}{c|}{true label}  \\ \hline
-% & long & not-long \\ \hline
-%  long &  TP & FP \\\hline
-%  not-long & FN & TN
-%\end{tabular}
-%\end{table}
-%
-%(this table can be better plotted than here)
-%
-%Compute precision, recall, F1-score, or the Matthews Correlation Coefficient %MCC (see \url{https://en.wikipedia.org/wiki/Phi_coefficient#Confusion_matrix})
-% 
-%Then find $c^{long}$ that maximizes F1-score or MCC
-%
-%
-%\subsection{Short/not-short classification}
-%
-%We want to learn the binary label
-%
-%\[ L^{short}_{i,t} = \begin{cases} \text{short}, &\text{if %$L_{i,t}=\text{bearish}$,}\\\text{not-short}, &\text{if %$L_{i,t}\in\{\text{bullish},\text{neutral},\text{no}\}$,} \end{cases}
-%\]
-%Thereto, we choose a threshold $c^{short}$ and predict $\hat %L^{short}_{i,t}=\text{short}$ if $CAP_{i,t}^{(R)}<c^{short}$.
-%
-%Doing so gives a $2\times 2$ confusion matrix 
-%
-%\begin{table}[H]
-%\centering
-%\begin{tabular}{c|c|c|}
-%predicted label & \multicolumn{2}{c|}{true label}  \\ \hline
-% & short & not-short \\ \hline
-%  short &  TP & FP \\\hline
-%  not-short & FN & TN
-%\end{tabular}
-%\end{table}
-%
-%(this table can be better plotted than here)
-%
-%Compute precision, recall, F1-score, or the Matthews Correlation Coefficient %MCC (see \url{https://en.wikipedia.org/wiki/Phi_coefficient#Confusion_matrix})
-% 
-%Then find $c^{short}$ that maximizes F1-score or MCC
+Based on these signals, we now construct reset CAP-sorted portfolios.
+Formally, we define the ticker sets $I^{bull}_t=\{ i\mid CAP_{i,t}^{(R)} > U_t(x)\}$ and $I^{bear}_t=\{ i\mid CAP_{i,t}^{(R)} < L_t(x)\}$.
+At the close of any day $t$, we form the equally weighted bullish (bearish) portfolio consisting of tickers in $I^{bull}_t$ (in $I^{bear}_t$), and realize the 1-day returns.
+If any of the index sets is empty, we set the corresponding return to zero.
+Top-left plot of Figure \ref{fig:pf} shows the cumulative log returns of bullish and bearish portfolios as well as the S\&P500.
+Overall, the portfolio performance is consistent with our approach: the bullish (bearish) portfolio outperforms (under-performs) the market.
+Remarkably, the upward (downward) steps suggests that our portfolio strategy succeeds to take the right positions just before an event.
+The remaining plots of Figure \ref{fig:pf} show the number of positions across time of our portfolios.
+Most of the returns are earned with portfolios consisting of very few tickers.
+This is a result of our market timing and stock picking strategy: we only invest in the top/bottom percentiles of CAP, whenever our signal is strong enough.
 
-\subsection{Portfolio and returns}
+% \begin{figure*}[h]
+% \centering
+% \begin{subfigure}{0.5\textwidth}  
+%             \centering 
+%             \includegraphics[width=\textwidth]{cross_sectional_cap.jpg}
+%         \end{subfigure}
+%         \quad
+%         \begin{subfigure}{0.5\textwidth}   
+%             \centering 
+%             \includegraphics[width=\textwidth]{cumulative_log_return.jpg}
+%         \end{subfigure}
+%         \caption{Top graph shows the cross-sectional statistics of $CAP_{i,t}^{(R)}$.
+The blue lines shows the cross-sectional mean of the reset CAP across the 19 tickers over the year.
+The green and red lines show the up and down 99\% confidence band.
+We use them as up and lower thresholds for the portfolio construction.
+Bottom graph shows the cumulative log returns of bullish and bearish portfolios for $x =2.58$, and S\&P500.}
+%         \label{fig:pfconstruction258}
+% \end{figure*}
 
-We look at two types of portfolios: long only, and short only that we hold for one day.
+%\begin{figure}[h!]
+%    \centering
+%    \includegraphics[width=0.5\textwidth]{cumulative_log_return.jpg}
+%    \caption{Cumulative log returns of bullish and bearish portfolios for $x =2.58$, and S\&P500.}
+%    \label{fig:cumretpf}
+%\end{figure}
 
-Denote by $R_{i,t+1}=(S_{i,t+1}-S_{i,t})/S_{i,t} - R_{f,t}$ the excess return of ticker $i$ over $[t,t+1]$.
-
-\paragraph{Long only:}
-
-At the end of any business day $t$, we go long all tickers $i$ with $CAP_{i,t}^{(R)} > U_t(x)$.
-Denote by $I^{long}_t=\{ i\mid CAP_{i,t}^{(R)} > U_t(x)\}$ the corresponding index set, which could be empty.
-We then form an equally weighted portfolio and consider the 1-day excess return
-\[ R^{long}_{t,t+1} = \begin{cases} \frac{1}{|I^{long}_t|}\sum_{i\in I^{long}_t} R_{i,t,t+1} ,&\text{if $I^{long}_t\neq \emptyset$.} \\
-0, &\text{if $I^{long}_t = \emptyset$.} \end{cases}\]
-
-\paragraph{Short only:}
-
-At the end of any business day $t$, we go long all tickers $i$ with $CAP_{i,t}^{(R)} < L_t(x)$.
-Denote by $I^{short}_t=\{ i\mid CAP_{i,t}^{(R)} < L_t(x)\}$ the corresponding index set, which could be empty.
-We then form an equally weighted portfolio and consider the 1-day excess return
-\[ R^{short}_{t,t+1} = \begin{cases} \frac{1}{|I^{short}_t|}\sum_{i\in I^{short}_t} R_{i,t,t+1} ,&\text{if $I^{short}_t\neq \emptyset$.} \\
-0, &\text{if $I^{short}_t = \emptyset$.} \end{cases}\]
-
-Now, we compute these returns for all $t$ and plot a time series.
-What we would like to see, of course, is significantly positive excess returns $R^{long}_{t,t+1}$ and negative excess returns $R^{short}_{t,t+1}$.
-
-Figure \ref{fig:cumretpf} shows the cumulative log returns of long and short portfolios as well as the S\&P500.
-Overall, the strategies are correct in both direction : the long portfolio shows positive returns and the short portfolio shows negative returns.
-Remarkably, cumulative log returns are most of the time shaped as step functions, suggesting that the strategy is able to take positions before an event, and close the position after the return is earned due to the CAP reset.
-Figure \ref{fig:pf} shows the number of positions across time of our portfolios.
-Most of the returns are earned with portfolios consisting of very few tickers.
-This is a result of our stock picking strategy : we only invest in the top/bottom percentile of CAP, whenever possible.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.75\textwidth]{cumulative_log_return.jpg}
-    \caption{Cumulative log returns of both portfolios (x=2.58) and the S\&P500.}
-    \label{fig:cumretpf}
-\end{figure}
-
-\begin{figure*}[]
+\begin{figure*}[h]
         \centering
@@
             \centering 
-            \includegraphics[width=\textwidth]{distribution_long_positions.JPG}
-            \caption*{Number of positions - Long portfolio}   
+            \includegraphics[width=\textwidth]{cumulative_log_return.jpg}
+            \caption*{Cumulative log returns}
         \end{subfigure}
@@
             \centering 
-            \includegraphics[width=\textwidth]{distribution_short_positions.JPG}
-            \caption*{Number of positions - Short portfolio}   
+            \includegraphics[width=\textwidth]{distribution_positions.JPG}
+            \caption*{Distribution of number of positions}   
         \end{subfigure}
@@
         \end{subfigure}
-        \caption{Long and Short portfolio for x=2.58.} 
+        \mycaption{Bullish and bearish portfolios for $x=2.58$}{Top left plot shows the cumulative log returns of the portfolios over the years, top right plot shows the distribution of the number of positions in the portfolios, bottom left plot is the number of positions over time and bottom right plot is the daily returns of both portfolios.}
         \label{fig:pf}
@@
 %\input{drawdowns}
-\newpage
+\clearpage
 \section{Conclusion} \label{S:conclusion}
 
-We believe that an accurate and timely estimation of investor sentiment on both firms and aggregate market is an excellent proxy of unobservable firm fundamentals.
-In particular, recent studies on \textit{nowcasting} shows that alternative sources of data can enhance traditional models of stock return and accounting earnings predictions (\citet{challet2013predicting}).
-In this paper, we scrape 90 million messages out of StockTwits during 2010 to 2020.
+We extract 90 million messages from StockTwits from January 2010 to March 2020, covering a large sample of US and Canadian stocks.
 Messages are either user-labeled as bullish or bearish or left unlabeled.
-Using the labeled messages as training set, we build a logistic regression on TFIDF vectorized messages to classify the unlabeled messages in either bullish, neutral or bearish class.
-We observe a 5-for-1 bullish-to-bearish ratio, indicating that investors are on average optimistic.
-Then, we build daily time-series of polarity for both individual firms and the aggregate market.
-We show that changes in daily polarity are strongly associated to changes of the same sign in contemporaneous stock returns, but this result loses its significance against next-day returns.
-However, focused around specific firm events (defined as sudden peak of message volume on a firm), we show that cumulative abnormal polarity has much more predictive power than cumulative abnormal returns.
-We also note that user's sentiment about a firm tend to be biased towards recent past events.
-Finally, as robustness check, we show that event studies on CAAR are consistent with previous literature on EMH.
+Using the user-labeled messages as training set, we run logistic regressions on TFIDF vectorized messages to classify all unlabeled messages as either bullish, neutral or bearish.
+We observe a 5-to-1 bullish-to-bearish ratio, indicating that investors are on average optimistic.
+We build time-series of daily sentiment polarity for individual tickers and the aggregate market.
+We show that daily polarity is positively associated to contemporaneous stock returns, but this result loses its significance against next-day returns.
+We then define events as days with sudden peak of message volume and relate them to corporate and stock market events.
+We show that cumulative abnormal polarity has significant predictive power on the type of event, in contract to cumulative abnormal returns.
+We also note that investor sentiment about a ticker tends to be biased towards recent past events.
+As robustness check, we show that our event study on cumulative abnormal returns is consistent with previous literature on the efficient market hypothesis.
+The performance of sentiment-sorted portfolios illustrates the economic relevance of our sentiment measure.

### 2022-07-24 tail/appendix.tex
Add files via upload

@@
 \section{Appendix to Chapter 1}
+\label{appendix1}
+
 \subsection{Relation between $g_{it}(\tau)$ and $\psi_{it}(\tau)$}
@@
 \begin{proof1}
-The last equation above can be computed with the first derivative of $\psi_{it}(\tau)$ with respect to time. Using definition \ref{eq:psi} we have :
+The last equation above can be computed with the first derivative of $\psi_{it}(\tau)$ with respect to time. Using Definition \ref{eq:psi} we have :
 \begin{align*}
@@
 \subsection{Likelihood function}
-\label{appendix1}
 
@@
 \subsection{Distance-to-Default estimation}
-\label{appendix2}
 
@@
 \end{equation}
-In the figure \ref{fig:DD} taken from \citet{KMV}, the distance to default is denoted by ``DD".
+In the Figure \ref{fig:DD} taken from \citet{KMV}, the distance-to-default is denoted by ``DD".
 The higher the asset value at horizon H the greater the distance-to-default will get, which lowers the default probability since the firm is more likely to be able to repay the debt owed.
@@
 \centering
-\includegraphics[width=6in]{DTD_Crosbie_Bohn.JPG} 
-\mycaption{Distance-to-default}{} 
+\includegraphics[width=0.75\textwidth]{DTD_Crosbie_Bohn.JPG} 
+\mycaption{Distance-to-default}{The distance-to-default is denoted by DD. It is defined as the difference between the expected value of the asset and the default point. } 
 \label{fig:DD}
@@
 \section{Appendix to Chapter 2}
+\label{appendix2}
 \subsection{Technical proofs}
@@
 \section{Appendix to Chapter 3}
-\subsection{Number of messages}
-As the same information sometimes refer to several companies, users are allowed to identify more than one ticker per message.
-Figure \ref{fig:histtic} shows the histogram of the number of tickers tagged per message.
+
+\subsection{Tutorial for StockTwits Messages Extraction}\label{app_tut}
+
+We use stock price data from CRSP/Compustat of all US and Canadian listed stocks from January 2010 to March 2020.
+From this dataset, we create the list of unique tickers for which we will extract messages.
+We will later be able to merge the two datasets using the date and ticker for every observation.
+We use the StockTwits Application Programming Interface (API) to download messages from StockTwits.
+One query on StockTwits API is called a JavaScript Object Notation (JSON) request.
+Every message on StockTwits has a unique identifier ("msg\_id") posted by a user with a unique identifier ("user\_id").
+JSON requests allow to query the database by ticker (called "symbol method") or by user (called "user method").
+We use the query by ticker.
+One query only outputs the latest 30 messages concerning that ticker.
+However, it is possible to set a parameter ("max") to output the latest 30 messages up to this particular message identifier.
+This parameter allows us to crawl the message history of a ticker by recursively changing the "max" parameter to the oldest message identifier in the query.
+To perform a JSON request for Apple (AAPL) up to the message identifier 30'000'000, simply enter the following URL in a browser : \url{https://api.stocktwits.com/api/2/streams/symbol/AAPL.json?\&max=30000000}.
+The page we get looks unreadable but it has always the same structure : several pairs of keys and values.
+The structure of JSON can easily be interpreted by modern programming languages.
+We create a Python script to query the API and extract the message history of every ticker in the ticker list.
+We store the output of every JSON request in .txt files in dedicated ticker folders.
+
+\subsection{Message Count}\label{app_counting}
+
+A StockTwits message can refer to multiple tickers.
+Figure \ref{fig:tickerpermessage} shows the histogram of the number of tickers tagged per message.
 As the vast majority of message includes only one ticker, we only show on this plot messages referring to more than one ticker.
-However, many messages refer to several tickers and this creates duplicates in the database because we consider the same message for all tickers tagged in the message.
-Figure \ref{fig:uniq} shows the number of messages with and without double counting.
+The maximum number of tickers per message amounts to 28 and corresponds to 11 messages in the sample.
+Many messages refer to several tickers and this creates duplicates in the database because we consider the same message for all tickers tagged in the message.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.75\textwidth]{hist_tic.jpg}
+    \mycaption{Histogram of the number of tickers per message}{Histogram of the number of tickers per message, across all messages referring to more than one ticker.}
+    \label{fig:tickerpermessage}
+\end{figure}
+
+Left plot of Figure \ref{fig:messagecount} shows the number of messages with and without double counting.
 In our sample, the number of messages without double counting is 76 million, as opposed to 90 million messages with double counting.
-Figure \ref{fig:ratio} shows the ratio between the number of messages with double counting divided by the number of messages without double counting.
+Right plot of Figure \ref{fig:messagecount} shows the ratio between the number of messages with double counting and the number of messages without double counting.
 Throughout this paper, we only refer to the number of messages with double counting.
 
-\begin{figure}
+\begin{figure*}[h!]
 \centering
-    \includegraphics[width=0.75\textwidth]{hist_tic.JPG}
-    \mycaption{Number of tickers per message}{Histogram of the number of tickers per message, across all messages referring to more than one ticker.
+        \begin{subfigure}{0.49\textwidth}   
+            \centering 
+            \includegraphics[width=\textwidth]{uniq.jpg}
+        \end{subfigure}
+        \begin{subfigure}{0.49\textwidth}   
+            \centering 
+            \includegraphics[width=\textwidth]{ratio.jpg}
+        \end{subfigure}
+        \mycaption{Number of messages}{Left plot shows the total number of messages with double counting (red) compared to the total number of messages without double counting (blue).
+Right plot shows the ratio between the number of messages with double counting and the number of messages without double counting.
+Numbers are aggregated daily.}
+        \label{fig:messagecount}
+\end{figure*}
+
+%\begin{figure}[h]
+%    \centering
+%    \includegraphics[width=0.75\textwidth]{hist_tic.JPG}
+%    \caption{Histogram of the number of tickers per message, across all messages referring to more than one ticker.
 The maximum number of tickers per message amounts to 28 and corresponds to 11 messages in the sample.}
-    \label{fig:histtic}
-\end{figure}
+%    \label{fig:histtic}
+%\end{figure}
 
-\begin{figure}
+%\begin{figure}[h]
+%    \centering
+%    \includegraphics[width=0.75\textwidth]{uniq.JPG}
+%    \caption{Total number of messages with double counting (red) compared to the total number of messages without double counting (blue). Numbers are aggregated daily.}
+%    \label{fig:uniq}
+%\end{figure}
+
+%\begin{figure}[h]
+%    \centering
+%    \includegraphics[width=0.75\textwidth]{ratio.JPG}
+%    \caption{Ratio between the number of messages with double counting and the number of messages without double counting. Numbers are aggreaged daily.}
+%    \label{fig:ratio}
+%\end{figure}
+
+%\newpage
+
+%\clearpage
+\subsection{User Summary Statistics}\label{app_userstats}
+
+Left plot of Figure \ref{fig:data-stocktwits1} shows a log-log histogram of the number of followers per user and a right plot shows a log-log histogram of the number of messages posted by users.
+There are a few users with many followers (they can be seen as ``influencers"), and many users with a few followers.
+In addition, most users seem to post on average between 10 and 400 messages and a few post a lot more.
+Overall, this appears to be a well balanced network structure.
+A more detailed study of the network effects on market sentiment is beyond the scope of this paper.
+
+\begin{figure*}[h]
 \centering
-    \includegraphics[width=0.75\textwidth]{uniq.JPG}
-    \mycaption{Number of messages}{Total number of messages with double counting (red) compared to the total number of messages without double counting (blue). Numbers are aggregated daily.}
-    \label{fig:uniq}
-\end{figure}
+        \begin{subfigure}{0.49\textwidth}   
+            \centering 
+            \includegraphics[width=\textwidth]{Followers.JPG}
+        \end{subfigure}
+        \begin{subfigure}{0.49\textwidth}
+            \centering
+            \includegraphics[width=\textwidth]{Tweets.jpg}
+        \end{subfigure}
 
-\begin{figure}
-    \centering
-    \includegraphics[width=0.75\textwidth]{ratio.JPG}
-    \mycaption{Number of messages II}{Ratio between the number of messages with double counting and the number of messages without double counting. Numbers are aggreaged daily.}
-    \label{fig:ratio}
-\end{figure}
+        \mycaption{User summary statistics}{Left graph is a log-log histogram of the number of followers per user and the right graph shows the log-log histogram of the number of messages posted by users.}
+        \label{fig:data-stocktwits1}
+\end{figure*}
 
-\newpage
+%\clearpage
+\subsection{Anomalies} \label{app_ano}
 
-\subsection{Coverage}
+We discuss here two anomalies that appear in the word clouds in Figure \ref{wordclouds}.
+
+The term ``aldox'' in the bullish cloud caught our attention.
+After some research, it is an abbreviation for Aldoxorubicin, a drug against tumors and is associated with pharmaceutical messages where investors were very enthusiastic about it.
+An example of a related message is ``aldox is on the slide.
+have great faith this is truly world change".
+That is why the term is appearing almost exclusively in bullish messages, hence in the bullish cloud.
+
+The bearish cloud contains the term "long position open", which seems like a bullish signal.
+Closer inspection shows that this term frequently appears in bearish user-labeled messages of intraday alerts such as ``sell \$labd close labd long position.
+open labd short position.
+time: 14:53 ny price: \$13.64 zquant intraday alerts''.
+However, this anomaly is not an issue.
+We tested what happened when "long position open" is fed as a message into our sentiment classifier.
+As a message, it consists of the trigram "long position open", the two bigrams ``long position'' and ``position open'', and the single words as unigrams.
+This results in a bullish score of 0.91 and the message is---correctly---classified as bullish.
+
+%\clearpage
+\subsection{Coverage}\label{app_coverage}
 
 Stocktwits is neither regulated nor moderated, so one needs to filter the information that we use.
-Even if Stocktwits has valuable information from respected contributors, a blog\footnote{\url{https://www.warriortrading.com/stocktwits-review}} describes the concerns that may rise when using Stocktwits as a financial information provider, namely self-promotion, lack of credibility and other noise.
+Even if Stocktwits has valuable information from respected contributors, a blog\footnote{\url{https://www.warriortrading.com/stocktwits-review}, last accessed on 1st of July 2022.} describes the concerns that may rise when using Stocktwits as a financial information provider, namely self-promotion, lack of credibility and other noise.
 To diversify noise and better extract information, we exclude from our sample tickers that are rarely discussed.
@@
 We chose a median threshold of 50 as a balanced trade-off between noise and coverage.
-Table \ref{tab:listcov} shows the list of tickers covered and associated market capitalization as of 31st of December 2019.
-To avoid bias in the polarity, we restrict the universe and focus on the most discussed tickers.
-On top of that, it appears that the most discussed tickers are not always big firms, so it is has the advantage of covering all sizes of firm and we avoid big-firm bias.
-Also, we do not consider only firms but also ETFs on alternative investments.
-Finally, we cover several sectors so even if we have a restrictive universe, it is well diversified.
 
@@
     \includegraphics[width=0.75\textwidth]{trim.JPG}
-    \mycaption{Coverage}{Coverage as a function of the median threshold. A lower threshold increases the coverage at the expense of a bigger bias in the polarity.}
+    \mycaption{Ticker coverage}{Coverage as a function of the median threshold. A lower threshold increases the coverage at the expense of a bigger bias in the polarity.}
     \label{fig:coverage}
 \end{figure}
+
+Table \ref{tab:listcov} shows the list of the 19 tickers above this threshold and their associated market capitalization as of 31st of December 2019.
+It appears that these most discussed tickers cover all sizes of stock, and hence we avoid big-firm bias.
+Also, it includes not only single firms but also ETFs on alternative investments.
+Finally, we cover several sectors so even if we have a restrictive universe, it is well diversified.
 
@@
 \end{tabular}
-\mycaption{Coverage}{Coverage after the trimming process. List of tickers and corresponding market capitalization as of 31st of December 2019.}
+\mycaption{Ticker coverage}{Coverage after the trimming process. List of tickers and corresponding market capitalization as of 31st of December 2019.}
 \label{tab:listcov}
@@
 
-\newpage
-\subsection{Tutorial for StockTwits messages extraction}
-We use CRSP to get stock prices of all US and Canadian listed firms from 1990 to 2020.
-Out of this dataset, we create the list of unique tickers for which we will extract messages.
-We will later be able to merge the two datasets using the date and ticker for every observation.
-We use the StockTwits Application Programming Interface (API) to collect messages from StockTwits.
-One query on StockTwits API is called a JavaScript Object Notation (JSON) request.
-Every message on StockTwits has a unique identifier ("msg\_id") posted by a user with a unique identifier ("user\_id").
-JSON requests allow to query the database by ticker (called "symbol method") or by user (called "user method").
-We use the query by ticker.
-One query only outputs the latest 30 messages concerning that ticker.
-However, it is possible to set a parameter ("max") to output the latest 30 messages up to this particular message identifier.
-This parameter allows us to crawl the message history of a ticker by recursively changing the "max" parameter to the oldest message identifier in the query.
-To perform a JSON request for Apple (AAPL) up to the message identifier 30'000'000, simply enter the following URL in a browser : https://api.stocktwits.com/api/2/streams/symbol/AAPL.json?\&max=30000000.
-The page we get looks unreadable but it has always the same structure : several pairs of keys and values.
-The structure of JSON can easily be interpreted by modern programming languages.
-We create a Python script to query the API and extract the message history of every ticker in the ticker list.
-We store the output of every JSON request in .txt files in dedicated ticker folders.
+\clearpage
+\subsection{Classifier Performance}\label{app_class}
+
+We first recap the definition of the basic performance measures for a binary classifier.
+First, one has to choose one class as the positive class.
+Instances (messages) are then divided according to their predicted and actual labels into true positives TP (predicted positive, actual positive), false positives FP (predicted positive, actual negative), true negatives TN (predicted negative, actual negative), and false negatives FN (predicted negative, actual positive).
+Precision $PRE=\dfrac{TP}{TP+FP}$ is the proportion of true positives among the predicted positives.
+Recall $REC = \dfrac{TP}{TP+FN}$ is the proportion of true positives among the actual positives.
+The precision-recall trade-off is captured by the F1 score, $\dfrac{2 \cdot PRE \cdot REC}{PRE + REC}$, the harmonic mean of precision and recall.
+
+Tables \ref{classif-out} and \ref{classif-in} show the confusion matrices of our combined classifier out-of-sample and in-sample, respectively.
+We define accuracy as the fraction of correct predictions, omitting the messages with a predicted neutral sentiment.
+We thus obtain an out-of-sample accuracy of 85.9\%.
+The in-sample accuracy is 87.4\%.
+
+\begin{table}[H]
+\centering
+\begin{tabular}{cl|ll|}
+\cline{3-4}
+\multicolumn{1}{l}{}                        &         & \multicolumn{2}{c|}{True}               \\ \cline{3-4}
+\multicolumn{1}{l}{}                        &         & \multicolumn{1}{l|}{Bullish} & Bearish  \\ \cline{1-4}
+\multicolumn{1}{|c|}{\multirow{3}{*}{Predicted}} & Bullish & 3'555'896                      & 155'280   \\ \cline{2-2}
+\multicolumn{1}{|c|}{}                      & Neutral & 663'541                       & 160'423    \\ \cline{2-2}
+\multicolumn{1}{|c|}{}                      & Bearish & 550'928                       & 746'261  \\ \cline{1-4}
+\end{tabular}
+\mycaption{Confusion matrix for the combined classifier out-of-sample}{Rows are the predicted class of the combined classifier, columns are the user-labels.
+Values are the number of messages in the corresponding category.}
+\label{classif-out}
+\end{table}
+
+\begin{table}[H]
+\centering
+\begin{tabular}{cl|ll|}
+\cline{3-4}
+\multicolumn{1}{l}{}                        &         & \multicolumn{2}{c|}{True}               \\ \cline{3-4}
+\multicolumn{1}{l}{}                        &         & \multicolumn{1}{l|}{Bullish} & Bearish   \\ \cline{1-4}
+\multicolumn{1}{|c|}{\multirow{3}{*}{Predicted}} & Bullish &  14'433'375                            & 532'509    \\ \cline{2-2}
+\multicolumn{1}{|c|}{}                      & Neutral &  2'656'730                            &  642'329  \\ \cline{2-2}
+\multicolumn{1}{|c|}{}                      & Bearish &  1'992'535                            &  3'071'837 \\ \cline{1-4}
+\end{tabular}
+\mycaption{Confusion matrix for the combined classifier in-sample}{Rows are the predicted class of the combined classifier, columns are the user-labels.
+Values are the number of messages in the corresponding category.}
+\label{classif-in}
+\end{table}
 
@@
 %So the risky profit seeking investor prefers a large recall.
-\clearpage
-\newpage
 
-\subsection{Portfolio construction - various $x$}
-As the thresholds $U_t(x)$ and $L_t(x)$ are functions of the hyperparameter $x$, we provide for robustness check the results of our portfolio construction with $x$ = 1.96  (95\% confidence band) and $x$ = 2.81 (99.5\% confidence band).
-\subsubsection{x = 1.96}
+%\clearpage
+\subsection{Examples of Classified Messages}\label{app_exampleclass}
 
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.75\textwidth]{cumulative_log_return1.jpg}
-    \mycaption{Cumulative log returns of both portfolios (x=1.96) and the S\&P500.}{}
-    \label{fig:cumpf1}
-\end{figure}
+Here are some representative examples of classifications.
+Typical messages classified as bullish contain terms such as “buy buy” or “hope the pump come soon”.
+Whereas typical bearish messages contain terms such as “sell everything” or “start short position here”.
+Neutral messages are either empty, or irrelevant to finance (e.g., ``political posturing friend''\footnote{This is a reply to the message ``honestly, how dumb can you be to believe that china was going to buy significant amount of agricultural products after the breakdown in trade talks.
+Even if they buy it will be just a little bit and not significant''}), or ambiguous (e.g., “lol wow”).
+
+%\clearpage
+\subsection{Sentiment-Sorted Portfolios for Various Thresholds}\label{app_CAPport}
+As the thresholds $U_t(x)$ and $L_t(x)$ are functions of the hyperparameter $x$, we provide for robustness check the results of our CAP-sorted portfolios for the values $x = 1.96$  (95\% confidence band) in Figure \ref{fig:pf1}, and for $x = 2.81$ (99.5\% confidence band) in Figure \ref{fig:pf3}.
+The portfolio performance arguably depends on the choice of $x$.
+In particular, the smaller $x$ the more likely the bearish portfolio exhibits positive returns.
+On the other hand, the larger $x$ the more likely the bullish portfolio misses the opportunities of positive returns.
+A careful gauging of $x$, possibly asymmetric in bullish and bearish, is therefore required for a real-world implementation of these strategies.
+Results could also improve for a larger cross-section of stocks than the 19 of our reduced sample.
+
+%\begin{figure}[h]
+%    \centering
+%    \includegraphics[width=0.5\textwidth]{cumulative_log_return1.jpg}
+%    \caption{Cumulative log returns of bullish and bearish portfolios for $x=1.96$, and S\&P500.}
+%    \label{fig:cumpf1}
+%\end{figure}
 
@@
             \centering 
-            \includegraphics[width=\textwidth]{distribution_long_positions1.jpg}
-            \caption*{Number of positions - Long portfolio}   
+    \includegraphics[width=\textwidth]{cumulative_log_return1.jpg}
+    \caption*{Cumulative log returns}
         \end{subfigure}
@@
             \centering 
-            \includegraphics[width=\textwidth]{distribution_short_positions1.jpg}
-            \caption*{Number of positions - Short portfolio}   
+            \includegraphics[width=\textwidth]{distribution_positions1.jpg}
+            \caption*{Distribution of number of positions}   
         \end{subfigure}
@@
         \end{subfigure}
-        \mycaption{Long and Short portfolios for x = 1.96.}{}
+        \mycaption{Bullish and bearish portfolios for $x = 1.96$}{Top left plot shows the cumulative log returns of the portfolios over the years, top right plot shows the distribution of the number of positions in the portfolios, bottom left plot is the number of positions over time and bottom right plot is the daily returns of both portfolios.}
         \label{fig:pf1}
@@
 
-\clearpage
-\subsubsection{x = 2.81}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.75\textwidth]{cumulative_log_return3.jpg}
-    \mycaption{Cumulative log returns of both portfolios (x=2.81) and the S\&P500.}{}
-    \label{fig:cumpf3}
-\end{figure}
+%\begin{figure}[h]
+%    \centering
+%    \includegraphics[width=0.5\textwidth]{cumulative_log_return3.jpg}
+%    \caption{Cumulative log returns of bullish and bearish portfolios for $x=2.81$, and S\&P500.}
+%    \label{fig:cumpf3}
+%\end{figure}
 
@@
             \centering 
-            \includegraphics[width=\textwidth]{distribution_long_positions3.jpg}
-            \caption*{Number of positions - Long portfolio}   
+        \includegraphics[width=\textwidth]{cumulative_log_return3.jpg}
+        \caption*{Cumulative log returns}
         \end{subfigure}
@@
             \centering 
-            \includegraphics[width=\textwidth]{distribution_short_positions3.jpg}
-            \caption*{Number of positions - Short portfolio}   
+            \includegraphics[width=\textwidth]{distribution_positions3.jpg}
+            \caption*{Distribution of number of positions}   
         \end{subfigure}
@@
         \end{subfigure}
-        \mycaption{Long and Short portfolio for x=2.81.}{}
+        \mycaption{Bullish and bearish portfolios for $x=2.81$}{Top left plot shows the cumulative log returns of the portfolios over the years, top right plot shows the distribution of the number of positions in the portfolios, bottom left plot is the number of positions over time and bottom right plot is the daily returns of both portfolios.}
         \label{fig:pf3}
     \end{figure*}
+
+% \newpage
+
+% \subsection*{To discuss}
+
+% \begin{figure}[H]
+%     \centering
+%     \includegraphics[width=0.75\textwidth]{opti_thresh_monday.JPG}
+%     \caption{F1-score of long/not-long classifier in green, short/not-short classifier in red.
+Blue circles are optimal thresholds, maximizing F1 score.
+CAP is computed over the last 15 days and portfolio is rebalanced every Monday.}
+%     \label{fig:my_label}
+% \end{figure}
+
+% \begin{table}[H]
+% \centering
+% \begin{tabular}{ll|lll|l}
+% \cline{3-5}
+%                               &       & \multicolumn{3}{c|}{event type}                                       &  \\ \cline{3-5}
+%                               &       & \multicolumn{1}{l|}{bullish} & \multicolumn{1}{l|}{neutral} & bearish &  \\ \cline{1-5}
+% \multicolumn{1}{|c|}{}         & long  & \multicolumn{1}{l|}{}      & \multicolumn{1}{l|}{20}      & 21      &  \\ \cline{2-5}
+% \multicolumn{1}{|c|}{position} & none  & \multicolumn{1}{l|}{275}     & \multicolumn{1}{l|}{275}     & 252     &  \\ \cline{2-5}
+% \multicolumn{1}{|l|}{}         & short & \multicolumn{1}{l|}{33}      & \multicolumn{1}{l|}{19}      & 88      &  \\ \cline{1-5}
+% \end{tabular}
+% \caption{Confusion matrix for the portfolio rebalanced every Monday.
+We build a long-short portfolio based on CAP over the last 15 days.
+Long-short positions are determined using optimal CAP thresholds, obtained by maximizing F1-Score of two distinct classifiers.}
+% \end{table}
+
+% \begin{figure}[H]
+%     \centering
+%     \includegraphics[width=0.75\textwidth]{opti_thresh_tuesday.JPG}
+%     \caption{F1-score of long/not-long classifier in green, short/not-short classifier in red.
+Blue circles are optimal thresholds, maximizing F1 score.
+CAP is computed over the last 15 days and portfolio is rebalanced every Tuesday.}
+%     \label{fig:my_label}
+% \end{figure}
+
+% \begin{table}[H]
+% \centering
+% \begin{tabular}{ll|lll|l}
+% \cline{3-5}
+%                               &       & \multicolumn{3}{c|}{event type}                                       &  \\ \cline{3-5}
+%                               &       & \multicolumn{1}{l|}{bullish} & \multicolumn{1}{l|}{neutral} & bearish &  \\ \cline{1-5}
+% \multicolumn{1}{|c|}{}         & long  & \multicolumn{1}{l|}{71}      & \multicolumn{1}{l|}{34}      & 13      &  \\ \cline{2-5}
+% \multicolumn{1}{|c|}{position} & none  & \multicolumn{1}{l|}{304}     & \multicolumn{1}{l|}{296}     & 278     &  \\ \cline{2-5}
+% \multicolumn{1}{|l|}{}         & short & \multicolumn{1}{l|}{21}      & \multicolumn{1}{l|}{36}      & 70      &  \\ \cline{1-5}
+% \end{tabular}
+% \caption{Confusion matrix for the portfolio rebalanced every Tuesday.
+We build a long-short portfolio based on CAP over the last 15 days.
+Long-short positions are determined using optimal CAP thresholds, obtained by maximizing F1-Score of two distinct classifiers.}
+% \end{table}
+
+% \begin{figure}[H]
+%     \centering
+%     \includegraphics[width=0.75\textwidth]{opti_thresh_wednesday.JPG}
+%     \caption{F1-score of long/not-long classifier in green, short/not-short classifier in red.
+Blue circles are optimal thresholds, maximizing F1 score.
+CAP is computed over the last 15 days and portfolio is rebalanced every Wednesday.}
+%     \label{fig:my_label}
+% \end{figure}
+
+% \begin{table}[H]
+% \centering
+% \begin{tabular}{ll|lll|l}
+% \cline{3-5}
+%                               &       & \multicolumn{3}{c|}{event type}                                       &  \\ \cline{3-5}
+%                               &       & \multicolumn{1}{l|}{bullish} & \multicolumn{1}{l|}{neutral} & bearish &  \\ \cline{1-5}
+% \multicolumn{1}{|c|}{}         & long  & \multicolumn{1}{l|}{88}      & \multicolumn{1}{l|}{50}      & 21      &  \\ \cline{2-5}
+% \multicolumn{1}{|c|}{position} & none  & \multicolumn{1}{l|}{296}     & \multicolumn{1}{l|}{299}     & 281     &  \\ \cline{2-5}
+% \multicolumn{1}{|l|}{}         & short & \multicolumn{1}{l|}{20}      & \multicolumn{1}{l|}{17}      & 59      &  \\ \cline{1-5}
+% \end{tabular}
+% \caption{Confusion matrix for the portfolio rebalanced every Wednesday.
+We build a long-short portfolio based on CAP over the last 15 days.
+Long-short positions are determined using optimal CAP thresholds, obtained by maximizing F1-Score of two distinct classifiers.}
+% \end{table}
+
+% \begin{figure}[H]
+%     \centering
+%     \includegraphics[width=0.75\textwidth]{opti_thresh_thursday.JPG}
+%     \caption{F1-score of long/not-long classifier in green, short/not-short classifier in red.
+Blue circles are optimal thresholds, maximizing F1 score.
+CAP is computed over the last 15 days and portfolio is rebalanced every Thursday.}
+%     \label{fig:my_label}
+% \end{figure}
+
+% \begin{table}[H]
+% \centering
+% \begin{tabular}{ll|lll|l}
+% \cline{3-5}
+%                               &       & \multicolumn{3}{c|}{event type}                                       &  \\ \cline{3-5}
+%                               &       & \multicolumn{1}{l|}{bullish} & \multicolumn{1}{l|}{neutral} & bearish &  \\ \cline{1-5}
+% \multicolumn{1}{|c|}{}         & long  & \multicolumn{1}{l|}{188}     & \multicolumn{1}{l|}{123}     & 73      &  \\ \cline{2-5}
+% \multicolumn{1}{|c|}{position} & none  & \multicolumn{1}{l|}{196}     & \multicolumn{1}{l|}{223}     & 228     &  \\ \cline{2-5}
+% \multicolumn{1}{|l|}{}         & short & \multicolumn{1}{l|}{20}      & \multicolumn{1}{l|}{20}      & 60      &  \\ \cline{1-5}
+% \end{tabular}
+% \caption{Confusion matrix for the portfolio rebalanced every Thursday.
+We build a long-short portfolio based on CAP over the last 15 days.
+Long-short positions are determined using optimal CAP thresholds, obtained by maximizing F1-Score of two distinct classifiers.}
+% \end{table}
+
+% \begin{figure}[H]
+%     \centering
+%     \includegraphics[width=0.75\textwidth]{opti_thresh_friday.JPG}
+%     \caption{F1-score of long/not-long classifier in green, short/not-short classifier in red.
+Blue circles are optimal thresholds, maximizing F1 score.
+CAP is computed over the last 15 days and portfolio is rebalanced every Friday.}
+%     \label{fig:my_label}
+% \end{figure}
+
+% \begin{table}[H]
+% \centering
+% \begin{tabular}{ll|lll|l}
+% \cline{3-5}
+%                               &       & \multicolumn{3}{c|}{event type}                                       &  \\ \cline{3-5}
+%                               &       & \multicolumn{1}{l|}{bullish} & \multicolumn{1}{l|}{neutral} & bearish &  \\ \cline{1-5}
+% \multicolumn{1}{|c|}{}         & long  & \multicolumn{1}{l|}{94}      & \multicolumn{1}{l|}{43}      & 21      &  \\ \cline{2-5}
+% \multicolumn{1}{|c|}{position} & none  & \multicolumn{1}{l|}{284}     & \multicolumn{1}{l|}{283}     & 263     &  \\ \cline{2-5}
+% \multicolumn{1}{|l|}{}         & short & \multicolumn{1}{l|}{26}      & \multicolumn{1}{l|}{40}      & 77      &  \\ \cline{1-5}
+% \end{tabular}
+% \caption{Confusion matrix for the portfolio rebalanced every Friday.
+We build a long-short portfolio based on CAP over the last 15 days.
+Long-short positions are determined using optimal CAP thresholds, obtained by maximizing F1-Score of two distinct classifiers.}
+% \end{table}
+
+%\begin{figure}
+%    \centering
+%    \includegraphics[width=\textwidth]{marketvol_relchange.JPG}
+%    \caption{Caption}
+%    \label{fig:my_label}
+%\end{figure}
+%
+%\begin{figure}
+%    \centering
+%    \includegraphics[width=\textwidth]{aapl_vol_relchange.JPG}
+%    \caption{Caption}
+%    \label{fig:my_label}
+%\end{figure}
+%
+%\begin{figure}
+%    \centering
+%    \includegraphics[width=\textwidth]{rel_activity_volume.JPG}
+%    \caption{Caption}
+%    \label{fig:my_label}
+%\end{figure}

### 2022-07-24 head/abstracts.tex
Add files via upload

@@
 The third chapter, titled `StockTwits Classified Sentiment and Stock Returns` is a co-authored work with Damir Filipović (SFI \& EPFL).
-We scrape 90 million messages from StockTwits over 10 years and classify them into bullish, bearish or neutral classes to create firm-individual sentiment polarity time-series.
+We classify the sentiment of a large sample of StockTwits messages as bullish, bearish or neutral, and create a stock-aggregate daily sentiment polarity measure.
 Polarity is positively associated with contemporaneous stock returns.
-On average, polarity is not able to predict next-day stock returns but when we focus on specific events (defined as sudden peak of message volume), polarity has predictive power on abnormal returns.
+On average, polarity is not able to predict next-day stock returns.
+But when we focus on specific events, defined as sudden peaks of message volume, polarity has predictive power on abnormal returns.
+Polarity-sorted portfolios illustrate the economic relevance of our sentiment measure.
 \\
@@
 Le troisième chapitre, intitulé `StockTwits Classified Sentiment and Stock Returns` est un travail conjoint avec Damir Filipović (SFI \& EPFL).
-Nous récupérons 90 millions de messages (couvrant 10 ans) provenant de StockTwits et les classifions dans la classe haussière, baissière ou neutre pour créer des séries temporelles de polarité propres à chaque entreprise.
+Nous classifions le sentiment d'un large échantillon de messages provenant de StockTwits dans la classe haussière, baissière ou neutre pour créer des séries temporelles de polarité propres à chaque entreprise.
 La polarité est associée positivement aux rendements d'actions contemporains.
-En moyenne, la polarité n'est pas capable de prédire les rendements du jour suivant mais lorsque nous nous focalisons sur des évenements spécifiques (définis par une augmentation soudaine du volume de messages), la polarité a de la puissance prédictive sur les rendements anormaux.
-\\
+En moyenne, la polarité n'est pas capable de prédire les rendements du jour suivant mais lorsque nous nous focalisons sur des évenements spécifiques, définis par une augmentation soudaine du volume de messages, la polarité a de la puissance prédictive sur les rendements anormaux.
+Nous illustrons la pertinence économique de notre mesure de sentiment avec des portefeuilles triés par polarité.\\
 

### 2022-07-24 head/titlepage.tex
Add files via upload

@@
 \null\vspace{2cm}
-{\huge Machine Learning and Risk Management} \\[24pt] 
+{\huge Financial Risk Management with Machine Learning} \\[24pt] 
 \textcolor{gray}{\small{THIS IS A TEMPORARY TITLE PAGE \\ It will be replaced for the final print by a version \\ provided by the registrar's office.}}
@@
 %
-    Prof Name Surname, président du jury\\
-    Prof Name Surname, directeur de thèse\\
-    Prof Name Surname, rapporteur\\
-    Prof Name Surname, rapporteur\\
-    Prof Name Surname, rapporteur\\[12pt]
+    Prof Pierre Collin-Dufresne, président du jury\\
+    Prof Damir Filipović, directeur de thèse\\
+    Prof Michael Rockinger, rapporteur\\
+    Prof Simon Scheidegger, rapporteur\\
+    Prof Andreas Fuster, rapporteur\\[12pt]
 %

### 2022-08-08 main/ch1_introduction.tex
Add files via upload

@@
 \citet{DSW} generate future random values for the covariates using a VAR process while \citet{Duan2012} relax this assumption and use forward intensities.
-The latter specifies the intensities as a linear function of state variables and uses maximum likelihood to estimate the parameters.
+The latter specify the intensities as a linear function of state variables and uses maximum likelihood to estimate the parameters.
 The first chapter of this thesis extends existing literature by removing the assumption of linear intensities and uses artificial neural networks to estimate the intensities of the Poisson processes.
@@
 Increasing the network's width and depth allows for non-linearities and out-of-sample Lorenz Curves show that the neural network's approach outperforms the linear assumption for every horizon.
-Finally, I show what are the most important predictors of default in the short and longer term.
+Finally, I show what the most important predictors of default in the short and longer term are.
 Interdepencies in a network are at the heart of systemic risk.
@@
 The output of this approach is a directed graph that visualizes the interconnections among a set of time series.
-Computing DI has both computational and sample complexity which makes it not suitable for inferring the causal structure of large networks.
-To overcome this problem, we develop a novel approach based on recurrent neural networks that reduces complexity of evaluating DI in high-dimensional settings.
-We show that our approach performs well both in a linear and non-linear simulated enviornments, then apply it to infer the causal relationships among US firms from 1990 to 2020.
+Computing DI has high computational and sample complexity which makes it not suitable for inferring the causal structure of large networks.
+To overcome this problem, we develop a novel approach based on recurrent neural networks that reduces the complexity of evaluating DI in high-dimensional settings.
+We show that our approach performs well both in linear and non-linear simulated environments, then apply it to infer the causal relationships among US firms from 1990 to 2020.
 \\
@@
 We scrape a large sample of messages from Stocktwits, a microblogging platform similar to Twitter but designed for finance professionals.
-One of the challenge in this context is to create a classifier that understands the vocabulary of the message posted by the users.
-After preprocessing steps, we use TFIDF vectorization to compute the importance of each word in the message.
+One of the challenges in this context is to create a classifier that understands the vocabulary of the messages posted by the users.
+After preprocessing steps, we use TFIDF vectorization to compute the importance of each word in a message.
 This transforms the messages from words to a vector of numbers which is now in the same dimension as the vocabulary.
@@
 This procedure allows us to create an artificial neutral class that absorbs all the tweets that do not convey financial information.
-Finally, with daily intervals, we aggregate the predicted sentiments per ticker to compute daily polarity time-series.
-We then use the daily volume of messages on a given firm to identify sudden peak of activity, indicating a firm event.
+Finally, with daily intervals, we aggregate the predicted sentiments per ticker to compute daily polarity time series.
+We then use the daily volume of messages on a given firm to identify sudden peaks of activity, indicating a firm event.
 Computing cumulative average abnormal return and cumulative abnormal polarity in a 41 days window centered at the identified event, we show that abnormal polarities have significant predictive power on the type of event.

### 2022-08-08 main/conclusion.tex
Add files via upload

\cleardoublepage
\chapter*{Conclusion}
\markboth{Conclusion}{Conclusion}
\addcontentsline{toc}{chapter}{Conclusion}

The three chapters presented in this thesis use neural networks and natural language processing to develop or improve financial risk models.

The first chapter contributes to the literature on reduced-form models for multiperiod corporate default prediction using a doubly stochastic formulation. The probability of default in a specific horizon is a function of the forward default intensity and of the forward combined intensity. Previous literature proposes a model to predict corporate default at multiple horizons by estimating these forward intensities via maximum likelihood. To do so, they use a linear assumption in the relationship between the variables and the forward intensities. I show in the first chapter that a significant improvement is achieved by relaxing the linear assumption and using an artificial neural network instead. To allow comparison with the benchmark model, I choose to work with a similar set of features consisting of firm-specific accounting variables and macroeconomic variables. However, a growing area of study is building market sentiment measure and is showing their predictive power on the stock market. A potential and interesting venue for future research would be to gauge the predictive power of such features for default prediction.

The second chapter investigates the interconnections among a set of financial instutions. We propose a recurrent neural network approach to reduce the computational complexity of computing directed information. This approach is well-suited to infer the causal structure of large networks. Future research could focus on using this new methodology as a preliminary feature selection of another predictive model. Feature selection is a process used in machine learning which consists of keeping only a subset of relevant features to avoid overfitting or reduce dimensionality.

The third chapter classifies the sentiment of a large sample of social media messages to create stock-aggregate daily sentiment polarity measure. We show that conditionally on a stock market event happening, investors are on average able to anticipate the type of the event. Future research could focus on understanding the causes of this result at the user-level. In particular, some users have more influence than others because they either have more followers or they are more active on the platform. Are these users performing on average better than less active users ?

### 2022-08-08 main/paper1.tex
Add files via upload

@@
 This work led to the so-called Altman Z-score formula which uses accounting data to compute the default probability of a firm in the next two years.
-However, when used for financial firms, Altman's Z-score formula needs to be used with care  because, as I will discuss in this chapter, financial firms have to be treated carefully due to their frequent use of off-balance sheet financing.
+However, when used for financial firms, Altman's Z-score formula needs to be used with care  because, as I discuss in this chapter, financial firms have to be treated carefully due to their frequent use of off-balance sheet financing.
 Twenty years later, a second generation of reduced-form models used econometrical tools such as maximum likelihood, probit, and logit regressions.
 The major drawback of these models is that they do not provide multi-period forecasts.
-One innovative recent development is the use of doubly stochastic Poisson intensity model combined with multiple logit regression to account for multi-period default probability estimation.
+One innovative recent development is the use of a doubly stochastic Poisson intensity model combined with multiple logit regressions to account for multi-period default probability estimation.
 This model is proposed by \citet{DSW} in \textit{Multi-period corporate default prediction with stochastic covariates}.
-The main contribution over prior work is to exploit the time-series dynamics of the explanatory covariates in order to estimate the likelihood of default over several future periods.
+Their main contribution over prior work is to take advantage of the explanatory covariates' dynamics in order to estimate the multiperiod likelihood of default.
 Their model employs firm-specific and macroeconomic data to create a Markov state vector $X_t$ in order to compute independent firm default intensities $\lambda(t)$ and other types of exit intensities $\phi(t)$.
 The model proposed in \citet{DSW} is the first one capable of multi-period default probability estimation using time dynamics of covariates $X_t$.
-Applications of \citet{DSW} works are various.
+The applications of \citet{DSW} are various.
 We can find them in credit rating by credit rating agencies, banks who want to calculate the minimal amount of capital to be held and other researches analyzing the link between macroeconomic cycles and firm's default probabilities.
-Covariates used by \citet{DSW} are firm's Distance-to-Default, firm's trailing one-year stock return, three-month Treasury bill rate and trailing one-year return on S\&P500.
+Covariates used by \citet{DSW} are firm's trailing one-year stock return, Distance-to-Default, trailing one-year return on S\&P500 and three-month Treasury bill rate.
 Estimating their model on US-listed industrial firms between 1980 and 2004, they find that Distance-to-Default and the current state of the economy have a significant impact on default hazard rates.
@@
 \clearpage
-The two papers the closest to this chapter are the work from \citet{DSW} and \citet{Duan2012}.
+The two papers closest to this chapter are from \citet{DSW} and \citet{Duan2012}.
 The first model uses the doubly stochastic argument to derive multi-period default probabilities.
 To do so, it requires strong assumptions (e.g.
-vector autoregressive process) regarding the behavior of the time-series of covariates to generate future random values for the covariates.
+vector autoregressive process) regarding the behavior of the time series of covariates to generate future random values for the covariates.
 If the process is misspecified, biases are introduced both in the forecasted covariates and in the future default probabilities.
 Five years later, \citet{Duan2012} show that we can relax the VAR assumption with the use of forward intensities.
-This chapter explains how we can reduce biases by projecting current event realizations on past data.
-For convenience, they specify the intensities as a linear function of state variables.
+Their paper explains how we can reduce biases by projecting current event realizations on past data.
+For convenience, the authors specify intensities as a linear function of state variables.
 I wish to extend the latter by removing the assumption of linear intensities and use an artificial neural network to estimate the intensities of the Poisson processes governing both default and other exits.
 Machine learning techniques for default probabilities estimation are increasingly drawing attention.
-\citet{barboza2017machine} test several machine learning models (support vector machines, bagging, boosting, and random forest) to predict bankruptcy one year prior to the event.
-They report a substantial improvement in prediction accuracy using these machine learning techniques especially when, in addition to the original Altman’s Z-score variables, six complementary financial indicators are included.
-However, pure data-driven models often lacks economic relevance and this is where the forward intensity model can contribute to the literature.
-It is able to provide multi-period predictions while being supported by an economic and econometric background.
+\citet{barboza2017machine} test several machine learning models to predict bankruptcy one year prior to the event.
+They document a substantial improvement in prediction accuracy using the Z-score as well as six complementary financial variables.
+However, pure data-driven models often lack economic relevance and this is where the forward intensity model can contribute to the literature.
+The forward intensity model is able to provide multi-period predictions while being supported by an economic and econometric background.
 
-In \citet{DSW}, one of the main assumption is that the covariates governing both default and other exits intensities follow a high-dimensional vector autoregressive (VAR) process.
+In \citet{DSW}, one of the main assumptions is that the covariates governing both default and other exits intensities follow a high-dimensional vector autoregressive (VAR) process.
 Using this type of process forces the model to greatly reduce either the number of firms in the sample or the number of state variables explaining firm attributes; if one does not restrict the number of firms or variables in the estimation, the dimension of the model will simply be too high and it will considerably increase computational time.
-A major step forward made in \citet{Duan2012} was to get rid off the VAR process in order to reduce computational time by using a new reduced-form approach based on a forward intensity model'.
+A major step forward made in \citet{Duan2012} is to get rid of the VAR process in order to reduce computational time by using a new reduced-form approach based on a forward intensity model.
 These forward intensities produce a term structure of bankruptcy probabilities without using any sort of high-dimensional process.
@@
 Regarding covariates used, both \citet{Duan2012} and \citet{DSW} estimate their own Distance-to-Default (hereafter DtD).
-An important aspect to highlight is that Distances-to-Default specified in \citet{DSW} differ from those estimated in \citet{Duan2012} since the former estimates DtD using the variance restriction method (see \citep{Duan2012DTD}) while the latter uses the transformed-data maximum likelihood estimation method to account for financial firms.
+An important aspect to highlight is that Distances-to-Default specified in \citet{DSW} differ from those estimated in \citet{Duan2012} since the former estimate DtD using the variance restriction method (see \citep{Duan2012DTD}) while the latter use the transformed-data maximum likelihood estimation method to account for financial firms.
 The variance restriction method is a popular way to implement the \citet{Merton1974} model but fails at estimating properly the default point for financial firms.
@@
 Hence, to include financial firms in the sample, the default point has to be adjusted to the sum of short-term debt, one half of long-term debt and a fraction $\delta$ of other liabilities.
-The method used by \citet{Duan2012} is using a maximum likelihood estimation in order to estimate the unknown fraction $\delta$.
+\citet{Duan2012} use a maximum likelihood estimation in order to compute the unknown fraction $\delta$.
 The Appendix provides additional methodological information on the DtD estimation using the variance restriction method and the maximum likelihood estimation.
@@
 
-The model adds to the literature of reduced-form models of default for multiperiod corporate prediction using the doubly stochastic formulation as in \citet{DSW} (DSW henceforth) and \citet{Duan2012} (Duan henceforth).
+The model adds to the literature on reduced-form models of default for multiperiod corporate prediction using the doubly stochastic formulation as in \citet{DSW} and \citet{Duan2012} (Duan henceforth).
 The default's time is modeled as the stopping time
@@
 
-\noindent Naturally, we have the following :
+\noindent We have the following :
 \begin{equation*}
@@
 
-Let us denote by $Z_{it}$ the set of firm-specific variables at time t for the firm i and $Y_t$ the set of macroeconomic variables at time t.
+Let us denote by $Z_{it}$ the set of firm-specific variables at time $t$ for the firm $i$ and $Y_t$ the set of macroeconomic variables at time $t$.
 Let  $t_i^0$ be the first time of entry of firm $i$ in the dataset.
-The econometrician's information set $\mathscr{F}_t$ at time t is thus
+The econometrician's information set $\mathscr{F}_t$ at time $t$ is thus
 \begin{equation}
@@
 \mycaption{Example of the lifespan of a firm}{Each dot corresponds to a time period where the econometrician gathers firm-specific (if available) and macroeconomic variables.
-$t^0_i$ is the entry time of the firm i and $\tau_{D_i}$ denotes the default time of the firm i.}
+$t^0_i$ is the entry time of the firm $i$ and $\tau_{D_i}$ denotes the default time of firm $i$.}
 \label{fig:lifespan}
@@
 At each period, the econometrician gathers firm-specific variables (DtD, Cash/Total Assets, Net Income/Total Assets, ...) and macroeconomic variables (S\&P500 return, treasury rate).
-For a particular point in time $t$, the econometrician knows the whole time-serie of macroeconomic variables until $t$ irrespective of $t_i^0$, and the time-series of firm-specific variables from $t_i^0$ to $\tau_{D_i}$ if $t_i^0 \leq t$ and $t \leq \tau_{D_i}$.
-Following Proposition 1 in \citet{DSW}, the conditional probability of default within $s$ years can be computed as 
+For a particular point in time $t$, the econometrician knows the time series of macroeconomic variables until $t$ irrespective of $t_i^0$, and the time series of firm-specific variables from $t_i^0$ to $\tau_{D_i}$ if $t_i^0 \leq t$ and $t \leq \tau_{D_i}$.
+Following \citet{DSW}, the conditional probability of default within $s$ years can be computed as 
 \begin{equation} \label{eq:3.6}
@@
 However, these intensities are unknown and unobservable.
-In DSW model, the state variables governing Poisson intensities are assumed to follow a specific vector autoregressive (VAR) process.
+In \citet{DSW}, the state variables governing Poisson intensities are assumed to follow a specific vector autoregressive (VAR) process.
 This assumption is relaxed in Duan's paper by using forward intensity rates.
-Instead of modeling $\lambda_{it}$ and $\phi_{it}$ as some function of state variables available at time t, \citet{Duan2012} propose to deal with $f_{it}(\tau)$ and $g_{it}(\tau)$ directly as functions of state variables available at time t and the forward starting time of interest $\tau$.
+Instead of modeling $\lambda_{it}$ and $\phi_{it}$ as some functions of state variables available at time $t$, \citet{Duan2012} propose to deal with $f_{it}(\tau)$ and $g_{it}(\tau)$ directly as functions of state variables available at time $t$ and the forward starting time of interest $\tau$.
 The analogy to interest rates would be that $\lambda_t$ is the short rate and $f_t(u)$ is the forward rate for horizon $u$.
@@
 I propose to use an artificial neural network (ANN) to find the set of weights governing the process $f_{it}$ and $g_{it}$.
-I show that I am able to capture potential non-linear relationships between the state variables and the forward intensities which significantly improves forecasts.
+I show that I am able to capture potential non-linear relationships between the state variables and the forward intensities, which significantly improve forecasts.
 
@@
 However, the default probability given in equation \ref{eq:3.6} (which depends on spot intensities) needs to be translated into a formula that depends on forward intensities.
-To do so, we first compute the survival probability as a function of forward combined intensity, which will be later used to compute the probability of default as a function of both combined and default forward intensities.
-Let us denote by $F_{it}(\tau)$ the conditional distribution function of the combined (default and other exits) exit time evaluated at $t+\tau$.
+To do so, we first compute the survival probability as a function of forward combined intensity, which will later be used to compute the probability of default as a function of both combined and default forward intensities.
+Let us denote $F_{it}(\tau)$ the conditional distribution function of the combined (default and other exits) exit time evaluated at $t+\tau$.
 Hence, 1-$F_{it}(\tau)$ is the probability of surviving in the interval $[t, t+\tau ]$.
@@
 In particular, in our doubly-stochastic framework, hazard rates and intensities are equivalent.
-Hence, in \citet{Duan2012} and this chapter, $\lambda(t) = h(t)$ and the distinction between hazard rates and intensity is not made.
+Hence, in \citet{Duan2012} and this chapter, $\lambda(t) = h(t)$ and thus the distinction between hazard rates and intensity is not made.
 \\
@@
 
-\noindent Next, we can compute\footnote{Proofs of the following formulations can be found in the Appendix \ref{appendix1}.} the quantity $\psi_{it}(\tau) \tau$ that we were looking for as :
+\noindent Next, we can compute\footnote{Proofs of the following formulations can be found in Appendix \ref{appendix1}.} the quantity $\psi_{it}(\tau) \tau$ that we were looking for as :
 \begin{equation}
@@
 
-Hence, the probability of defaulting between t and t+$\tau$ is given by :
+Hence, the probability of defaulting between $t$ and $t+\tau$ is given by :
 
@@
 In this setup, the likelihood function depends on three types of probabilities (default, other exit and surviving) which themselves depend on two types of intensities (default and other exits).
-The negative log-likelihood function has to be adjusted to the neural network framework and can be used as objective function to be minimized as long as we feed mini-batches and not single data points (``online learning'').
-Mini-batch feeding is a common practice in machine learning and consists of splitting the available data in subsets of fixed size.
-Next, each backward pass takes one batch to perform gradient descent to update the parameters of the model.
+The negative log-likelihood function has to be adjusted to the neural network framework and can be used as an objective function to be minimized as long as we feed mini-batches and not single data points (``online learning'').
+Mini-batch feeding is a common practice in machine learning and consists in splitting the available data in subsets of fixed size.
+Next, each backward pass takes one batch to perform a gradient descent to update the parameters of the model.
 \\
 
-To allow further comparison with \citet{Duan2012},  I employ the same discretization of time: $t = 0, 1, 2,...$ and $\tau = 0, 1, 2,...$ are time sequences of one month increment.
-Similarly, $f_{it}(\tau)$ and $g_{it}(\tau)$ are forward intensities computed at time t for the period [t+$\tau$, t+$\tau$+1].
+To allow further comparison with \citet{Duan2012},  I employ the same discretization of time: $t = 0, 1, 2,...$ and $\tau = 0, 1, 2,...$ are time sequences of one month increments.
+Similarly, $f_{it}(\tau)$ and $g_{it}(\tau)$ are forward intensities computed at time $t$ for the period [t+$\tau$, t+$\tau$+1].
 The use of the $\tau$ index is to account for multiperiod prediction.
 When $\tau = 0$, the forward intensity model computes spot intensities.
-When we set $\tau = 1$, the forward intensity model produces estimates one step ahead, and so on so forth.
+When we set $\tau = 1$, the forward intensity model produces estimates one step ahead, and so forth.
 I denote $X_{it} = (x_{it,1}, x_{it,2} ,...)$ the set of firm-specific and macroeconomic variables explaining both default and combined exit intensities.
-As specified in \citet{Duan2012}, $f_{it}(\tau)$ and $g_{it}(\tau)$ are functions of $X_{it}$ and can be specified as any form of function as long as they satisfy the following constraints.
-Since combined exit intensity has to be greater or equal than default intensity, we need to make sure that the forms specified for $f_{it}(\tau)$ and $g_{it}(\tau)$ satisfy the following conditions : $f_{it}(\tau) \leq g_{it}(\tau)$, $f_{it}(\tau) > 0$, $g_{it}(\tau) > 0$.
+As specified in \citet{Duan2012}, $f_{it}(\tau)$ and $g_{it}(\tau)$ are functions of $X_{it}$ and can be specified as any form of function as long as they satisfy the constraints that follow.
+Since combined exit intensity has to be greater than or equal to default intensity, we need to make sure that the forms specified for $f_{it}(\tau)$ and $g_{it}(\tau)$ satisfy the following conditions : $f_{it}(\tau) \leq g_{it}(\tau)$, $f_{it}(\tau) > 0$, $g_{it}(\tau) > 0$.
 
-I have designed two neural networks, one will be trained to compute $f_{it}$ and one will be trained to output $h_{it}$ where $g_{it} = f_{it} + h_{it}$.
+I design two neural networks.
+One is trained to compute $f_{it}$ and the other is trained to output $h_{it}$ where $g_{it} = f_{it} + h_{it}$.
 I impose non-negativity on outputs of both models such that the combined exit intensity will never be smaller than the default intensity for all horizons.
-Let us denote by $\lambda$ and $\mu$ the set of parameters (weights) tuned in the neural network for $f_{it}$ and $h_{it}$ respectively.
-$N^{(\lambda)}$ and $N^{(\mu)}$ stand for the output of the neural network for $f_{it}$ and $h_{it}$ respectively.
-The log-likelihood for prediction horizon $\tau$ is expressed\footnote{Proof of the above formulation can be found in the Appendix \ref{appendix1}.} as
+Let us denote $\lambda$ and $\mu$ the set of parameters (weights) tuned in the neural network for $f_{it}$ and $h_{it}$ respectively.
+$N^{(\lambda)}$ and $N^{(\mu)}$ represent the output of the neural network for $f_{it}$ and $h_{it}$ respectively.
+The log-likelihood for prediction horizon $\tau$ is expressed\footnote{Proof of the above formulation can be found in Appendix \ref{appendix1}.} as
 \begin{equation} \label{decomp1}
@@
 
-The likelihoods above are a sum of indicator functions multiplied by their respective probability.
-The indicator function $\textbf{1}_{A<B}$ is equal to one if $A<B$ or zero if $A \geq B$.
-Hence, the likelihoods above are specifying three mutually exclusive indicator functions which define three independent cases during time interval $[t,t+\tau+1]$:
+The likelihoods \eqref{small_lik_f} and \eqref{small_lik_h} are the sum of the products between event indicator functions and their respective occurence probability.
+I define the indicator function $\textbf{1}_{A<B}$ to be one if $A<B$ or zero otherwise.
+These likelihoods specify three mutually exclusive indicator functions, defining three cases over the time interval $[t,t+\tau+1]$:
 \begin{enumerate}
-\item The firm does not exit the sample between t and $t+\tau$ and is considered as surviving.
+\item The firm does not exit the sample between t and $t+\tau$ and is classified as surviving.
 This case is specified as (1) in the likelihood because the combined exit time $\tau_{Ci}$ is not in the interval $[t,t+\tau+1]$.
-\item The firm exits the sample due to default during the interval and is considered as defaulted.
-This case is specified as (2) because $\tau_{Ci}=\tau_{Di}$ when the firm exits due to default conjointly with $\tau_{Di}$ being in the interval $[t,t+\tau+1]$.
-\item The firm exits due to other reasons specified as (3) during the interval since the stopping time $\tau_{Di} \neq \tau_{Ci}$ conjointly with $\tau_{Ci} \leq t+\tau+1$.
+\item The firm defaults and exits the sample during the interval.
+This case is specified as (2) because $\tau_{Ci}=\tau_{Di}$ when the firm exits due to default jointly with $\tau_{Di}$ being in the interval $[t,t+\tau+1]$.
+\item The firm exits the sample for other reasons. This case is specified in (3) since the stopping time $\tau_{Di} \neq \tau_{Ci}$ jointly with $\tau_{Ci} \leq t+\tau+1$.
 \end{enumerate}
@@
 In \citet{Duan2012} the mapping is made with a linear assumption, whereas in the framework of this chapter, the mapping depends on the whole architecture of the neural network.
-When the neural network has only one hidden layer of one neuron coupled with an exponential activation function, the model boils down to \citet{Duan2012} as the intensities would be a linear combination of the covariates.
+When the neural network has only one hidden layer of one neuron coupled with an exponential activation function, the model boils down to \citet{Duan2012} since the intensities would be a linear combination of the covariates.
 As the width and depth of the network increases, we depart more and more from the linear assumption and we allow more non-linearities to be incorporated in these intensities.
@@
 
-Neural networks can be seen as a very general function to map a given input (in this case a set of macroeconomic and firm-specific variables) into a desired output (forward intensities).
-It learns how to compute the output by tuning weights in order to minimize a given loss function.
+Neural networks can be seen as a very general function to map a given input (in this case firm-specific and macroeconomic variables) into a desired output (forward intensities). 
+They learn how to compute the output by tuning weights in order to minimize a given loss function.
 A neural network is constructed by juxtaposing several hidden layers of neurons.
@@
 Then, the training process starts and consists of many iterations of a forward pass and a backward pass.
-The forward pass takes as input a batch of data and computes the loss value, the backward pass then computes the gradient and adjust the weights of the network based on a learning rate hyperparameter.
+The forward pass takes as input a batch of data and computes the loss value; the backward pass then computes the gradient and adjusts the weights of the network based on a learning rate hyperparameter.
 As an illustration, Figure \ref{fig:net} shows a neural network with 2 hidden layers : 5 neurons in the first layer and 3 neurons in the second layer.
@@
 The use of neural networks can be motivated twofold.
-First, neural networks are well suited to approximate a function (in our case forward intensities) with the advantage of having different degrees of modularity.
+First, neural networks are well suited to approximating a function (in our case forward intensities) with the advantage of having different degrees of modularity.
 By definition, the architecture of the network generates the form of the function approximated.
@@
 For instance, suppose that every observation comes with 12 input features (i.e.
-: $x$ is a vector of shape 1x12), a [5, 3] (i.e.
+$x$ is a vector of shape 1x12), a neural network [5, 3] (i.e.
 2 layers neural network with 5 neurons in the first hidden layer and 3 neurons in the second hidden layer) can be viewed as a function computing the output $N_{it}^{(\lambda)}$ in the following way :
@@
 \\
-Neural networks are also well-suited for this chapter because they allow an easy replication of the benchmark model \citet{Duan2012}.
+Second, neural networks are well-suited for this chapter because they allow an easy replication of the benchmark model \citet{Duan2012}.
 More specifically, if the activation function $\phi(x)$ is chosen as being an exponential $\exp(x)$, and the network architecture is [1] (i.e.
@@
 
-\noindent Results of this architecture are described in section 4.
+\noindent The results from this architecture are described in Section 4.
 
@@
 
-The accounting data is taken from Wharton Research Data Services (WRDS) using the CRSP/Compustat merged database.
+The accounting data is taken from the Wharton Research Data Services (WRDS) using the CRSP/Compustat merged database.
 The macroeconomic data is taken from CRSP, the Federal Reserve Bank Reports and Datastream.
 The bankrupcty data is taken from the Compustat database, using the DLRSN item for the reason of deletion and the DLDTE item for the date of deletion.
-DLRSN contains the code that indicates the reason a company become inactive on the database.
+DLRSN contains the code that indicates the reason why a company becomes inactive on the database.
 I consider firms with a DLRSN code 2 (bankruptcy) or 3 (liquidation) to be defaulted, any other DLRSN code as other exits and no DLRSN code as surviving.
@@
 I focus on the period from 1991 to 2018 to match the accounting data with the bankruptcy data.
-Using WRDS database, I have downloaded accounting information for every company that has been listed someday on either NYSE, AMEX or NASDAQ between 1991 and 2018.
+Using the WRDS database, I download accounting information for every company that has been listed someday on either NYSE, AMEX or NASDAQ between 1991 and 2018.
 The dataset spans 27 years of data where firms entered and/or exited anywhere in this sample.
 Using this kind of sample brings a problem of cylindric data : firm's entering/exiting time are not the same for each company.
-The dataset is represented as a three dimensions matrix with the x-axis being features (i.e.
-variables), the y-axis being time, and the z-axis being firm.
-I fill the matrix with missing values for elements where the firm $i$ doesn't exist or already left at time $t$.
-Since neural networks need a lot of data points to be well trained, I chose not to remove firms even if it has a short lifespan.
+The dataset is represented as a three dimensional matrix with the x-axis being features (i.e.
+variables), the y-axis being time, and the z-axis being firms.
+I fill the matrix with missing values for elements where firm $i$ does not exist or already left at time $t$.
+Since neural networks need a lot of data points to be well trained, I choose not to remove firms even if they have a short lifespan.
 When a firm has a variable completely missing, I drop the whole firm because the likelihood is not specified if a variable is fully missing.
@@
 I winsorize all variables at the 2.5 and 97.5 percentile.
-Before dropping firms, the dataset has in total 12527 surviving firms, 1453 defaulted firms and 17950 firms that exited the sample for other reasons.
+%Before dropping firms, the dataset has in total 12527 surviving firms, 1453 defaulted firms and 17950 firms that exited the sample for other reasons. 
 Finally, I standardize all variables by substracting the mean of the variable and dividing the result by its standard deviation.
 The variables in the test set are also standardized using their respective mean and standard deviation from the training set.
-Table \ref{tab:observ} shows the number of firms in the three categories for each horizon of prediction $\tau$ which corresponds to the firm-month observations used in the likelihood for horizon $\tau$ (see equations \ref{small_lik_f} and \ref{small_lik_h}).
+Table \ref{tab:observ} shows the number of firms in the three categories for each horizon of prediction $\tau$, which corresponds to the firm-month observations used in the likelihood for horizon $\tau$ (see equations \ref{small_lik_f} and \ref{small_lik_h}).
 Figure \ref{fig:defsurv} shows the total number of firms that defaulted, survived or exited for other reasons plotted on a year on year basis.
@@
     \includegraphics[width=0.75\textwidth]{def_surv_oe_plot.jpg}
-    \mycaption{Defaults, exits for other reasons and survivals each year}{Top plot shows the number of defaulted firms, middle plot shows the number of exits for other reasons and bottom plot shows the number of firms surviving each year.}
+    \mycaption{Defaults, exits for other reasons and survivals each year}{The top plot shows the number of defaulted firms, the middle plot shows the number of exits for other reasons and the bottom plot shows the number of firms surviving each year.}
     \label{fig:defsurv}
@@
     \hline
-0 &  2'025'094 & 499 & 10'392\\
+0 &  2'025'094 & 514 & 10'746\\
 %1 &  2'007'311 & 497 & 10'384\\
 %2 &  1'989'548 & 483 & 10'173\\
-3 &  1'972'063 & 514 & 10'746\\
+3 &  1'972'063 & 499 & 10'713\\
 %4 &  1'953'935 & 511 & 10'743\\
 %5 &  1'935'839 & 481 & 10'503\\
-6 &  1'918'061 & 499 & 10'713\\
+6 &  1'918'061 & 499 & 10'392\\
 %7 &  1'900'041 & 495 & 10'712\\
@@
 The exhaustive list of variables is the following :
+
+\clearpage
+
 \begin{enumerate}
@@
 \item Treasury : 3-month annualized US Treasury bill rate. I use this variable as the risk-free rate $r$ in the model. 
-\item DtD : the Distance-to-Default is a volatility adjusted leverage measure for gauging how far is the firm from default.
-It was first introduced by \citet{Merton1974} model which treat firm's equity value as a call option on the underlying asset and the strike being the debt level.
-The DtD can be viewed as the number of standard deviations of annual asset growth by which the asset level exceeds the firm's liabilities.
+\item CASH/TA : cash and short-term investment divided by total assets. Both quantities are taken from the balance sheet of the company.
+\item NI/TA : net income divided by total assets. 
+\item SIZE : log of the market equity value divided by the cross-sectional average market equity value. The number of shares outsanding times the stock price gives the market equity value.
+\item DtD : the Distance-to-Default is a measure introduced in \citet{Merton1974} to gauge how close a firm is from default.
 In this chapter, the DtD is estimated using a maximum likelihood taking into account other liabilities of each firm to handle financial firm's bias.
-It is well-known in the literature that DtD is a significant measure to estimate default probabilities but has to be used conjointly with other variables.
+It is well-known in the literature that DtD is a significant measure to estimate default probabilities but has to be used jointly with other variables.
 See the Appendix \ref{appendix1} for additional information regarding the estimation procedure of DtD.
-\item CASH/TA : ratio between cash and short-term investment to total assets. Both quantities are taken from the balance sheet of the company.
-\item NI/TA : ratio between net income to total assets. A loss is registered as a negative net income.
-\item SIZE : logarithm of the ratio between market equity value to the cross-sectional average market equity value of the stocks in the S\&P500.
-The market equity value is computed as the stock price multiplied by the number of outstanding shares at the end of each quarter.
-SIZE is negative if the firm has a smaller market capitalization than the average market capitalization and positive otherwise.
-\item M/B : market-to-book ratio computed using the ratio between the asset market value from the DtD maximum likelihood estimation to the total book asset value.
+\item M/B : asset market value divded by the total book asset value.
 \end{enumerate}
 
-To capture momentum of variables, I also compute one step rolling window differences for each firm-specific variables.
-These variables are called ``$\Delta$" followed by the name of variable.
-They are telling whether the firm has been improving or deteriorating with respect to this particular variable comparing to the last period performance.
-Given this model specification, the $\Delta$ is particularly interesting because if a firm shows many consecutive negative delta values, it really means the company is in danger.
+To capture momentum of variables, I also compute one step rolling window differences for each firm-specific variable.
+These variables are called ``$\Delta$" followed by the name of the variable.
+They represent whether the firm has been improving or deteriorating with respect to this particular variable comparing to the last period performance.
+Given this model specification, the $\Delta$ is particularly interesting because if a firm shows many consecutive negative delta values, it means the company is in danger.
 However, it is also important to look at the level value to compare a defaulted firm with a non-defaulted firm.
@@
 This section depicts the summary statistics of the dataset.
-Table \ref{tab:meanvar} shows a summary of the mean of each variable for the three categories of firm.
-Please note that we have to be very careful when comparing two means in this table.
-Comparing means needs to be done with confidence intervals and significance tests which involves standard deviations.
+Table \ref{tab:meanvar} shows a summary of the mean of each variable for the three categories of firms.
+Please note that we must be very careful when comparing two means in this table.
+Comparing means needs to be done with confidence intervals and significance tests, which involves standard deviations.
 This table is presented to give a rough idea without performing any statistical inference.
-The table shows that the average defaulted firm is smaller, has a lower DtD, a smaller Market-to-book ratio, loses more money and has less cash.
+The table shows that the average defaulted firm is smaller, has a lower DtD, a smaller Market-to-book ratio, loses more money and has less cash than the average surviving firm.
 The prefixes $\Delta$ in front of the variables stand for the one-lag differences.
@@
     \end{tabular}
-    \mycaption{Mean of variables for surviving firms, defaulted firms and other exits}{The prefix $\Delta$ stands for one period lagged differences.}
+    \mycaption{Mean of variables for surviving firms, defaulted firms and other exits}{The prefix $\Delta$ stands for a one period lagged difference.}
     \label{tab:meanvar}
@@
 
-As every neural network, we need to chose the optimal hyperparameters to achieve the highest performance.
+As for every neural network, we need to chose the optimal hyperparameters to achieve the highest performance.
 In my setup, this consists mainly of choosing the architecture of the model (i.e.
@@
 To avoid any overfitting, I perform a 5-fold cross validation for each horizon of prediction.
-I cut 15\% of the dataset as test set, all results that I will talk about in this chapter are out-of-sample and performed on the observations of the test set that the model has never seen before.
+I cut 15\% of the dataset as test set, all results that I will talk about in this chapter are out-of-sample and performed on the observations from the test set that the model has never seen before.
 The remaining set of observations is partitioned into smaller subsets so that in every fold of the validation a different subset is used as validation set and the rest is used as training set.
-Finally, to measure the discriminatory power of the models, I use the Lorenz curve (\citet{lorenz}) and I use the Gini coefficient as a scalar performance measure to aggregate across folds to get the  measure that I will use to discriminate models.
+Finally, to measure the discriminatory power of the models, I use the Lorenz curve (\citet{lorenz}) and I use the Gini coefficient as a scalar performance measure to aggregate across folds to get the  measure that I use to discriminate the models (see Definition \ref{lorenzdef} (\citet{leippold})).
 \\
@@
 
-\begin{definition}[Lorenz curve]
+\begin{definition}[Lorenz curve]\label{lorenzdef}
 The Lorenz curve of a predictor P is the two-dimensional graph
@@
 power curves, cumulative accuracy profiles).} are often used in the literature for default prediction (for instance \citet{leippold}, \citet{Duan2012}).
-They are different from ROC curves and precision-recall curves because they don't rely on thresholds to discriminate between true positives and false positives.
+They are different from ROC curves and precision-recall curves because they do not rely on thresholds to discriminate between true positives and false positives.
 Hence, they are particularly well-suited as performance measure in this model because of the multiperiod framework involved.
-The idea is that if the model is outputting a false positive for an horizon $\tau$ but the true positive is horizon $\tau+1$, the model should not be too much penalized.
+The idea is that if the model is outputting a false positive for an horizon $\tau$ but the true positive is horizon $\tau+1$, the model should not be too hardly penalized.
 The ROC and precision-recall curves would treat this as a false positive even though the prediction was not that far from the target.
@@
 
-Finally, the Gini coefficient is used as the scalar summary statistic to compare models.
+Finally, the Gini coefficient is used as the scalar summary statistic to compare the models.
 It measures the degree of inequality of the Lorenz curve.
@@
 
-The choice of the optimal architecture of the neural network has been made using k-fold cross-validation.
-The average Gini coefficient across all folds for every horizon is used as comparison tool to determine the best model architecture.
+The choice of the optimal architecture of the neural network is made using k-fold cross-validation.
+The average Gini coefficient across all folds for every horizon is used as a comparison tool to determine the best model architecture.
 If the architecture is too deep, the implicit function computed in the network to output forward intensities incorporates too many parameters and the risk of overfitting is larger, resulting in a lower accuracy measure.
@@
 I compare the scores obtained with different network architectures with those using the linear assumption.
-In this setup, one can interpret the architecture model as the non-linearity degree because higher architecture involves more weights in the output function to be estimated.
+In this setup, one can interpret the architecture model as the non-linearity degree because a higher architecture involves more weights in the output function to be estimated.
 By disantangling the spaghettis, we can clearly see the bias-variance tradeoff.
@@
     \includegraphics[width=0.75\textwidth]{duan_vs_NN.jpeg}
-    \mycaption{Gini coefficient : Linear vs Neural Networks}{Yellow line shows the Gini coefficient of the linear model for every horizon.
+    \mycaption{Gini coefficient : Linear vs Neural Networks}{The yellow line shows the Gini coefficient of the linear model for every horizon.
 The remaining lines show the Gini coefficient for different neural network architectures.
-Green line is the best performing model as it exhibits the highest Gini for every horizon.}
+The green line is the best performing model as it exhibits the highest Gini for every horizon.}
     \label{fig:ginicomp}
@@
 
-Figure \ref{fig:perf} shows Lorenz curves for horizon 0, 3, 6, 12, 24 and 35 but all results generalize well to all others horizons.
+Figure \ref{fig:perf} shows Lorenz curves for horizons 0, 3, 6, 12, 24 and 35 but all results generalize well to all other horizons.
 The curves are completely out-of-sample since they are computed on the test set that the model has never seen before.
@@
 Overall, as expected the neural network clearly outperforms the linear assumption, suggesting that the linear assumption from \citet{Duan2012} can be greatly improved by adding non-linearities in the specification of the intensities.
-Unfortunately, it is difficult to tell which kind of non-linearities should be taken into account since neural networks are often seen as a black box.
+Unfortunately, it is difficult to tell which kinds of non-linearities should be taken into account since neural networks are often seen as a black box.
 However, I will still try to answer this question by looking how the average intensity outputted by the model changes when we change a feature ceteris paribus (see Section \ref{sec4-4} dedicated to sensitivities).
@@
 
-For comparison purposes, Figure \ref{fig:lor0duan} exhibits the Lorenz curves of the linear assumption, its replication in the neural network framework, and the [5, 3] model.
-The linear assumption and its replication have similar performance, showing that the neural network ``NN+exp+[1]" is indeed able to replicate the linear framework depicted in \citet{Duan2012}.
+For comparison purposes, Figure \ref{fig:lor0duan} exhibits the Lorenz curves of the linear assumption, the replication of the linear assumption in the neural network framework, and the [5, 3] model.
+The linear assumption and its replication have a similar performance, showing that the neural network ``NN+exp+[1]" is indeed able to replicate the linear framework depicted in \citet{Duan2012}.
 
@@
         \end{subfigure}
-        \mycaption{Out-of-sample Lorenz Curves for each horizon}{The orange line shows the performance of Duan's model and the red line shows the performance of the [5,3] neural network.} 
+        \mycaption{Out-of-sample Lorenz Curves for each horizon}{The orange line shows the performance of Duan's model and the red line shows the performance of the [5, 3] neural network.} 
         \label{fig:perf}
@@
     \end{tabular}
-    \mycaption{Gini coefficients}{Comparison of the out-of-sample Gini coefficients associated to the [5,3 ] neural network and the linear assumption.}
+    \mycaption{Gini coefficients}{Comparison of the out-of-sample Gini coefficients associated to the [5, 3] neural network and the linear assumption.}
     \label{tab:ginicoeff}
@@
 a single hidden layer with a single neuron) with an exponential activation function.
-The orange line shows the performance of Duan's model and the red line shows the performance of the [5,3] neural network.}
+The orange line shows the performance of Duan's model and the red line shows the performance of the [5, 3] neural network.}
     \label{fig:lor0duan}
@@
 Neural networks are often seen as black boxes because their outputs are coming from a general function involving many parameters.
-It is incorporating non-linearities via the layers and the activation functions.
+They are incorporating non-linearities via the layers and the activation functions.
 Figure \ref{fig:sensi} is an attempt at gauging how the model reacts to a change of an input variable.
@@
     \item Compute the forward default intensity for all the observations in the test set and average the result.
-Keeping everything else equal, change the value of one feature by an absolute value and feed the updated observations to the network as new input and compute the new average forward default intensity.
+Keeping everything else equal, change the value of one feature by an absolute value and feed the updated observations to the network as new inputs and compute the new average forward default intensity.
     \item Repeat step 1 for all absolute values in some interval.
     \item Plot the average intensity against the absolute change.
-    \item Repeat step 1-3 for all 11 other features.
+    \item Repeat steps 1-3 for all 11 other features.
 \end{enumerate}
@@
 Moreover, all relationships are intuitive and expected.
-Forward intensities in both Size and $\Delta$Size are decreasing, suggesting that small firms tend to have higher likelihood of defaulting which is consistent with the ``too big to fail" paradigm.
-Similarly, firms with decreasing cash ($\Delta$CASH/TA $<$ 0) or low levels of CASH/TA appear to have higher probability of default.
-The model also predicts that firms with low NI/TA or decreasing NI/TA should have higher probability of default.
+Forward intensities in both Size and $\Delta$Size are decreasing, suggesting that small firms tend to have a higher likelihood of defaulting, which is consistent with the ``too big to fail" paradigm.
+Similarly, firms with decreasing cash ($\Delta$CASH/TA $<$ 0) or low levels of CASH/TA appear to have higher probabilities of default.
+The model also predicts that firms with low NI/TA or decreasing NI/TA should have higher probabilities of default.
 Finally, forward intensities in DtD should be decreasing for all horizons to reflect that a higher Distance-to-Default makes the firm less likely to default.
@@
 Overall,
-it appears that the most important predictors of default in this model are in short term Market-to-book ratio, DtD and CASH/TA, and in the long run NI/TA, CASH/TA, $\Delta$CASH/TA and DtD.
+it appears that the most important predictors of default in this model are in the short term Market-to-book ratio, DtD and CASH/TA, and in the long run NI/TA, CASH/TA, $\Delta$CASH/TA and DtD.
 
@@
         \end{subfigure}
-        \mycaption{Sensitivities for each horizon}{These plots show the average default forward intensity against a change in the specified variable, ceteris paribus}
+        \mycaption{Sensitivities for each horizon}{These plots show the average default forward intensity against a change in the specified variable, ceteris paribus.}
         \label{fig:sensi}
@@
 
-I propose an approach to estimate default forward intensities which relies on using machine learning techniques.
+I propose an approach to estimate forward default intensities, which relies on using machine learning techniques.
 The key improvement over previous estimation methods is the introduction of possible highly non-linear relationships between covariates and forward intensities.
@@
 : one layer and one neuron) and the activation function is chosen as being an exponential, the neural network boils down to a logit regression.
-The dataset used in this chapter is the same as previous literature to allow easier comparison.
+The dataset used in this chapter is the same as in previous literature to allow for an easier comparison.
 It consists of 5 firm-specific variables computed from accounting data and 2 macroeconomic variables to control for the health of the economy.
 I also account for momentum of these variables by feeding the model the one-lagged differences of each variable.
-Looking at summary statistics only, the average defaulted firm is small, has low cash, low market-to-book ratio, low Distance-to-Default and has large and negative profits.
-To measure the discriminatory power of the models, I follow previous literature and use Lorenz curves (also known as ``cumulative accuracy profile" or ``power curves").
-The idea behind Lorenz curves is to order default probabilities and look how they are distributed across defaulted and non-defaulted firms.
-The average Gini coefficient across all folds of the cross-validation is used as comparison tool to gauge the accuracy of the model.
-Results show that the architecture [5, 3] (i.e.
-2 layers with 5 neurons in the first hidden layer and 3 neurons in the second hidden layer) seems to outperform others architecture.
-In this setup, one can interpret the architecture as the non-linearity degree as higher architecture involves more weights in the output function to be estimated.
+Looking at summary statistics only, the average defaulted firm is small, has low cash, low market-to-book ratio, low Distance-to-Default and has large negative profits.
+To measure the discriminatory power of the models, I follow the previous literature and use Lorenz curves (also known as ``cumulative accuracy profile" or ``power curves").
+The idea behind Lorenz curves is to order default probabilities and look at how they are distributed across defaulted and non-defaulted firms.
+The average Gini coefficient across all folds of the cross-validation is used as a comparison tool to gauge the accuracy of the model.
+The results show that the architecture [5, 3] (i.e.
+2 layers with 5 neurons in the first hidden layer and 3 neurons in the second hidden layer) seems to outperform other architectures.
+In this setup, one can interpret the architecture as the non-linearity degree since a higher architecture involves more weights in the output function to be estimated.
 Out-of-sample Lorenz Curves and Gini coefficients show that the neural network approach outperforms the linear assumption for every horizon, suggesting the presence of non-linearities in forward default intensities.
@@
 Further works could involve more variables in the estimation.
-In particular, a challenging (due to lack of data) but nonetheless exciting study would be to gauge the effect of market sentiment on default intensities.
+In particular, a challenging (due to lack of data) but nonetheless exciting study would be to explore the effect of market sentiment on default intensities.

### 2022-08-08 main/paper2.tex
Add files via upload

@@
 There have been many attempts during the past decades to capture and visualize the network of interconnections among a set of financial institutions. 
-The most widely used concept of causality in time series econometrics is due to \citet{granger}. 
-This is based on statistical analysis of the financial series such as their stock prices over a finite time period. 
-Granger’s definition of causality states that a time series $X$ is a cause of another time series $Y$, if the mean squared error of the 1-step ahead forecast for $Y$ is smaller when the history of $X$ is included in the forecasting information set.
+A well-known time series causality measure used in econometrics is Granger-causality (\citet{granger}).
+This is based on statistical analysis of the financial time series such as their stock prices over a finite time period. 
+Granger’s definition of causality states that a time serie $X$ is a cause of another time serie $Y$, if the one-step future forecast of $Y$ is more precise when the forecasting information set includes X.
 Otherwise, when the forecast does not improve by including the information of $X$, then it is declared that $X$ does not cause $Y$.
@@
 
-In the great majority of practical applications, Granger-causality has been studied in the context of Vector Autoregressive (VAR) models. 
-For instance, \citet{billio2012econometric} propose several measures  based on Granger-causality to capture the connections between the monthly returns of different financial institutions. 
+Most empirical applications of Granger-causality have been studied with Vector Autoregressive (VAR) models. 
+For instance, \citet{billio2012econometric} propose several measures based on Granger-causality to capture the interconnections between the returns of financial institutions on a monthly basis. 
 It uses principle component analysis and pairwise Granger-causality tests to identify the causal networks. 
@@
 \begin{itemize}
-    \item We use an information-theoretic measure known as directed information (DI) to infer the Granger-causalities among a set of time series. 
+    \item We use directed information (DI), an information measure to infer the Granger-causalities among a set of time series. 
 This measure is non-parametric (i.e.
@@
 \item Computing DI has high computational and sample complexity which makes it not suitable for inferring the causal structure of large networks. 
-To overcome this problem, we develop a novel approach based on Recurrent Neural Networks (RNNs) that reduces the complexity of evaluating DIs in high-dimensional settings. 
+To solve this problem, we develop a novel approach based on Recurrent Neural Networks (RNNs) that reduces the complexity of evaluating DIs in high-dimensional settings. 
 \end{itemize}
@@
 \citet{tsai2009} states that feature selection preprocessing is not addressed carefully enough in the bankruptcy prediction literature.
-They compare five feature selection methods used in bankruptcy prediction: t-test, correlation matrix, step-wise regression, Principle Component Analysis (PCA) and factor analysis and show that any of these methods improves performance.
+They compare five feature selection methods used in bankruptcy prediction: step-wise regressions, correlation matrix, principal component analysis, t-tests and factor analysis and show that any of these methods improves performance.
 \citet{fs_stock} uses a Sequential Forward Selection algorithm to select relevant features predicting the Turkish market index.
@@
 To mention a few, \citet{psaradakis2005markov} introduce different terminologies for causality based on Granger’s ideas and provide a set of parametric non-causality constraints in the context of Markov switching VAR models.
-In a similar context, \citet{bianchi2019modeling} investigate time-varying systemic risk based on a range of multi-factor asset pricing models and develop a Markov Chain Monte Carlo (MCMC) scheme to infer their model parameters and consequently obtain their corresponding networks.
+In a similar context, \citet{bianchi2019modeling} investigate time-varying systemic risk based on a range of multi-factor asset pricing models and develop a Markov Chain Monte Carlo scheme to infer their model parameters and consequently obtain their corresponding networks.
 Another attempt is \citet{bonaccolto2019estimation} in which the authors explore quantile-based methods of Granger-causality.
@@
 Often, the problem of network identification in high dimensional settings requires more considerations and even its own techniques. 
-For instance, \citet{billio2019bayesian} propose a Bayesian non-parametric Lasso prior (BNP-Lasso) for high-dimensional VAR models that can improve efficiency and accuracy. 
-In order to overcome overparametrization and overfitting issues in large VAR models, BNP-Lasso clusters the VAR coefficients into groups and shrinks the coefficients of each group toward a common location.
+For instance, \citet{billio2019bayesian} propose a Bayesian non-parametric Lasso (BNPL) prior for high-dimensional vector autoregressive models that improve efficiency and accuracy. 
+To overcome overfitting in large VAR models, BNPL clusters the vector autoregressive coefficients and shrinks the coefficients of each cluster toward a common location.
 %Although, the proposed method in this work is suitable for inference in high dimensional VAR models but 
@@
 The work in \citet{iacopini2019bayesian} tackles the curse of dimensionality by a two-stage approach.
-In the first stage, a spike-and-slab prior distribution is used for each entry of the coefficient matrix which also identifies the interconnection network.
+First, a spike-and-slab prior distribution is used for each entry of the coefficient matrix which also identifies the interconnection network.
 In the second stage, it imposes prior dependence on the coefficients by specifying a Markov process for their random distribution. 
@@
 
-The rest of the chapter is organized as follows. 
+The remainder of this chapter is strutured as follows. 
 Section \ref{Sec1} reviews the notion of Granger-causality and formally introduces directed information graphs which are suitable for linear and nonlinear systems. 
-In Section \ref{sec:method}, we introduce a novel approach for inferring the Granger-causal network of high dimensional nonlinear systems. 
-Finally, in Section \ref{sec:exp}, we apply our method to learn the causal network of both synthetic and real-world dataset. For the real-world experiment, we use the daily stock prices of US firms.
+Section \ref{sec:method} introduces a novel approach for inferring the Granger-causal network of high dimensional nonlinear systems. 
+Section \ref{sec:exp} applies our methodology to learn the causal network of both synthetic and real-world dataset. For the real-world experiment, we use the daily stock prices of US firms.
 
@@
 
-In this section, we present a statistical approach to learn the causal interconnections in a dynamical systems based on Granger-causality (see \citet{granger}). 
+We present in this section our econometric approach to learn the causal dependencies in a dynamical systems based on Granger-causality (see \citet{granger}). 
 We begin by introducing some notations.
@@
 
-Researchers from different fields have developed various frameworks and graphical models to capture and represent interconnections among variables or processes.
-One of the most popular and widely used frameworks in economics is the notion of Granger-causality.
-The basic idea in this framework was originally introduced by \citet{wiener1956theory}, and later formalized by \citet{granger}. 
-The idea is as follows: ``we say that $X$ is causing $Y$ if we are better able to predict the future of $Y$ using all available information than if the information apart from the past of $X$ had been used."
+Various frameworks and graphical models have been developed to capture and represent interconnections among variables or processes.
+One of the most popular and widely used frameworks in economics is the notion of Granger-causality originally introduced by \citet{wiener1956theory} and
+generalized by \citet{granger}.
+We say that $X$ Granger-causes $Y$ if the prediction of the future of $Y$ is more precise using all available information (including $X$) than using all available information except $X$.
 
-Despite broad philosophical viewpoint of \citet{granger1963economic}, his formulation for practical implementation was done using multivariate autoregressive (MVAR) models and linear regression which has been widely adopted in econometric and other disciplines.
-More precisely, in order to identify the influence of $X_t$ on $Y_t$ in a MVAR comprises of three time series $\{X,Y,Z\}$, Granger's idea is to compare the performance of two linear regressions: the first one predicts $Y_t$ given $\{X^{t-1},Y^{t-1},Z^{t-1}\}$ and the second one predicts $Y_t$ given $\{Y^{t-1},Z^{t-1}\}$.
-Clearly, the performance of the second predictor is bounded by the first predictor.
+This formulation was originally implemented using multivariate autoregressive (MVAR) models and linear regressions.
+In particular, let $\{X,Y,Z\}$ be three time series.
+In order to identify the influence of $X_t$ on $Y_t$, Granger's idea is to compare the performance of two linear regressions\footnote{Note that this formulation is only applicable in linear systems.}: the first one predicts $Y_t$ given $\{X^{t-1},Y^{t-1},Z^{t-1}\}$ and the second one predicts $Y_t$ given $\{Y^{t-1},Z^{t-1}\}$.
 If they have the same performance, then we say $X$ does not Granger-cause $Y$.
-It is important to emphasize that this formulation is only applicable in linear systems.
 
 To go beyond linear systems, works such as \citet{directed} and \citet{massey} use information-theoretical measures and generalize Granger-causality. 
-In this work, we introduce and apply directed information (see \citet{directed}), an information-theoretical tools to measure interconnections among firms.
+In this work, we introduce and apply directed information (see \citet{directed}), an information-theoretical tool to measure interconnections among firms.
 Directed information (DI) has been used in many applications to infer causal relationships.
-For example, \citet{spike} and \citet{di12} use it for analyzing neuroscience data and \citet{acc2014}  and \citet{etesami2018econometric} apply to market data.
+For example, \citet{spike} and \citet{di12} use it for analyzing neuroscience data and \citet{acc2014}  and \citet{etesami2018econometric} apply it to market data.
 
@@
 
-In the rest of this section, we describe how the DI can capture the interconnections in causal dynamical systems (linear or non-linear) and formally define DIGs.
+We describe in this section how the directed information is able to capture the connections in causal networks. Next, we formally define directed information graphs (DIG).
 
-Consider a dynamical system comprised of three time series $\{X,Y,Z\}$ that we assume they have a joint probability density function $p(X,Y,Z)$. 
-To answer whether $X$ has influence on $Y$ or not over time horizon $[1,T]$,  following the idea of Granger, we compare the average performance of two particular predictors over this time horizon. 
-The first predictor uses the history of all three time series while the second one uses the history of all processes excluding process $X$. 
-On average, the performance of the predictor with less information (the second one) is upper bounded by the performance of the predictor with more information  (the first one). 
-However,  if the prediction of both predictors are close over time horizon $[1,T]$,  it is an indication that $X$ does not cause $Y$ in this time horizon.
+We define a dynamical system constituted of three time series $\{X,Y,Z\}$ that we assume have a joint probability density function $p(X,Y,Z)$. 
+Granger-causality states that to know whether $X$ influences $Y$ or not, we need to compare the performance of two predictors of $Y$.
+The first predictor uses the history of all information available (i.e. $\{X,Y,Z\}$) while the second predictor uses only the history of $\{Y,Z\}$
+If the former performs better than the latter, $X$ has information on $Y$. However, if they perform equally, it is an indication that $X$ is not causing $Y$ in this time interval.
 To rigorously formalize this idea, we need the predictors and a measure to compare their performances.
@@
 According to the above measures of goodness, the better the predictor is, the smaller its log-loss will be.
-This loss function has meaningful information-theoretical interpretations.
+This loss function also has meaningful information-theoretical interpretations.
 Namely, the log-loss is the Shannon's code length\footnote{It is also called the description length of $y_t$.
-For more information see \citet{cover2012elements}.}, i.e., the number of bits required to efficiently represent $y_t$.
+For more information see \citet{cover2012elements}.}, i.e., the number of bits required to efficiently represent $y_t$ (see \citet{etesami2018econometric}).
 
@@
 \end{align}
-Note that the regrets are non-negative for all $t$ and all outcomes $y_t$.
-The average regret over the time horizon $[1,T]$ is given by 
+Regrets are always positive for all $t$ and all outcomes $y_t$.
+Over the time interval $[1,T]$, the average regret is
 \begin{align}\label{eq:ave}
@@
 \end{align}
-The average regret in \eqref{eq:ave} is called directed information (DI) and will be our measure of causation in this work.
+The average regret in \eqref{eq:ave} is the directed information (DI). We use it as the measure of causation in this chapter.
 This measure is always positive and if it is zero, it indicates that the history of time series $X$ contains no significant information that would help in predicting the future of time series $Y$ given the history of $Y$ and $Z$.
 This definition can be generalized to more than three time series as follows,
+
+\clearpage
 
@@
 
-It is important to emphasize that the definition of DI does not rely on any model assumption, thus DI is capable of inferring the causal relationships in general (linear or non-linear) dynamical systems.
+Note that the definition of DI does not rely on any model assumption, thus DI is capable of inferring the causal relationships in general (linear or non-linear) dynamical systems.
 Next, we define the graphical model that we use in this work to visualize the causal network among firms.
@@
 Example \ref{r1} demonstrates the DIG of a simple linear system.
+
+\clearpage
+
 \begin{example}\label{r1}
-Consider a network of three times series $\{X,Y,Z\}$ with the following linear dynamic,
+Let $\{X,Y,Z\}$ be a network of three time series with the following dynamics,
 \begin{small}
@@
 where $N_X$, $N_Y$, and $N_Z$ are three independent stationary Gaussian processes with zero mean and a diagonal covariance matrix (1, 0.9, 1). 
-Since the dynamic is linear and the exogenous noises are Gaussian, we can compute the DIs using the following expression\footnote{Equation \eqref{simueq} does not hold in more general settings.}(see \citet{acc2014}).
+Since the dynamics is linear and the exogenous noises are Gaussian, we can compute the DIs using the following expression\footnote{Equation \eqref{simueq} does not hold in more general settings.}(see \citet{acc2014}).
 \begin{equation}\label{simueq}
@@
 
+\begin{comment}
 \begin{remark}
@@
 \end{remark}
+\end{comment}
 
@@
 Therefore, parametric and non-parametric estimators for the conditional mutual information can be used to estimate the DIs. 
-There are different methods that can be used to estimate the terms in (\ref{estdi}) given i.i.d.
-samples of the time series such as plug-in empirical estimator and k-nearest neighbor estimator.
-For an overview of such estimators see Appendix \ref{appendix2} and the articles in \citet{paninski2003estimation}, \citet{noshad2019scalable}, and \citet{jiao2013universal}.
+Given i.i.d.
+samples of the time series, it exists two main methods to estimate the terms in (\ref{estdi}) :  the plug-in empirical estimator and the k-nearest neighbor estimator.
+For an overview of such estimators, we refer to the articles in \citet{paninski2003estimation}, \citet{noshad2019scalable}, \citet{sricharan2011k} and \citet{jiao2013universal}.
 
- In general, estimating the DI in (\ref{estdi}) is a complicated task and has high sample complexity.
- This is due to the fact that it requires estimating high dimensional conditional distributions. 
-However, knowing some side information about the underlying dynamic can simplify the learning task of the DIG. 
-For instance, in Example \ref{r1}, since the underlying dynamic is linear with Gaussian exogenous noises, the DIs can be computed via the covariance matrices \eqref{simueq}. 
+ In general, estimating the DI in (\ref{estdi}) has high sample complexity because it requires estimating high dimensional conditional distributions. 
+However, information about the underlying dynamics can simplify the learning task of the DIG. 
+For instance, in Example \ref{r1}, since the underlying dynamics is linear with Gaussian exogenous noises, the DIs can be computed via the covariance matrices \eqref{simueq}. 
 Clearly, the covariance matrix can be estimated with lower complexity compared to conditional mutual information. 
-For our experimental results, we used \eqref{simueq} for the linear Gaussian experiment and  the k-nearest method in \citet{sricharan2011k} for the non-linear experiment. 
-The main reason for selecting k-nearest method is because it has shown relatively better performance compared to the other non-parametric estimators. 
+For our experimental results, we used \eqref{simueq} for the linear Gaussian experiment and the k-nearest neighbor estimator in for the non-linear experiment. 
+The main reason for selecting k-nearest method is because it usually shows better performance in comparison to the other estimators. 
 %More precisely, we estimated each of the above conditional mutual information using the $k$-nearest method of \citet{sricharan2011k}. 
-For the sake of completeness, we describe the steps of this method in Appendix \ref{appendix2}.
+% For the sake of completeness, we describe the steps of this method in Appendix \ref{appendix2}.
 
 Side information can also help to infer the DIG of a dynamical system without directly estimating the DIs but instead providing an alternative approach to identify the DIG. 
-For example, if it is given that the underlying dynamic is linear, i.e., $\textbf{X}_t = \textbf{AX}_{t-1} + \textbf{N}_t$, then it has been shown in \citet{acc2014} that the support\footnote{The support of a matrix $\textbf{B}=[b_{i,j}]$ is a binary matrix of the same dimension as $\textbf{B}$ such that its entry $(i,j)$ is one if and only if $b_{i,j}\neq0$.} of the coefficient matrix $\textbf{A}$ is equal to the adjacency matrix of its corresponding DIG.
+For example, if it is given that the underlying dynamics is linear, i.e., $\textbf{X}_t = \textbf{AX}_{t-1} + \textbf{N}_t$, then it has been shown in \citet{acc2014} that the support\footnote{The support of a matrix $\textbf{B}=[b_{i,j}]$ is a binary matrix of the same dimension as $\textbf{B}$ such that its entry $(i,j)$ is one if and only if $b_{i,j}\neq0$.} of the coefficient matrix $\textbf{A}$ is equal to the adjacency matrix of its corresponding DIG.
 This result implies that in linear systems, one can obtain the DIG by estimating the coefficient matrix. The latter problem has lower complexity and it can be done using e.g., linear regression. 
@@
 
-One such approaches is proposed by \citet{quinn2013efficient}, in which they developed an efficient algorithm to identify the best directed tree approximation of a given network.
+One such approach is proposed by \citet{quinn2013efficient}, in which they developed an efficient algorithm to identify the best directed tree approximation of a given network.
 This means reducing the size of the conditioning set to zero, i.e., no conditioning.  
@@
 Next section explains this idea in more details.
+
+\clearpage
 
@@
 Inferring the causal network of non-linear systems is a challenging problem that its complexity increases exponentially with the dimension of the network. 
-In this section, we study the causal inference problem in non-linear systems whose dynamic can be captured by
+In this section, we study the causal inference problem in non-linear systems whose dynamics can be captured by
 \begin{align}\label{nonlinear}
@@
 We call this model non-linear with \textit{additive noise} due to the noise term that is added to the non-linear term\footnote{In contrary to additive noise, there are systems in which the exogenous terms are multiplicative, e.g., $X_{j,t}=X_{i,t-1}\varepsilon_{j,t}+X_{j,t-1}$.}.
-This is a general non-linear dynamic that can be used to model the behavior of wide range of physical dynamical systems.
-The dynamic is called \textit{Markovian} if $\mathcal{X}^{t-1}$ is replaced by $\mathcal{X}_{t-1}=\{X_{1,t-1},...,X_{m,t-1}\}$.
+This is a general non-linear dynamics that can be used to model the behavior of wide range of physical dynamical systems.
+The dynamics is called \textit{Markovian} if $\mathcal{X}^{t-1}$ is replaced by $\mathcal{X}_{t-1}=\{X_{1,t-1},...,X_{m,t-1}\}$.
 
@@
 \subsubsection{Koopman-based lifting technique}
-Consider a particular sub-class of the non-linear system in \eqref{nonlinear} whose dynamic is defined by
+Consider a particular sub-class of the non-linear system in \eqref{nonlinear} whose dynamics is defined by
 \begin{align}\label{nonlin}
@@
 \end{align}
-However, in this formulation, the weights $\{w_{,k}\}$ are unknown. 
+However, in this formulation, the weights $\{w_{j,k}\}$ are unknown. 
 An approach to obtain the weights is a non-linear filtering technique known as Koopman-based lifting (see \citet{koopman1931hamiltonian}). 
-This technique takes observational data and a set of library functions as inputs and obtains the unknown coefficients $\{w_{,k}\}$.
+This technique takes observational data and a set of library functions as inputs and obtains the unknown coefficients $\{w_{j,k}\}$.
 The main steps of this technique are transforming the data (lifting the data),  applying a linear identification on the lifted data, and finally applying another transformation to bring down the results into the original vector field.
@@
 
-%To simulate our VAR, we need that the matrix $\textbf{A}$ is composed of assets that influence others (i.e $a_{i,j} \neq 0$) and some assets that don't ($a_{i,j} = 0$).
+%To simulate our VAR, we need that the matrix $\textbf{A}$ is composed of assets that influence others (i.e $a_{i,j} \neq 0$) and some assets that do not ($a_{i,j} = 0$).
 $\textbf{A}$ is constructed to allow all cases of potential influences between assets: i $\leftrightharpoons$ j, i $\rightarrow$ j, i $\leftarrow$ j, i independent of j.
@@
     \item \textit{Ideal portfolio:} In this scenario, $\mathcal{C}_j$ contains the portfolio $Q$, where $Q$ is defined in Lemma \ref{lemma:lin}.
-    For further discussion see Appendix. 
+    For further discussion see Appendix \ref{appendix2}. 
     %comprises of assets that influence $j$ except asset $i$, i.e., $\mathcal{C}_j=\{U\}$, where $Q=(Q_t)$ and $Q_t=\sum_{k\neq i}a_{i,k}X_{k,t}$.
@@
 Hence, we could not compare its performance with the other methods. 
-In this experiment, since the dynamic is linear and the noises are Gaussian, we use Equation \eqref{simueq} to estimate the DIs. 
+In this experiment, since the dynamics is linear and the noises are Gaussian, we use Equation \eqref{simueq} to estimate the DIs. 
 Finally, we obtain the adjacency matrix of the corresponding DIGs by comparing the estimated DIs with a threshold $\alpha>0$, i.e., 
@@
 where $\widehat{I}(X_i\rightarrow X_j||\mathcal{C}_j)$ denotes the estimated DI from $X_i$ to $X_j$ given the conditioning set $\mathcal{C}_j$. 
-In order to compare the performances of the aforementioned four approaches, we use the precision and recall measure between the true DIG (obtained from $\textbf{A}$) and their estimated DIGs. 
+In order to compare the performances of the aforementioned four approaches, we use the precision and recall measures between the true DIG (obtained from $\textbf{A}$) and their estimated DIGs. 
 Formally, the precision and the recall are defined by
@@
 \subsection{Non-Linear framework} \label{nonlinfr}
-To compare the performances of the different approaches from the previous section in non-linear environment, we simulate a set of quadratic processes whose dynamic is given below, 
+To compare the performances of the different approaches from the previous section in a non-linear environment, we simulate a set of quadratic processes whose dynamics is given below, 
 \begin{equation}\label{eq:ex_non}
@@
 This result once more fortifies our claim that with a RNN we can summarize the information of the network into one time series and use it for detecting the causal relationships. 
-This claim is due to Lemma \ref{lemma:nonlinear} and the universal approximation theorem which states that a neural network with enough hidden layers can approximate any non-linear function  \citet{HORNIK1989359}.
+This claim is due to Lemma \ref{lemma:nonlinear} and the universal approximation theorem which states that a neural network with enough hidden layers can approximate any non-linear function (see \citet{HORNIK1989359}).
 The slight difference between the performance of the \textit{RNN} and the \textit{true parents} is because of the estimation error in the recurrent neural network.
@@
 This result suggests that the market return is influencing a high number of firms, but the converse is not necessarily true.
-One way to intuitively interpret this finding is that bullish and/or bearish markets are likely to influence next period stock returns (as a persistence effect known in business cycles), but individual stock returns struggle at predicting next period market returns.
 %In each of the 5-years subsample, we use the excess returns of the 1000 biggest firms as conditional information to compute the DIs of the 30 biggest firms.

### 2022-08-08 main/paper3.tex
Add files via upload

@@
 In their pioneering paper, \citet{antweiler2004all} compute a bullishness measure out of 1.5 million messages posted on Yahoo!
-Finance and Raging Bull and finds that stock messages help predict market volatility.
+Finance and Raging Bull and find that stock messages help predict market volatility.
 Their results clearly reject the hypothesis that all that talk is just noise.
@@
 How to create a sentiment classifier that understands the vocabulary of the messages posted by the investors?
-For instance, ``bull" is an animal in everyday language but it is someone optimistic in the financial jargon.
-\citet{loughran2011liability} create a word list, which helps classify tone in a financial document.
+For instance, ``bull" is an animal in everyday language but refers to someone optimistic in the financial jargon.
+\citet{loughran2011liability} create a words list, which helps classify tone in a financial document.
 However, this might not be sufficient in the context of social media because messages posted present many typos, abbreviations and slang, so one needs to have an additional layer of data preprocessing.
@@
 
-Our paper overcomes these limitations.
+This chapter overcomes these limitations.
 We develop a machine learning algorithm to classify the sentiment of a large sample of StockTwits messages as bullish, bearish, or neutral.
@@
 Our method scales and performs very well.
-It achieves an out-of-sample accuracy of 85.9\%, which compares well to the anecdotal 80 to 85\% probability that human annotators agree on the sentiment of a document, see, e.g., \citet{wilson2005recognizing} and \citet{chen2020large}.
+It achieves an out-of-sample accuracy of 85.9\%, which compares well to the anecdotal 80 to 85\% probability that human annotators agree on the sentiment of a document (see, e.g., \citet{wilson2005recognizing} and \citet{chen2020large}).
 As a side product, we generate a vocabulary of one million investor sentiment-labeled terms consisting of up to three words that frequently appear in StockTwits messages.
@@
 This approach is very simple and efficient, and eliminates the class imbalance bias at the same time.
-It builds on any traditional binary classifiers.
+It builds on any traditional binary classifier.
 We use logistic regression on Term Frequency-Inverse Document Frequency (TFIDF)-vectorized messages.
@@
 
-Our paper contributes to the growing literature on machine learning classification of social media and its interaction with the stock market.
+This chapter contributes to the growing literature on machine learning classification of social media and its interaction with the stock market.
 Most previous financial studies use Twitter as their primary source of data.
@@
 
-Our paper is the first that analyzes the predictive power of StockTwits messages on stock returns unconditionally and around specific events.
+This chapter is the first work that analyzes the predictive power of StockTwits messages on stock returns unconditionally and around specific events.
 \citet{renault2017intraday} builds an intraday investor sentiment indicator using messages and finds that the change in investor sentiment of the first half-hour of a trading day helps forecast the last half-hour market return of that trading day.
@@
 They train a  sentiment score directly on returns.
-In contrast, we use user sentiment-labeled StockTwits messages as training and validation set for our sentiment classifier.
+In contrast, we use user sentiment-labeled StockTwits messages as the training and validation set for our sentiment classifier.
 
-Our paper also contributes to the EMH literature by gauging how cumulative average abnormal returns and abnormal polarities behave around sudden peaks of message activity.
+This chapter also contributes to the EMH literature by gauging how cumulative average abnormal returns and abnormal polarities behave around sudden peaks of message activity.
 
-The remainder of the paper is structured as follows.
+The remainder of the chapter is structured as follows.
 Section \ref{S:data} discusses StockTwits and stock market data.
@@
 In addition to that, social media slang is an additional layer of language complexity.
-To this extent, the functionality to self-tag bullish and bearish messages that StockTwits is extremely valuable as it allows the creation of a specific labeled vocabulary out of labeled messages.
+To this extent, the functionality to self-tag bullish and bearish messages on StockTwits is extremely valuable as it allows the creation of a specific labeled vocabulary out of labeled messages.
 We are not aware of any other social media platform in finance offering this functionality.
@@
 Every StockTwits message includes eight features: (1) the reference ticker(s), (2) a timestamp, (3) a unique message identifier, (4) the body of the message, (5) the sentiment label (bearish, bullish, or none) entered by the user, (6) a unique identifier of the user who posted the message, (7) the number of messages published by the user who posted the message, and (8) the number of followers of the user who posted the message.
-Our sentiment analyses builds on the first five features.
+Our sentiment analysis builds on the first five features.
 The last three provide additional information on the network structure, which we briefly discuss in Appendix \ref{app_userstats}.
@@
 
-Left plot of Figure \ref{fig:data-stocktwits2} shows the top 30 most discussed tickers on StockTwits.
+The left plot of Figure \ref{fig:data-stocktwits2} shows the top 30 most discussed tickers on StockTwits.
 SPY, a large ETF that tracks the S\&P 500 stock market index, is the most discussed ticker, followed by Apple and other big tickers.
 The messages about the 15 (30) biggest tickers represent 20\% (25\%) of the total number of messages, which indicates that users talk about a wide panel of tickers and not only big firms.
-Right graph shows a histogram of the number of messages per ticker.
+The right graph shows a histogram of the number of messages per ticker.
 The x-axis is log-scaled because due to extreme values the distribution is highly skewed.
@@
 
-Left plot of Figure \ref{fig:classif_time} shows the proportions of user sentiment-labeled messages across time.
+The left plot of Figure \ref{fig:classif_time} shows the proportions of user sentiment-labeled messages across time.
 In the early years of the platform, most messages were unlabeled, presumably because users were not familiar with the sentiment label yet.
@@
     \mycaption{Message classification}{Left plot shows the proportions of user-labeled messages: bullish (green), bearish (red), and unlabeled (gray) across time.
-Right plot shows the proportions of machine learning classified messages: bullish (light green predicted, green user-labeled), bearish (light red predicted, red user-labeled), and neutral (gray) across time.
+The right plot shows the proportions of machine learning classified messages: bullish (light green predicted, green user-labeled), bearish (light red predicted, red user-labeled), and neutral (gray) across time.
 Proportions are aggregated monthly.}
@@
 This ratio indicates that investors are on average optimistic about the market, which is consistent with findings in the literature, e.g., \citet{renault2017intraday}.
-Such an imbalance is a well-known issue in machine learning classification as it creates a bias towards the over-represented class, see \citet{chawla2004special}.
+Such an imbalance is a well-known issue in machine learning classification as it creates a bias towards the over-represented class (see \citet{chawla2004special}).
 There are various ways to tackle class imbalance.
@@
 As not every message carries a substantial stock market relevant sentiment, we deviate from the bullish--bearish dichotomy.
-Instead, we create an auxiliary neutral sentiment class to account of messages that do not take a clear stand.
+Instead, we create an auxiliary neutral sentiment class to account for messages that do not take a clear stand.
 See Appendix \ref{app_exampleclass} for examples of such neutral messages.
@@
 
-Right plot of Figure \ref{fig:classif_time} shows the proportions of our machine learning classifications across time.
+The right plot of Figure \ref{fig:classif_time} shows the proportions of our machine learning classifications across time.
 Percentages of bearish (user-labeled and classified as bearish) and bullish (user-labeled and classified as bearish) messages are stable over time, suggesting that our classification method is robust.
@@
 We next aggregate message sentiments on a daily ticker-level and across the market.
-Thereto, we denote by $C_{i,t,j}=1,0,-1$ for bullish, neutral, bearish, respectively, the sentiment of the $j$th message about ticker $i$ on day $t$.\footnote{We follow the close-to-close convention.
+Thereto, we denote $C_{i,t,j}=1,0,-1$ for bullish, neutral, bearish, respectively, the sentiment of the $j$th message about ticker $i$ on day $t$.\footnote{We follow the close-to-close convention.
 First, we remove all non-business days from the sample, whereby messages posted on non-business days count towards the next business day.
@@
 
-For the following time-series analysis and event study we restrict our sample.
+For the following time series analysis and event study we restrict our sample.
 There are two reasons for doing so.
@@
 
-Event studies constitute a statistical method widely used in financial econometrics, see, e.g., \citet{mackinlay1997event}.
+Event studies constitute a statistical method widely used in financial econometrics (see, e.g., \citet{mackinlay1997event}).
 In general, they are used to measure the effect of events on the market value of stocks.
@@
 \end{figure}
+
+\clearpage
 
@@
     \includegraphics[width=0.75\textwidth]{abnpol_distr_all.JPG}
-    \mycaption{Empirical distribution of abnormal polarities on event dates}{Red dashed lines show the one-third and two-third percentiles.}
+    \mycaption{Empirical distribution of abnormal polarities on event dates}{The red dashed lines show the one-third and two-third percentiles.}
     \label{fig:abnpol_all}
@@
     \includegraphics[width=0.75\textwidth]{events_across_time.JPG}
-    \mycaption{Number of events of each type across time}{Green area shows bullish events, gray area shows neutral events and red area shows bearish events. Numbers are aggregated monthly.}
+    \mycaption{Number of events of each type across time}{The green area shows bullish events, the gray area shows neutral events and the red area shows bearish events.
+Numbers are aggregated monthly.}
     \label{fig:events}
@@
 
-As an illustration, Figure \ref{fig:activityaapl} shows for Apple the time-series of message volume and the the corresponding events.
+As an illustration, Figure \ref{fig:activityaapl} shows for Apple the time series of message volume and the corresponding events.
 Between January 2011 and March 2020, our algorithm identified 73 events for Apple.
@@
     \mycaption{Daily message volume for Apple}{Events are days with an unusual high number of messages.
-Green upper-triangles show bullish events, gray circles are neutral events and red down-triangles represent bearish events.}
+The green upper-triangles show bullish events, gray circles are neutral events and red down-triangles represent bearish events.}
     \label{fig:activityaapl}
@@
 CAR has no predictive power on the type of the event: five days before an event, the median of the CAR distribution of the bullish events is not statistically different from the median of the neutral events.
-Same holds for the bearish events.
+The same holds for the bearish events.
 
@@
 How does sentiment polarity behave around events?
-Similar to the above, we use as benchmark model a one-year rolling window regression of the the daily polarity of ticker $i$ on the daily market polarity defined in \eqref{marketpoleq},
+Similar to the above, we use as benchmark model a one-year rolling window regression of the daily polarity of ticker $i$ on the daily market polarity defined in \eqref{marketpoleq},
 \begin{equation}
@@
 
-Right plot of Figure \ref{fig:caarcaap} shows the CAAP around the events.
+The right plot of Figure \ref{fig:caarcaap} shows the CAAP around the events.
 There are two main findings.
@@
 
+\clearpage
 To check statistical significance, we use the Mann-Whitney U-test (see \citet{utest} and \citet{sheskin}) to test whether the three samples (bullish, neutral and bearish) represent populations with different median values.\footnote{This interpretation only holds under stringent assumptions on the populations, namely that the two population distributions are equal up to a shift.
@@
 At the event date, the medians of the CAR shift as the abnormal returns jump for both bullish and bearish events.
-Also this is consistent with the EMH.
+This is also consistent with the EMH.
 Finally, 5 days after the event, the distributions of the CAR are very similar to the ones at the event date.
@@
 \end{equation}
-which is the running CAP over the last 14 days plus the current day $t$ (we rebalance the portfolio at the close on day $t$).\footnote{As above, we work here with the the restricted sample of 19 tickers and dates $t$ ranging through all business days of the sample period, excluding the first 14 days (for the CAP) and the last day (for the last portfolio holding period).} Note the difference to \eqref{eq_CAPtaut}.
-While we cannot predict the arrival of an event, we assume that the more $CAP_{i,t}$ deviates from zero the more likely there will be a an event on the next day.
+which is the running CAP over the last 14 days plus the current day $t$ (we rebalance the portfolio at the close on day $t$).\footnote{As above, we work here with the restricted sample of 19 tickers and dates $t$ ranging through all business days of the sample period, excluding the first 14 days (for the CAP) and the last day (for the last portfolio holding period).} Note the difference to \eqref{eq_CAPtaut}.
+While we cannot predict the arrival of an event, we assume that the more $CAP_{i,t}$ deviates from zero the more likely there will be an event on the next day.
 We will thus use $CAP_{i,t}$ as a baseline signal for market timing.
@@
 As a robustness check of our approach, we observe that the mean is well centered at zero.
-We also see a regime change in the early 2015.
+We also see a regime change in early 2015.
 In the first regime the standard deviation is much larger (and more volatile) than in the second regime.\footnote{We could not find an exogenous cause for this regime change.} Appendix~\ref{app_CAPport} contains the results for the 95\% ($x = 1.96$) and 99.5\% ($x = 2.81$) confidence intervals.
@@
 
-Based on these signals, we now construct reset CAP-sorted portfolios.
+Based on these signals, we now construct reset-CAP-sorted portfolios.
 Formally, we define the ticker sets $I^{bull}_t=\{ i\mid CAP_{i,t}^{(R)} > U_t(x)\}$ and $I^{bear}_t=\{ i\mid CAP_{i,t}^{(R)} < L_t(x)\}$.
@@
 If any of the index sets is empty, we set the corresponding return to zero.
-Top-left plot of Figure \ref{fig:pf} shows the cumulative log returns of bullish and bearish portfolios as well as the S\&P500.
+The top-left plot of Figure \ref{fig:pf} shows the cumulative log returns of bullish and bearish portfolios as well as the S\&P500.
 Overall, the portfolio performance is consistent with our approach: the bullish (bearish) portfolio outperforms (under-performs) the market.
-Remarkably, the upward (downward) steps suggests that our portfolio strategy succeeds to take the right positions just before an event.
+Remarkably, the upward (downward) steps suggest that our portfolio strategy succeeds to take the right positions just before an event.
 The remaining plots of Figure \ref{fig:pf} show the number of positions across time of our portfolios.
@@
 
-We extract 90 million messages from StockTwits from January 2010 to March 2020, covering a large sample of US and Canadian stocks.
+We extract a large sample of messages from StockTwits from January 2010 to March 2020, covering US and Canadian stocks.
 Messages are either user-labeled as bullish or bearish or left unlabeled.
@@
 We observe a 5-to-1 bullish-to-bearish ratio, indicating that investors are on average optimistic.
-We build time-series of daily sentiment polarity for individual tickers and the aggregate market.
+We build time series of daily sentiment polarity for individual tickers and the aggregate market.
 We show that daily polarity is positively associated to contemporaneous stock returns, but this result loses its significance against next-day returns.
-We then define events as days with sudden peak of message volume and relate them to corporate and stock market events.
-We show that cumulative abnormal polarity has significant predictive power on the type of event, in contract to cumulative abnormal returns.
+We then define events as days with sudden peaks of message volume and relate them to corporate and stock market events.
+We show that cumulative abnormal polarity has significant predictive power on the type of event, in contrast to cumulative abnormal returns.
 We also note that investor sentiment about a ticker tends to be biased towards recent past events.

### 2022-08-08 head/abstracts.tex
Add files via upload

@@
 These causal graphs can be used as preliminary feature selection for another predictive model or for systemic risk management.
-We evaluate the performance of our algorithm using both synthetic linear and nonlinear experiments and apply it to the daily stock returns of U.S.
-listed firms and infer their interconnections from 1990 to 2020.
+We evaluate the performance of our algorithm using both synthetic linear and nonlinear experiments and apply it to the daily stock returns of US listed firms and infer their interconnections from 1990 to 2020.
 \\
@@
 
-Cette thèse consiste en trois applications de techniques d'apprentissage automatique à la gestion des risques. \\
+Cette thèse est composée de trois applications de techniques d'apprentissage automatique à la gestion des risques. \\
 

### 2022-08-08 head/acknowledgements.tex
Add files via upload

@@
 
-I wish to express my deepest gratitude to my supervisor, Damir Filipović, for his availability and continuous support during these five years.
+I wish to express my deepest gratitude to my supervisor, Professor Damir Filipović, for his availability and continuous support during these five years.
 This thesis greatly benefited from his countless valuable comments and flow of smart ideas.
@@
 
-Special thanks go to professor Michael Rockinger who convinced me to pursue a doctorate degree.
+Special thanks go to Professor Michael Rockinger who convinced me to pursue a doctorate degree.
 He took me under his wing since my Bachelor studies and helped me becoming the researcher I am today.
@@
 
-I would like to acknowledge the rest of my thesis committee, the professors Pierre Collin-Dufresne, Andreas Fuster and Simon Scheidegger for their advice and comments.
+I would like to acknowledge the rest of my thesis committee, Professors Pierre Collin-Dufresne, Andreas Fuster and Simon Scheidegger for their advice and comments.
 
-I am also thankful to the professors Elena Perazzi and Damien Challet. I was their TA for several years and they helped me become a better teacher.
+I am also thankful to Professors Elena Perazzi and Damien Challet. I was their TA for several years and they helped me become a better teacher.
 
@@
 I am thankful to my colleagues and friends Antoine Didisheim and Coralie Jaunin for the many chess games and `bridge tours` around the campus.
+I am also very grateful to my longstanding friend Christophe George for the proofreading of this thesis.
 

### 2022-08-08 tail/appendix.tex
Add files via upload

@@
 However, the likelihoods have to be slightly updated to be compatible with the neural network framework.
-That is, ($\lambda$, $\mu$) is the set of parameters in the neural network for the forward intensity of default $f_{it}$ and one for $h_{it}$ with the combined exit forward intensity being $g_{it} = f_{it} + h_{it}$.
+One neural network is trained to compute $f_{it}$ and the other is trained to output $h_{it}$ where $g_{it} = f_{it} + h_{it}$.
+Let us denote $\lambda$ and $\mu$ the set of parameters (weights) tuned in the neural network for $f_{it}$ and $h_{it}$ respectively.
 I impose non-negativity on both $f_{it}$ and $h_{it}$ to ensure that the combined exit intensity is at least bigger than the forward default intensity.
@@
     &+ \textbf{1}_{t_{0_i} \leq t, \tau_{D_i}=\tau_{C_i} \leq t+\tau} \cdot \mathbb{P}_t(t+\tau < \tau_{D_i} = \tau_{C_i} \leq t + \tau + 1) \\ \nonumber
-    &+ \textbf{1}_{t_{0_i} \leq t, \tau_{C_i} \leq t+\tau, \tau_{D_i} \neq \tau_{C_i}} \cdot \mathbb{P}_t(t+\tau < \tau_{C_i} \neq \tau_{C_i} \leq t+\tau + 1) \\ \nonumber
+    &+ \textbf{1}_{t_{0_i} \leq t, \tau_{C_i} \leq t+\tau, \tau_{D_i} \neq \tau_{C_i}} \cdot \mathbb{P}_t(t+\tau < \tau_{D_i} \neq \tau_{C_i} \leq t+\tau + 1) \\ \nonumber
     &+ \textbf{1}_{t_{0_i}>t} + \textbf{1}_{\tau_{C_i} \leq t},
@@
     \end{cases} \\
-    \mathbb{P}_t(t+\tau < \tau_{C_i} \neq \tau_{C_i} \leq t+\tau + 1)&=
+    \mathbb{P}_t(t+\tau < \tau_{D_i} \neq \tau_{C_i} \leq t+\tau + 1)&=
     \begin{cases}
@@
 The first decomposition consists of separating terms involving $f$ and terms involving $g$.
-The second decomposition consists on separating terms corresponding to different $\tau$.
+The second decomposition consists of separating terms corresponding to different $\tau$.
 In the end, we get two likelihood functions to estimate for each horizon.
 For each horizon the parameters of the likelihood function involving $f$ are estimated in a neural network with output $N_{it}^{(\lambda)}$.
-$g$ is assumed to be of the form $f + h$ as in previous literature.
 The parameters of the likelihood function involving $h$ are estimated in a neural network with output $N_{it}^{(\mu)}$.
+The combined exit forward intensity $g$ is assumed to be of the form $f + h$ as in previous literature.
 Since $\exp(-f)-\exp(-g) = \exp(-f)-\exp(-f-h) = \exp(-f)*(1-\exp(-h))$, the first decomposition is the following :
@@
 \end{align}
-Similar to previous literature, we can still decompose the likelihoods into terms involving different horizon of prediction $\tau$.
+Similar to previous literature, we can still decompose the likelihoods into terms involving different horizons of prediction $\tau$.
 For each $\tau$, we can consider as constant all terms involving previous horizons forward intensities.
@@
 
-Let $V_T$ be the firm value at time T, $L$ the amount of debt to be repaid and $E_T$ the equity value at time T.
-In case of bankruptcy debt holders receive money before shareholders in case of bankruptcy.
+Let $V_T$ be the firm value at time $T$, $L$ the amount of debt to be repaid and $E_T$ the equity value at time $T$.
+In case of bankruptcy, debt holders receive money before shareholders.
 The payoff to shareholders is then given by $E_T = \max(V_T - L,0)$.
 This payoff is the same as a call option $E_T$ on the underlying $V_T$ with the strike being $L$.
-\citet{Merton1974} model considers $V_t$ is following a geometric Brownian motion :
+\citet{Merton1974} model considers that $V_t$ is following a geometric Brownian motion :
 \begin{equation*}
@@
 \end{equation*}
-We can then use Black-Scholes formula to get the option price and firm's equity value $E_t$ at any time t :
+We can then use Black-Scholes formula to get the option price and firm's equity value $E_t$ at any time $t$ :
 \begin{equation} \label{blackscholes}
@@
 \end{equation}
-In the Figure \ref{fig:DD} taken from \citet{KMV}, the distance-to-default is denoted by ``DD".
+In Figure \ref{fig:DD} taken from \citet{KMV}, the distance-to-default is denoted by ``DD".
 The higher the asset value at horizon H the greater the distance-to-default will get, which lowers the default probability since the firm is more likely to be able to repay the debt owed.
@@
 The variance restriction method is used in \citet{DSW} to estimate the distance-to-default (DtD).
-This method is based on \citet{Merton1974} model which states that firm's equity value can be seen as a call option on the underlying asset and the strike being the amount of debt.
+This method is based on \citet{Merton1974} model which states that the firm's equity value can be seen as a call option on the underlying asset and the strike being the amount of debt.
 This holds because stockholders receive money only once debt holders are fully paid.
@@
 \end{equation}
-However, many papers agree to say that $\mu$ is very tedious to estimate. Hence, DtD is often computed as
+However, many papers agree that $\mu$ is very tedious to estimate. Hence, DtD is often computed as
 \begin{equation}
@@
 $\delta$ for a whole industry).
-However, we will improve the methodology by computing $\delta$ for each firm individually.
+However, we improve the methodology by computing $\delta$ for each firm individually.
 We should obtain higher deltas for financial firms than for non-financial firms.
@@
 Using this kind of likelihood is complex and very time consuming because we have to solve many inverse Black-Scholes formulas to get the time values of implied asset value $\hat{V}_t(\sigma,\delta)$ for each firm by solving equation \ref{blackscholes}.
-However, inverse Black-Scholes formula doesn't have any closed form solution.
+However, inverse Black-Scholes formula does not have any closed form solution.
 The optimization is even more tedious since the implied asset value $\hat{V}_t(\sigma,\delta)$ depends on the final output of the likelihood $\delta$.
-$A_t$ is the book asset value and $h_t$ is the ``length of time between two consecutive equity values measured in trading days as a fraction of a year".
+$A_t$ is the book asset value and $h_t$ is the time interval as a fraction of a year between two observations.
 $h_t$ in the model is set to be 0.25 since we performed a linear interpolation when we had a missing value in the sample (see ``Missing information").
@@
 
+\begin{comment}
 \subsection{k-Nearest Neighbors estimator of mutual information}
@@
 \end{small}
-Similarly, we obtain k-nearest density estimators $\widehat{\PP}_{X,Z}, \widehat{\PP}_{Y,Z}$, and $\widehat{\PP}_{Z}$.
+\noindent Similarly, we obtain k-nearest density estimators $\widehat{\PP}_{X,Z}, \widehat{\PP}_{Y,Z}$, and $\widehat{\PP}_{Z}$.
 Subsequently, the $N$ samples of $\mathcal{S}_1$ is used to approximate the conditional mutual information:
@@
 \end{small}
-For more details corresponding this estimator including its bias, variance, and confidence, please see the works by \cite{sricharan2011k} and \cite{loftsgaarden1965nonparametric}.
+\noindent For more details corresponding this estimator including its bias, variance, and confidence, please see the works by \cite{sricharan2011k} and \cite{loftsgaarden1965nonparametric}.
+\end{comment}
 

### 2022-09-21 main/paper2.tex
Add files via upload

@@
 Otherwise, when the forecast does not improve by including the information of $X$, then it is declared that $X$ does not cause $Y$.
-This idea is reflected in the information-theoretic measure that we use in this work to infer the causal interactions among a network of time series.
+This idea is reflected in the information-theoretic measure that we use in this chapter to infer the causal interactions among a network of time series.
 
@@
 In order to identify the influence of $X_t$ on $Y_t$, Granger's idea is to compare the performance of two linear regressions\footnote{Note that this formulation is only applicable in linear systems.}: the first one predicts $Y_t$ given $\{X^{t-1},Y^{t-1},Z^{t-1}\}$ and the second one predicts $Y_t$ given $\{Y^{t-1},Z^{t-1}\}$.
-If they have the same performance, then we say $X$ does not Granger-cause $Y$.
+If they perform similarly, then we say $X$ does not Granger-cause $Y$.
 
 To go beyond linear systems, works such as \citet{directed} and \citet{massey} use information-theoretical measures and generalize Granger-causality. 
-In this work, we introduce and apply directed information (see \citet{directed}), an information-theoretical tool to measure interconnections among firms.
+In this chapter, we introduce and apply directed information (see \citet{directed}), an information-theoretical tool to measure interconnections among firms.
 Directed information (DI) has been used in many applications to infer causal relationships.
@@
 To rigorously formalize this idea, we need the predictors and a measure to compare their performances.
+
+\clearpage
 
@@
 The average regret in \eqref{eq:ave} is the directed information (DI). We use it as the measure of causation in this chapter.
-This measure is always positive and if it is zero, it indicates that the history of time series $X$ contains no significant information that would help in predicting the future of time series $Y$ given the history of $Y$ and $Z$.
-This definition can be generalized to more than three time series as follows,
+This measure is always positive and if it is zero, it is an indication that the history of the time serie $X$ does not contain significant information helping in the prediction of the time serie $Y$ given the past of the time series $Y$ and $Z$.
+We can generalize this definition to more than three time series as follows,
 
@@
  \begin{definition}\label{cdi}
-Consider a network of $m$ time series $\mathcal{R}=\{R_1,...,R_m\}$ with the joint probability density function $p$. The directed information from $R_i$ to $R_j$ over time horizon $[1,T]$ is given by
+Consider a network of $m$ time series $\mathcal{R}=\{R_1,...,R_m\}$ with the joint probability density function $p$.
+The directed information from $R_i$ to $R_j$ over the time interval $[1,T]$ is given by
 \begin{align}\label{cdif}
@@
 where $\mathcal{R}_{-\{i,j\}}^{t-1}\!:=\!\{R_{1}^{t-1},...,R_{m}^{t-1}\}\!\setminus\!\{R_{i}^{t-1},R_{j}^{t-1}\}$.
-We declare $R_i$ influences $R_j$ over time horizon $[1,T]$, if and only if
+We say that $R_i$ has influence on $R_j$ over the time interval $[1,T]$, if and only if
 \begin{align}\label{cdif1}
@@
  In another words, if $R_i$ does not influence $R_j$, then varying $R_i$ would not change $R_j$ when the values of the remaining times series are fixed. 
- This can be seen from the fact that DI compares two conditional distributions of $R_j$ over a time horizon of length $T$; one is given the history of all the time series while the other one is given all the history except the history of $R_i$.
+ This can be seen from the fact that DI compares two conditional distributions of $R_j$ over a time horizon of length $T$; one is conditioned on the past of all time series while the other one is conditioned on all the history except $R_i$.
 Thus, if DI in \eqref{cdif} is zero, then these two conditional distributions are equal over this time horizon.
@@
 Note that the definition of DI does not rely on any model assumption, thus DI is capable of inferring the causal relationships in general (linear or non-linear) dynamical systems.
-Next, we define the graphical model that we use in this work to visualize the causal network among firms.
+Next, we define the graphical model that we use in this chapter to visualize the causal network among firms.
 
 \begin{definition}\label{DIG1}
-Directed information graph (DIG) of a set of $m$ time series $\mathcal{R}=\{R_1,...,R_m\}$ is a directed graph $G=(\mathcal{V},\mathcal{E})$, where nodes represent time series ($\mathcal{V}=\mathcal{R}$) and arrow $(R_i,R_j)\in \mathcal{E}$ denotes that $R_i$ influences $R_j$.
+We denote by DIG the directed information graph of a set of $m$ time series $\mathcal{R}=\{R_1,...,R_m\}$ which consists of a directed graph $G=(\mathcal{V},\mathcal{E})$, where $\mathcal{V}$ stands for the set of nodes representing time series $\mathcal{R}$ while $\mathcal{E}$ stands for the set of arrows indicating influences between time series (i.e.
+an arrow from $R_i$ to  $R_j$ indicates that $R_i$ has an influence on $R_j$).
 %with weight $w_{i,j}:=I(R_i\rightarrow R_j||\mathcal{R}_{-\{i,j\}})$. 
@@
 
-In this work, we propose a new method that reduces the size of the conditioning set in \eqref{cdif} to only one for any given network while introducing less approximation error compared to the prior works.
+In this chapter, we propose a new method that reduces the size of the conditioning set in \eqref{cdif} to only one for any given network while introducing less approximation error compared to the prior works.
 In this method, we estimate the directed information from $R_i$ to $R_j$ by conditioning on an auxiliary time series.
@@
 
- Given the aforementioned result, we train a RNN using the observational data to estimate the time series $Q$ defined in Lemma \ref{lemma:nonlinear}.
+Given the aforementioned result, we train a RNN with LSTM layers using the observational data to estimate the time series $Q$ defined in Lemma \ref{lemma:nonlinear}.
  More precisely, our RNN maps $\mathcal{X}_{-\{i\}}^{t-1}$ as the inputs to $X_{j,t}$ as the output.
+We choose the architecture of the network to be pyramid-like.
+That is, the width of layer $k$ is strictly bigger than the width of layer $k+1$.
+The pyramidal structure is known to retain and improve accuracy while reducing the number of parameters (see \citet{ullah2016pyramid} and \citet{TRIPATHY2018565}).
+In our case, the first hidden layer has 100 units, the second hidden layer has 50 units, the third hidden layer has 40 units and the fourth hidden layer has 30 units.
+In every layer, the activation function that we use for the recurrent step (i.e.
+: forget, input and output gates) is the sigmoid function while the activation function for the cell and hidden states is the hyperbolic tangent.
+Finally, to avoid overfitting and to allow the network to generalize better on out-of-sample data, we use dropout on every layer.
+Dropout is a regularization technique often used in neural networks where connections within the LSTM network are randomly selected and excluded from updates in the training process.
+This has the effect of introducing noise in the training process because every training step is performed with a different network layout and it allows more nodes to be involved.
+In our case, the fraction of units that we drop (i.e : the dropout rate) in each layer is 0.5, 0.3, 0.2, 0.1, respectively.
+
 Let $R_j(\mathcal{X}_{-\{i\}}^{t-1};\Theta^*)$ denotes the trained RNN with parameters $\Theta^*$. 
@@
 To study the effect of the conditioning set on detecting the influences, in our experiments, we consider four different conditioning sets.
-More precisely, to measure whether asset $i$ influences asset $j$, we estimate $I(X_i\rightarrow X_j||\mathcal{C}_j)$ for the following choices of the conditioning set; 
+More precisely, to measure whether asset $i$ influences asset $j$, we estimate $I(X_i\rightarrow X_j||\mathcal{C}_j)$ for the following choices of the conditioning set:
+
+\clearpage
+
 \begin{enumerate}
@@
 Note that in this experiment, the true DIGs of the market are not known, hence, we could not compute the precision-recall curves.
+
+\clearpage
 
